# Prerequisites {-}

This book assumes a basic understanding of [the R language](http://www.r-project.org/).  If you like my style of pedagogy, you could try watching [my introductory video lectures](http://www.twotorials.com/).  Otherwise, review the R learning options at [flowingdata.com](http://flowingdata.com/2012/06/04/resources-for-getting-started-with-r/).

The R `lodown` package depends on most of the packages used in this text, so these three lines should install many of the R packages presented here, including the `survey` and `RSQLite` R packages.  Windows users may need to install the external software [Rtools](http://cran.r-project.org/bin/windows/Rtools/) prior to installing `lodown` from [github](http://github.com/).

```{r eval = FALSE}
install.packages( "devtools" , repos = "http://cran.rstudio.com/" )
library(devtools)
install_github( "ajdamico/lodown" , dependencies = TRUE )
```

The survey microdata presented in this book require the [R survey package](http://r-survey.r-forge.r-project.org/survey/) by [Dr. Thomas Lumley](https://www.stat.auckland.ac.nz/people/tlum005) at the University of Auckland.  Dr. Lumley wrote [a textbook](http://r-survey.r-forge.r-project.org/svybook/) to showcase that software.

The R `convey` package estimates measures of inequality, poverty and wellbeing.  Guilherme Jacob, Dr. Djalma Pessoa, and I have written [this book](https://guilhermejacob.github.io/context/) about inequality measurement with complex survey microdata to accompany the software.

```{r eval = FALSE}
install.packages( "convey" , repos = "http://cran.rstudio.com/" )
```

The R `srvyr` package by [Greg Freedman Ellis](https://github.com/gergness/) allows [dplyr](https://cran.r-project.org/web/packages/dplyr/README.html)-like syntax for many `survey` package commands.  For detailed usage examples, review his [vignettes](https://cran.r-project.org/web/packages/srvyr/index.html).

```{r eval = FALSE}
install.packages( "srvyr" , repos = "http://cran.rstudio.com/" )
```
