# Area Health Resource File (AHRF) {-}

<a href="https://github.com/asdfree/ahrf/actions"><img src="https://github.com/asdfree/ahrf/actions/workflows/r.yml/badge.svg" alt="Github Actions Badge"></a>

Though not a survey data set itself, useful to merge onto other microdata.

* One table with one row per county and a second table with one row per state.

* Replaced annually with the latest available county- and state-level statistics.

* Compiled by the United States [Health Services and Resources Administration (HRSA)](http://www.hrsa.gov/).

## Simplified Download and Importation {-}

The R `lodown` package easily downloads and imports all available AHRF microdata by simply specifying `"ahrf"` with an `output_dir =` parameter in the `lodown()` function. Depending on your internet connection and computer processing speed, you might prefer to run this step overnight.

```{r eval = FALSE }
library(lodown)
lodown( "ahrf" , output_dir = file.path( path.expand( "~" ) , "AHRF" ) )
```

## Analysis Examples with base R \ {-}

Load a data frame:

```{r eval = FALSE }
ahrf_df <- readRDS( file.path( path.expand( "~" ) , "AHRF" , "county/AHRF_2016-2017.rds" ) )
```

```{r eval = FALSE }

```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE }
ahrf_df <- 
	transform( 
		ahrf_df , 
		
		cbsa_indicator_code = 
			factor( 
				1 + as.numeric( f1406715 ) , 
				labels = c( "not metro" , "metro" , "micro" ) 
			) ,
			
		mhi_2014 = f1322614 ,
		
		whole_county_hpsa_2016 = as.numeric( f0978716 == 1 ) ,
		
		census_region = 
			factor( 
				as.numeric( f04439 ) , 
				labels = c( "northeast" , "midwest" , "south" , "west" ) 
			)

	)
	
```

### Unweighted Counts {-}

Count the unweighted number of records in the table, overall and by groups:
```{r eval = FALSE , results = "hide" }
nrow( ahrf_df )

table( ahrf_df[ , "cbsa_indicator_code" ] , useNA = "always" )
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
mean( ahrf_df[ , "mhi_2014" ] , na.rm = TRUE )

tapply(
	ahrf_df[ , "mhi_2014" ] ,
	ahrf_df[ , "cbsa_indicator_code" ] ,
	mean ,
	na.rm = TRUE 
)
```

Calculate the distribution of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
prop.table( table( ahrf_df[ , "census_region" ] ) )

prop.table(
	table( ahrf_df[ , c( "census_region" , "cbsa_indicator_code" ) ] ) ,
	margin = 2
)
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
sum( ahrf_df[ , "mhi_2014" ] , na.rm = TRUE )

tapply(
	ahrf_df[ , "mhi_2014" ] ,
	ahrf_df[ , "cbsa_indicator_code" ] ,
	sum ,
	na.rm = TRUE 
)
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
quantile( ahrf_df[ , "mhi_2014" ] , 0.5 , na.rm = TRUE )

tapply(
	ahrf_df[ , "mhi_2014" ] ,
	ahrf_df[ , "cbsa_indicator_code" ] ,
	quantile ,
	0.5 ,
	na.rm = TRUE 
)
```

### Subsetting {-}

Limit your `data.frame` to California:
```{r eval = FALSE , results = "hide" }
sub_ahrf_df <- subset( ahrf_df , f12424 == "CA" )
```
Calculate the mean (average) of this subset:
```{r eval = FALSE , results = "hide" }
mean( sub_ahrf_df[ , "mhi_2014" ] , na.rm = TRUE )
```

### Measures of Uncertainty {-}

Calculate the variance, overall and by groups:
```{r eval = FALSE , results = "hide" }
var( ahrf_df[ , "mhi_2014" ] , na.rm = TRUE )

tapply(
	ahrf_df[ , "mhi_2014" ] ,
	ahrf_df[ , "cbsa_indicator_code" ] ,
	var ,
	na.rm = TRUE 
)
```

### Regression Models and Tests of Association {-}

Perform a t-test:
```{r eval = FALSE , results = "hide" }
t.test( mhi_2014 ~ whole_county_hpsa_2016 , ahrf_df )
```

Perform a chi-squared test of association:
```{r eval = FALSE , results = "hide" }
this_table <- table( ahrf_df[ , c( "whole_county_hpsa_2016" , "census_region" ) ] )

chisq.test( this_table )
```

Perform a generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	glm( 
		mhi_2014 ~ whole_county_hpsa_2016 + census_region , 
		data = ahrf_df
	)

summary( glm_result )
```

## Analysis Examples with `dplyr` \ {-}

The R `dplyr` library offers an alternative grammar of data manipulation to base R and SQL syntax. [dplyr](https://github.com/tidyverse/dplyr/) offers many verbs, such as `summarize`, `group_by`, and `mutate`, the convenience of pipe-able functions, and the `tidyverse` style of non-standard evaluation. [This vignette](https://cran.r-project.org/web/packages/dplyr/vignettes/dplyr.html) details the available features. As a starting point for AHRF users, this code replicates previously-presented examples:

```{r eval = FALSE , results = "hide" }
library(dplyr)
ahrf_tbl <- tbl_df( ahrf_df )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
ahrf_tbl %>%
	summarize( mean = mean( mhi_2014 , na.rm = TRUE ) )

ahrf_tbl %>%
	group_by( cbsa_indicator_code ) %>%
	summarize( mean = mean( mhi_2014 , na.rm = TRUE ) )
```


