# American National Election Study (ANES) {-}

[![Build Status](https://travis-ci.org/asdfree/anes.svg?branch=master)](https://travis-ci.org/asdfree/anes) [![Build status](https://ci.appveyor.com/api/projects/status/github/asdfree/anes?svg=TRUE)](https://ci.appveyor.com/project/ajdamico/anes)

The American National Election Study (ANES) collects information on political belief and behavior from eligible voters in the United States.

* Most tables contain one row per sampled eligible voter.

* A complex sample survey designed to generalize to eligible voters in the United States.

* Time series studies released biennially.

* Administered by [Stanford University](https://iriss.stanford.edu/) and the [University of Michigan](https://www.isr.umich.edu/cps/index.html) and funded by the [National Science Foundation](https://www.nsf.gov).

## Simplified Download and Importation {-}

The R `lodown` package easily downloads and imports all available ANES microdata by simply specifying `"anes"` with an `output_dir =` parameter in the `lodown()` function. Depending on your internet connection and computer processing speed, you might prefer to run this step overnight.

```{r eval = FALSE }
library(lodown)
lodown( "anes" , output_dir = file.path( path.expand( "~" ) , "ANES" ) , 
	your_email = "email@address.com" )
```

`lodown` also provides a catalog of available microdata extracts with the `get_catalog()` function. After requesting the ANES catalog, you could pass a subsetted catalog through the `lodown()` function in order to download and import specific extracts (rather than all available extracts).

```{r eval = FALSE , results = "hide" }
library(lodown)
# examine all available ANES microdata files
anes_cat <-
	get_catalog( "anes" ,
		output_dir = file.path( path.expand( "~" ) , "ANES" ) , 
		your_email = "email@address.com" )

# 2016 only
anes_cat <- subset( anes_cat , directory == "2016 Time Series Study" )
# download the microdata to your local computer
lodown( "anes" , anes_cat , 
	your_email = "email@address.com" )
```

## Analysis Examples with the `survey` library {-}

Construct a complex sample survey design:

```{r eval = FALSE }

```

```{r eval = FALSE }
library(survey)

anes_df <- 
	readRDS( 
		file.path( path.expand( "~" ) , "ANES" , 
			"2016 Time Series Study/anes_timeseries_2016_.rds" )
	)

anes_design <-
	svydesign( 
		~v160202 , 
		strata = ~v160201 , 
		data = anes_df , 
		weights = ~v160102 , 
		nest = TRUE 
	)

```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE }
anes_design <- 
	update( 
		anes_design , 
		
		one = 1 ,
		
		pope_francis_score = ifelse( v162094 %in% 0:100 , v162094 , NA ) ,

		christian_fundamentalist_score = ifelse( v162095 %in% 0:100 , v162095 , NA ) ,
		
		primary_voter = ifelse( v161021 %in% 1:2 , as.numeric( v161021 == 1 ) , NA ) ,

		think_gov_spend =
			factor( v161514 , levels = 1:4 , labels =
				c( 'foreign aid' , 'medicare' , 'national defense' , 'social security' )
			) ,
		
		undoc_kids =
			factor( v161195x , levels = 1:6 , labels =
				c( 'should sent back - favor a great deal' ,
					'should sent back - favor a moderate amount' ,
					'should sent back - favor a little' ,
					'should allow to stay - favor a little' ,
					'should allow to stay - favor a moderate amount' ,
					'should allow to stay - favor a great deal' )
			)

	)
```

### Unweighted Counts {-}

Count the unweighted number of records in the survey sample, overall and by groups:
```{r eval = FALSE , results = "hide" }
sum( weights( anes_design , "sampling" ) != 0 )

svyby( ~ one , ~ undoc_kids , anes_design , unwtd.count )
```

### Weighted Counts {-}
Count the weighted size of the generalizable population, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ one , anes_design )

svyby( ~ one , ~ undoc_kids , anes_design , svytotal )
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ pope_francis_score , anes_design , na.rm = TRUE )

svyby( ~ pope_francis_score , ~ undoc_kids , anes_design , svymean , na.rm = TRUE )
```

Calculate the distribution of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ think_gov_spend , anes_design , na.rm = TRUE )

svyby( ~ think_gov_spend , ~ undoc_kids , anes_design , svymean , na.rm = TRUE )
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ pope_francis_score , anes_design , na.rm = TRUE )

svyby( ~ pope_francis_score , ~ undoc_kids , anes_design , svytotal , na.rm = TRUE )
```

Calculate the weighted sum of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ think_gov_spend , anes_design , na.rm = TRUE )

svyby( ~ think_gov_spend , ~ undoc_kids , anes_design , svytotal , na.rm = TRUE )
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svyquantile( ~ pope_francis_score , anes_design , 0.5 , na.rm = TRUE )

svyby( 
	~ pope_francis_score , 
	~ undoc_kids , 
	anes_design , 
	svyquantile , 
	0.5 ,
	ci = TRUE ,
	keep.var = TRUE ,
	na.rm = TRUE
)
```

Estimate a ratio:
```{r eval = FALSE , results = "hide" }
svyratio( 
	numerator = ~ christian_fundamentalist_score , 
	denominator = ~ pope_francis_score , 
	anes_design ,
	na.rm = TRUE
)
```

### Subsetting {-}

Restrict the survey design to party id: independent:
```{r eval = FALSE , results = "hide" }
sub_anes_design <- subset( anes_design , v161158x == 4 )
```
Calculate the mean (average) of this subset:
```{r eval = FALSE , results = "hide" }
svymean( ~ pope_francis_score , sub_anes_design , na.rm = TRUE )
```

### Measures of Uncertainty {-}

Extract the coefficient, standard error, confidence interval, and coefficient of variation from any descriptive statistics function result, overall and by groups:
```{r eval = FALSE , results = "hide" }
this_result <- svymean( ~ pope_francis_score , anes_design , na.rm = TRUE )

coef( this_result )
SE( this_result )
confint( this_result )
cv( this_result )

grouped_result <-
	svyby( 
		~ pope_francis_score , 
		~ undoc_kids , 
		anes_design , 
		svymean ,
		na.rm = TRUE 
	)
	
coef( grouped_result )
SE( grouped_result )
confint( grouped_result )
cv( grouped_result )
```

Calculate the degrees of freedom of any survey design object:
```{r eval = FALSE , results = "hide" }
degf( anes_design )
```

Calculate the complex sample survey-adjusted variance of any statistic:
```{r eval = FALSE , results = "hide" }
svyvar( ~ pope_francis_score , anes_design , na.rm = TRUE )
```

Include the complex sample design effect in the result for a specific statistic:
```{r eval = FALSE , results = "hide" }
# SRS without replacement
svymean( ~ pope_francis_score , anes_design , na.rm = TRUE , deff = TRUE )

# SRS with replacement
svymean( ~ pope_francis_score , anes_design , na.rm = TRUE , deff = "replace" )
```

Compute confidence intervals for proportions using methods that may be more accurate near 0 and 1. See `?svyciprop` for alternatives:
```{r eval = FALSE , results = "hide" }
svyciprop( ~ primary_voter , anes_design ,
	method = "likelihood" , na.rm = TRUE )
```

### Regression Models and Tests of Association {-}

Perform a design-based t-test:
```{r eval = FALSE , results = "hide" }
svyttest( pope_francis_score ~ primary_voter , anes_design )
```

Perform a chi-squared test of association for survey data:
```{r eval = FALSE , results = "hide" }
svychisq( 
	~ primary_voter + think_gov_spend , 
	anes_design 
)
```

Perform a survey-weighted generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	svyglm( 
		pope_francis_score ~ primary_voter + think_gov_spend , 
		anes_design 
	)

summary( glm_result )
```

## Analysis Examples with `srvyr` {-}

The R `srvyr` library calculates summary statistics from survey data, such as the mean, total or quantile using [dplyr](https://github.com/tidyverse/dplyr/)-like syntax. [srvyr](https://github.com/gergness/srvyr) allows for the use of many verbs, such as `summarize`, `group_by`, and `mutate`, the convenience of pipe-able functions, the `tidyverse` style of non-standard evaluation and more consistent return types than the `survey` package. [This vignette](https://cran.r-project.org/web/packages/srvyr/vignettes/srvyr-vs-survey.html) details the available features. As a starting point for ANES users, this code replicates previously-presented examples:

```{r eval = FALSE , results = "hide" }
library(srvyr)
anes_srvyr_design <- as_survey( anes_design )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
anes_srvyr_design %>%
	summarize( mean = survey_mean( pope_francis_score , na.rm = TRUE ) )

anes_srvyr_design %>%
	group_by( undoc_kids ) %>%
	summarize( mean = survey_mean( pope_francis_score , na.rm = TRUE ) )
```

---

## Replication Example {-}

```{r eval = FALSE , results = "hide" }

```

