# Basic Stand Alone Public Use File (BSAPUF) {-}

[![Build status](https://sisyphus.project.cwi.nl/?badge=bsapuf&host=stones06fedora)](http://sisyphus.project.cwi.nl/) [![Build Status](https://sisyphus.project.cwi.nl/?badge=bsapuf&host=stones14win)](http://sisyphus.project.cwi.nl/)

The CMS Basic Stand Alone Public Use File (BSAPUF) contains a five percent sample of Medicare beneficiary spending and utilization in the enrolled population.

* Multiple non-linkable tables, each with one row per beneficiary event.

* The population of elderly and disabled individuals covered by [fee-for-service](https://en.wikipedia.org/wiki/Fee-for-service) Medicare in the United States.

* No listed update frequency.

* Maintained by the United States [Centers for Medicare & Medicaid Services (CMS)](http://www.cms.gov/)

## Simplified Download and Importation {-}

The R `lodown` package easily downloads and imports all available BSAPUF microdata by simply specifying `"bsapuf"` with an `output_dir =` parameter in the `lodown()` function. Depending on your internet connection and computer processing speed, you might prefer to run this step overnight.

```{r eval = FALSE }
library(lodown)
lodown( "bsapuf" , output_dir = file.path( path.expand( "~" ) , "BSAPUF" ) )
```

## Analysis Examples with SQL and `MonetDBLite` {-}

Connect to a database:

```{r eval = FALSE }
library(DBI)
dbdir <- file.path( path.expand( "~" ) , "BSAPUF" , "MonetDB" )
db <- dbConnect( MonetDBLite::MonetDBLite() , dbdir )
```

```{r eval = FALSE }

```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE }
dbSendQuery( 
	db , 
	"ALTER TABLE bsa_partd_events_2008 ADD COLUMN brand_name_drug INTEGER" 
)

dbSendQuery( db , 
	"UPDATE bsa_partd_events_2008 
	SET brand_name_drug = 
		CASE 
			WHEN pde_drug_type_cd = 1 THEN 1 
			WHEN pde_drug_type_cd = 2 THEN 0 
			ELSE NULL 
		END" 
)
```

### Unweighted Counts {-}

Count the unweighted number of records in the SQL table, overall and by groups:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db , "SELECT COUNT(*) FROM bsa_partd_events_2008" )

dbGetQuery( db ,
	"SELECT
		bene_sex_ident_cd ,
		COUNT(*) 
	FROM bsa_partd_events_2008
	GROUP BY bene_sex_ident_cd"
)
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db , "SELECT AVG( pde_drug_cost ) FROM bsa_partd_events_2008" )

dbGetQuery( db , 
	"SELECT 
		bene_sex_ident_cd , 
		AVG( pde_drug_cost ) AS mean_pde_drug_cost
	FROM bsa_partd_events_2008 
	GROUP BY bene_sex_ident_cd" 
)
```

Initiate a function that allows division by zero:
```{r eval = FALSE , results = "hide" }
dbSendQuery( db , 
	"CREATE FUNCTION 
		div_noerror(l DOUBLE, r DOUBLE) 
	RETURNS DOUBLE 
	EXTERNAL NAME calc.div_noerror" 
)
```

Calculate the distribution of a categorical variable:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db , 
	"SELECT 
		bene_age_cat_cd , 
		div_noerror( 
			COUNT(*) , 
			( SELECT COUNT(*) FROM bsa_partd_events_2008 ) 
		) AS share_bene_age_cat_cd
	FROM bsa_partd_events_2008 
	GROUP BY bene_age_cat_cd" 
)
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db , "SELECT SUM( pde_drug_cost ) FROM bsa_partd_events_2008" )

dbGetQuery( db , 
	"SELECT 
		bene_sex_ident_cd , 
		SUM( pde_drug_cost ) AS sum_pde_drug_cost 
	FROM bsa_partd_events_2008 
	GROUP BY bene_sex_ident_cd" 
)
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db , "SELECT QUANTILE( pde_drug_cost , 0.5 ) FROM bsa_partd_events_2008" )

dbGetQuery( db , 
	"SELECT 
		bene_sex_ident_cd , 
		QUANTILE( pde_drug_cost , 0.5 ) AS median_pde_drug_cost
	FROM bsa_partd_events_2008 
	GROUP BY bene_sex_ident_cd" 
)
```

### Subsetting {-}

Limit your SQL analysis to events where patient paid 100% of drug's cost with `WHERE`:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db ,
	"SELECT
		AVG( pde_drug_cost )
	FROM bsa_partd_events_2008
	WHERE pde_drug_pat_pay_cd = 3"
)
```

### Measures of Uncertainty {-}

Calculate the variance and standard deviation, overall and by groups:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db , 
	"SELECT 
		VAR_SAMP( pde_drug_cost ) , 
		STDDEV_SAMP( pde_drug_cost ) 
	FROM bsa_partd_events_2008" 
)

dbGetQuery( db , 
	"SELECT 
		bene_sex_ident_cd , 
		VAR_SAMP( pde_drug_cost ) AS var_pde_drug_cost ,
		STDDEV_SAMP( pde_drug_cost ) AS stddev_pde_drug_cost
	FROM bsa_partd_events_2008 
	GROUP BY bene_sex_ident_cd" 
)
```

### Regression Models and Tests of Association {-}

Calculate the correlation between two variables, overall and by groups:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db , 
	"SELECT 
		CORR( CAST( brand_name_drug AS DOUBLE ) , CAST( pde_drug_cost AS DOUBLE ) )
	FROM bsa_partd_events_2008" 
)

dbGetQuery( db , 
	"SELECT 
		bene_sex_ident_cd , 
		CORR( CAST( brand_name_drug AS DOUBLE ) , CAST( pde_drug_cost AS DOUBLE ) )
	FROM bsa_partd_events_2008 
	GROUP BY bene_sex_ident_cd" 
)
```

## Analysis Examples with `dplyr` {-}

The R `dplyr` library offers an alternative grammar of data manipulation to base R and SQL syntax. [dplyr](https://github.com/tidyverse/dplyr/) offers many verbs, such as `summarize`, `group_by`, and `mutate`, the convenience of pipe-able functions, and the `tidyverse` style of non-standard evaluation. [This vignette](https://cran.r-project.org/web/packages/dplyr/vignettes/dplyr.html) details the available features. As a starting point for BSAPUF users, this code replicates previously-presented examples:

```{r eval = FALSE , results = "hide" }
library(dplyr)
dplyr_db <- MonetDBLite::src_monetdblite( dbdir )
bsapuf_tbl <- tbl( dplyr_db , 'bsa_partd_events_2008' )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
bsapuf_tbl %>%
	summarize( mean = mean( pde_drug_cost ) )

bsapuf_tbl %>%
	group_by( bene_sex_ident_cd ) %>%
	summarize( mean = mean( pde_drug_cost ) )
```

---

## Replication Example {-}

```{r eval = FALSE , results = "hide" }
dbGetQuery( db , "SELECT COUNT(*) FROM bsa_partd_events_2008" )
```

## Database Shutdown {-}

```{r eval = FALSE , results = "hide" }
dbDisconnect( db , shutdown = TRUE )
```
