# Brazilian Censo Demografico (CENSO) {-}

[![Build status](https://sisyphus.project.cwi.nl/?badge=censo&host=stones06fedora)](http://sisyphus.project.cwi.nl/) [![Build Status](https://sisyphus.project.cwi.nl/?badge=censo&host=stones14win)](http://sisyphus.project.cwi.nl/)

*Contributed by Dr. Djalma Pessoa <<pessoad@gmail.com>>*

Brazil's decennial census.

* One table with one row per household and a second table with one row per individual within each household. The 2000 Censo also includes a table with one record per family inside each household.

* An enumeration of the civilian non-institutional population of Brazil.

* Released decennially by IBGE since 2000, however earlier extracts are available from IPUMS International.

* Administered by the [Instituto Brasileiro de Geografia e Estatistica](http://www.ibge.gov.br/).

## Simplified Download and Importation {-}

The R `lodown` package easily downloads and imports all available CENSO microdata by simply specifying `"censo"` with an `output_dir =` parameter in the `lodown()` function. Depending on your internet connection and computer processing speed, you might prefer to run this step overnight.

```{r eval = FALSE }
library(lodown)
lodown( "censo" , output_dir = file.path( path.expand( "~" ) , "CENSO" ) )
```

`lodown` also provides a catalog of available microdata extracts with the `get_catalog()` function. After requesting the CENSO catalog, you could pass a subsetted catalog through the `lodown()` function in order to download and import specific extracts (rather than all available extracts).

```{r eval = FALSE , results = "hide" }
library(lodown)
# examine all available CENSO microdata files
censo_cat <-
	get_catalog( "censo" ,
		output_dir = file.path( path.expand( "~" ) , "CENSO" ) )

# 2010 only
censo_cat <- subset( censo_cat , year == 2010 )
# download the microdata to your local computer
lodown( "censo" , censo_cat )
```

## Analysis Examples with the `survey` library {-}

Construct a database-backed complex sample survey design:

```{r eval = FALSE }

```

```{r eval = FALSE }
library(DBI)
library(MonetDBLite)
library(survey)

options( survey.lonely.psu = "adjust" )

censo_design <- readRDS( file.path( path.expand( "~" ) , "CENSO" , "pes 2010 design.rds" ) )

censo_design <- open( censo_design , driver = MonetDBLite() )
```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE }
censo_design <-
	update(
		
		censo_design ,
		
		nmorpob1 = ifelse( v6531 >= 0 , as.numeric( v6531 < 70 ) , NA ) ,
		nmorpob2 = ifelse( v6531 >= 0 , as.numeric( v6531 < 80 ) , NA ) , 
		nmorpob3 = ifelse( v6531 >= 0 , as.numeric( v6531 < 90 ) , NA ) , 
		nmorpob4 = ifelse( v6531 >= 0 , as.numeric( v6531 < 100 ) , NA ) , 
		nmorpob5 = ifelse( v6531 >= 0 , as.numeric( v6531 < 140 ) , NA ) , 
		nmorpob6 = ifelse( v6531 >= 0 , as.numeric( v6531 < 272.50 ) , NA ) ,
		
		sexo = factor( v0601 , labels = c( "masculino" , "feminino" ) ) ,
		
		state_name = 
			factor( 
				v0001 , 
				levels = c( 11:17 , 21:29 , 31:33 , 35 , 41:43 , 50:53 ) ,
				labels = c( "Rondonia" , "Acre" , "Amazonas" , 
				"Roraima" , "Para" , "Amapa" , "Tocantins" , 
				"Maranhao" , "Piaui" , "Ceara" , "Rio Grande do Norte" , 
				"Paraiba" , "Pernambuco" , "Alagoas" , "Sergipe" , 
				"Bahia" , "Minas Gerais" , "Espirito Santo" , 
				"Rio de Janeiro" , "Sao Paulo" , "Parana" , 
				"Santa Catarina" , "Rio Grande do Sul" , 
				"Mato Grosso do Sul" , "Mato Grosso" , "Goias" , 
				"Distrito Federal" )
			)
	)
```

### Unweighted Counts {-}

Count the unweighted number of records in the survey sample, overall and by groups:
```{r eval = FALSE , results = "hide" }
sum( weights( censo_design , "sampling" ) != 0 )

svyby( ~ one , ~ state_name , censo_design , unwtd.count )
```

### Weighted Counts {-}
Count the weighted size of the generalizable population, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ one , censo_design )

svyby( ~ one , ~ state_name , censo_design , svytotal )
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ v6033 , censo_design )

svyby( ~ v6033 , ~ state_name , censo_design , svymean )
```

Calculate the distribution of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ sexo , censo_design )

svyby( ~ sexo , ~ state_name , censo_design , svymean )
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ v6033 , censo_design )

svyby( ~ v6033 , ~ state_name , censo_design , svytotal )
```

Calculate the weighted sum of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ sexo , censo_design )

svyby( ~ sexo , ~ state_name , censo_design , svytotal )
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svyquantile( ~ v6033 , censo_design , 0.5 )

svyby( 
	~ v6033 , 
	~ state_name , 
	censo_design , 
	svyquantile , 
	0.5 ,
	ci = TRUE ,
	keep.var = TRUE 
)
```

Estimate a ratio:
```{r eval = FALSE , results = "hide" }
svyratio( 
	numerator = ~ nmorpob1 , 
	denominator = ~ nmorpob1 + one , 
	censo_design ,
	na.rm = TRUE
)
```

### Subsetting {-}

Restrict the survey design to married persons:
```{r eval = FALSE , results = "hide" }
sub_censo_design <- subset( censo_design , v0640 == 1 )
```
Calculate the mean (average) of this subset:
```{r eval = FALSE , results = "hide" }
svymean( ~ v6033 , sub_censo_design )
```

### Measures of Uncertainty {-}

Extract the coefficient, standard error, confidence interval, and coefficient of variation from any descriptive statistics function result, overall and by groups:
```{r eval = FALSE , results = "hide" }
this_result <- svymean( ~ v6033 , censo_design )

coef( this_result )
SE( this_result )
confint( this_result )
cv( this_result )

grouped_result <-
	svyby( 
		~ v6033 , 
		~ state_name , 
		censo_design , 
		svymean 
	)
	
coef( grouped_result )
SE( grouped_result )
confint( grouped_result )
cv( grouped_result )
```

Calculate the degrees of freedom of any survey design object:
```{r eval = FALSE , results = "hide" }
degf( censo_design )
```

Calculate the complex sample survey-adjusted variance of any statistic:
```{r eval = FALSE , results = "hide" }
svyvar( ~ v6033 , censo_design )
```

Include the complex sample design effect in the result for a specific statistic:
```{r eval = FALSE , results = "hide" }
# SRS without replacement
svymean( ~ v6033 , censo_design , deff = TRUE )

# SRS with replacement
svymean( ~ v6033 , censo_design , deff = "replace" )
```

Compute confidence intervals for proportions using methods that may be more accurate near 0 and 1. See `?svyciprop` for alternatives:
```{r eval = FALSE , results = "hide" }
svyciprop( ~ nmorpob6 , censo_design ,
	method = "likelihood" , na.rm = TRUE )
```

### Regression Models and Tests of Association {-}

Perform a design-based t-test:
```{r eval = FALSE , results = "hide" }
svyttest( v6033 ~ nmorpob6 , censo_design )
```

Perform a chi-squared test of association for survey data:
```{r eval = FALSE , results = "hide" }
svychisq( 
	~ nmorpob6 + sexo , 
	censo_design 
)
```

Perform a survey-weighted generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	svyglm( 
		v6033 ~ nmorpob6 + sexo , 
		censo_design 
	)

summary( glm_result )
```

## Poverty and Inequality Estimation with `convey` {-}

The R `convey` library estimates measures of income concentration, poverty, inequality, and wellbeing. [This textbook](https://guilhermejacob.github.io/context/) details the available features. As a starting point for CENSO users, this code calculates the gini coefficient on complex sample survey data:

```{r eval = FALSE , results = "hide" }
library(convey)
censo_design <- convey_prep( censo_design )

sub_censo_design <- 
	subset( censo_design , v6531 >= 0 )

svygini( ~ v6531 , sub_censo_design , na.rm = TRUE )
```

---

## Replication Example {-}

```{r eval = FALSE , results = "hide" }

```

## Database Shutdown {-}

```{r eval = FALSE , results = "hide" }
close( censo_design , shutdown = TRUE )
```
