# Censo Escolar (CENSOESCOLAR) {-}

[![Build Status](https://travis-ci.org/asdfree/censoescolar.svg?branch=master)](https://travis-ci.org/asdfree/censoescolar) [![Build status](https://ci.appveyor.com/api/projects/status/github/asdfree/censoescolar?svg=TRUE)](https://ci.appveyor.com/project/ajdamico/censoescolar)

*Contributed by Guilherme Jacob <<guilhermejacob91@gmail.com>>*

The Brazilian School Census (Censo Escolar) is a massive source of information about basic education. Synthetic tables can be produced using the [InepData interface](http://inepdata.inep.gov.br/analytics/saw.dll?Dashboard).

* Since 2007, each year is composed of 4 tables, containing information about students, teachers, schools and classes.

* A census of schools, students, teachers and classes in Brazil.

* Released annually since 1995.

* Administered by the [Anisio Teixeira National Institute for Educational Study and Research (INEP)](http://portal.inep.gov.br), a branch of the Brazilian Ministry of Education.

## Simplified Download and Importation {-}

The R `lodown` package easily downloads and imports all available CENSOESCOLAR microdata by simply specifying `"censoescolar"` with an `output_dir =` parameter in the `lodown()` function. Depending on your internet connection and computer processing speed, you might prefer to run this step overnight.

```{r eval = FALSE }
library(lodown)
lodown( "censoescolar" , output_dir = file.path( path.expand( "~" ) , "CENSOESCOLAR" ) )
```

## Analysis Examples with SQL and `RSQLite` \ {-}

Connect to a database:

```{r eval = FALSE }
library(DBI)
dbdir <- file.path( path.expand( "~" ) , "CENSOESCOLAR" , "SQLite.db" )
db <- dbConnect( RSQLite::SQLite() , dbdir )
```

```{r eval = FALSE }

```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE }
dbSendQuery( db , "ALTER TABLE matricula_2015 ADD COLUMN id_indigenous_area INTEGER" )

dbSendQuery( db ,
	"UPDATE matricula_2015
	SET id_indigenous_area =
		CASE WHEN ( tp_localizacao_diferenciada IN (2,5) ) THEN 1 ELSE 0 END"
)

dbSendQuery( db , "ALTER TABLE matricula_2015 ADD COLUMN id_public INTEGER" )

dbSendQuery( db ,
	"UPDATE matricula_2015
	SET id_public =
		CASE WHEN ( tp_dependencia <> 4 ) THEN 1 ELSE 0 END"
)
```

### Unweighted Counts {-}

Count the unweighted number of records in the SQL table, overall and by groups:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db , "SELECT COUNT(*) FROM matricula_2015" )

dbGetQuery( db ,
	"SELECT
		tp_localizacao ,
		COUNT(*) 
	FROM matricula_2015
	GROUP BY tp_localizacao"
)
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db , "SELECT AVG( nu_idade ) FROM matricula_2015" )

dbGetQuery( db , 
	"SELECT 
		tp_localizacao , 
		AVG( nu_idade ) AS mean_nu_idade
	FROM matricula_2015 
	GROUP BY tp_localizacao" 
)
```

Calculate the distribution of a categorical variable:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db , 
	"SELECT 
		tp_categoria_escola_privada , 
		COUNT(*) / ( SELECT COUNT(*) FROM matricula_2015 ) 
			AS share_tp_categoria_escola_privada
	FROM matricula_2015 
	GROUP BY tp_categoria_escola_privada" 
)
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db , "SELECT SUM( nu_idade ) FROM matricula_2015" )

dbGetQuery( db , 
	"SELECT 
		tp_localizacao , 
		SUM( nu_idade ) AS sum_nu_idade 
	FROM matricula_2015 
	GROUP BY tp_localizacao" 
)
```

Calculate the 25th, median, and 75th percentiles of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
RSQLite::initExtension( db )

dbGetQuery( db , 
	"SELECT 
		LOWER_QUARTILE( nu_idade ) , 
		MEDIAN( nu_idade ) , 
		UPPER_QUARTILE( nu_idade ) 
	FROM matricula_2015" 
)

dbGetQuery( db , 
	"SELECT 
		tp_localizacao , 
		LOWER_QUARTILE( nu_idade ) AS lower_quartile_nu_idade , 
		MEDIAN( nu_idade ) AS median_nu_idade , 
		UPPER_QUARTILE( nu_idade ) AS upper_quartile_nu_idade
	FROM matricula_2015 
	GROUP BY tp_localizacao" 
)
```

### Subsetting {-}

Limit your SQL analysis to students enrolled in public schools with `WHERE`:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db ,
	"SELECT
		AVG( nu_idade )
	FROM matricula_2015
	WHERE id_public = 1"
)
```

### Measures of Uncertainty {-}

Calculate the variance and standard deviation, overall and by groups:
```{r eval = FALSE , results = "hide" }
RSQLite::initExtension( db )

dbGetQuery( db , 
	"SELECT 
		VARIANCE( nu_idade ) , 
		STDEV( nu_idade ) 
	FROM matricula_2015" 
)

dbGetQuery( db , 
	"SELECT 
		tp_localizacao , 
		VARIANCE( nu_idade ) AS var_nu_idade ,
		STDEV( nu_idade ) AS stddev_nu_idade
	FROM matricula_2015 
	GROUP BY tp_localizacao" 
)
```

### Regression Models and Tests of Association {-}

Perform a t-test:
```{r eval = FALSE , results = "hide" }
censoescolar_three_columns_df <- 
	dbGetQuery( db , 
		"SELECT 
			nu_idade , 
			id_public ,
			tp_categoria_escola_privada
		FROM matricula_2015" 
	)

t.test( nu_idade ~ id_public , censoescolar_three_columns_df )
```

Perform a chi-squared test of association:
```{r eval = FALSE , results = "hide" }
this_table <- table( censoescolar_three_columns_df[ , c( "id_public" , "tp_categoria_escola_privada" ) ] )

chisq.test( this_table )
```

Perform a generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	glm( 
		nu_idade ~ id_public + tp_categoria_escola_privada , 
		data = censoescolar_three_columns_df
	)

summary( glm_result )
```

## Analysis Examples with `dplyr` \ {-}

The R `dplyr` library offers an alternative grammar of data manipulation to base R and SQL syntax. [dplyr](https://github.com/tidyverse/dplyr/) offers many verbs, such as `summarize`, `group_by`, and `mutate`, the convenience of pipe-able functions, and the `tidyverse` style of non-standard evaluation. [This vignette](https://cran.r-project.org/web/packages/dplyr/vignettes/dplyr.html) details the available features. As a starting point for CENSOESCOLAR users, this code replicates previously-presented examples:

```{r eval = FALSE , results = "hide" }
library(dplyr)
library(dbplyr)
dplyr_db <- dplyr::src_sqlite( dbdir )
censoescolar_tbl <- tbl( dplyr_db , 'matricula_2015' )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
censoescolar_tbl %>%
	summarize( mean = mean( nu_idade ) )

censoescolar_tbl %>%
	group_by( tp_localizacao ) %>%
	summarize( mean = mean( nu_idade ) )
```

---

## Replication Example {-}

This snippet replicates the first three rows of total enrollment in basic education, as in the Table 1.1 of [this spreadsheet](http://download.inep.gov.br/informacoes_estatisticas/sinopses_estatisticas/sinopses_educacao_basica/sinopse_estatistica_educacao_basica_2015.zip).

```{r eval = FALSE , results = "hide" }
# first row:
dbGetQuery( db ,"SELECT COUNT(*) AS n_mat_tot
 FROM matricula_2015
 WHERE TP_TIPO_TURMA NOT IN (4,5) " )

# second row:
dbGetQuery( db ,"SELECT COUNT(*) AS n_mat_tot
 FROM matricula_2015
 WHERE TP_TIPO_TURMA NOT IN (4,5) AND CO_REGIAO = 1" )

# third row:
dbGetQuery( db ,"SELECT COUNT(*) AS n_mat_tot
 FROM matricula_2015
 WHERE TP_TIPO_TURMA NOT IN (4,5) AND CO_UF = 11 " )
```

