# Censo Escolar (CENSO_ESCOLAR) {-}

[![Build Status](https://sisyphus.project.cwi.nl/?badge=censo_escolar&host=stones14win)](http://sisyphus.project.cwi.nl/) [![Build status](https://sisyphus.project.cwi.nl/?badge=censo_escolar&host=stones06fedora)](http://sisyphus.project.cwi.nl/)

*Contributed by Guilherme Jacob <<guilhermejacob91@gmail.com>>*

The Brazilian School Census (Censo Escolar) is a massive source of information about basic education. Synthetic tables can be produced using the [InepData interface](http://inepdata.inep.gov.br/analytics/saw.dll?Dashboard).

* Since 2007, each year is composed of 4 tables, containing information about students, teachers, schools and classes.

* A census of schools, students, teachers and classes in Brazil.

* Released annually since 1995.

* Administered by the [Anisio Teixeira National Institute for Educational Study and Research (INEP)](http://portal.inep.gov.br), a branch of the Brazilian Ministry of Education.

## Simplified Download and Importation {-}

The R `lodown` package easily downloads and imports all available CENSO_ESCOLAR microdata by simply specifying `"censo_escolar"` with an `output_dir =` parameter in the `lodown()` function. Depending on your internet connection and computer processing speed, you might prefer to run this step overnight.

```{r eval = FALSE }
library(lodown)
lodown( "censo_escolar" , output_dir = file.path( path.expand( "~" ) , "CENSO_ESCOLAR" ) )
```

## Analysis Examples with SQL and `MonetDBLite` {-}

Connect to a database:

```{r eval = FALSE }
library(DBI)
dbdir <- file.path( path.expand( "~" ) , "CENSO_ESCOLAR" , "MonetDB" )
db <- dbConnect( MonetDBLite::MonetDBLite() , dbdir )
```

```{r eval = FALSE }

```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE }
dbSendQuery( db , "ALTER TABLE matricula_2015 ADD COLUMN id_indigenous_area INTEGER" )

dbSendQuery( db ,
	"UPDATE matricula_2015
	SET id_indigenous_area =
		CASE WHEN ( tp_localizacao_diferenciada IN (2,5) ) THEN 1 ELSE 0 END"
)

dbSendQuery( db , "ALTER TABLE matricula_2015 ADD COLUMN id_public INTEGER" )

dbSendQuery( db ,
	"UPDATE matricula_2015
	SET id_public =
		CASE WHEN ( tp_dependencia <> 4 ) THEN 1 ELSE 0 END"
)
```

### Unweighted Counts {-}

Count the unweighted number of records in the SQL table, overall and by groups:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db , "SELECT COUNT(*) FROM matricula_2015" )

dbGetQuery( db ,
	"SELECT
		tp_localizacao ,
		COUNT(*) 
	FROM matricula_2015
	GROUP BY tp_localizacao"
)
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db , "SELECT AVG( nu_idade ) FROM matricula_2015" )

dbGetQuery( db , 
	"SELECT 
		tp_localizacao , 
		AVG( nu_idade ) AS mean_nu_idade
	FROM matricula_2015 
	GROUP BY tp_localizacao" 
)
```

Initiate a function that allows division by zero:
```{r eval = FALSE , results = "hide" }
dbSendQuery( db , 
	"CREATE FUNCTION 
		div_noerror(l DOUBLE, r DOUBLE) 
	RETURNS DOUBLE 
	EXTERNAL NAME calc.div_noerror" 
)
```

Calculate the distribution of a categorical variable:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db , 
	"SELECT 
		tp_categoria_escola_privada , 
		div_noerror( 
			COUNT(*) , 
			( SELECT COUNT(*) FROM matricula_2015 ) 
		) AS share_tp_categoria_escola_privada
	FROM matricula_2015 
	GROUP BY tp_categoria_escola_privada" 
)
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db , "SELECT SUM( nu_idade ) FROM matricula_2015" )

dbGetQuery( db , 
	"SELECT 
		tp_localizacao , 
		SUM( nu_idade ) AS sum_nu_idade 
	FROM matricula_2015 
	GROUP BY tp_localizacao" 
)
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db , "SELECT QUANTILE( nu_idade , 0.5 ) FROM matricula_2015" )

dbGetQuery( db , 
	"SELECT 
		tp_localizacao , 
		QUANTILE( nu_idade , 0.5 ) AS median_nu_idade
	FROM matricula_2015 
	GROUP BY tp_localizacao" 
)
```

### Subsetting {-}

Limit your SQL analysis to students enrolled in public schools with `WHERE`:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db ,
	"SELECT
		AVG( nu_idade )
	FROM matricula_2015
	WHERE id_public = 1"
)
```

### Measures of Uncertainty {-}

Calculate the variance and standard deviation, overall and by groups:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db , 
	"SELECT 
		VAR_SAMP( nu_idade ) , 
		STDDEV_SAMP( nu_idade ) 
	FROM matricula_2015" 
)

dbGetQuery( db , 
	"SELECT 
		tp_localizacao , 
		VAR_SAMP( nu_idade ) AS var_nu_idade ,
		STDDEV_SAMP( nu_idade ) AS stddev_nu_idade
	FROM matricula_2015 
	GROUP BY tp_localizacao" 
)
```

### Regression Models and Tests of Association {-}

Calculate the correlation between two variables, overall and by groups:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db , 
	"SELECT 
		CORR( CAST( id_public AS DOUBLE ) , CAST( nu_idade AS DOUBLE ) )
	FROM matricula_2015" 
)

dbGetQuery( db , 
	"SELECT 
		tp_localizacao , 
		CORR( CAST( id_public AS DOUBLE ) , CAST( nu_idade AS DOUBLE ) )
	FROM matricula_2015 
	GROUP BY tp_localizacao" 
)
```

## Analysis Examples with `dplyr` {-}

The R `dplyr` library offers an alternative grammar of data manipulation to base R and SQL syntax. [dplyr](https://github.com/tidyverse/dplyr/) offers many verbs, such as `summarize`, `group_by`, and `mutate`, the convenience of pipe-able functions, and the `tidyverse` style of non-standard evaluation. [This vignette](https://cran.r-project.org/web/packages/dplyr/vignettes/dplyr.html) details the available features. As a starting point for CENSO_ESCOLAR users, this code replicates previously-presented examples:

```{r eval = FALSE , results = "hide" }
library(dplyr)
dplyr_db <- MonetDBLite::src_monetdblite( dbdir )
censo_escolar_tbl <- tbl( dplyr_db , 'matricula_2015' )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
censo_escolar_tbl %>%
	summarize( mean = mean( nu_idade ) )

censo_escolar_tbl %>%
	group_by( tp_localizacao ) %>%
	summarize( mean = mean( nu_idade ) )
```

---

## Replication Example {-}

This snippet replicates the first three rows of total enrollment in basic education, as in the Table 1.1 of [this spreadsheet](http://download.inep.gov.br/informacoes_estatisticas/sinopses_estatisticas/sinopses_educacao_basica/sinopse_estatistica_educacao_basica_2015.zip).

```{r eval = FALSE , results = "hide" }
# first row:
dbGetQuery( db ,"SELECT COUNT(*) AS n_mat_tot
 FROM matricula_2015
 WHERE TP_TIPO_TURMA NOT IN (4,5) " )

# second row:
dbGetQuery( db ,"SELECT COUNT(*) AS n_mat_tot
 FROM matricula_2015
 WHERE TP_TIPO_TURMA NOT IN (4,5) AND CO_REGIAO = 1" )

# third row:
dbGetQuery( db ,"SELECT COUNT(*) AS n_mat_tot
 FROM matricula_2015
 WHERE TP_TIPO_TURMA NOT IN (4,5) AND CO_UF = 11 " )
```

## Database Shutdown {-}

```{r eval = FALSE , results = "hide" }
dbDisconnect( db , shutdown = TRUE )
```
