# Current Population Survey - Basic Monthly (CPSBASIC) {-}

[![Build Status](https://travis-ci.org/asdfree/cpsbasic.svg?branch=master)](https://travis-ci.org/asdfree/cpsbasic) [![Build status](https://ci.appveyor.com/api/projects/status/github/asdfree/cpsbasic?svg=TRUE)](https://ci.appveyor.com/project/ajdamico/cpsbasic)

The Current Population Survey - Basic Monthly is the monthly labor force survey of the United States.

* One table with one row per sampled youth respondent.

* A complex sample survey designed to generalize to the civilian non-institutional population of the United States

* Released monthly since 1994.

* Administered jointly by the [US Census Bureau](http://www.census.gov/) and the [Bureau of Labor Statistics](http://www.bls.gov/).

## Simplified Download and Importation {-}

The R `lodown` package easily downloads and imports all available CPSBASIC microdata by simply specifying `"cpsbasic"` with an `output_dir =` parameter in the `lodown()` function. Depending on your internet connection and computer processing speed, you might prefer to run this step overnight.

```{r eval = FALSE }
library(lodown)
lodown( "cpsbasic" , output_dir = file.path( path.expand( "~" ) , "CPSBASIC" ) )
```

`lodown` also provides a catalog of available microdata extracts with the `get_catalog()` function. After requesting the CPSBASIC catalog, you could pass a subsetted catalog through the `lodown()` function in order to download and import specific extracts (rather than all available extracts).

```{r eval = FALSE , results = "hide" }
library(lodown)
# examine all available CPSBASIC microdata files
cpsbasic_cat <-
	get_catalog( "cpsbasic" ,
		output_dir = file.path( path.expand( "~" ) , "CPSBASIC" ) )

# march 2017 only
cpsbasic_cat <- subset( cpsbasic_cat , year == 2017 & month == 3 )
# download the microdata to your local computer
lodown( "cpsbasic" , cpsbasic_cat )
```

## Analysis Examples with the `survey` library {-}

Construct a complex sample survey design:

```{r eval = FALSE }

```

```{r eval = FALSE }
library(survey)

cpsbasic_df <- 
	readRDS( file.path( path.expand( "~" ) , "CPSBASIC" , "2017 03 cps basic.rds" ) )

# construct a fake survey design
warning( "this survey design produces correct point estimates
but incorrect standard errors." )
cpsbasic_design <- 
	svydesign( 
		~ 1 , 
		data = cpsbasic_df , 
		weights = ~ pwsswgt
	)
```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE }
cpsbasic_design <- 
	update( 
		cpsbasic_design , 
		
		one = 1 ,
		
		pesex = factor( pesex , levels = 1:2 , labels = c( 'male' , 'female' ) ) ,
		
		weekly_earnings = ifelse( prernwa == -.01 , NA , prernwa ) ,
		
		# exclude anyone whose hours vary
		weekly_hours = ifelse( pehrusl1 < 0 , NA , pehrusl1 ) ,
		
		class_of_worker =
			factor( peio1cow , levels = 1:8 ,
				labels = 
					c( "government - federal" , "government - state" ,
					"government - local" , "private, for profit" ,
					"private, nonprofit" , "self-employed, incorporated" ,
					"self-employed, unincorporated" , "without pay" )
			) ,
			
		part_time = ifelse( pemlr == 1 , as.numeric( pehruslt < 35 ) , NA )
	)
```

### Unweighted Counts {-}

Count the unweighted number of records in the survey sample, overall and by groups:
```{r eval = FALSE , results = "hide" }
sum( weights( cpsbasic_design , "sampling" ) != 0 )

svyby( ~ one , ~ pesex , cpsbasic_design , unwtd.count )
```

### Weighted Counts {-}
Count the weighted size of the generalizable population, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ one , cpsbasic_design )

svyby( ~ one , ~ pesex , cpsbasic_design , svytotal )
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ weekly_earnings , cpsbasic_design , na.rm = TRUE )

svyby( ~ weekly_earnings , ~ pesex , cpsbasic_design , svymean , na.rm = TRUE )
```

Calculate the distribution of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ class_of_worker , cpsbasic_design , na.rm = TRUE )

svyby( ~ class_of_worker , ~ pesex , cpsbasic_design , svymean , na.rm = TRUE )
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ weekly_earnings , cpsbasic_design , na.rm = TRUE )

svyby( ~ weekly_earnings , ~ pesex , cpsbasic_design , svytotal , na.rm = TRUE )
```

Calculate the weighted sum of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ class_of_worker , cpsbasic_design , na.rm = TRUE )

svyby( ~ class_of_worker , ~ pesex , cpsbasic_design , svytotal , na.rm = TRUE )
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svyquantile( ~ weekly_earnings , cpsbasic_design , 0.5 , na.rm = TRUE )

svyby( 
	~ weekly_earnings , 
	~ pesex , 
	cpsbasic_design , 
	svyquantile , 
	0.5 ,
	ci = TRUE ,
	keep.var = TRUE ,
	na.rm = TRUE
)
```

Estimate a ratio:
```{r eval = FALSE , results = "hide" }
svyratio( 
	numerator = ~ weekly_earnings , 
	denominator = ~ weekly_hours , 
	cpsbasic_design ,
	na.rm = TRUE
)
```

### Subsetting {-}

Restrict the survey design to california residents:
```{r eval = FALSE , results = "hide" }
sub_cpsbasic_design <- subset( cpsbasic_design , gestfips == 6 )
```
Calculate the mean (average) of this subset:
```{r eval = FALSE , results = "hide" }
svymean( ~ weekly_earnings , sub_cpsbasic_design , na.rm = TRUE )
```

### Measures of Uncertainty {-}

Extract the coefficient, standard error, confidence interval, and coefficient of variation from any descriptive statistics function result, overall and by groups:
```{r eval = FALSE , results = "hide" }
this_result <- svymean( ~ weekly_earnings , cpsbasic_design , na.rm = TRUE )

coef( this_result )
SE( this_result )
confint( this_result )
cv( this_result )

grouped_result <-
	svyby( 
		~ weekly_earnings , 
		~ pesex , 
		cpsbasic_design , 
		svymean ,
		na.rm = TRUE 
	)
	
coef( grouped_result )
SE( grouped_result )
confint( grouped_result )
cv( grouped_result )
```

Calculate the degrees of freedom of any survey design object:
```{r eval = FALSE , results = "hide" }
degf( cpsbasic_design )
```

Calculate the complex sample survey-adjusted variance of any statistic:
```{r eval = FALSE , results = "hide" }
svyvar( ~ weekly_earnings , cpsbasic_design , na.rm = TRUE )
```

Include the complex sample design effect in the result for a specific statistic:
```{r eval = FALSE , results = "hide" }
# SRS without replacement
svymean( ~ weekly_earnings , cpsbasic_design , na.rm = TRUE , deff = TRUE )

# SRS with replacement
svymean( ~ weekly_earnings , cpsbasic_design , na.rm = TRUE , deff = "replace" )
```

Compute confidence intervals for proportions using methods that may be more accurate near 0 and 1. See `?svyciprop` for alternatives:
```{r eval = FALSE , results = "hide" }
svyciprop( ~ part_time , cpsbasic_design ,
	method = "likelihood" , na.rm = TRUE )
```

### Regression Models and Tests of Association {-}

Perform a design-based t-test:
```{r eval = FALSE , results = "hide" }
svyttest( weekly_earnings ~ part_time , cpsbasic_design )
```

Perform a chi-squared test of association for survey data:
```{r eval = FALSE , results = "hide" }
svychisq( 
	~ part_time + class_of_worker , 
	cpsbasic_design 
)
```

Perform a survey-weighted generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	svyglm( 
		weekly_earnings ~ part_time + class_of_worker , 
		cpsbasic_design 
	)

summary( glm_result )
```

## Analysis Examples with `srvyr` {-}

The R `srvyr` library calculates summary statistics from survey data, such as the mean, total or quantile using [dplyr](https://github.com/tidyverse/dplyr/)-like syntax. [srvyr](https://github.com/gergness/srvyr) allows for the use of many verbs, such as `summarize`, `group_by`, and `mutate`, the convenience of pipe-able functions, the `tidyverse` style of non-standard evaluation and more consistent return types than the `survey` package. [This vignette](https://cran.r-project.org/web/packages/srvyr/vignettes/srvyr-vs-survey.html) details the available features. As a starting point for CPSBASIC users, this code replicates previously-presented examples:

```{r eval = FALSE , results = "hide" }
library(srvyr)
cpsbasic_srvyr_design <- as_survey( cpsbasic_design )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
cpsbasic_srvyr_design %>%
	summarize( mean = survey_mean( weekly_earnings , na.rm = TRUE ) )

cpsbasic_srvyr_design %>%
	group_by( pesex ) %>%
	summarize( mean = survey_mean( weekly_earnings , na.rm = TRUE ) )
```

---

## Replication Example {-}

```{r eval = FALSE , results = "hide" }

```


