# Brazilian Vital Statistics System (DATASUS) {-}

[![Build status](https://sisyphus.project.cwi.nl/?badge=datasus&host=stones06fedora)](http://sisyphus.project.cwi.nl/) [![Build Status](https://sisyphus.project.cwi.nl/?badge=datasus&host=stones14win)](http://sisyphus.project.cwi.nl/)

*Contributed by Guilherme Jacob <<guilhermejacob91@gmail.com>>*

The Brazilian Health System (also known as SUS) provides information about births, deaths and prenatal care through DataSUS. Preset tabulation can be done online using [TabNet](http://tabnet.datasus.gov.br), the online interface of the DataSUS.

* The system is composed of 14 tables, three tables containing birth information, ten tables containing information about mortality information and one containg prenatal care information.

* A census of births, deaths and pregnancies ocurred and registered by the Brazilian health system.

* Released annually since 1979.

* Administered by the [Brazilian Ministry of Health](http://www.saude.gov.br/).

## Simplified Download and Importation {-}

The R `lodown` package easily downloads and imports all available DATASUS microdata by simply specifying `"datasus"` with an `output_dir =` parameter in the `lodown()` function. Depending on your internet connection and computer processing speed, you might prefer to run this step overnight.

```{r eval = FALSE }
library(lodown)
lodown( "datasus" , output_dir = file.path( path.expand( "~" ) , "DATASUS" ) )
```

## Analysis Examples with SQL and `MonetDBLite` {-}

Connect to a database:

```{r eval = FALSE }
library(DBI)
dbdir <- file.path( path.expand( "~" ) , "DATASUS" , "MonetDB" )
db <- dbConnect( MonetDBLite::MonetDBLite() , dbdir )
```

```{r eval = FALSE }

```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE }
dbSendQuery( db , "ALTER TABLE nasc_cid10 ADD COLUMN birthyear INTEGER" )

dbSendQuery( db , 
	"UPDATE nasc_cid10 
	SET birthyear = 
		RIGHT( DTNASC , 4 )" 
)

dbSendQuery( db , "ALTER TABLE nasc_cid10 ADD COLUMN race INTEGER" )

dbSendQuery( db , 
	"UPDATE nasc_cid10 
	SET race = 
		CASE WHEN ( racacor = 1 ) THEN 1 ELSE 0 END" 
)
```

### Unweighted Counts {-}

Count the unweighted number of records in the SQL table, overall and by groups:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db , "SELECT COUNT(*) FROM nasc_cid10" )

dbGetQuery( db ,
	"SELECT
		semagestac ,
		COUNT(*) 
	FROM nasc_cid10
	GROUP BY semagestac"
)
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db , "SELECT AVG( peso ) FROM nasc_cid10" )

dbGetQuery( db , 
	"SELECT 
		semagestac , 
		AVG( peso ) AS mean_peso
	FROM nasc_cid10 
	GROUP BY semagestac" 
)
```

Initiate a function that allows division by zero:
```{r eval = FALSE , results = "hide" }
dbSendQuery( db , 
	"CREATE FUNCTION 
		div_noerror(l DOUBLE, r DOUBLE) 
	RETURNS DOUBLE 
	EXTERNAL NAME calc.div_noerror" 
)
```

Calculate the distribution of a categorical variable:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db , 
	"SELECT 
		sexo , 
		div_noerror( 
			COUNT(*) , 
			( SELECT COUNT(*) FROM nasc_cid10 ) 
		) AS share_sexo
	FROM nasc_cid10 
	GROUP BY sexo" 
)
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db , "SELECT SUM( peso ) FROM nasc_cid10" )

dbGetQuery( db , 
	"SELECT 
		semagestac , 
		SUM( peso ) AS sum_peso 
	FROM nasc_cid10 
	GROUP BY semagestac" 
)
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db , "SELECT QUANTILE( peso , 0.5 ) FROM nasc_cid10" )

dbGetQuery( db , 
	"SELECT 
		semagestac , 
		QUANTILE( peso , 0.5 ) AS median_peso
	FROM nasc_cid10 
	GROUP BY semagestac" 
)
```

### Subsetting {-}

Limit your SQL analysis to births occurred in 2015 with `WHERE`:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db ,
	"SELECT
		AVG( peso )
	FROM nasc_cid10
	WHERE birthyear = 2015"
)
```

### Measures of Uncertainty {-}

Calculate the variance and standard deviation, overall and by groups:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db , 
	"SELECT 
		VAR_SAMP( peso ) , 
		STDDEV_SAMP( peso ) 
	FROM nasc_cid10" 
)

dbGetQuery( db , 
	"SELECT 
		semagestac , 
		VAR_SAMP( peso ) AS var_peso ,
		STDDEV_SAMP( peso ) AS stddev_peso
	FROM nasc_cid10 
	GROUP BY semagestac" 
)
```

### Regression Models and Tests of Association {-}

Calculate the correlation between two variables, overall and by groups:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db , 
	"SELECT 
		CORR( CAST( race AS DOUBLE ) , CAST( peso AS DOUBLE ) )
	FROM nasc_cid10" 
)

dbGetQuery( db , 
	"SELECT 
		semagestac , 
		CORR( CAST( race AS DOUBLE ) , CAST( peso AS DOUBLE ) )
	FROM nasc_cid10 
	GROUP BY semagestac" 
)
```

## Analysis Examples with `dplyr` {-}

The R `dplyr` library offers an alternative grammar of data manipulation to base R and SQL syntax. [dplyr](https://github.com/tidyverse/dplyr/) offers many verbs, such as `summarize`, `group_by`, and `mutate`, the convenience of pipe-able functions, and the `tidyverse` style of non-standard evaluation. [This vignette](https://cran.r-project.org/web/packages/dplyr/vignettes/dplyr.html) details the available features. As a starting point for DATASUS users, this code replicates previously-presented examples:

```{r eval = FALSE , results = "hide" }
library(dplyr)
dplyr_db <- MonetDBLite::src_monetdblite( dbdir )
datasus_tbl <- tbl( dplyr_db , 'nasc_cid10' )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
datasus_tbl %>%
	summarize( mean = mean( peso ) )

datasus_tbl %>%
	group_by( semagestac ) %>%
	summarize( mean = mean( peso ) )
```

---

## Replication Example {-}

```{r eval = FALSE , results = "hide" }
dbGetQuery( db , "SELECT COUNT(*) FROM nasc_cid10 GROUP BY birthyear ORDER BY birthyear" )
```

## Database Shutdown {-}

```{r eval = FALSE , results = "hide" }
dbDisconnect( db , shutdown = TRUE )
```
