# Health and Retirement Study (HRS) {-}

[![Build Status](https://travis-ci.org/asdfree/hrs.svg?branch=master)](https://travis-ci.org/asdfree/hrs) [![Build status](https://ci.appveyor.com/api/projects/status/github/asdfree/hrs?svg=TRUE)](https://ci.appveyor.com/project/ajdamico/hrs)

needs_sisyphus_build_status_line

The Health and Retirement Study interviews Americans aged 50+ for their entire life. Allows for findings like, "Among Americans who were 50-74 years old in 1998, X% lived in nursing homes by 2010."

* Many tables, most with one row per sampled respondent and linkable over time. Use the RAND HRS data file for a cleaner, cross-wave data set.

* A complex sample survey designed to generalize to Americans aged 50+ at each interview, but longitudinal analysts can observe outcomes.

* Released biennially since 1992.

* Administered by the [University of Michigan's Institute for Social Research](http://isr.umich.edu/) with data management by the [RAND Corporation](http://www.rand.org/). Funded by the [National Institute on Aging](https://www.nia.nih.gov/) and the [Social Security Administration](https://www.ssa.gov/).

## Simplified Download and Importation {-}

The R `lodown` package easily downloads and imports all available HRS microdata by simply specifying `"hrs"` with an `output_dir =` parameter in the `lodown()` function. Depending on your internet connection and computer processing speed, you might prefer to run this step overnight.

```{r eval = FALSE }
library(lodown)
lodown( "hrs" , output_dir = file.path( path.expand( "~" ) , "HRS" ) , 
	your_username = "username" , 
	your_password = "password" )
```

`lodown` also provides a catalog of available microdata extracts with the `get_catalog()` function. After requesting the HRS catalog, you could pass a subsetted catalog through the `lodown()` function in order to download and import specific extracts (rather than all available extracts).

```{r eval = FALSE , results = "hide" }
library(lodown)
# examine all available HRS microdata files
hrs_cat <-
	get_catalog( "hrs" ,
		output_dir = file.path( path.expand( "~" ) , "HRS" ) , 
		your_username = "username" , 
		your_password = "password" )

# RAND consolidated file only
hrs_cat <- subset( hrs_cat , grepl( 'rand([a-z]+)stata\\.zip' , file_name ) )
# download the microdata to your local computer
lodown( "hrs" , hrs_cat , 
	your_username = "username" , 
	your_password = "password" )
```

## Analysis Examples with the `survey` library {-}

Construct a complex sample survey design:

```{r eval = FALSE }

```

```{r eval = FALSE }
library(survey)

hrs_df <- 
	readRDS( list.files( hrs_cat$output_folder , full.names = TRUE ) )
	
# RAM cleanup
keep_vars <- 
	c( "raehsamp" , "raestrat" , "r3wtresp" , 
		"r3work" , "r12work" , "h12ahous" ,
		"r3mstat" , "r12mstat" , "h4ahous" )

hrs_df <- hrs_df[ keep_vars ]
	
# community residents aged 50+ in 1996
hrs_design <- 
	svydesign(
		id = ~ raehsamp ,
		strata = ~ raestrat ,
		weights = ~ r3wtresp , 
		nest = TRUE ,
		data = subset( hrs_df , r3wtresp > 0 )
	)
```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE }
hrs_design <- 
	update( 
		hrs_design , 

		one = 1 ,
		
		working_in_1996 = r3work ,

		working_in_2014 = r12work ,

		marital_status_in_1996 =
			factor( r3mstat , levels = 1:8 , labels =
				c( "Married" , "Married, spouse absent" ,
				"Partnered" , "Separated" , "Divorced" ,
				"Separated/divorced" , "Widowed" ,
				"Never married" ) ) ,
				
		marital_status_in_2014 =
			factor( r12mstat , levels = 1:8 , labels =
				c( "Married" , "Married, spouse absent" ,
				"Partnered" , "Separated" , "Divorced" ,
				"Separated/divorced" , "Widowed" ,
				"Never married" ) )
	)
```

### Unweighted Counts {-}

Count the unweighted number of records in the survey sample, overall and by groups:
```{r eval = FALSE , results = "hide" }
sum( weights( hrs_design , "sampling" ) != 0 )

svyby( ~ one , ~ marital_status_in_1996 , hrs_design , unwtd.count )
```

### Weighted Counts {-}
Count the weighted size of the generalizable population, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ one , hrs_design )

svyby( ~ one , ~ marital_status_in_1996 , hrs_design , svytotal )
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ h12ahous , hrs_design , na.rm = TRUE )

svyby( ~ h12ahous , ~ marital_status_in_1996 , hrs_design , svymean , na.rm = TRUE )
```

Calculate the distribution of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ marital_status_in_2014 , hrs_design , na.rm = TRUE )

svyby( ~ marital_status_in_2014 , ~ marital_status_in_1996 , hrs_design , svymean , na.rm = TRUE )
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ h12ahous , hrs_design , na.rm = TRUE )

svyby( ~ h12ahous , ~ marital_status_in_1996 , hrs_design , svytotal , na.rm = TRUE )
```

Calculate the weighted sum of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ marital_status_in_2014 , hrs_design , na.rm = TRUE )

svyby( ~ marital_status_in_2014 , ~ marital_status_in_1996 , hrs_design , svytotal , na.rm = TRUE )
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svyquantile( ~ h12ahous , hrs_design , 0.5 , na.rm = TRUE )

svyby( 
	~ h12ahous , 
	~ marital_status_in_1996 , 
	hrs_design , 
	svyquantile , 
	0.5 ,
	ci = TRUE ,
	keep.var = TRUE ,
	na.rm = TRUE
)
```

Estimate a ratio:
```{r eval = FALSE , results = "hide" }
svyratio( 
	numerator = ~ h4ahous , 
	denominator = ~ h12ahous , 
	hrs_design ,
	na.rm = TRUE
)
```

### Subsetting {-}

Restrict the survey design to :
```{r eval = FALSE , results = "hide" }
sub_hrs_design <- subset( hrs_design , working_in_1996 == 1 )
```
Calculate the mean (average) of this subset:
```{r eval = FALSE , results = "hide" }
svymean( ~ h12ahous , sub_hrs_design , na.rm = TRUE )
```

### Measures of Uncertainty {-}

Extract the coefficient, standard error, confidence interval, and coefficient of variation from any descriptive statistics function result, overall and by groups:
```{r eval = FALSE , results = "hide" }
this_result <- svymean( ~ h12ahous , hrs_design , na.rm = TRUE )

coef( this_result )
SE( this_result )
confint( this_result )
cv( this_result )

grouped_result <-
	svyby( 
		~ h12ahous , 
		~ marital_status_in_1996 , 
		hrs_design , 
		svymean ,
		na.rm = TRUE 
	)
	
coef( grouped_result )
SE( grouped_result )
confint( grouped_result )
cv( grouped_result )
```

Calculate the degrees of freedom of any survey design object:
```{r eval = FALSE , results = "hide" }
degf( hrs_design )
```

Calculate the complex sample survey-adjusted variance of any statistic:
```{r eval = FALSE , results = "hide" }
svyvar( ~ h12ahous , hrs_design , na.rm = TRUE )
```

Include the complex sample design effect in the result for a specific statistic:
```{r eval = FALSE , results = "hide" }
# SRS without replacement
svymean( ~ h12ahous , hrs_design , na.rm = TRUE , deff = TRUE )

# SRS with replacement
svymean( ~ h12ahous , hrs_design , na.rm = TRUE , deff = "replace" )
```

Compute confidence intervals for proportions using methods that may be more accurate near 0 and 1. See `?svyciprop` for alternatives:
```{r eval = FALSE , results = "hide" }
svyciprop( ~ working_in_2014 , hrs_design ,
	method = "likelihood" , na.rm = TRUE )
```

### Regression Models and Tests of Association {-}

Perform a design-based t-test:
```{r eval = FALSE , results = "hide" }
svyttest( h12ahous ~ working_in_2014 , hrs_design )
```

Perform a chi-squared test of association for survey data:
```{r eval = FALSE , results = "hide" }
svychisq( 
	~ working_in_2014 + marital_status_in_2014 , 
	hrs_design 
)
```

Perform a survey-weighted generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	svyglm( 
		h12ahous ~ working_in_2014 + marital_status_in_2014 , 
		hrs_design 
	)

summary( glm_result )
```

## Analysis Examples with `srvyr` {-}

The R `srvyr` library calculates summary statistics from survey data, such as the mean, total or quantile using [dplyr](https://github.com/tidyverse/dplyr/)-like syntax. [srvyr](https://github.com/gergness/srvyr) allows for the use of many verbs, such as `summarize`, `group_by`, and `mutate`, the convenience of pipe-able functions, the `tidyverse` style of non-standard evaluation and more consistent return types than the `survey` package. [This vignette](https://cran.r-project.org/web/packages/srvyr/vignettes/srvyr-vs-survey.html) details the available features. As a starting point for HRS users, this code replicates previously-presented examples:

```{r eval = FALSE , results = "hide" }
library(srvyr)
hrs_srvyr_design <- as_survey( hrs_design )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
hrs_srvyr_design %>%
	summarize( mean = survey_mean( h12ahous , na.rm = TRUE ) )

hrs_srvyr_design %>%
	group_by( marital_status_in_1996 ) %>%
	summarize( mean = survey_mean( h12ahous , na.rm = TRUE ) )
```

---

## Replication Example {-}

```{r eval = FALSE , results = "hide" }

```


