# National Plan and Provider Enumeration System (MTPS) {-}

The National Plan and Provider Enumeration System (NPPES) contains information about every medical provider, insurance plan, and clearinghouse actively operating in the United States healthcare industry.

* A single large table with one row per enumerated health care provider.

* A census of individuals and organizations who bill for medical services in the United States.

* Updated monthly with new providers.

* Maintained by the United States [Centers for Medicare & Medicaid Services (CMS)](http://www.cms.gov/)

## Simplified Download and Importation {-}

The R `lodown` package easily downloads and imports all available MTPS microdata by simply specifying `"mtps"` with an `output_dir =` parameter in the `lodown()` function. Depending on your internet connection and computer processing speed, you might prefer to run this step overnight.

```{r eval = FALSE }
library(lodown)
lodown( "mtps" , output_dir = file.path( path.expand( "~" ) , "MTPS" ) )
```

## Analysis Examples with SQL and `MonetDBLite` {-}

Connect to a database:

```{r eval = FALSE }
library(DBI)
dbdir <- file.path( path.expand( "~" ) , "MTPS" , "MonetDB" )
db <- dbConnect( MonetDBLite::MonetDBLite() , dbdir )
```

```{r eval = FALSE }

```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE }
dbSendQuery( db , "ALTER TABLE npi ADD COLUMN individual INTEGER" )

dbSendQuery( db , 
	"UPDATE npi 
	SET individual = 
		CASE WHEN entity_type_code = 1 THEN 1 ELSE 0 END" 
)

dbSendQuery( db , "ALTER TABLE npi ADD COLUMN provider_enumeration_year INTEGER" )

dbSendQuery( db , 
	"UPDATE npi 
	SET provider_enumeration_year = 
		CAST( SUBSTRING( provider_enumeration_date , 7 , 10 ) AS INTEGER )" 
)
```

### Unweighted Counts {-}

Count the unweighted number of records in the SQL table, overall and by groups:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db , "SELECT COUNT(*) FROM npi" )

dbGetQuery( db ,
	"SELECT
		provider_gender_code ,
		COUNT(*) 
	FROM npi
	GROUP BY provider_gender_code"
)
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db , "SELECT AVG( provider_enumeration_year ) FROM npi" )

dbGetQuery( db , 
	"SELECT 
		provider_gender_code , 
		AVG( provider_enumeration_year ) AS mean_provider_enumeration_year
	FROM npi 
	GROUP BY provider_gender_code" 
)
```

Initiate a function that allows division by zero:
```{r eval = FALSE , results = "hide" }
dbSendQuery( db , 
	"CREATE FUNCTION 
		div_noerror(l DOUBLE, r DOUBLE) 
	RETURNS DOUBLE 
	EXTERNAL NAME calc.div_noerror" 
)
```

Calculate the distribution of a categorical variable:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db , 
	"SELECT 
		is_sole_proprietor , 
		div_noerror( 
			COUNT(*) , 
			( SELECT COUNT(*) FROM npi ) 
		) AS share_is_sole_proprietor
	FROM npi 
	GROUP BY is_sole_proprietor" 
)
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db , "SELECT SUM( provider_enumeration_year ) FROM npi" )

dbGetQuery( db , 
	"SELECT 
		provider_gender_code , 
		SUM( provider_enumeration_year ) AS sum_provider_enumeration_year 
	FROM npi 
	GROUP BY provider_gender_code" 
)
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db , "SELECT QUANTILE( provider_enumeration_year , 0.5 ) FROM npi" )

dbGetQuery( db , 
	"SELECT 
		provider_gender_code , 
		QUANTILE( provider_enumeration_year , 0.5 ) AS median_provider_enumeration_year
	FROM npi 
	GROUP BY provider_gender_code" 
)
```

### Subsetting {-}

Limit your SQL analysis to California with `WHERE`:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db ,
	"SELECT
		AVG( provider_enumeration_year )
	FROM npi
	WHERE provider_business_practice_location_address_state_name = 'CA'"
)
```

### Measures of Uncertainty {-}

Calculate the variance and standard deviation, overall and by groups:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db , 
	"SELECT 
		VAR_SAMP( provider_enumeration_year ) , 
		STDDEV_SAMP( provider_enumeration_year ) 
	FROM npi" 
)

dbGetQuery( db , 
	"SELECT 
		provider_gender_code , 
		VAR_SAMP( provider_enumeration_year ) AS var_provider_enumeration_year ,
		STDDEV_SAMP( provider_enumeration_year ) AS stddev_provider_enumeration_year
	FROM npi 
	GROUP BY provider_gender_code" 
)
```

### Regression Models and Tests of Association {-}

Calculate the correlation between two variables, overall and by groups:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db , 
	"SELECT 
		CORR( CAST( individual AS DOUBLE ) , CAST( provider_enumeration_year AS DOUBLE ) )
	FROM npi" 
)

dbGetQuery( db , 
	"SELECT 
		provider_gender_code , 
		CORR( CAST( individual AS DOUBLE ) , CAST( provider_enumeration_year AS DOUBLE ) )
	FROM npi 
	GROUP BY provider_gender_code" 
)
```

## Analysis Examples with `dplyr` {-}

The R `dplyr` library offers an alternative grammar of data manipulation to base R and SQL syntax. [dplyr](https://github.com/tidyverse/dplyr/) offers many verbs, such as `summarize`, `group_by`, and `mutate`, the convenience of pipe-able functions, and the `tidyverse` style of non-standard evaluation. [This vignette](https://cran.r-project.org/web/packages/dplyr/vignettes/dplyr.html) details the available features. As a starting point for MTPS users, this code replicates previously-presented examples:

```{r eval = FALSE , results = "hide" }
library(dplyr)
dplyr_db <- MonetDBLite::src_monetdblite( dbdir )
mtps_tbl <- tbl( dplyr_db , 'npi' )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
mtps_tbl %>%
	summarize( mean = mean( provider_enumeration_year ) )

mtps_tbl %>%
	group_by( provider_gender_code ) %>%
	summarize( mean = mean( provider_enumeration_year ) )
```

---

## Replication Example {-}

```{r eval = FALSE , results = "hide" }
dbGetQuery( db , "SELECT COUNT(*) FROM npi" )
```

## Database Shutdown {-}

```{r eval = FALSE , results = "hide" }
dbDisconnect( db , shutdown = TRUE )
```
