# National Health Interview Survey (NHIS) {-}

[![Build Status](https://travis-ci.org/asdfree/nhis.svg?branch=master)](https://travis-ci.org/asdfree/nhis) [![Build status](https://ci.appveyor.com/api/projects/status/github/asdfree/nhis?svg=TRUE)](https://ci.appveyor.com/project/ajdamico/nhis)

The National Health Interview Survey (NHIS) is America's most detailed household survey of health status and medical experience.

* A main table with one row for each person within each sampled household, mergeable other tables like the sample child table with a more detailed questionnaire for only one child (when available) within each sampled household.

* A complex sample survey designed to generalize to the civilian non-institutionalized population of the United States.

* Released annually since 1963, the most recent major re-design in 1997.

* Administered by the [Centers for Disease Control and Prevention](http://www.cdc.gov/).

## Simplified Download and Importation {-}

The R `lodown` package easily downloads and imports all available NHIS microdata by simply specifying `"nhis"` with an `output_dir =` parameter in the `lodown()` function. Depending on your internet connection and computer processing speed, you might prefer to run this step overnight.

```{r eval = FALSE }
library(lodown)
lodown( "nhis" , output_dir = file.path( path.expand( "~" ) , "NHIS" ) )
```

`lodown` also provides a catalog of available microdata extracts with the `get_catalog()` function. After requesting the NHIS catalog, you could pass a subsetted catalog through the `lodown()` function in order to download and import specific extracts (rather than all available extracts).

```{r eval = FALSE , results = "hide" }
library(lodown)
# examine all available NHIS microdata files
nhis_cat <-
	get_catalog( "nhis" ,
		output_dir = file.path( path.expand( "~" ) , "NHIS" ) )

# 2016 only
nhis_cat <- subset( nhis_cat , year == 2016 )
# download the microdata to your local computer
lodown( "nhis" , nhis_cat )
```

## Analysis Examples with the `survey` library {-}

Construct a multiply-imputed, complex sample survey design:

```{r eval = FALSE }

```

```{r eval = FALSE }
options( survey.lonely.psu = "adjust" )

library(survey)
library(mitools)

nhis_personsx_df <- 
	readRDS( file.path( path.expand( "~" ) , "NHIS" , "2016/personsx.rds" ) )

nhis_income_list <- 
	readRDS( file.path( path.expand( "~" ) , "NHIS" , "2016/incmimp.rds" ) )

merge_variables <- c( "hhx" , "fmx" , "fpx" )

nhis_personsx_df[ merge_variables ] <- 
	sapply( nhis_personsx_df[ merge_variables ] , as.numeric )

inc_vars_to_keep <- 
	c( 
		merge_variables , 
		setdiff( 
			names( nhis_income_list[[ 1 ]] ) , 
			names( nhis_personsx_df )
		)
	)

# personsx variables to keep
vars_to_keep <- 
	c( merge_variables , "ppsu" , "pstrat" , "wtfa" ,
		'phstat' , 'sex' , 'hospno' , 'age_p' , 'hinotmyr' , 'notcov' )

nhis_personsx_df <- nhis_personsx_df[ vars_to_keep ]
	
nhis_personsx_list <-
	lapply( nhis_income_list ,
		function( w ){
			w <- w[ inc_vars_to_keep ]
			w[ merge_variables ] <- sapply( w[ merge_variables ] , as.numeric )
			result <- merge( nhis_personsx_df , w )
			stopifnot( nrow( result ) == nrow( nhis_personsx_df ) )
			result
		} )

# personsx design		
nhis_design <- 
	svydesign( 
		id = ~ppsu , 
		strata = ~pstrat ,
		nest = TRUE ,
		weights = ~wtfa ,
		data = imputationList( nhis_personsx_list )
	)

rm( nhis_personsx_list ) ; gc()

nhis_samadult_df <- 
	readRDS( file.path( path.expand( "~" ) , "NHIS" , "2016/samadult.rds" ) )

nhis_samadult_df[ merge_variables ] <- 
	sapply( nhis_samadult_df[ merge_variables ] , as.numeric )

samadult_vars_to_keep <- 
	c( 
		merge_variables , 
		setdiff( 
			names( nhis_samadult_df ) , 
			names( nhis_personsx_df ) 
		) 
	)

nhis_personsx_samadult_df <-
	merge( nhis_personsx_df , nhis_samadult_df[ samadult_vars_to_keep ] )

stopifnot( nrow( nhis_personsx_samadult_df ) == nrow( nhis_samadult_df ) )

rm( nhis_personsx_df , nhis_samadult_df ) ; gc()

nhis_samadult_list <-
	lapply( nhis_income_list ,
		function( w ){
			w <- w[ inc_vars_to_keep ]
			w[ merge_variables ] <- sapply( w[ merge_variables ] , as.numeric )
			result <- merge( nhis_personsx_samadult_df , w )
			stopifnot( nrow( result ) == nrow( nhis_personsx_samadult_df ) )
			result
		} )

rm( nhis_income_list , nhis_personsx_samadult_df ) ; gc()

# sample adult design (commented out)
# nhis_samadult_design <- 
	# svydesign( 
		# id = ~ppsu , 
		# strata = ~pstrat ,
		# nest = TRUE ,
		# weights = ~wtfa_sa ,
		# data = imputationList( nhis_samadult_list )
	# )
	
rm( nhis_samadult_list ) ; gc()
```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE }
nhis_design <- 
	update( 
		nhis_design , 
		
		one = 1 ,
		
		poverty_category =
			factor( 
				findInterval( povrati3 , 1:4 ) ,
				labels = 
					c( "below poverty" , "100-199%" , "200-299%" , "300-399%" , "400%+" )
			) ,
			
		fair_or_poor_reported_health = 
			ifelse( phstat %in% 1:5 , as.numeric( phstat >= 4 ) , NA ) ,
			
		sex = factor( sex , labels = c( "male" , "female" ) ) ,
		
		hospno = ifelse( hospno > 366 , NA , hospno )

	)
```

### Unweighted Counts {-}

Count the unweighted number of records in the survey sample, overall and by groups:
```{r eval = FALSE , results = "hide" }
MIcombine( with( nhis_design , svyby( ~ one , ~ one , unwtd.count ) ) )

MIcombine( with( nhis_design , svyby( ~ one , ~ poverty_category , unwtd.count ) ) )
```

### Weighted Counts {-}
Count the weighted size of the generalizable population, overall and by groups:
```{r eval = FALSE , results = "hide" }
MIcombine( with( nhis_design , svytotal( ~ one ) ) )

MIcombine( with( nhis_design ,
	svyby( ~ one , ~ poverty_category , svytotal )
) )
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
MIcombine( with( nhis_design , svymean( ~ age_p ) ) )

MIcombine( with( nhis_design ,
	svyby( ~ age_p , ~ poverty_category , svymean )
) )
```

Calculate the distribution of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
MIcombine( with( nhis_design , svymean( ~ sex ) ) )

MIcombine( with( nhis_design ,
	svyby( ~ sex , ~ poverty_category , svymean )
) )
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
MIcombine( with( nhis_design , svytotal( ~ age_p ) ) )

MIcombine( with( nhis_design ,
	svyby( ~ age_p , ~ poverty_category , svytotal )
) )
```

Calculate the weighted sum of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
MIcombine( with( nhis_design , svytotal( ~ sex ) ) )

MIcombine( with( nhis_design ,
	svyby( ~ sex , ~ poverty_category , svytotal )
) )
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
MIcombine( with( nhis_design , svyquantile( ~ age_p , 0.5 , se = TRUE ) ) )

MIcombine( with( nhis_design ,
	svyby( 
		~ age_p , ~ poverty_category , svyquantile , 0.5 ,
		se = TRUE , keep.var = TRUE , ci = TRUE 
) ) )
```

Estimate a ratio:
```{r eval = FALSE , results = "hide" }
MIcombine( with( nhis_design ,
	svyratio( numerator = ~ hinotmyr , denominator = ~ hospno , na.rm = TRUE )
) )
```

### Subsetting {-}

Restrict the survey design to uninsured:
```{r eval = FALSE , results = "hide" }
sub_nhis_design <- subset( nhis_design , notcov == 1 )
```
Calculate the mean (average) of this subset:
```{r eval = FALSE , results = "hide" }
MIcombine( with( sub_nhis_design , svymean( ~ age_p ) ) )
```

### Measures of Uncertainty {-}

Extract the coefficient, standard error, confidence interval, and coefficient of variation from any descriptive statistics function result, overall and by groups:
```{r eval = FALSE , results = "hide" }
this_result <-
	MIcombine( with( nhis_design ,
		svymean( ~ age_p )
	) )

coef( this_result )
SE( this_result )
confint( this_result )
cv( this_result )

grouped_result <-
	MIcombine( with( nhis_design ,
		svyby( ~ age_p , ~ poverty_category , svymean )
	) )

coef( grouped_result )
SE( grouped_result )
confint( grouped_result )
cv( grouped_result )
```

Calculate the degrees of freedom of any survey design object:
```{r eval = FALSE , results = "hide" }
degf( nhis_design$designs[[1]] )
```

Calculate the complex sample survey-adjusted variance of any statistic:
```{r eval = FALSE , results = "hide" }
MIcombine( with( nhis_design , svyvar( ~ age_p ) ) )
```

Include the complex sample design effect in the result for a specific statistic:
```{r eval = FALSE , results = "hide" }
# SRS without replacement
MIcombine( with( nhis_design ,
	svymean( ~ age_p , deff = TRUE )
) )

# SRS with replacement
MIcombine( with( nhis_design ,
	svymean( ~ age_p , deff = "replace" )
) )
```

Compute confidence intervals for proportions using methods that may be more accurate near 0 and 1. See `?svyciprop` for alternatives:
```{r eval = FALSE , results = "hide" }
lodown:::MIsvyciprop( ~ fair_or_poor_reported_health , nhis_design ,
	method = "likelihood" , na.rm = TRUE )
```

### Regression Models and Tests of Association {-}

Perform a design-based t-test:
```{r eval = FALSE , results = "hide" }
lodown:::MIsvyttest( age_p ~ fair_or_poor_reported_health , nhis_design )
```

Perform a chi-squared test of association for survey data:
```{r eval = FALSE , results = "hide" }
lodown:::MIsvychisq( ~ fair_or_poor_reported_health + sex , nhis_design )
```

Perform a survey-weighted generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	MIcombine( with( nhis_design ,
		svyglm( age_p ~ fair_or_poor_reported_health + sex )
	) )
	
summary( glm_result )
```

---

## Replication Example {-}

```{r eval = FALSE , results = "hide" }

```


