# National Survey of OAA Participants (NPS) {-}

<a href="https://github.com/asdfree/nps/actions"><img src="https://github.com/asdfree/nps/actions/workflows/r.yml/badge.svg" alt="Github Actions Badge"></a>

The National Survey of OAA Participants measures program satisfaction with state agency community services for American seniors.

* One table with one row per sampled senior respondent.

* A complex sample survey designed to generalize to non-institutionalized beneficiaries of Area Agencies on Aging (AAA) within the United States.

* Released annually since 2003.

* Administered by the [U.S. Administration on Aging](http://www.aoa.gov/).

## Simplified Download and Importation {-}

The R `lodown` package easily downloads and imports all available NPS microdata by simply specifying `"nps"` with an `output_dir =` parameter in the `lodown()` function. Depending on your internet connection and computer processing speed, you might prefer to run this step overnight.

```{r eval = FALSE }
library(lodown)
lodown( "nps" , output_dir = file.path( path.expand( "~" ) , "NPS" ) )
```

`lodown` also provides a catalog of available microdata extracts with the `get_catalog()` function. After requesting the NPS catalog, you could pass a subsetted catalog through the `lodown()` function in order to download and import specific extracts (rather than all available extracts).

```{r eval = FALSE , results = "hide" }
library(lodown)
# examine all available NPS microdata files
nps_cat <-
	get_catalog( "nps" ,
		output_dir = file.path( path.expand( "~" ) , "NPS" ) )

# 2015 only 
nps_cat <- subset( nps_cat , year == 2015 )
# download the microdata to your local computer
nps_cat <- lodown( "nps" , nps_cat )
```

## Analysis Examples with the `survey` library \ {-}

Construct a complex sample survey design:

```{r eval = FALSE }

```

```{r eval = FALSE }
library(survey)

nps_df <- 
	readRDS( 
		file.path( path.expand( "~" ) , "NPS" , 
			"2015 transportation.rds" ) )

nps_design <- 
	svrepdesign( 
		data = nps_df , 
		repweights = "pstotwgt[0-9]" , 
		weights = ~ pstotwgt , 
		type = "Fay" , 
		rho = 0.29986 , 
		mse = TRUE
	)
```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE }
nps_design <- 
	update( 
		nps_design , 
		
		age_category =
			factor( agec , levels = 2:5 , labels =
			c( "60-64" , "65-74" , "75-84" , "85+" ) ) ,
		
		gender = factor( gender , labels = c( "male" , "female" ) ) ,
		
		trip_this_week = as.numeric( trdays %in% 1:2 )

	)
```

### Unweighted Counts {-}

Count the unweighted number of records in the survey sample, overall and by groups:
```{r eval = FALSE , results = "hide" }
sum( weights( nps_design , "sampling" ) != 0 )

svyby( ~ one , ~ age_category , nps_design , unwtd.count )
```

### Weighted Counts {-}
Count the weighted size of the generalizable population, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ one , nps_design )

svyby( ~ one , ~ age_category , nps_design , svytotal )
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ adlaoa6p , nps_design , na.rm = TRUE )

svyby( ~ adlaoa6p , ~ age_category , nps_design , svymean , na.rm = TRUE )
```

Calculate the distribution of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ gender , nps_design )

svyby( ~ gender , ~ age_category , nps_design , svymean )
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ adlaoa6p , nps_design , na.rm = TRUE )

svyby( ~ adlaoa6p , ~ age_category , nps_design , svytotal , na.rm = TRUE )
```

Calculate the weighted sum of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ gender , nps_design )

svyby( ~ gender , ~ age_category , nps_design , svytotal )
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svyquantile( ~ adlaoa6p , nps_design , 0.5 , na.rm = TRUE )

svyby( 
	~ adlaoa6p , 
	~ age_category , 
	nps_design , 
	svyquantile , 
	0.5 ,
	ci = TRUE ,
	keep.var = TRUE ,
	na.rm = TRUE
)
```

Estimate a ratio:
```{r eval = FALSE , results = "hide" }
svyratio( 
	numerator = ~ adlaoa6p , 
	denominator = ~ iadlaoa7 , 
	nps_design ,
	na.rm = TRUE
)
```

### Subsetting {-}

Restrict the survey design to beneficiaries who live alone:
```{r eval = FALSE , results = "hide" }
sub_nps_design <- subset( nps_design , livealone == 1 )
```
Calculate the mean (average) of this subset:
```{r eval = FALSE , results = "hide" }
svymean( ~ adlaoa6p , sub_nps_design , na.rm = TRUE )
```

### Measures of Uncertainty {-}

Extract the coefficient, standard error, confidence interval, and coefficient of variation from any descriptive statistics function result, overall and by groups:
```{r eval = FALSE , results = "hide" }
this_result <- svymean( ~ adlaoa6p , nps_design , na.rm = TRUE )

coef( this_result )
SE( this_result )
confint( this_result )
cv( this_result )

grouped_result <-
	svyby( 
		~ adlaoa6p , 
		~ age_category , 
		nps_design , 
		svymean ,
		na.rm = TRUE 
	)
	
coef( grouped_result )
SE( grouped_result )
confint( grouped_result )
cv( grouped_result )
```

Calculate the degrees of freedom of any survey design object:
```{r eval = FALSE , results = "hide" }
degf( nps_design )
```

Calculate the complex sample survey-adjusted variance of any statistic:
```{r eval = FALSE , results = "hide" }
svyvar( ~ adlaoa6p , nps_design , na.rm = TRUE )
```

Include the complex sample design effect in the result for a specific statistic:
```{r eval = FALSE , results = "hide" }
# SRS without replacement
svymean( ~ adlaoa6p , nps_design , na.rm = TRUE , deff = TRUE )

# SRS with replacement
svymean( ~ adlaoa6p , nps_design , na.rm = TRUE , deff = "replace" )
```

Compute confidence intervals for proportions using methods that may be more accurate near 0 and 1. See `?svyciprop` for alternatives:
```{r eval = FALSE , results = "hide" }
svyciprop( ~ trip_this_week , nps_design ,
	method = "likelihood" )
```

### Regression Models and Tests of Association {-}

Perform a design-based t-test:
```{r eval = FALSE , results = "hide" }
svyttest( adlaoa6p ~ trip_this_week , nps_design )
```

Perform a chi-squared test of association for survey data:
```{r eval = FALSE , results = "hide" }
svychisq( 
	~ trip_this_week + gender , 
	nps_design 
)
```

Perform a survey-weighted generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	svyglm( 
		adlaoa6p ~ trip_this_week + gender , 
		nps_design 
	)

summary( glm_result )
```

## Analysis Examples with `srvyr` \ {-}

The R `srvyr` library calculates summary statistics from survey data, such as the mean, total or quantile using [dplyr](https://github.com/tidyverse/dplyr/)-like syntax. [srvyr](https://github.com/gergness/srvyr) allows for the use of many verbs, such as `summarize`, `group_by`, and `mutate`, the convenience of pipe-able functions, the `tidyverse` style of non-standard evaluation and more consistent return types than the `survey` package. [This vignette](https://cran.r-project.org/web/packages/srvyr/vignettes/srvyr-vs-survey.html) details the available features. As a starting point for NPS users, this code replicates previously-presented examples:

```{r eval = FALSE , results = "hide" }
library(srvyr)
nps_srvyr_design <- as_survey( nps_design )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
nps_srvyr_design %>%
	summarize( mean = survey_mean( adlaoa6p , na.rm = TRUE ) )

nps_srvyr_design %>%
	group_by( age_category ) %>%
	summarize( mean = survey_mean( adlaoa6p , na.rm = TRUE ) )
```

---

## Replication Example {-}

```{r eval = FALSE , results = "hide" }

```

