# Public Libraries Survey (PLS) {-}

needs_travis_build_status_line

An annual census of public libraries in the United States.

* One table with one row per state, a second table with one row per library system, and a third table with one row per library building or bookmobile.

* Released annually since 1992.

* Conducted by the [Institute of Museum and Library Services (IMLS)](https://www.imls.gov/) and collected by the [US Census Bureau](http://www.census.gov/).

## Simplified Download and Importation {-}

The R `lodown` package easily downloads and imports all available PLS microdata by simply specifying `"pls"` with an `output_dir =` parameter in the `lodown()` function. Depending on your internet connection and computer processing speed, you might prefer to run this step overnight.

```{r eval = FALSE }
library(lodown)
lodown( "pls" , output_dir = file.path( path.expand( "~" ) , "PLS" ) )
```

## Analysis Examples with base R \ {-}

Load a data frame:

```{r eval = FALSE }
pls_df <- readRDS( file.path( path.expand( "~" ) , "PLS" , "2014/pls_fy_ae_puplda.rds" ) )
```

```{r eval = FALSE }

```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE }
pls_df <- 
	transform( 
		pls_df , 
		
		c_relatn = 
			factor( c_relatn , levels = c( "HQ" , "ME" , "NO" ) ,
				c( "HQ-Headquarters of a federation or cooperative" ,
				"ME-Member of a federation or cooperative" ,
				"NO-Not a member of a federation or cooperative" )
			) ,
			
		more_than_one_librarian = as.numeric( libraria > 1 )
				
	)	
```

### Unweighted Counts {-}

Count the unweighted number of records in the table, overall and by groups:
```{r eval = FALSE , results = "hide" }
nrow( pls_df )

table( pls_df[ , "stabr" ] , useNA = "always" )
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
mean( pls_df[ , "popu_lsa" ] )

tapply(
	pls_df[ , "popu_lsa" ] ,
	pls_df[ , "stabr" ] ,
	mean 
)
```

Calculate the distribution of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
prop.table( table( pls_df[ , "c_relatn" ] ) )

prop.table(
	table( pls_df[ , c( "c_relatn" , "stabr" ) ] ) ,
	margin = 2
)
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
sum( pls_df[ , "popu_lsa" ] )

tapply(
	pls_df[ , "popu_lsa" ] ,
	pls_df[ , "stabr" ] ,
	sum 
)
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
quantile( pls_df[ , "popu_lsa" ] , 0.5 )

tapply(
	pls_df[ , "popu_lsa" ] ,
	pls_df[ , "stabr" ] ,
	quantile ,
	0.5 
)
```

### Subsetting {-}

Limit your `data.frame` to more than one million annual visits:
```{r eval = FALSE , results = "hide" }
sub_pls_df <- subset( pls_df , visits > 1000000 )
```
Calculate the mean (average) of this subset:
```{r eval = FALSE , results = "hide" }
mean( sub_pls_df[ , "popu_lsa" ] )
```

### Measures of Uncertainty {-}

Calculate the variance, overall and by groups:
```{r eval = FALSE , results = "hide" }
var( pls_df[ , "popu_lsa" ] )

tapply(
	pls_df[ , "popu_lsa" ] ,
	pls_df[ , "stabr" ] ,
	var 
)
```

### Regression Models and Tests of Association {-}

Perform a t-test:
```{r eval = FALSE , results = "hide" }
t.test( popu_lsa ~ more_than_one_librarian , pls_df )
```

Perform a chi-squared test of association:
```{r eval = FALSE , results = "hide" }
this_table <- table( pls_df[ , c( "more_than_one_librarian" , "c_relatn" ) ] )

chisq.test( this_table )
```

Perform a generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	glm( 
		popu_lsa ~ more_than_one_librarian + c_relatn , 
		data = pls_df
	)

summary( glm_result )
```

## Analysis Examples with `dplyr` \ {-}

The R `dplyr` library offers an alternative grammar of data manipulation to base R and SQL syntax. [dplyr](https://github.com/tidyverse/dplyr/) offers many verbs, such as `summarize`, `group_by`, and `mutate`, the convenience of pipe-able functions, and the `tidyverse` style of non-standard evaluation. [This vignette](https://cran.r-project.org/web/packages/dplyr/vignettes/dplyr.html) details the available features. As a starting point for PLS users, this code replicates previously-presented examples:

```{r eval = FALSE , results = "hide" }
library(dplyr)
pls_tbl <- tbl_df( pls_df )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
pls_tbl %>%
	summarize( mean = mean( popu_lsa ) )

pls_tbl %>%
	group_by( stabr ) %>%
	summarize( mean = mean( popu_lsa ) )
```


