# Program for International Student Assessment (PISA) {-}

[![Build Status](https://travis-ci.org/asdfree/pisa.svg?branch=master)](https://travis-ci.org/asdfree/pisa) [![Build status](https://ci.appveyor.com/api/projects/status/github/asdfree/pisa?svg=TRUE)](https://ci.appveyor.com/project/ajdamico/pisa)

The authoritative source for evaluating educational achievement across nations, the Program(me) for International Student Assessment ranks the math, science, and reading skills of high school students across the developed world.

* A large table with one row per student, a smaller table with one row per school, then multiple (optional) tables such as one row per parent or per teacher.

* A complex sample survey designed to generalize to 15-year-old schoolchildren in more than sixty countries.

* Released triennially since 2000.

* Administered by the [OECD](www.oecd.org/).

## Simplified Download and Importation {-}

The R `lodown` package easily downloads and imports all available PISA microdata by simply specifying `"pisa"` with an `output_dir =` parameter in the `lodown()` function. Depending on your internet connection and computer processing speed, you might prefer to run this step overnight.

```{r eval = FALSE }
library(lodown)
lodown( "pisa" , output_dir = file.path( path.expand( "~" ) , "PISA" ) )
```

`lodown` also provides a catalog of available microdata extracts with the `get_catalog()` function. After requesting the PISA catalog, you could pass a subsetted catalog through the `lodown()` function in order to download and import specific extracts (rather than all available extracts).

```{r eval = FALSE , results = "hide" }
library(lodown)
# examine all available PISA microdata files
pisa_cat <-
	get_catalog( "pisa" ,
		output_dir = file.path( path.expand( "~" ) , "PISA" ) )

# 2015 only
pisa_cat <- subset( pisa_cat , year == 2015 )
# download the microdata to your local computer
lodown( "pisa" , pisa_cat )
```

## Analysis Examples with the `survey` library \ {-}

Construct a multiply-imputed, database-backed complex sample survey design:

```{r eval = FALSE }

```

```{r eval = FALSE }
library(DBI)
library(RSQLite)
library(survey)
library(mitools)

pisa_design <- readRDS( file.path( path.expand( "~" ) , "PISA" , "2015 cmb_stu_qqq design.rds" ) )

pisa_design <- lodown:::svyMDBdesign( pisa_design )
```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE }
pisa_design <- 
	update( 
		pisa_design , 
		
		gender = factor( st004d01t , labels = c( "male" , "female" ) ) ,
		
		how_many_computers_at_home = 
			factor( 
				st012q06na , 
				labels = c( "none" , "one" , "two" , "three or more" ) 
			)
 
	)
```

### Unweighted Counts {-}

Count the unweighted number of records in the survey sample, overall and by groups:
```{r eval = FALSE , results = "hide" }
MIcombine( with( pisa_design , svyby( ~ one , ~ one , unwtd.count ) ) )

MIcombine( with( pisa_design , svyby( ~ one , ~ gender , unwtd.count ) ) )
```

### Weighted Counts {-}
Count the weighted size of the generalizable population, overall and by groups:
```{r eval = FALSE , results = "hide" }
MIcombine( with( pisa_design , svytotal( ~ one ) ) )

MIcombine( with( pisa_design ,
	svyby( ~ one , ~ gender , svytotal )
) )
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
MIcombine( with( pisa_design , svymean( ~ scie ) ) )

MIcombine( with( pisa_design ,
	svyby( ~ scie , ~ gender , svymean )
) )
```

Calculate the distribution of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
MIcombine( with( pisa_design , svymean( ~ how_many_computers_at_home ) ) )

MIcombine( with( pisa_design ,
	svyby( ~ how_many_computers_at_home , ~ gender , svymean )
) )
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
MIcombine( with( pisa_design , svytotal( ~ scie ) ) )

MIcombine( with( pisa_design ,
	svyby( ~ scie , ~ gender , svytotal )
) )
```

Calculate the weighted sum of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
MIcombine( with( pisa_design , svytotal( ~ how_many_computers_at_home ) ) )

MIcombine( with( pisa_design ,
	svyby( ~ how_many_computers_at_home , ~ gender , svytotal )
) )
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
MIcombine( with( pisa_design , svyquantile( ~ scie , 0.5 , se = TRUE ) ) )

MIcombine( with( pisa_design ,
	svyby( 
		~ scie , ~ gender , svyquantile , 0.5 ,
		se = TRUE , keep.var = TRUE , ci = TRUE 
) ) )
```

Estimate a ratio:
```{r eval = FALSE , results = "hide" }
MIcombine( with( pisa_design ,
	svyratio( numerator = ~ math , denominator = ~ reading )
) )
```

### Subsetting {-}

Restrict the survey design to Albania:
```{r eval = FALSE , results = "hide" }
sub_pisa_design <- subset( pisa_design , cnt == "ALB" )
```
Calculate the mean (average) of this subset:
```{r eval = FALSE , results = "hide" }
MIcombine( with( sub_pisa_design , svymean( ~ scie ) ) )
```

### Measures of Uncertainty {-}

Extract the coefficient, standard error, confidence interval, and coefficient of variation from any descriptive statistics function result, overall and by groups:
```{r eval = FALSE , results = "hide" }
this_result <-
	MIcombine( with( pisa_design ,
		svymean( ~ scie )
	) )

coef( this_result )
SE( this_result )
confint( this_result )
cv( this_result )

grouped_result <-
	MIcombine( with( pisa_design ,
		svyby( ~ scie , ~ gender , svymean )
	) )

coef( grouped_result )
SE( grouped_result )
confint( grouped_result )
cv( grouped_result )
```

Calculate the degrees of freedom of any survey design object:
```{r eval = FALSE , results = "hide" }
degf( pisa_design$designs[[1]] )
```

Calculate the complex sample survey-adjusted variance of any statistic:
```{r eval = FALSE , results = "hide" }
MIcombine( with( pisa_design , svyvar( ~ scie ) ) )
```

Include the complex sample design effect in the result for a specific statistic:
```{r eval = FALSE , results = "hide" }
# SRS without replacement
MIcombine( with( pisa_design ,
	svymean( ~ scie , deff = TRUE )
) )

# SRS with replacement
MIcombine( with( pisa_design ,
	svymean( ~ scie , deff = "replace" )
) )
```

Compute confidence intervals for proportions using methods that may be more accurate near 0 and 1. See `?svyciprop` for alternatives:
```{r eval = FALSE , results = "hide" }
lodown:::MIsvyciprop( ~ oecd , pisa_design ,
	method = "likelihood" )
```

### Regression Models and Tests of Association {-}

Perform a design-based t-test:
```{r eval = FALSE , results = "hide" }
lodown:::MIsvyttest( scie ~ oecd , pisa_design )
```

Perform a chi-squared test of association for survey data:
```{r eval = FALSE , results = "hide" }
lodown:::MIsvychisq( ~ oecd + how_many_computers_at_home , pisa_design )
```

Perform a survey-weighted generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	MIcombine( with( pisa_design ,
		svyglm( scie ~ oecd + how_many_computers_at_home )
	) )
	
summary( glm_result )
```

## Analysis Examples with `srvyr` \ {-}

The R `srvyr` library calculates summary statistics from survey data, such as the mean, total or quantile using [dplyr](https://github.com/tidyverse/dplyr/)-like syntax. [srvyr](https://github.com/gergness/srvyr) allows for the use of many verbs, such as `summarize`, `group_by`, and `mutate`, the convenience of pipe-able functions, the `tidyverse` style of non-standard evaluation and more consistent return types than the `survey` package. [This vignette](https://cran.r-project.org/web/packages/srvyr/vignettes/srvyr-vs-survey.html) details the available features. As a starting point for PISA users, this code replicates previously-presented examples:

```{r eval = FALSE , results = "hide" }
library(dbplyr)
library(srvyr)
pisa_srvyr_design <- as_survey( pisa_design )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
pisa_srvyr_design %>%
	summarize( mean = survey_mean( scie ) )

pisa_srvyr_design %>%
	group_by( gender ) %>%
	summarize( mean = survey_mean( scie ) )
```

---

## Replication Example {-}

```{r eval = FALSE , results = "hide" }

```
