# Progress in International Reading Literacy Study (PIRLS) {-}

[![Build Status](https://travis-ci.org/asdfree/pirls.svg?branch=master)](https://travis-ci.org/asdfree/pirls) [![Build status](https://ci.appveyor.com/api/projects/status/github/asdfree/pirls?svg=TRUE)](https://ci.appveyor.com/project/ajdamico/pirls)

The Progress in International Reading Literacy Study (PIRLS) tracks the reading competency of fourth graders across about fifty nations.

* A series of tables with one record per school (ACG), per student (ASG), per teacher (ATG), as well as files containing student achievement (ASA), home background (ASH), student-teacher linkage (AST), and within-country scoring reliability (ASR).

* A complex sample survey designed to generalize to the fourth-grade student population of participating countries.

* Released quinquennially since 2001.

* Funded by the [International Association for the Evaluation of Educational Achievement](https://www.iea.nl/) and compiled by the [Lynch School of Education at Boston College](http://www.bc.edu/bc-web/schools/lsoe.html).

## Simplified Download and Importation {-}

The R `lodown` package easily downloads and imports all available PIRLS microdata by simply specifying `"pirls"` with an `output_dir =` parameter in the `lodown()` function. Depending on your internet connection and computer processing speed, you might prefer to run this step overnight.

```{r eval = FALSE }
library(lodown)
lodown( "pirls" , output_dir = file.path( path.expand( "~" ) , "PIRLS" ) )
```

`lodown` also provides a catalog of available microdata extracts with the `get_catalog()` function. After requesting the PIRLS catalog, you could pass a subsetted catalog through the `lodown()` function in order to download and import specific extracts (rather than all available extracts).

```{r eval = FALSE , results = "hide" }
library(lodown)
# examine all available PIRLS microdata files
pirls_cat <-
	get_catalog( "pirls" ,
		output_dir = file.path( path.expand( "~" ) , "PIRLS" ) )

# 2011 only
pirls_cat <- subset( pirls_cat , year == 2011 )
# download the microdata to your local computer
lodown( "pirls" , pirls_cat )
```

## Analysis Examples with the `survey` library {-}

Construct a multiply-imputed, complex sample survey design:

```{r eval = FALSE }

```

```{r eval = FALSE }
library(survey)
library(mitools)

# load the ASG (student background) + ASH (home background) merged design
pirls_design <- readRDS( file.path( path.expand( "~" ) , "PIRLS" , "2011/asg_design.rds" ) )

# optional step to limit memory usage
variables_to_keep <-
	c( 'idcntry' , 'itsex' , 'itbirthy' , 'asrrea' , 'asrlit' )
	
pirls_design$designs <-
	lapply( 
		pirls_design$designs ,
		function( w ) {
			w$variables <- w$variables[ variables_to_keep ]
			w
		}
	)

gc()

```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE }
pirls_design <- 
	update( 
		pirls_design , 
		
		one = 1 ,
		
		idcntry = factor( idcntry ) ,
		
		sex = factor( itsex , labels = c( "male" , "female" ) ) ,
		
		born_2001_or_later = as.numeric( itbirthy >= 2001 )

	)
```

### Unweighted Counts {-}

Count the unweighted number of records in the survey sample, overall and by groups:
```{r eval = FALSE , results = "hide" }
lodown:::pirls_MIcombine( with( pirls_design , svyby( ~ one , ~ one , unwtd.count ) ) )

lodown:::pirls_MIcombine( with( pirls_design , svyby( ~ one , ~ idcntry , unwtd.count ) ) )
```

### Weighted Counts {-}
Count the weighted size of the generalizable population, overall and by groups:
```{r eval = FALSE , results = "hide" }
lodown:::pirls_MIcombine( with( pirls_design , svytotal( ~ one ) ) )

lodown:::pirls_MIcombine( with( pirls_design ,
	svyby( ~ one , ~ idcntry , svytotal )
) )
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
lodown:::pirls_MIcombine( with( pirls_design , svymean( ~ asrrea ) ) )

lodown:::pirls_MIcombine( with( pirls_design ,
	svyby( ~ asrrea , ~ idcntry , svymean )
) )
```

Calculate the distribution of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
lodown:::pirls_MIcombine( with( pirls_design , svymean( ~ sex , na.rm = TRUE ) ) )

lodown:::pirls_MIcombine( with( pirls_design ,
	svyby( ~ sex , ~ idcntry , svymean , na.rm = TRUE )
) )
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
lodown:::pirls_MIcombine( with( pirls_design , svytotal( ~ asrrea ) ) )

lodown:::pirls_MIcombine( with( pirls_design ,
	svyby( ~ asrrea , ~ idcntry , svytotal )
) )
```

Calculate the weighted sum of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
lodown:::pirls_MIcombine( with( pirls_design , svytotal( ~ sex , na.rm = TRUE ) ) )

lodown:::pirls_MIcombine( with( pirls_design ,
	svyby( ~ sex , ~ idcntry , svytotal , na.rm = TRUE )
) )
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
lodown:::pirls_MIcombine( with( pirls_design , svyquantile( ~ asrrea , 0.5 , se = TRUE ) ) )

lodown:::pirls_MIcombine( with( pirls_design ,
	svyby( 
		~ asrrea , ~ idcntry , svyquantile , 0.5 ,
		se = TRUE , keep.var = TRUE , ci = TRUE 
) ) )
```

Estimate a ratio:
```{r eval = FALSE , results = "hide" }
lodown:::pirls_MIcombine( with( pirls_design ,
	svyratio( numerator = ~ asrlit , denominator = ~ asrrea )
) )
```

### Subsetting {-}

Restrict the survey design to Australia, Austria, Azerbaijan, Belgium (French):
```{r eval = FALSE , results = "hide" }
sub_pirls_design <- subset( pirls_design , idcntry %in% c( 36 , 40 , 31 , 957 ) )
```
Calculate the mean (average) of this subset:
```{r eval = FALSE , results = "hide" }
lodown:::pirls_MIcombine( with( sub_pirls_design , svymean( ~ asrrea ) ) )
```

### Measures of Uncertainty {-}

Extract the coefficient, standard error, confidence interval, and coefficient of variation from any descriptive statistics function result, overall and by groups:
```{r eval = FALSE , results = "hide" }
this_result <-
	lodown:::pirls_MIcombine( with( pirls_design ,
		svymean( ~ asrrea )
	) )

coef( this_result )
SE( this_result )
confint( this_result )
cv( this_result )

grouped_result <-
	lodown:::pirls_MIcombine( with( pirls_design ,
		svyby( ~ asrrea , ~ idcntry , svymean )
	) )

coef( grouped_result )
SE( grouped_result )
confint( grouped_result )
cv( grouped_result )
```

Calculate the degrees of freedom of any survey design object:
```{r eval = FALSE , results = "hide" }
degf( pirls_design$designs[[1]] )
```

Calculate the complex sample survey-adjusted variance of any statistic:
```{r eval = FALSE , results = "hide" }
lodown:::pirls_MIcombine( with( pirls_design , svyvar( ~ asrrea ) ) )
```

Include the complex sample design effect in the result for a specific statistic:
```{r eval = FALSE , results = "hide" }
# SRS without replacement
lodown:::pirls_MIcombine( with( pirls_design ,
	svymean( ~ asrrea , deff = TRUE )
) )

# SRS with replacement
lodown:::pirls_MIcombine( with( pirls_design ,
	svymean( ~ asrrea , deff = "replace" )
) )
```

Compute confidence intervals for proportions using methods that may be more accurate near 0 and 1. See `?svyciprop` for alternatives:
```{r eval = FALSE , results = "hide" }
lodown:::MIsvyciprop( ~ born_2001_or_later , pirls_design ,
	method = "likelihood" , na.rm = TRUE )
```

### Regression Models and Tests of Association {-}

Perform a design-based t-test:
```{r eval = FALSE , results = "hide" }
lodown:::MIsvyttest( asrrea ~ born_2001_or_later , pirls_design )
```

Perform a chi-squared test of association for survey data:
```{r eval = FALSE , results = "hide" }
lodown:::MIsvychisq( ~ born_2001_or_later + sex , pirls_design )
```

Perform a survey-weighted generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	lodown:::pirls_MIcombine( with( pirls_design ,
		svyglm( asrrea ~ born_2001_or_later + sex )
	) )
	
summary( glm_result )
```

---

## Replication Example {-}

```{r eval = FALSE , results = "hide" }

```


