# Pesquisa Mensal de Emprego (PME) {-}

[![Build Status](https://travis-ci.org/asdfree/pme.svg?branch=master)](https://travis-ci.org/asdfree/pme) [![Build status](https://ci.appveyor.com/api/projects/status/github/asdfree/pme?svg=TRUE)](https://ci.appveyor.com/project/ajdamico/pme)

*Contributed by Dr. Djalma Pessoa <<pessoad@gmail.com>>*

The Pesquisa Mensal de Emprego (PME) is the monthly labor force survey covering the six largest Brazilian cities.

* One table with one row per individual within each sampled household.

* A complex sample survey designed to generalize to the civilian population of Brazil's six largest cities.

* Released monthly since March 2002.

* Administered by the [Instituto Brasileiro de Geografia e Estatistica](http://www.ibge.gov.br/).

## Simplified Download and Importation {-}

The R `lodown` package easily downloads and imports all available PME microdata by simply specifying `"pme"` with an `output_dir =` parameter in the `lodown()` function. Depending on your internet connection and computer processing speed, you might prefer to run this step overnight.

```{r eval = FALSE }
library(lodown)
lodown( "pme" , output_dir = file.path( path.expand( "~" ) , "PME" ) )
```

`lodown` also provides a catalog of available microdata extracts with the `get_catalog()` function. After requesting the PME catalog, you could pass a subsetted catalog through the `lodown()` function in order to download and import specific extracts (rather than all available extracts).

```{r eval = FALSE , results = "hide" }
library(lodown)
# examine all available PME microdata files
pme_cat <-
	get_catalog( "pme" ,
		output_dir = file.path( path.expand( "~" ) , "PME" ) )

# 2016 only
pme_cat <- subset( pme_cat , year == 2016 )
# download the microdata to your local computer
lodown( "pme" , pme_cat )
```

## Analysis Examples with the `survey` library \ {-}

Construct a complex sample survey design:

```{r eval = FALSE }

```

```{r eval = FALSE }
options( survey.lonely.psu = "adjust" )

library(survey)

pme_df <- readRDS( file.path( path.expand( "~" ) , "PME" , "pme 2016 01.rds" ) )

# throw out records missing their cluster variable
pme_df <- subset( pme_df , !is.na( v113 ) )

pop_totals <- unique( pme_df[ , c( 'v035' , 'v114' ) ] )

prestratified_design <- 
	svydesign( 
		~ v113 , 
		strata = ~ v112 , 
		data = pme_df ,
		weights = ~ v211 , 
		nest = TRUE
	)

pme_design <- 
	postStratify( prestratified_design , ~ v035 , pop_totals )
```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE }
pme_design <- 
	update( 
		pme_design , 

		one = 1 ,
		
		# calculate whether each person is at least ten years of age
		pia = as.numeric( v234 >= 10 ) ,

		# determine individuals who are employed
		ocup_c = as.numeric( v401 == 1 | v402 == 1 | v403 == 1 ) ,
		
		sexo = factor( v203 , labels = c( "male" , "female" ) ) ,
		
		region = 
			factor( 
				v035 , 
				levels = c( 26 , 29 , 31 , 33 , 35 , 43 ) , 
				labels = c( "Recife" , "Salvador" , "Belo Horizonte" , 
					"Rio de Janeiro" , "Sao Paulo" , "Porto Alegre" )
			)
	)
	
pme_design <-
	update(
		pme_design ,
		
		# determine individuals who are unemployed
		desocup30 = as.numeric( ocup_c == 0 & !is.na( v461 ) & v465 == 1 )
	)
		
pme_design <-
	update(
		pme_design ,
		
		# determine individuals who are either working or not working
		pea_c = as.numeric( ocup_c == 1 | desocup30 == 1 )

	)
```

### Unweighted Counts {-}

Count the unweighted number of records in the survey sample, overall and by groups:
```{r eval = FALSE , results = "hide" }
sum( weights( pme_design , "sampling" ) != 0 )

svyby( ~ one , ~ region , pme_design , unwtd.count )
```

### Weighted Counts {-}
Count the weighted size of the generalizable population, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ one , pme_design )

svyby( ~ one , ~ region , pme_design , svytotal )
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ vd25 , pme_design , na.rm = TRUE )

svyby( ~ vd25 , ~ region , pme_design , svymean , na.rm = TRUE )
```

Calculate the distribution of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ sexo , pme_design )

svyby( ~ sexo , ~ region , pme_design , svymean )
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ vd25 , pme_design , na.rm = TRUE )

svyby( ~ vd25 , ~ region , pme_design , svytotal , na.rm = TRUE )
```

Calculate the weighted sum of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ sexo , pme_design )

svyby( ~ sexo , ~ region , pme_design , svytotal )
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svyquantile( ~ vd25 , pme_design , 0.5 , na.rm = TRUE )

svyby( 
	~ vd25 , 
	~ region , 
	pme_design , 
	svyquantile , 
	0.5 ,
	ci = TRUE ,
	keep.var = TRUE ,
	na.rm = TRUE
)
```

Estimate a ratio:
```{r eval = FALSE , results = "hide" }
svyratio( 
	numerator = ~ desocup30 , 
	denominator = ~ pea_c , 
	pme_design ,
	na.rm = TRUE
)
```

### Subsetting {-}

Restrict the survey design to teenagers:
```{r eval = FALSE , results = "hide" }
sub_pme_design <- subset( pme_design , v234 %in% 13:19 )
```
Calculate the mean (average) of this subset:
```{r eval = FALSE , results = "hide" }
svymean( ~ vd25 , sub_pme_design , na.rm = TRUE )
```

### Measures of Uncertainty {-}

Extract the coefficient, standard error, confidence interval, and coefficient of variation from any descriptive statistics function result, overall and by groups:
```{r eval = FALSE , results = "hide" }
this_result <- svymean( ~ vd25 , pme_design , na.rm = TRUE )

coef( this_result )
SE( this_result )
confint( this_result )
cv( this_result )

grouped_result <-
	svyby( 
		~ vd25 , 
		~ region , 
		pme_design , 
		svymean ,
		na.rm = TRUE 
	)
	
coef( grouped_result )
SE( grouped_result )
confint( grouped_result )
cv( grouped_result )
```

Calculate the degrees of freedom of any survey design object:
```{r eval = FALSE , results = "hide" }
degf( pme_design )
```

Calculate the complex sample survey-adjusted variance of any statistic:
```{r eval = FALSE , results = "hide" }
svyvar( ~ vd25 , pme_design , na.rm = TRUE )
```

Include the complex sample design effect in the result for a specific statistic:
```{r eval = FALSE , results = "hide" }
# SRS without replacement
svymean( ~ vd25 , pme_design , na.rm = TRUE , deff = TRUE )

# SRS with replacement
svymean( ~ vd25 , pme_design , na.rm = TRUE , deff = "replace" )
```

Compute confidence intervals for proportions using methods that may be more accurate near 0 and 1. See `?svyciprop` for alternatives:
```{r eval = FALSE , results = "hide" }
svyciprop( ~ ocup_c , pme_design ,
	method = "likelihood" , na.rm = TRUE )
```

### Regression Models and Tests of Association {-}

Perform a design-based t-test:
```{r eval = FALSE , results = "hide" }
svyttest( vd25 ~ ocup_c , pme_design )
```

Perform a chi-squared test of association for survey data:
```{r eval = FALSE , results = "hide" }
svychisq( 
	~ ocup_c + sexo , 
	pme_design 
)
```

Perform a survey-weighted generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	svyglm( 
		vd25 ~ ocup_c + sexo , 
		pme_design 
	)

summary( glm_result )
```

## Analysis Examples with `srvyr` \ {-}

The R `srvyr` library calculates summary statistics from survey data, such as the mean, total or quantile using [dplyr](https://github.com/tidyverse/dplyr/)-like syntax. [srvyr](https://github.com/gergness/srvyr) allows for the use of many verbs, such as `summarize`, `group_by`, and `mutate`, the convenience of pipe-able functions, the `tidyverse` style of non-standard evaluation and more consistent return types than the `survey` package. [This vignette](https://cran.r-project.org/web/packages/srvyr/vignettes/srvyr-vs-survey.html) details the available features. As a starting point for PME users, this code replicates previously-presented examples:

```{r eval = FALSE , results = "hide" }
library(srvyr)
pme_srvyr_design <- as_survey( pme_design )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
pme_srvyr_design %>%
	summarize( mean = survey_mean( vd25 , na.rm = TRUE ) )

pme_srvyr_design %>%
	group_by( region ) %>%
	summarize( mean = survey_mean( vd25 , na.rm = TRUE ) )
```

---

## Replication Example {-}

```{r eval = FALSE , results = "hide" }

```

