# Pesquisa de Orcamentos Familiares (POF) {-}

[![Build Status](https://travis-ci.org/asdfree/pof.svg?branch=master)](https://travis-ci.org/asdfree/pof) [![Build status](https://ci.appveyor.com/api/projects/status/github/asdfree/pof?svg=TRUE)](https://ci.appveyor.com/project/ajdamico/pof)

*Contributed by Dr. Djalma Pessoa <<pessoad@gmail.com>>*

The Pesquisa de Orcamentos Familiares is Brazil's national survey of household budgets.

* One table of survey responses per sampled household. Additional tables, many containing one record per expenditure.

* A complex sample survey designed to generalize to the civilian population of Brazil.

* Released at irregular intervals, with only 2002-2003 and 2008-2009 microdata available.

* Administered by the [Instituto Brasileiro de Geografia e Estatistica](http://www.ibge.gov.br/).

## Simplified Download and Importation {-}

The R `lodown` package easily downloads and imports all available POF microdata by simply specifying `"pof"` with an `output_dir =` parameter in the `lodown()` function. Depending on your internet connection and computer processing speed, you might prefer to run this step overnight.

```{r eval = FALSE }
library(lodown)
lodown( "pof" , output_dir = file.path( path.expand( "~" ) , "POF" ) )
```

`lodown` also provides a catalog of available microdata extracts with the `get_catalog()` function. After requesting the POF catalog, you could pass a subsetted catalog through the `lodown()` function in order to download and import specific extracts (rather than all available extracts).

```{r eval = FALSE , results = "hide" }
library(lodown)
# examine all available POF microdata files
pof_cat <-
	get_catalog( "pof" ,
		output_dir = file.path( path.expand( "~" ) , "POF" ) )

# 2008-2009 only
pof_cat <- subset( pof_cat , period == "2008_2009" )
# download the microdata to your local computer
lodown( "pof" , pof_cat )
```

## Analysis Examples with the `survey` library \ {-}

Construct a complex sample survey design:

```{r eval = FALSE }

```

```{r eval = FALSE }
options( survey.lonely.psu = "adjust" )

library(survey)

poststr <- 
	readRDS( 
		file.path( path.expand( "~" ) , "POF" , 
			"2008_2009/poststr.rds" ) 
		)

		
t_morador_s <- 
	readRDS( 
		file.path( path.expand( "~" ) , "POF" , 
			"2008_2009/t_morador_s.rds" ) 
		)

t_morador_s <-
	transform(
		t_morador_s , 
		control = paste0( cod_uf , num_seq , num_dv ) 
	)
	
pof_df <- merge( t_morador_s , poststr )

stopifnot( nrow( pof_df ) == nrow( t_morador_s ) )

pre_stratified_design <- 
	svydesign(
		id = ~control , 
		strata = ~estrato_unico ,
		weights = ~fator_expansao1 ,
		data = pof_df ,
		nest = TRUE
	)

population_totals <- 
	data.frame(
		pos_estrato = unique( pof_df$pos_estrato ) , 
		Freq = unique( pof_df$tot_pop ) 
	)

pof_design <-
	postStratify(
		pre_stratified_design , 
		~ pos_estrato , 
		population_totals
	)
```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE }
pof_design <- 
	update(
		pof_design , 
		
		one = 1 ,
		
		# centimeters instead of meters
		altura_imputado = altura_imputado / 100 ,
		
		age_categories =
			factor( 
				1 + findInterval( idade_anos , 
					c( 20 , 25 , 30 , 35 , 45 , 55 , 65 , 75 ) ) ,
				levels = 1:9 , labels = c( "under 20" , "20-24" , "25-29" ,
				"30-34" , "35-44" , "45-54" , "55-64" , "65-74" , "75+" )
			) ,
		
		# create a body mass index (bmi) variable, excluding babies (who have altura_imputado==0)			
		body_mass_index = ifelse( altura_imputado == 0 , 0 , peso_imputado / ( altura_imputado ^ 2 ) ) ,
		
		sexo = ifelse( cod_sexo == '01' , "masculino" , ifelse( cod_sexo == '02' , "feminino" , NA ) )
		
		
	)

pof_design <-
	transform(
		pof_design ,
		
		# individuals with a low bmi - underweight
		underweight = ifelse( body_mass_index < 18.5 , 1 , 0 ) ,
		
		# individuals with a high bmi - overweight
		overweight = ifelse( body_mass_index >= 25 , 1 , 0 ) ,
		
		# individuals with a very high bmi - obese
		obese = ifelse( body_mass_index >= 30 , 1 , 0 )
	)
```

### Unweighted Counts {-}

Count the unweighted number of records in the survey sample, overall and by groups:
```{r eval = FALSE , results = "hide" }
sum( weights( pof_design , "sampling" ) != 0 )

svyby( ~ one , ~ sexo , pof_design , unwtd.count )
```

### Weighted Counts {-}
Count the weighted size of the generalizable population, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ one , pof_design )

svyby( ~ one , ~ sexo , pof_design , svytotal )
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ body_mass_index , pof_design , na.rm = TRUE )

svyby( ~ body_mass_index , ~ sexo , pof_design , svymean , na.rm = TRUE )
```

Calculate the distribution of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ age_categories , pof_design )

svyby( ~ age_categories , ~ sexo , pof_design , svymean )
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ body_mass_index , pof_design , na.rm = TRUE )

svyby( ~ body_mass_index , ~ sexo , pof_design , svytotal , na.rm = TRUE )
```

Calculate the weighted sum of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ age_categories , pof_design )

svyby( ~ age_categories , ~ sexo , pof_design , svytotal )
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svyquantile( ~ body_mass_index , pof_design , 0.5 , na.rm = TRUE )

svyby( 
	~ body_mass_index , 
	~ sexo , 
	pof_design , 
	svyquantile , 
	0.5 ,
	ci = TRUE ,
	keep.var = TRUE ,
	na.rm = TRUE
)
```

Estimate a ratio:
```{r eval = FALSE , results = "hide" }
svyratio( 
	numerator = ~ peso_imputado , 
	denominator = ~ altura_imputado , 
	pof_design ,
	na.rm = TRUE
)
```

### Subsetting {-}

Restrict the survey design to :
```{r eval = FALSE , results = "hide" }
sub_pof_design <- subset( pof_design , underweight == 1 )
```
Calculate the mean (average) of this subset:
```{r eval = FALSE , results = "hide" }
svymean( ~ body_mass_index , sub_pof_design , na.rm = TRUE )
```

### Measures of Uncertainty {-}

Extract the coefficient, standard error, confidence interval, and coefficient of variation from any descriptive statistics function result, overall and by groups:
```{r eval = FALSE , results = "hide" }
this_result <- svymean( ~ body_mass_index , pof_design , na.rm = TRUE )

coef( this_result )
SE( this_result )
confint( this_result )
cv( this_result )

grouped_result <-
	svyby( 
		~ body_mass_index , 
		~ sexo , 
		pof_design , 
		svymean ,
		na.rm = TRUE 
	)
	
coef( grouped_result )
SE( grouped_result )
confint( grouped_result )
cv( grouped_result )
```

Calculate the degrees of freedom of any survey design object:
```{r eval = FALSE , results = "hide" }
degf( pof_design )
```

Calculate the complex sample survey-adjusted variance of any statistic:
```{r eval = FALSE , results = "hide" }
svyvar( ~ body_mass_index , pof_design , na.rm = TRUE )
```

Include the complex sample design effect in the result for a specific statistic:
```{r eval = FALSE , results = "hide" }
# SRS without replacement
svymean( ~ body_mass_index , pof_design , na.rm = TRUE , deff = TRUE )

# SRS with replacement
svymean( ~ body_mass_index , pof_design , na.rm = TRUE , deff = "replace" )
```

Compute confidence intervals for proportions using methods that may be more accurate near 0 and 1. See `?svyciprop` for alternatives:
```{r eval = FALSE , results = "hide" }
svyciprop( ~ obese , pof_design ,
	method = "likelihood" , na.rm = TRUE )
```

### Regression Models and Tests of Association {-}

Perform a design-based t-test:
```{r eval = FALSE , results = "hide" }
svyttest( body_mass_index ~ obese , pof_design )
```

Perform a chi-squared test of association for survey data:
```{r eval = FALSE , results = "hide" }
svychisq( 
	~ obese + age_categories , 
	pof_design 
)
```

Perform a survey-weighted generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	svyglm( 
		body_mass_index ~ obese + age_categories , 
		pof_design 
	)

summary( glm_result )
```

## Analysis Examples with `srvyr` \ {-}

The R `srvyr` library calculates summary statistics from survey data, such as the mean, total or quantile using [dplyr](https://github.com/tidyverse/dplyr/)-like syntax. [srvyr](https://github.com/gergness/srvyr) allows for the use of many verbs, such as `summarize`, `group_by`, and `mutate`, the convenience of pipe-able functions, the `tidyverse` style of non-standard evaluation and more consistent return types than the `survey` package. [This vignette](https://cran.r-project.org/web/packages/srvyr/vignettes/srvyr-vs-survey.html) details the available features. As a starting point for POF users, this code replicates previously-presented examples:

```{r eval = FALSE , results = "hide" }
library(srvyr)
pof_srvyr_design <- as_survey( pof_design )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
pof_srvyr_design %>%
	summarize( mean = survey_mean( body_mass_index , na.rm = TRUE ) )

pof_srvyr_design %>%
	group_by( sexo ) %>%
	summarize( mean = survey_mean( body_mass_index , na.rm = TRUE ) )
```

---

## Replication Example {-}

```{r eval = FALSE , results = "hide" }

```

