# Pesquisa Nacional de Saude (PNS) {-}

[![Build Status](https://travis-ci.org/asdfree/pns.svg?branch=master)](https://travis-ci.org/asdfree/pns) [![Build status](https://ci.appveyor.com/api/projects/status/github/asdfree/pns?svg=TRUE)](https://ci.appveyor.com/project/ajdamico/pns)

*Contributed by Dr. Djalma Pessoa <<pessoad@gmail.com>>*

The Pesquisa Nacional de Saude (PNS) is Brazil's healthcare survey.

* One table with one row per long-questionnaire respondent and a second table with one row for all respondents.

* A complex sample survey designed to generalize to Brazil's civilian population.

* First released 2013.

* Administered by the [Instituto Brasileiro de Geografia e Estatistica](http://www.ibge.gov.br/).

## Simplified Download and Importation {-}

The R `lodown` package easily downloads and imports all available PNS microdata by simply specifying `"pns"` with an `output_dir =` parameter in the `lodown()` function. Depending on your internet connection and computer processing speed, you might prefer to run this step overnight.

```{r eval = FALSE }
library(lodown)
lodown( "pns" , output_dir = file.path( path.expand( "~" ) , "PNS" ) )
```

`lodown` also provides a catalog of available microdata extracts with the `get_catalog()` function. After requesting the PNS catalog, you could pass a subsetted catalog through the `lodown()` function in order to download and import specific extracts (rather than all available extracts).

```{r eval = FALSE , results = "hide" }
library(lodown)
# examine all available PNS microdata files
pns_cat <-
	get_catalog( "pns" ,
		output_dir = file.path( path.expand( "~" ) , "PNS" ) )

# download the microdata to your local computer
lodown( "pns" , pns_cat )
```

## Analysis Examples with the `survey` library {-}

Construct a complex sample survey design:

```{r eval = FALSE }

```

```{r eval = FALSE }
options( survey.lonely.psu = "adjust" )

library(survey)

pns_design <- 
	readRDS( 
		file.path( 
			path.expand( "~" ) , "PNS" , 
			"2013 long questionnaire survey design.rds" ) 
		)
```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE }
pns_design <- 
	update( 
		pns_design , 

		one = 1 ,
		
		health_insurance = as.numeric( i001 == 1 )
	)
```

### Unweighted Counts {-}

Count the unweighted number of records in the survey sample, overall and by groups:
```{r eval = FALSE , results = "hide" }
sum( weights( pns_design , "sampling" ) != 0 )

svyby( ~ one , ~ uf , pns_design , unwtd.count )
```

### Weighted Counts {-}
Count the weighted size of the generalizable population, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ one , pns_design )

svyby( ~ one , ~ uf , pns_design , svytotal )
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ w00101 , pns_design , na.rm = TRUE )

svyby( ~ w00101 , ~ uf , pns_design , svymean , na.rm = TRUE )
```

Calculate the distribution of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ c006 , pns_design )

svyby( ~ c006 , ~ uf , pns_design , svymean )
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ w00101 , pns_design , na.rm = TRUE )

svyby( ~ w00101 , ~ uf , pns_design , svytotal , na.rm = TRUE )
```

Calculate the weighted sum of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ c006 , pns_design )

svyby( ~ c006 , ~ uf , pns_design , svytotal )
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svyquantile( ~ w00101 , pns_design , 0.5 , na.rm = TRUE )

svyby( 
	~ w00101 , 
	~ uf , 
	pns_design , 
	svyquantile , 
	0.5 ,
	ci = TRUE ,
	keep.var = TRUE ,
	na.rm = TRUE
)
```

Estimate a ratio:
```{r eval = FALSE , results = "hide" }
svyratio( 
	numerator = ~ w00203 , 
	denominator = ~ w00101 , 
	pns_design ,
	na.rm = TRUE
)
```

### Subsetting {-}

Restrict the survey design to at least 30 minutes of physical activity:
```{r eval = FALSE , results = "hide" }
sub_pns_design <- subset( pns_design , atfi04 == 1 )
```
Calculate the mean (average) of this subset:
```{r eval = FALSE , results = "hide" }
svymean( ~ w00101 , sub_pns_design , na.rm = TRUE )
```

### Measures of Uncertainty {-}

Extract the coefficient, standard error, confidence interval, and coefficient of variation from any descriptive statistics function result, overall and by groups:
```{r eval = FALSE , results = "hide" }
this_result <- svymean( ~ w00101 , pns_design , na.rm = TRUE )

coef( this_result )
SE( this_result )
confint( this_result )
cv( this_result )

grouped_result <-
	svyby( 
		~ w00101 , 
		~ uf , 
		pns_design , 
		svymean ,
		na.rm = TRUE 
	)
	
coef( grouped_result )
SE( grouped_result )
confint( grouped_result )
cv( grouped_result )
```

Calculate the degrees of freedom of any survey design object:
```{r eval = FALSE , results = "hide" }
degf( pns_design )
```

Calculate the complex sample survey-adjusted variance of any statistic:
```{r eval = FALSE , results = "hide" }
svyvar( ~ w00101 , pns_design , na.rm = TRUE )
```

Include the complex sample design effect in the result for a specific statistic:
```{r eval = FALSE , results = "hide" }
# SRS without replacement
svymean( ~ w00101 , pns_design , na.rm = TRUE , deff = TRUE )

# SRS with replacement
svymean( ~ w00101 , pns_design , na.rm = TRUE , deff = "replace" )
```

Compute confidence intervals for proportions using methods that may be more accurate near 0 and 1. See `?svyciprop` for alternatives:
```{r eval = FALSE , results = "hide" }
svyciprop( ~ health_insurance , pns_design ,
	method = "likelihood" , na.rm = TRUE )
```

### Regression Models and Tests of Association {-}

Perform a design-based t-test:
```{r eval = FALSE , results = "hide" }
svyttest( w00101 ~ health_insurance , pns_design )
```

Perform a chi-squared test of association for survey data:
```{r eval = FALSE , results = "hide" }
svychisq( 
	~ health_insurance + c006 , 
	pns_design 
)
```

Perform a survey-weighted generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	svyglm( 
		w00101 ~ health_insurance + c006 , 
		pns_design 
	)

summary( glm_result )
```

## Analysis Examples with `srvyr` {-}

The R `srvyr` library calculates summary statistics from survey data, such as the mean, total or quantile using [dplyr](https://github.com/tidyverse/dplyr/)-like syntax. [srvyr](https://github.com/gergness/srvyr) allows for the use of many verbs, such as `summarize`, `group_by`, and `mutate`, the convenience of pipe-able functions, the `tidyverse` style of non-standard evaluation and more consistent return types than the `survey` package. [This vignette](https://cran.r-project.org/web/packages/srvyr/vignettes/srvyr-vs-survey.html) details the available features. As a starting point for PNS users, this code replicates previously-presented examples:

```{r eval = FALSE , results = "hide" }
library(srvyr)
pns_srvyr_design <- as_survey( pns_design )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
pns_srvyr_design %>%
	summarize( mean = survey_mean( w00101 , na.rm = TRUE ) )

pns_srvyr_design %>%
	group_by( uf ) %>%
	summarize( mean = survey_mean( w00101 , na.rm = TRUE ) )
```

---

## Replication Example {-}

```{r eval = FALSE , results = "hide" }

```


