# Survey of Health, Ageing and Retirement in Europe (SHARE) {-}

[![Build Status](https://travis-ci.org/asdfree/share.svg?branch=master)](https://travis-ci.org/asdfree/share) [![Build status](https://ci.appveyor.com/api/projects/status/github/asdfree/share?svg=TRUE)](https://ci.appveyor.com/project/ajdamico/share)

The Survey of Health, Ageing and Retirement in Europe interviews senior citizens across the continent for their entire life. Allows for findings like, "Among Belgians who were 50-74 years old in 2004, X% lived in nursing homes by 2010."

* Many tables, most with one row per sampled respondent for the period.

* A complex sample longitudinal survey designed to generalize to the civilian, non-institutionalized population of participating European countries aged 50 or older.

* Released every two or three years since 2004.

* Coordinated at the [Max Planck Institute](http://www.share-project.org/organisation/coordination.html) and [funded by consortium](http://www.share-project.org/organisation/funding.html).

## Simplified Download and Importation {-}

The R `lodown` package easily downloads and imports all available SHARE microdata by simply specifying `"share"` with an `output_dir =` parameter in the `lodown()` function. Depending on your internet connection and computer processing speed, you might prefer to run this step overnight.

```{r eval = FALSE }
library(lodown)
lodown( "share" , output_dir = file.path( path.expand( "~" ) , "SHARE" ) , 
	your_username = "username" , 
	your_password = "password" )
```

`lodown` also provides a catalog of available microdata extracts with the `get_catalog()` function. After requesting the SHARE catalog, you could pass a subsetted catalog through the `lodown()` function in order to download and import specific extracts (rather than all available extracts).

```{r eval = FALSE , results = "hide" }
library(lodown)
# examine all available SHARE microdata files
share_cat <-
	get_catalog( "share" ,
		output_dir = file.path( path.expand( "~" ) , "SHARE" ) , 
		your_username = "username" , 
		your_password = "password" )

# wave 1, wave 6, and longitudinal weights only
share_cat <- subset( share_cat , grepl( "ave 1|ave 6|ongitudinal" , output_folder ) )
# download the microdata to your local computer
share_cat <- lodown( "share" , share_cat , 
	your_username = "username" , 
	your_password = "password" )
```

## Analysis Examples with the `survey` library \ {-}

Construct a complex sample survey design:

```{r eval = FALSE }

```

```{r eval = FALSE }
options( survey.lonely.psu = "adjust" )

library(survey)

available_files <-
	list.files( 
		file.path( path.expand( "~" ) , "SHARE" ) , 
		recursive = TRUE , 
		full.names = TRUE 
	)

# wave six demographics file
share_dn6_df <-
	readRDS( grep( "6\\.0\\.0(.*)sharew6(.*)dn\\.rds" , available_files , value = TRUE ) )

share_dn6_df <-
	share_dn6_df[ c( "mergeid" , "country" , "dn042_" , "dn004_" ) ]
	
# wave six physical health file
share_ph1_df <-
	readRDS( grep( "sharew1(.*)ph\\.rds" , available_files , value = TRUE ) )

share_ph1_df$weight_in_2004 <-
		ifelse( share_ph1_df$ph012_ < 0 , NA , share_ph1_df$ph012_ )
		
share_ph1_df <-
	share_ph1_df[ c( "mergeid" , "weight_in_2004" , "ph005_" ) ]
	
# wave six physical health file
share_ph6_df <-
	readRDS( grep( "6\\.0\\.0(.*)sharew6(.*)ph\\.rds" , available_files , value = TRUE ) )

share_ph6_df$weight_in_2015 <-
		ifelse( share_ph6_df$ph012_ < 0 , NA , share_ph6_df$ph012_ )
		
share_ph6_df <-
	share_ph6_df[ c( "mergeid" , "weight_in_2015" , "ph003_" ) ]
	

# longitudinal weights file
share_longwt_df <-
	readRDS( grep( "longitudinal_weights_w1\\-(.*)\\.rds" , available_files , value = TRUE ) )

# france only longitudinal weights
france_df <- subset( share_longwt_df , country == 17 & ( cliw_a > 0 ) )

nrow_check <- nrow( france_df )

# merge on each of the tables
france_df <- merge( france_df , share_dn6_df )
france_df <- merge( france_df , share_ph1_df )
france_df <- merge( france_df , share_ph6_df )

# confirm no change in records
stopifnot( nrow( france_df ) == nrow_check )

share_design <- 
	svydesign( 
		~ psu + ssu , 
		strata = ~ stratum1 + stratum2 , 
		data = france_df , 
		weights = ~ cliw_a , 
		nest = TRUE 
	)
```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE }
share_design <- 
	update( 
		share_design , 
		
		one = 1 ,
		
		sexe = factor( dn042_ , levels = 1:2 , labels = c( 'masculin' , 'feminin' ) ) ,
		
		health_in_general_2015 =
			factor( ph003_ , levels = 1:5 , labels =
				c( "excellente" , "tres bonne" , "bonne" , "acceptable" , "mediocre" )
			) ,
			
		fortemente_limite_2004 = ifelse( ph005_ %in% 1:3 , as.numeric( ph005_ == 1 ) , NA )

	)
```

### Unweighted Counts {-}

Count the unweighted number of records in the survey sample, overall and by groups:
```{r eval = FALSE , results = "hide" }
sum( weights( share_design , "sampling" ) != 0 )

svyby( ~ one , ~ sexe , share_design , unwtd.count )
```

### Weighted Counts {-}
Count the weighted size of the generalizable population, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ one , share_design )

svyby( ~ one , ~ sexe , share_design , svytotal )
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ weight_in_2015 , share_design , na.rm = TRUE )

svyby( ~ weight_in_2015 , ~ sexe , share_design , svymean , na.rm = TRUE )
```

Calculate the distribution of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ health_in_general_2015 , share_design , na.rm = TRUE )

svyby( ~ health_in_general_2015 , ~ sexe , share_design , svymean , na.rm = TRUE )
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ weight_in_2015 , share_design , na.rm = TRUE )

svyby( ~ weight_in_2015 , ~ sexe , share_design , svytotal , na.rm = TRUE )
```

Calculate the weighted sum of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ health_in_general_2015 , share_design , na.rm = TRUE )

svyby( ~ health_in_general_2015 , ~ sexe , share_design , svytotal , na.rm = TRUE )
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svyquantile( ~ weight_in_2015 , share_design , 0.5 , na.rm = TRUE )

svyby( 
	~ weight_in_2015 , 
	~ sexe , 
	share_design , 
	svyquantile , 
	0.5 ,
	ci = TRUE ,
	keep.var = TRUE ,
	na.rm = TRUE
)
```

Estimate a ratio:
```{r eval = FALSE , results = "hide" }
svyratio( 
	numerator = ~ weight_in_2015 , 
	denominator = ~ weight_in_2004 , 
	share_design ,
	na.rm = TRUE
)
```

### Subsetting {-}

Restrict the survey design to persons born in france:
```{r eval = FALSE , results = "hide" }
sub_share_design <- subset( share_design , dn004_ == 1 )
```
Calculate the mean (average) of this subset:
```{r eval = FALSE , results = "hide" }
svymean( ~ weight_in_2015 , sub_share_design , na.rm = TRUE )
```

### Measures of Uncertainty {-}

Extract the coefficient, standard error, confidence interval, and coefficient of variation from any descriptive statistics function result, overall and by groups:
```{r eval = FALSE , results = "hide" }
this_result <- svymean( ~ weight_in_2015 , share_design , na.rm = TRUE )

coef( this_result )
SE( this_result )
confint( this_result )
cv( this_result )

grouped_result <-
	svyby( 
		~ weight_in_2015 , 
		~ sexe , 
		share_design , 
		svymean ,
		na.rm = TRUE 
	)
	
coef( grouped_result )
SE( grouped_result )
confint( grouped_result )
cv( grouped_result )
```

Calculate the degrees of freedom of any survey design object:
```{r eval = FALSE , results = "hide" }
degf( share_design )
```

Calculate the complex sample survey-adjusted variance of any statistic:
```{r eval = FALSE , results = "hide" }
svyvar( ~ weight_in_2015 , share_design , na.rm = TRUE )
```

Include the complex sample design effect in the result for a specific statistic:
```{r eval = FALSE , results = "hide" }
# SRS without replacement
svymean( ~ weight_in_2015 , share_design , na.rm = TRUE , deff = TRUE )

# SRS with replacement
svymean( ~ weight_in_2015 , share_design , na.rm = TRUE , deff = "replace" )
```

Compute confidence intervals for proportions using methods that may be more accurate near 0 and 1. See `?svyciprop` for alternatives:
```{r eval = FALSE , results = "hide" }
svyciprop( ~ fortemente_limite_2004 , share_design ,
	method = "likelihood" , na.rm = TRUE )
```

### Regression Models and Tests of Association {-}

Perform a design-based t-test:
```{r eval = FALSE , results = "hide" }
svyttest( weight_in_2015 ~ fortemente_limite_2004 , share_design )
```

Perform a chi-squared test of association for survey data:
```{r eval = FALSE , results = "hide" }
svychisq( 
	~ fortemente_limite_2004 + health_in_general_2015 , 
	share_design 
)
```

Perform a survey-weighted generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	svyglm( 
		weight_in_2015 ~ fortemente_limite_2004 + health_in_general_2015 , 
		share_design 
	)

summary( glm_result )
```

## Analysis Examples with `srvyr` \ {-}

The R `srvyr` library calculates summary statistics from survey data, such as the mean, total or quantile using [dplyr](https://github.com/tidyverse/dplyr/)-like syntax. [srvyr](https://github.com/gergness/srvyr) allows for the use of many verbs, such as `summarize`, `group_by`, and `mutate`, the convenience of pipe-able functions, the `tidyverse` style of non-standard evaluation and more consistent return types than the `survey` package. [This vignette](https://cran.r-project.org/web/packages/srvyr/vignettes/srvyr-vs-survey.html) details the available features. As a starting point for SHARE users, this code replicates previously-presented examples:

```{r eval = FALSE , results = "hide" }
library(srvyr)
share_srvyr_design <- as_survey( share_design )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
share_srvyr_design %>%
	summarize( mean = survey_mean( weight_in_2015 , na.rm = TRUE ) )

share_srvyr_design %>%
	group_by( sexe ) %>%
	summarize( mean = survey_mean( weight_in_2015 , na.rm = TRUE ) )
```

---

## Replication Example {-}

```{r eval = FALSE , results = "hide" }

```

