# Surveillance Epidemiology and End Results (SEER) {-}

[![Build Status](https://travis-ci.org/asdfree/seer.svg?branch=master)](https://travis-ci.org/asdfree/seer) [![Build status](https://ci.appveyor.com/api/projects/status/github/asdfree/seer?svg=TRUE)](https://ci.appveyor.com/project/ajdamico/seer)

The Surveillance Epidemiology and End Results (SEER) aggregates person-level information for more than a quarter of cancer incidence in the United States.

* A series of both individual- and population-level tables, grouped by site of cancer diagnosis.

* A registry covering various geographies across the US population, standardized by SEER*Stat to produce nationally-representative estimates.

* Updated every spring based on the previous November's submission of data.

* Maintained by the United States [National Cancer Institute (NCI)](http://www.cancer.gov/)

## Simplified Download and Importation {-}

The R `lodown` package easily downloads and imports all available SEER microdata by simply specifying `"seer"` with an `output_dir =` parameter in the `lodown()` function. Depending on your internet connection and computer processing speed, you might prefer to run this step overnight.

```{r eval = FALSE }
library(lodown)
lodown( "seer" , output_dir = file.path( path.expand( "~" ) , "SEER" ) , 
	your_username = "username" , 
	your_password = "password" )
```

## Analysis Examples with SQL and `RSQLite` \ {-}

Connect to a database:

```{r eval = FALSE }
library(DBI)
seer_df <- 
	file.path( 
		path.expand( "~" ) , "SEER" , 
		"incidence/yr1973_2014.seer9/LYMYLEUK.rds" 
	)
```

```{r eval = FALSE }

```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE }
seer_df <- 
	transform( 
		seer_df , 
		
		survival_months = ifelse( srv_time_mon == 9999 , NA , as.numeric( srv_time_mon ) ) ,
		
		female = as.numeric( sex == 2 ) ,
		
		race_ethnicity =
			ifelse( race1v == 99 , "unknown" ,
			ifelse( nhiade > 0 , "hispanic" , 
			ifelse( race1v == 1 , "white non-hispanic" ,
			ifelse( race1v == 2 , "black non-hispanic" , 
				"other non-hispanic" ) ) ) ) ,
		
		marital_status_at_dx =
			factor( 
				as.numeric( mar_stat ) , 
				levels = c( 1:6 , 9 ) ,
				labels =
					c(
						"single (never married)" ,
						"married" ,
						"separated" ,
						"divorced" ,
						"widowed" ,
						"unmarried or domestic partner or unregistered" ,
						"unknown"
					)
			)
	)
	
```

### Unweighted Counts {-}

Count the unweighted number of records in the SQL table, overall and by groups:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db , "SELECT COUNT(*) FROM " )

dbGetQuery( db ,
	"SELECT
		race_ethnicity ,
		COUNT(*) 
	FROM 
	GROUP BY race_ethnicity"
)
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db , "SELECT AVG( survival_months ) FROM " )

dbGetQuery( db , 
	"SELECT 
		race_ethnicity , 
		AVG( survival_months ) AS mean_survival_months
	FROM 
	GROUP BY race_ethnicity" 
)
```

Calculate the distribution of a categorical variable:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db , 
	"SELECT 
		marital_status_at_dx , 
		COUNT(*) / ( SELECT COUNT(*) FROM ) 
			AS share_marital_status_at_dx
	FROM 
	GROUP BY marital_status_at_dx" 
)
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db , "SELECT SUM( survival_months ) FROM " )

dbGetQuery( db , 
	"SELECT 
		race_ethnicity , 
		SUM( survival_months ) AS sum_survival_months 
	FROM 
	GROUP BY race_ethnicity" 
)
```

Calculate the 25th, median, and 75th percentiles of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
RSQLite::initExtension( db )

dbGetQuery( db , 
	"SELECT 
		LOWER_QUARTILE( survival_months ) , 
		MEDIAN( survival_months ) , 
		UPPER_QUARTILE( survival_months ) 
	FROM " 
)

dbGetQuery( db , 
	"SELECT 
		race_ethnicity , 
		LOWER_QUARTILE( survival_months ) AS lower_quartile_survival_months , 
		MEDIAN( survival_months ) AS median_survival_months , 
		UPPER_QUARTILE( survival_months ) AS upper_quartile_survival_months
	FROM 
	GROUP BY race_ethnicity" 
)
```

### Subsetting {-}

Limit your SQL analysis to inpatient hospital reporting source with `WHERE`:
```{r eval = FALSE , results = "hide" }
dbGetQuery( db ,
	"SELECT
		AVG( survival_months )
	FROM 
	WHERE rept_src == 1"
)
```

### Measures of Uncertainty {-}

Calculate the variance and standard deviation, overall and by groups:
```{r eval = FALSE , results = "hide" }
RSQLite::initExtension( db )

dbGetQuery( db , 
	"SELECT 
		VARIANCE( survival_months ) , 
		STDEV( survival_months ) 
	FROM " 
)

dbGetQuery( db , 
	"SELECT 
		race_ethnicity , 
		VARIANCE( survival_months ) AS var_survival_months ,
		STDEV( survival_months ) AS stddev_survival_months
	FROM 
	GROUP BY race_ethnicity" 
)
```

### Regression Models and Tests of Association {-}

Perform a t-test:
```{r eval = FALSE , results = "hide" }
seer_slim_df <- 
	dbGetQuery( db , 
		"SELECT 
			survival_months , 
			female ,
			marital_status_at_dx
		FROM " 
	)

t.test( survival_months ~ female , seer_slim_df )
```

Perform a chi-squared test of association:
```{r eval = FALSE , results = "hide" }
this_table <-
	table( seer_slim_df[ , c( "female" , "marital_status_at_dx" ) ] )

chisq.test( this_table )
```

Perform a generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	glm( 
		survival_months ~ female + marital_status_at_dx , 
		data = seer_slim_df
	)

summary( glm_result )
```

## Analysis Examples with `dplyr` \ {-}

The R `dplyr` library offers an alternative grammar of data manipulation to base R and SQL syntax. [dplyr](https://github.com/tidyverse/dplyr/) offers many verbs, such as `summarize`, `group_by`, and `mutate`, the convenience of pipe-able functions, and the `tidyverse` style of non-standard evaluation. [This vignette](https://cran.r-project.org/web/packages/dplyr/vignettes/dplyr.html) details the available features. As a starting point for SEER users, this code replicates previously-presented examples:

```{r eval = FALSE , results = "hide" }
library(dplyr)
library(dbplyr)
dplyr_db <- dplyr::src_sqlite( dbdir )
seer_tbl <- tbl( dplyr_db , '' )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
seer_tbl %>%
	summarize( mean = mean( survival_months ) )

seer_tbl %>%
	group_by( race_ethnicity ) %>%
	summarize( mean = mean( survival_months ) )
```


