# Trends in International Mathematics and Science Study (TIMSS) {-}

<a href="https://github.com/asdfree/timss/actions"><img src="https://github.com/asdfree/timss/actions/workflows/r.yml/badge.svg" alt="Github Actions Badge"></a>

The Trends in International Mathematics and Science Study (TIMSS) tracks the math and science competency of fourth graders across about fifty nations.

* A series of tables with one record per school (ACG), per student (ASG), per teacher (ATG), as well as files containing student achievement (ASA), home background (ASH), student-teacher linkage (AST), and within-country scoring reliability (ASR).

* A complex sample survey designed to generalize to the fourth-grade student population of participating countries.

* Released quadrennially since 1995.

* Funded by the [International Association for the Evaluation of Educational Achievement](https://www.iea.nl/) and compiled by the [Lynch School of Education at Boston College](http://www.bc.edu/bc-web/schools/lsoe.html).

## Simplified Download and Importation {-}

The R `lodown` package easily downloads and imports all available TIMSS microdata by simply specifying `"timss"` with an `output_dir =` parameter in the `lodown()` function. Depending on your internet connection and computer processing speed, you might prefer to run this step overnight.

```{r eval = FALSE }
library(lodown)
lodown( "timss" , output_dir = file.path( path.expand( "~" ) , "TIMSS" ) )
```

`lodown` also provides a catalog of available microdata extracts with the `get_catalog()` function. After requesting the TIMSS catalog, you could pass a subsetted catalog through the `lodown()` function in order to download and import specific extracts (rather than all available extracts).

```{r eval = FALSE , results = "hide" }
library(lodown)
# examine all available TIMSS microdata files
timss_cat <-
	get_catalog( "timss" ,
		output_dir = file.path( path.expand( "~" ) , "TIMSS" ) )

# 2015 only
timss_cat <- subset( timss_cat , year == 2015 )
# download the microdata to your local computer
timss_cat <- lodown( "timss" , timss_cat )
```

## Analysis Examples with the `survey` library \ {-}

Construct a multiply-imputed, complex sample survey design:

```{r eval = FALSE }

```

```{r eval = FALSE }
library(survey)
library(mitools)
library(RSQLite)

# load the ASG (student background) + ASH (home background) merged design
timss_design <- readRDS( file.path( path.expand( "~" ) , "TIMSS" , "2015/asg_design.rds" ) )

design_weights <- readRDS( file.path( path.expand( "~" ) , "TIMSS" , "2015/asg_weights.rds" ) )

five_tablenames <- paste0( "asg_2015_" , 1:5 )

timss_design <- lodown:::svyMDBdesign( timss_design )
```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE }
timss_design <- 
	update( 
		timss_design , 
		
		one = 1 ,
		
		idcntry = factor( idcntry ) ,
		
		sex = factor( itsex , labels = c( "male" , "female" ) ) ,
		
		born_2005_or_later = as.numeric( itbirthy >= 2005 )

	)
```

### Unweighted Counts {-}

Count the unweighted number of records in the survey sample, overall and by groups:
```{r eval = FALSE , results = "hide" }
MIcombine( with( timss_design , svyby( ~ one , ~ one , unwtd.count ) ) )

MIcombine( with( timss_design , svyby( ~ one , ~ sex , unwtd.count ) ) )
```

### Weighted Counts {-}
Count the weighted size of the generalizable population, overall and by groups:
```{r eval = FALSE , results = "hide" }
MIcombine( with( timss_design , svytotal( ~ one ) ) )

MIcombine( with( timss_design ,
	svyby( ~ one , ~ sex , svytotal )
) )
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
MIcombine( with( timss_design , svymean( ~ asmmat ) ) )

MIcombine( with( timss_design ,
	svyby( ~ asmmat , ~ sex , svymean )
) )
```

Calculate the distribution of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
MIcombine( with( timss_design , svymean( ~ idcntry ) ) )

MIcombine( with( timss_design ,
	svyby( ~ idcntry , ~ sex , svymean )
) )
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
MIcombine( with( timss_design , svytotal( ~ asmmat ) ) )

MIcombine( with( timss_design ,
	svyby( ~ asmmat , ~ sex , svytotal )
) )
```

Calculate the weighted sum of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
MIcombine( with( timss_design , svytotal( ~ idcntry ) ) )

MIcombine( with( timss_design ,
	svyby( ~ idcntry , ~ sex , svytotal )
) )
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
MIcombine( with( timss_design ,
	svyquantile(
		~ asmmat ,
		0.5 , se = TRUE 
) ) )

MIcombine( with( timss_design ,
	svyby(
		~ asmmat , ~ sex , svyquantile ,
		0.5 , se = TRUE ,
		keep.var = TRUE , ci = TRUE 
) ) )
```

Estimate a ratio:
```{r eval = FALSE , results = "hide" }
MIcombine( with( timss_design ,
	svyratio( numerator = ~ asssci , denominator = ~ asmmat )
) )
```

### Subsetting {-}

Restrict the survey design to Australia, Austria, Azerbaijan, Belgium (French):
```{r eval = FALSE , results = "hide" }
sub_timss_design <- subset( timss_design , idcntry %in% c( 36 , 40 , 31 , 957 ) )
```
Calculate the mean (average) of this subset:
```{r eval = FALSE , results = "hide" }
MIcombine( with( sub_timss_design , svymean( ~ asmmat ) ) )
```

### Measures of Uncertainty {-}

Extract the coefficient, standard error, confidence interval, and coefficient of variation from any descriptive statistics function result, overall and by groups:
```{r eval = FALSE , results = "hide" }
this_result <-
	MIcombine( with( timss_design ,
		svymean( ~ asmmat )
	) )

coef( this_result )
SE( this_result )
confint( this_result )
cv( this_result )

grouped_result <-
	MIcombine( with( timss_design ,
		svyby( ~ asmmat , ~ sex , svymean )
	) )

coef( grouped_result )
SE( grouped_result )
confint( grouped_result )
cv( grouped_result )
```

Calculate the degrees of freedom of any survey design object:
```{r eval = FALSE , results = "hide" }
degf( timss_design$designs[[1]] )
```

Calculate the complex sample survey-adjusted variance of any statistic:
```{r eval = FALSE , results = "hide" }
MIcombine( with( timss_design , svyvar( ~ asmmat ) ) )
```

Include the complex sample design effect in the result for a specific statistic:
```{r eval = FALSE , results = "hide" }
# SRS without replacement
MIcombine( with( timss_design ,
	svymean( ~ asmmat , deff = TRUE )
) )

# SRS with replacement
MIcombine( with( timss_design ,
	svymean( ~ asmmat , deff = "replace" )
) )
```

Compute confidence intervals for proportions using methods that may be more accurate near 0 and 1. See `?svyciprop` for alternatives:
```{r eval = FALSE , results = "hide" }
MIsvyciprop( ~ born_2001_or_later , timss_design ,
	method = "likelihood" , na.rm = TRUE )
```

### Regression Models and Tests of Association {-}

Perform a design-based t-test:
```{r eval = FALSE , results = "hide" }
MIsvyttest( asmmat ~ born_2001_or_later , timss_design )
```

Perform a chi-squared test of association for survey data:
```{r eval = FALSE , results = "hide" }
MIsvychisq( ~ born_2001_or_later + idcntry , timss_design )
```

Perform a survey-weighted generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	MIcombine( with( timss_design ,
		svyglm( asmmat ~ born_2001_or_later + idcntry )
	) )
	
summary( glm_result )
```

---

## Replication Example {-}

```{r eval = FALSE , results = "hide" }

```
