# World Values Survey (WVS) {-}

[![Build Status](https://travis-ci.org/asdfree/wvs.svg?branch=master)](https://travis-ci.org/asdfree/wvs) [![Build status](https://ci.appveyor.com/api/projects/status/github/asdfree/wvs?svg=TRUE)](https://ci.appveyor.com/project/ajdamico/wvs)

The World Values Survey studies changing values and their impact on social and political life in almost one hundred nations.

* One table per country per wave, with one row per sampled respondent.

* A complex sample survey designed to generalize the population aged eighteen and older in participating countries.

* Released about twice per decade since 1981.

* Administered as a confederacy, guided by a [scientific advisory committee](http://www.worldvaluessurvey.org/WVSContents.jsp?CMSID=SAC) and funded by [consortium](http://www.worldvaluessurvey.org/WVSContents.jsp).

## Simplified Download and Importation {-}

The R `lodown` package easily downloads and imports all available WVS microdata by simply specifying `"wvs"` with an `output_dir =` parameter in the `lodown()` function. Depending on your internet connection and computer processing speed, you might prefer to run this step overnight.

```{r eval = FALSE }
library(lodown)
lodown( "wvs" , output_dir = file.path( path.expand( "~" ) , "WVS" ) )
```

`lodown` also provides a catalog of available microdata extracts with the `get_catalog()` function. After requesting the WVS catalog, you could pass a subsetted catalog through the `lodown()` function in order to download and import specific extracts (rather than all available extracts).

```{r eval = FALSE , results = "hide" }
library(lodown)
# examine all available WVS microdata files
wvs_cat <-
	get_catalog( "wvs" ,
		output_dir = file.path( path.expand( "~" ) , "WVS" ) )

# wave six only
wvs_cat <- subset( wvs_cat , grepl( "United(.*)States" , full_url ) & wave == 6 )
# download the microdata to your local computer
lodown( "wvs" , wvs_cat )
```

## Analysis Examples with the `survey` library {-}

Construct a complex sample survey design:

```{r eval = FALSE }

```

```{r eval = FALSE }
library(survey)

wvs_df <-
	readRDS( 
		file.path( path.expand( "~" ) , "WVS" , 
			"wave 6/F00003106-WV6_Data_United_States_2011_spss_v_2016-01-01.rds" ) 
	)

# construct a fake survey design
warning( "this survey design produces correct point estimates
but incorrect standard errors." )
wvs_design <- 
	svydesign( 
		~ 1 , 
		data = wvs_df , 
		weights = ~ v258
	)
```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE }
wvs_design <- 
	update( 
		wvs_design , 
		
		one = 1 ,
		
		language_spoken_at_home =
			factor( v247 , 
				levels = c( 101 , 128 , 144 , 208 , 426 , 800 ) , 
				labels = c( 'chinese' , 'english' , 'french' , 
					'japanese' , 'spanish; castilian' , 'other' )
			) ,

		citizen = as.numeric( v246 == 1 ) ,
		
		task_creativity_1_10 = as.numeric( v232 ) ,
		
		work_independence_1_10 = as.numeric( v233 ) ,
		
		family_importance =
			factor( v4 , 
				labels = c( 'very' , 'rather' , 'not very' , 'not at all' ) 
			)
	)
```

### Unweighted Counts {-}

Count the unweighted number of records in the survey sample, overall and by groups:
```{r eval = FALSE , results = "hide" }
sum( weights( wvs_design , "sampling" ) != 0 )

svyby( ~ one , ~ language_spoken_at_home , wvs_design , unwtd.count )
```

### Weighted Counts {-}
Count the weighted size of the generalizable population, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ one , wvs_design )

svyby( ~ one , ~ language_spoken_at_home , wvs_design , svytotal )
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ task_creativity_1_10 , wvs_design , na.rm = TRUE )

svyby( ~ task_creativity_1_10 , ~ language_spoken_at_home , wvs_design , svymean , na.rm = TRUE )
```

Calculate the distribution of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ family_importance , wvs_design , na.rm = TRUE )

svyby( ~ family_importance , ~ language_spoken_at_home , wvs_design , svymean , na.rm = TRUE )
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ task_creativity_1_10 , wvs_design , na.rm = TRUE )

svyby( ~ task_creativity_1_10 , ~ language_spoken_at_home , wvs_design , svytotal , na.rm = TRUE )
```

Calculate the weighted sum of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ family_importance , wvs_design , na.rm = TRUE )

svyby( ~ family_importance , ~ language_spoken_at_home , wvs_design , svytotal , na.rm = TRUE )
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svyquantile( ~ task_creativity_1_10 , wvs_design , 0.5 , na.rm = TRUE )

svyby( 
	~ task_creativity_1_10 , 
	~ language_spoken_at_home , 
	wvs_design , 
	svyquantile , 
	0.5 ,
	ci = TRUE ,
	keep.var = TRUE ,
	na.rm = TRUE
)
```

Estimate a ratio:
```{r eval = FALSE , results = "hide" }
svyratio( 
	numerator = ~ task_creativity_1_10 , 
	denominator = ~ work_independence_1_10 , 
	wvs_design ,
	na.rm = TRUE
)
```

### Subsetting {-}

Restrict the survey design to seniors:
```{r eval = FALSE , results = "hide" }
sub_wvs_design <- subset( wvs_design , v242 >= 65 )
```
Calculate the mean (average) of this subset:
```{r eval = FALSE , results = "hide" }
svymean( ~ task_creativity_1_10 , sub_wvs_design , na.rm = TRUE )
```

### Measures of Uncertainty {-}

Extract the coefficient, standard error, confidence interval, and coefficient of variation from any descriptive statistics function result, overall and by groups:
```{r eval = FALSE , results = "hide" }
this_result <- svymean( ~ task_creativity_1_10 , wvs_design , na.rm = TRUE )

coef( this_result )
SE( this_result )
confint( this_result )
cv( this_result )

grouped_result <-
	svyby( 
		~ task_creativity_1_10 , 
		~ language_spoken_at_home , 
		wvs_design , 
		svymean ,
		na.rm = TRUE 
	)
	
coef( grouped_result )
SE( grouped_result )
confint( grouped_result )
cv( grouped_result )
```

Calculate the degrees of freedom of any survey design object:
```{r eval = FALSE , results = "hide" }
degf( wvs_design )
```

Calculate the complex sample survey-adjusted variance of any statistic:
```{r eval = FALSE , results = "hide" }
svyvar( ~ task_creativity_1_10 , wvs_design , na.rm = TRUE )
```

Include the complex sample design effect in the result for a specific statistic:
```{r eval = FALSE , results = "hide" }
# SRS without replacement
svymean( ~ task_creativity_1_10 , wvs_design , na.rm = TRUE , deff = TRUE )

# SRS with replacement
svymean( ~ task_creativity_1_10 , wvs_design , na.rm = TRUE , deff = "replace" )
```

Compute confidence intervals for proportions using methods that may be more accurate near 0 and 1. See `?svyciprop` for alternatives:
```{r eval = FALSE , results = "hide" }
svyciprop( ~ citizen , wvs_design ,
	method = "likelihood" , na.rm = TRUE )
```

### Regression Models and Tests of Association {-}

Perform a design-based t-test:
```{r eval = FALSE , results = "hide" }
svyttest( task_creativity_1_10 ~ citizen , wvs_design )
```

Perform a chi-squared test of association for survey data:
```{r eval = FALSE , results = "hide" }
svychisq( 
	~ citizen + family_importance , 
	wvs_design 
)
```

Perform a survey-weighted generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	svyglm( 
		task_creativity_1_10 ~ citizen + family_importance , 
		wvs_design 
	)

summary( glm_result )
```

## Analysis Examples with `srvyr` {-}

The R `srvyr` library calculates summary statistics from survey data, such as the mean, total or quantile using [dplyr](https://github.com/tidyverse/dplyr/)-like syntax. [srvyr](https://github.com/gergness/srvyr) allows for the use of many verbs, such as `summarize`, `group_by`, and `mutate`, the convenience of pipe-able functions, the `tidyverse` style of non-standard evaluation and more consistent return types than the `survey` package. [This vignette](https://cran.r-project.org/web/packages/srvyr/vignettes/srvyr-vs-survey.html) details the available features. As a starting point for WVS users, this code replicates previously-presented examples:

```{r eval = FALSE , results = "hide" }
library(srvyr)
wvs_srvyr_design <- as_survey( wvs_design )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
wvs_srvyr_design %>%
	summarize( mean = survey_mean( task_creativity_1_10 , na.rm = TRUE ) )

wvs_srvyr_design %>%
	group_by( language_spoken_at_home ) %>%
	summarize( mean = survey_mean( task_creativity_1_10 , na.rm = TRUE ) )
```

---

## Replication Example {-}

```{r eval = FALSE , results = "hide" }

```

