# Area Health Resources Files (AHRF) {-}

[![License: GPL v3](https://img.shields.io/badge/License-GPLv3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0) <a href="https://github.com/asdfree/ahrf/actions"><img src="https://github.com/asdfree/ahrf/actions/workflows/r.yml/badge.svg" alt="Github Actions Badge"></a> <a href="https://www.eventbrite.com/e/1458850233319"><img src="https://img.shields.io/badge/Free%20Workshop%20New%20York%20NY-October%2022nd,%202025-orange" alt="New York NY: October 22nd, 2025"></a> <a href="https://www.eventbrite.com/e/1560552678069"><img src="https://img.shields.io/badge/Free%20Workshop%20Philadelphia%20PA-October%2029th,%202025-orange" alt="Philadelphia PA: October 29th, 2025"></a> <a href="https://www.eventbrite.com/e/1710882017169"><img src="https://img.shields.io/badge/Free%20Workshop%20Washington%20DC-November%2012th,%202025-orange" alt="Washington DC: November 12th, 2025"></a> <a href="https://www.eventbrite.com/e/1771826874879"><img src="https://img.shields.io/badge/Free%20R%20Training%20Washington%20DC-November%2019th,%202025-yellow" alt="Washington DC: November 19th, 2025"></a>

National, state, and county-level data on health care professions, health facilities, population characteristics, health workforce training, hospital utilization and expenditure, and the environment.

* One table with one row per county and a second table with one row per state.

* Replaced annually with the latest available county- and state-level statistics.

* Compiled by the [Bureau of Health Workforce](https://bhw.hrsa.gov/) at the [Health Services and Resources Administration](http://www.hrsa.gov/).

---

## Recommended Reading {-}

Two Methodology Documents:

> [User Documentation for the County Area Health Resources File (AHRF) 2021-2022 Release](https://data.hrsa.gov/DataDownload/AHRF/AHRF%202021-2022_User_Tech.zip)

> [Frequently Asked Questions](https://data.hrsa.gov/faq)

<br>

One Haiku:

```{r}
# local aggregates
# to spread merge join spline regress
# like fresh buttered bread
```

---

## Download, Import, Preparation {-}

Download and import the most current county-level file:
```{r eval = FALSE , results = "hide" }
library(haven)

tf <- tempfile()

ahrf_url <- "https://data.hrsa.gov//DataDownload/AHRF/AHRF_2021-2022_SAS.zip"

download.file( ahrf_url , tf , mode = 'wb' )

unzipped_files <- unzip( tf , exdir = tempdir() )

sas_fn <- grep( "\\.sas7bdat$" , unzipped_files , value = TRUE )

ahrf_tbl <- read_sas( sas_fn )

ahrf_df <- data.frame( ahrf_tbl )

names( ahrf_df ) <- tolower( names( ahrf_df ) )

```

### Save Locally \ {-}

Save the object at any point:

```{r eval = FALSE , results = "hide" }
# ahrf_fn <- file.path( path.expand( "~" ) , "AHRF" , "this_file.rds" )
# saveRDS( ahrf_df , file = ahrf_fn , compress = FALSE )
```

Load the same object:

```{r eval = FALSE , results = "hide" }
# ahrf_df <- readRDS( ahrf_fn )
```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE , results = "hide" }
ahrf_df <- 
	transform( 
		ahrf_df , 
		
		cbsa_indicator_code = 
			factor( 
				as.numeric( f1406720 ) , 
				levels = 0:2 ,
				labels = c( "not metro" , "metro" , "micro" ) 
			) ,
			
		mhi_2020 = f1322620 ,
		
		whole_county_hpsa_2022 = as.numeric( f0978722 ) == 1 ,
		
		census_region = 
			factor( 
				as.numeric( f04439 ) , 
				levels = 1:4 ,
				labels = c( "northeast" , "midwest" , "south" , "west" ) 
			)

	)
	
```

---

## Analysis Examples with base R \ {-}

### Unweighted Counts {-}

Count the unweighted number of records in the table, overall and by groups:
```{r eval = FALSE , results = "hide" }
nrow( ahrf_df )

table( ahrf_df[ , "cbsa_indicator_code" ] , useNA = "always" )
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
mean( ahrf_df[ , "mhi_2020" ] , na.rm = TRUE )

tapply(
	ahrf_df[ , "mhi_2020" ] ,
	ahrf_df[ , "cbsa_indicator_code" ] ,
	mean ,
	na.rm = TRUE 
)
```

Calculate the distribution of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
prop.table( table( ahrf_df[ , "census_region" ] ) )

prop.table(
	table( ahrf_df[ , c( "census_region" , "cbsa_indicator_code" ) ] ) ,
	margin = 2
)
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
sum( ahrf_df[ , "mhi_2020" ] , na.rm = TRUE )

tapply(
	ahrf_df[ , "mhi_2020" ] ,
	ahrf_df[ , "cbsa_indicator_code" ] ,
	sum ,
	na.rm = TRUE 
)
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
quantile( ahrf_df[ , "mhi_2020" ] , 0.5 , na.rm = TRUE )

tapply(
	ahrf_df[ , "mhi_2020" ] ,
	ahrf_df[ , "cbsa_indicator_code" ] ,
	quantile ,
	0.5 ,
	na.rm = TRUE 
)
```

### Subsetting {-}

Limit your `data.frame` to California:
```{r eval = FALSE , results = "hide" }
sub_ahrf_df <- subset( ahrf_df , f12424 == "CA" )
```
Calculate the mean (average) of this subset:
```{r eval = FALSE , results = "hide" }
mean( sub_ahrf_df[ , "mhi_2020" ] , na.rm = TRUE )
```

### Measures of Uncertainty {-}

Calculate the variance, overall and by groups:
```{r eval = FALSE , results = "hide" }
var( ahrf_df[ , "mhi_2020" ] , na.rm = TRUE )

tapply(
	ahrf_df[ , "mhi_2020" ] ,
	ahrf_df[ , "cbsa_indicator_code" ] ,
	var ,
	na.rm = TRUE 
)
```

### Regression Models and Tests of Association {-}

Perform a t-test:
```{r eval = FALSE , results = "hide" }
t.test( mhi_2020 ~ whole_county_hpsa_2022 , ahrf_df )
```

Perform a chi-squared test of association:
```{r eval = FALSE , results = "hide" }
this_table <- table( ahrf_df[ , c( "whole_county_hpsa_2022" , "census_region" ) ] )

chisq.test( this_table )
```

Perform a generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	glm( 
		mhi_2020 ~ whole_county_hpsa_2022 + census_region , 
		data = ahrf_df
	)

summary( glm_result )
```

---

## Replication Example {-}

Match the record count in row number 8,543 of `AHRF 2021-2022 Technical Documentation.xlsx`:

```{r eval = FALSE , results = "hide" }
stopifnot( nrow( ahrf_df ) == 3232 )
```

---

## Analysis Examples with `dplyr` \ {-}

The R `dplyr` library offers an alternative grammar of data manipulation to base R and SQL syntax. [dplyr](https://github.com/tidyverse/dplyr/) offers many verbs, such as `summarize`, `group_by`, and `mutate`, the convenience of pipe-able functions, and the `tidyverse` style of non-standard evaluation. [This vignette](https://cran.r-project.org/web/packages/dplyr/vignettes/dplyr.html) details the available features. As a starting point for AHRF users, this code replicates previously-presented examples:

```{r eval = FALSE , results = "hide" }
library(dplyr)
ahrf_tbl <- as_tibble( ahrf_df )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
ahrf_tbl %>%
	summarize( mean = mean( mhi_2020 , na.rm = TRUE ) )

ahrf_tbl %>%
	group_by( cbsa_indicator_code ) %>%
	summarize( mean = mean( mhi_2020 , na.rm = TRUE ) )
```

---

## Analysis Examples with `data.table` \ {-}

The R `data.table` library provides a high-performance version of base R's data.frame with syntax and feature enhancements for ease of use, convenience and programming speed. [data.table](https://r-datatable.com) offers concise syntax: fast to type, fast to read, fast speed, memory efficiency, a careful API lifecycle management, an active community, and a rich set of features. [This vignette](https://cran.r-project.org/web/packages/data.table/vignettes/datatable-intro.html) details the available features. As a starting point for AHRF users, this code replicates previously-presented examples:

```{r eval = FALSE , results = 'hide' }
library(data.table)
ahrf_dt <- data.table( ahrf_df )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = 'hide' }
ahrf_dt[ , mean( mhi_2020 , na.rm = TRUE ) ]

ahrf_dt[ , mean( mhi_2020 , na.rm = TRUE ) , by = cbsa_indicator_code ]
```

---

## Analysis Examples with `duckdb` \ {-}

The R `duckdb` library provides an embedded analytical data management system with support for the Structured Query Language (SQL). [duckdb](https://duckdb.org) offers a simple, feature-rich, fast, and free SQL OLAP management system. [This vignette](https://duckdb.org/docs/api/r) details the available features. As a starting point for AHRF users, this code replicates previously-presented examples:

```{r eval = FALSE , results = 'hide' }
library(duckdb)
con <- dbConnect( duckdb::duckdb() , dbdir = 'my-db.duckdb' )
dbWriteTable( con , 'ahrf' , ahrf_df )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = 'hide' }
dbGetQuery( con , 'SELECT AVG( mhi_2020 ) FROM ahrf' )

dbGetQuery(
	con ,
	'SELECT
		cbsa_indicator_code ,
		AVG( mhi_2020 )
	FROM
		ahrf
	GROUP BY
		cbsa_indicator_code'
)
```
