# Current Population Survey - Annual Social and Economic Supplement (CPSASEC) {-}

<a href="https://github.com/asdfree/cpsasec/actions"><img src="https://github.com/asdfree/cpsasec/actions/workflows/r.yml/badge.svg" alt="Github Actions Badge"></a>

The March Current Population Survey Annual Social and Economic Supplement has supplied the statistics for the US Census Bureau's report on income, poverty, and health insurance coverage since 1948.

* One table with one row per sampled household, a second table with one row per family within each sampled household, and a third table with one row per individual within each of those families.

* A complex sample survey designed to generalize to the civilian non-institutional population of the United States

* Released annually since 1998.

* Administered jointly by the [US Census Bureau](http://www.census.gov/) and the [Bureau of Labor Statistics](http://www.bls.gov/).

---

---

## Download, Import, Preparation {-}

Download and import the Alabama household file:

(switch `sas_hal` to `sas_hak` for Alaska or `sas_hus` for the entire country)
```{r eval = FALSE , results = "hide" }
library(haven)

tf_household <- tempfile()

this_url_household <-
	"https://www2.census.gov/programs-surveys/acs/data/pums/2021/1-Year/sas_hal.zip"

download.file( this_url_household , tf_household , mode = 'wb' )

unzipped_files_household <- unzip( tf_household , exdir = tempdir() )

cpsasec_sas_household <-
	grep( '\\.sas7bdat$' , unzipped_files_household , value = TRUE )

cpsasec_df_household <- read_sas( cpsasec_sas_household )

names( cpsasec_df_household ) <- tolower( names( cpsasec_df_household ) )
```

Download and import the Alabama person file:

(switch `sas_pal` to `sas_pak` for Alaska or `sas_pus` for the entire country)
```{r eval = FALSE , results = "hide" }
tf_person <- tempfile()

this_url_person <-
	"https://www2.census.gov/programs-surveys/acs/data/pums/2021/1-Year/sas_pal.zip"

download.file( this_url_person , tf_person , mode = 'wb' )

unzipped_files_person <- unzip( tf_person , exdir = tempdir() )

cpsasec_sas_person <-
	grep( '\\.sas7bdat$' , unzipped_files_person , value = TRUE )

cpsasec_df_person <- read_sas( cpsasec_sas_person )

names( cpsasec_df_person ) <- tolower( names( cpsasec_df_person ) )
```

Remove overlapping column and merge household + person files:
```{r eval = FALSE , results = "hide" }

cpsasec_df_household[ , 'rt' ] <- NULL

cpsasec_df_person[ , 'rt' ] <- NULL

cpsasec_df <- merge( cpsasec_df_household , cpsasec_df_person )
	
stopifnot( nrow( cpsasec_df ) == nrow( cpsasec_df_person ) )

cpsasec_df[ , 'one' ] <- 1
```

### Save locally \ {-}

Save the object at any point:

```{r eval = FALSE , results = "hide" }
# cpsasec_fn <- file.path( path.expand( "~" ) , "CPSASEC" , "this_file.rds" )
# saveRDS( cpsasec_df , file = cpsasec_fn , compress = FALSE )
```

Load the same object:

```{r eval = FALSE , results = "hide" }
# cpsasec_df <- readRDS( cpsasec_fn )
```

### Survey Design Definition {-}
Construct a complex sample survey design:

library(survey)

cpsasec_df <- 
	readRDS( file.path( path.expand( "~" ) , "CPSASEC" , "2016 cps asec.rds" ) )

variables_to_keep <-
	c( 'a_maritl' , 'gestfips' , 'a_sex' , 'ptotval' , 'moop' , 'a_age' , 'htotval' , 
	'one' , 'a_exprrp' , 'marsupwt' , 
	grep( "pwwgt" , names( cpsasec_df ) , value = TRUE ) )
	
cpsasec_df <- cpsasec_df[ variables_to_keep ] ; gc()
	
cpsasec_design <-
	svrepdesign(
		weights = ~ marsupwt ,
		repweights = "pwwgt[1-9]" ,
		type = "Fay" ,
		rho = ( 1 - 1 / sqrt( 4 ) ) ,
		data = cpsasec_df ,
		combined.weights = TRUE ,
		mse = TRUE
	)

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE , results = "hide" }
cpsasec_design <- 
	update( 
		cpsasec_design , 

		a_maritl = 
			factor( 
				a_maritl , 
				labels = 
					c( 
						"married - civilian spouse present" ,
						"married - AF spouse present" ,
						"married - spouse absent" ,
						"widowed" ,
						"divorced" , 
						"separated" , 
						"never married"
					)
			) ,
			
		state_name =
			factor(
				gestfips ,
				levels = 
					c(1L, 2L, 4L, 5L, 6L, 8L, 9L, 10L, 
					11L, 12L, 13L, 15L, 16L, 17L, 18L, 
					19L, 20L, 21L, 22L, 23L, 24L, 25L, 
					26L, 27L, 28L, 29L, 30L, 31L, 32L, 
					33L, 34L, 35L, 36L, 37L, 38L, 39L, 
					40L, 41L, 42L, 44L, 45L, 46L, 47L, 
					48L, 49L, 50L, 51L, 53L, 54L, 55L, 
					56L) ,
				labels =
					c("Alabama", "Alaska", "Arizona", "Arkansas", "California", 
					"Colorado", "Connecticut", "Delaware", "District of Columbia", 
					"Florida", "Georgia", "Hawaii", "Idaho", "Illinois", "Indiana", 
					"Iowa", "Kansas", "Kentucky", "Louisiana", "Maine", "Maryland", 
					"Massachusetts", "Michigan", "Minnesota", "Mississippi", "Missouri", 
					"Montana", "Nebraska", "Nevada", "New Hampshire", "New Jersey", 
					"New Mexico", "New York", "North Carolina", "North Dakota", "Ohio", 
					"Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", "South Carolina", 
					"South Dakota", "Tennessee", "Texas", "Utah", "Vermont", "Virginia", 
					"Washington", "West Virginia", "Wisconsin", "Wyoming")
			) ,

		male = as.numeric( a_sex == 1 )
	)
```

---

## Analysis Examples with the `survey` library \ {-}

### Unweighted Counts {-}

Count the unweighted number of records in the survey sample, overall and by groups:
```{r eval = FALSE , results = "hide" }
sum( weights( cpsasec_design , "sampling" ) != 0 )

svyby( ~ one , ~ state_name , cpsasec_design , unwtd.count )
```

### Weighted Counts {-}
Count the weighted size of the generalizable population, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ one , cpsasec_design )

svyby( ~ one , ~ state_name , cpsasec_design , svytotal )
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ ptotval , cpsasec_design )

svyby( ~ ptotval , ~ state_name , cpsasec_design , svymean )
```

Calculate the distribution of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ a_maritl , cpsasec_design )

svyby( ~ a_maritl , ~ state_name , cpsasec_design , svymean )
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ ptotval , cpsasec_design )

svyby( ~ ptotval , ~ state_name , cpsasec_design , svytotal )
```

Calculate the weighted sum of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ a_maritl , cpsasec_design )

svyby( ~ a_maritl , ~ state_name , cpsasec_design , svytotal )
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svyquantile( ~ ptotval , cpsasec_design , 0.5 )

svyby( 
	~ ptotval , 
	~ state_name , 
	cpsasec_design , 
	svyquantile , 
	0.5 ,
	ci = TRUE 
)
```

Estimate a ratio:
```{r eval = FALSE , results = "hide" }
svyratio( 
	numerator = ~ moop , 
	denominator = ~ ptotval , 
	cpsasec_design 
)
```

### Subsetting {-}

Restrict the survey design to persons aged 18-64:
```{r eval = FALSE , results = "hide" }
sub_cpsasec_design <- subset( cpsasec_design , a_age %in% 18:64 )
```
Calculate the mean (average) of this subset:
```{r eval = FALSE , results = "hide" }
svymean( ~ ptotval , sub_cpsasec_design )
```

### Measures of Uncertainty {-}

Extract the coefficient, standard error, confidence interval, and coefficient of variation from any descriptive statistics function result, overall and by groups:
```{r eval = FALSE , results = "hide" }
this_result <- svymean( ~ ptotval , cpsasec_design )

coef( this_result )
SE( this_result )
confint( this_result )
cv( this_result )

grouped_result <-
	svyby( 
		~ ptotval , 
		~ state_name , 
		cpsasec_design , 
		svymean 
	)
	
coef( grouped_result )
SE( grouped_result )
confint( grouped_result )
cv( grouped_result )
```

Calculate the degrees of freedom of any survey design object:
```{r eval = FALSE , results = "hide" }
degf( cpsasec_design )
```

Calculate the complex sample survey-adjusted variance of any statistic:
```{r eval = FALSE , results = "hide" }
svyvar( ~ ptotval , cpsasec_design )
```

Include the complex sample design effect in the result for a specific statistic:
```{r eval = FALSE , results = "hide" }
# SRS without replacement
svymean( ~ ptotval , cpsasec_design , deff = TRUE )

# SRS with replacement
svymean( ~ ptotval , cpsasec_design , deff = "replace" )
```

Compute confidence intervals for proportions using methods that may be more accurate near 0 and 1. See `?svyciprop` for alternatives:
```{r eval = FALSE , results = "hide" }
svyciprop( ~ male , cpsasec_design ,
	method = "likelihood" )
```

### Regression Models and Tests of Association {-}

Perform a design-based t-test:
```{r eval = FALSE , results = "hide" }
svyttest( ptotval ~ male , cpsasec_design )
```

Perform a chi-squared test of association for survey data:
```{r eval = FALSE , results = "hide" }
svychisq( 
	~ male + a_maritl , 
	cpsasec_design 
)
```

Perform a survey-weighted generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	svyglm( 
		ptotval ~ male + a_maritl , 
		cpsasec_design 
	)

summary( glm_result )
```

---

## Intermish {-}

<center>https://en.wikipedia.org/wiki/Lorem_ipsum</center>

---

## Replication Example {-}

```{r eval = FALSE , results = "hide" }

```

---

## Poverty and Inequality Estimation with `convey` \ {-}

The R `convey` library estimates measures of income concentration, poverty, inequality, and wellbeing. [This textbook](https://guilhermejacob.github.io/context/) details the available features. As a starting point for CPSASEC users, this code calculates the gini coefficient on complex sample survey data:

```{r eval = FALSE , results = "hide" }
library(convey)
cpsasec_design <- convey_prep( cpsasec_design )

sub_cpsasec_design <- 
	subset( 
		cpsasec_design , 
		a_exprrp %in% 1:2
	)

svygini( ~ htotval , sub_cpsasec_design )
```

---

## Analysis Examples with `srvyr` \ {-}

The R `srvyr` library calculates summary statistics from survey data, such as the mean, total or quantile using [dplyr](https://github.com/tidyverse/dplyr/)-like syntax. [srvyr](https://github.com/gergness/srvyr) allows for the use of many verbs, such as `summarize`, `group_by`, and `mutate`, the convenience of pipe-able functions, the `tidyverse` style of non-standard evaluation and more consistent return types than the `survey` package. [This vignette](https://cran.r-project.org/web/packages/srvyr/vignettes/srvyr-vs-survey.html) details the available features. As a starting point for CPSASEC users, this code replicates previously-presented examples:

```{r eval = FALSE , results = "hide" }
library(srvyr)
cpsasec_srvyr_design <- as_survey( cpsasec_design )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
cpsasec_srvyr_design %>%
	summarize( mean = survey_mean( ptotval ) )

cpsasec_srvyr_design %>%
	group_by( state_name ) %>%
	summarize( mean = survey_mean( ptotval ) )
```


