# Exame Nacional de Desempenho de Estudantes (ENADE) {-}

[![License: GPL v3](https://img.shields.io/badge/License-GPLv3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0) <a href="https://github.com/asdfree/enade/actions"><img src="https://github.com/asdfree/enade/actions/workflows/r.yml/badge.svg" alt="Github Actions Badge"></a> <a href="https://www.eventbrite.com/e/1601880590969"><img src="https://img.shields.io/badge/Free%20Workshop%20Washington%20DC-October%208th,%202025-orange" alt="Washington DC: October 8th, 2025"></a> <a href="https://www.eventbrite.com/e/1634269818019"><img src="https://img.shields.io/badge/Free%20R%20Training%20Washington%20DC-October%2015th,%202025-yellow" alt="Washington DC: October 15th, 2025"></a> <a href="https://www.eventbrite.com/e/1458850233319"><img src="https://img.shields.io/badge/Free%20Workshop%20New%20York%20NY-October%2022nd,%202025-orange" alt="New York NY: October 22nd, 2025"></a> <a href="https://www.eventbrite.com/e/1560552678069"><img src="https://img.shields.io/badge/Free%20Workshop%20Philadelphia%20PA-October%2029th,%202025-orange" alt="Philadelphia PA: October 29th, 2025"></a> <a href="https://www.eventbrite.com/e/1710882017169"><img src="https://img.shields.io/badge/Free%20Workshop%20Washington%20DC-November%2012th,%202025-orange" alt="Washington DC: November 12th, 2025"></a>

The nationwide mandatory examination of college graduates.

* One table with one row per individual undergraduate student in Brazil.

* An enumeration of undergraduate students in Brazil.

* Released annually since 2004.

* Compiled by the [Instituto Nacional de Estudos e Pesquisas Educacionais Anísio Teixeira (INEP)](http://www.inep.gov.br/).

---

## Recommended Reading {-}

Two Methodology Documents:

> [Cálculo da nota final do Exame Nacional de Desempenho dos Estudiantes](https://download.inep.gov.br/educacao_superior/enade/notas_tecnicas/2017/Nota_Tecnica_CGCQES_n12_2017_Calculo_da_nota_final_do_Enade.pdf)

> [Wikipedia Entry](https://pt.wikipedia.org/wiki/Exame_Nacional_de_Desempenho_de_Estudantes)

<br>

One Haiku:

```{r}
# undergraduates
# sit for standardized testing
# exit interview
```

---

## Download, Import, Preparation {-}

Download, import, and merge two of the 2021 files:

```{r eval = FALSE , results = "hide" }
library(httr)
library(archive)

tf <- tempfile()

this_url <-	"https://download.inep.gov.br/microdados/microdados_enade_2021.zip"

GET( this_url , write_disk( tf ) , progress() )

archive_extract( tf , dir = tempdir() )

read_enade_archive <-
	function( this_regular_expression , this_directory ){
		this_filename <- 
			grep( 
				this_regular_expression , 
				list.files( 
					this_directory ,
					recursive = TRUE ,
					full.names = TRUE 
				) , 
				value = TRUE 
			)

		this_df <-
			read.table(
				this_filename ,
				header = TRUE ,
				sep = ";" ,
				na.strings = ""
			)
			
		names( this_df ) <- tolower( names( this_df ) )
		
		this_df
	}

arq1_df <- read_enade_archive( 'arq1\\.txt$' , tempdir() )

arq1_df <- unique( arq1_df[ c( 'co_curso' , 'co_uf_curso' , 'co_categad' , 'co_grupo' ) ] )

arq3_df <- read_enade_archive( 'arq3\\.txt$' , tempdir() )

enade_df <- merge( arq3_df , arq1_df )

stopifnot( nrow( enade_df ) == nrow( arq3_df ) )

```

### Save Locally \ {-}

Save the object at any point:

```{r eval = FALSE , results = "hide" }
# enade_fn <- file.path( path.expand( "~" ) , "ENADE" , "this_file.rds" )
# saveRDS( enade_df , file = enade_fn , compress = FALSE )
```

Load the same object:

```{r eval = FALSE , results = "hide" }
# enade_df <- readRDS( enade_fn )
```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE , results = "hide" }
enade_df <- 
	transform( 
		enade_df , 
		
		# qual foi o tempo gasto por voce para concluir a prova?
		less_than_two_hours = as.numeric( co_rs_i9 %in% c( 'A' , 'B' ) ) ,
		
		administrative_category =
			factor(
				co_categad ,
				levels = c( 1:5 , 7 ) ,
				labels = c( '1. Pública Federal' , '2. Pública Estadual' , 
				'3. Pública Municipal' , '4. Privada com fins lucrativos' , 
				'5. Privada sem fins lucrativos' , '7. Especial' )
			) ,

		state_name = 
			factor( 
				co_uf_curso , 
				levels = c( 11:17 , 21:29 , 31:33 , 35 , 41:43 , 50:53 ) ,
				labels = c( "Rondonia" , "Acre" , "Amazonas" , 
				"Roraima" , "Para" , "Amapa" , "Tocantins" , 
				"Maranhao" , "Piaui" , "Ceara" , "Rio Grande do Norte" , 
				"Paraiba" , "Pernambuco" , "Alagoas" , "Sergipe" , 
				"Bahia" , "Minas Gerais" , "Espirito Santo" , 
				"Rio de Janeiro" , "Sao Paulo" , "Parana" , 
				"Santa Catarina" , "Rio Grande do Sul" , 
				"Mato Grosso do Sul" , "Mato Grosso" , "Goias" , 
				"Distrito Federal" )
			)

	)
	
```

---

## Analysis Examples with base R \ {-}

### Unweighted Counts {-}

Count the unweighted number of records in the table, overall and by groups:
```{r eval = FALSE , results = "hide" }
nrow( enade_df )

table( enade_df[ , "administrative_category" ] , useNA = "always" )
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
mean( enade_df[ , "nt_obj_fg" ] , na.rm = TRUE )

tapply(
	enade_df[ , "nt_obj_fg" ] ,
	enade_df[ , "administrative_category" ] ,
	mean ,
	na.rm = TRUE 
)
```

Calculate the distribution of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
prop.table( table( enade_df[ , "state_name" ] ) )

prop.table(
	table( enade_df[ , c( "state_name" , "administrative_category" ) ] ) ,
	margin = 2
)
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
sum( enade_df[ , "nt_obj_fg" ] , na.rm = TRUE )

tapply(
	enade_df[ , "nt_obj_fg" ] ,
	enade_df[ , "administrative_category" ] ,
	sum ,
	na.rm = TRUE 
)
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
quantile( enade_df[ , "nt_obj_fg" ] , 0.5 , na.rm = TRUE )

tapply(
	enade_df[ , "nt_obj_fg" ] ,
	enade_df[ , "administrative_category" ] ,
	quantile ,
	0.5 ,
	na.rm = TRUE 
)
```

### Subsetting {-}

Limit your `data.frame` to students reporting that the general training section was easy or very easy:
```{r eval = FALSE , results = "hide" }
sub_enade_df <- subset( enade_df , co_rs_i1 %in% c( "A" , "B" ) )
```
Calculate the mean (average) of this subset:
```{r eval = FALSE , results = "hide" }
mean( sub_enade_df[ , "nt_obj_fg" ] , na.rm = TRUE )
```

### Measures of Uncertainty {-}

Calculate the variance, overall and by groups:
```{r eval = FALSE , results = "hide" }
var( enade_df[ , "nt_obj_fg" ] , na.rm = TRUE )

tapply(
	enade_df[ , "nt_obj_fg" ] ,
	enade_df[ , "administrative_category" ] ,
	var ,
	na.rm = TRUE 
)
```

### Regression Models and Tests of Association {-}

Perform a t-test:
```{r eval = FALSE , results = "hide" }
t.test( nt_obj_fg ~ less_than_two_hours , enade_df )
```

Perform a chi-squared test of association:
```{r eval = FALSE , results = "hide" }
this_table <- table( enade_df[ , c( "less_than_two_hours" , "state_name" ) ] )

chisq.test( this_table )
```

Perform a generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	glm( 
		nt_obj_fg ~ less_than_two_hours + state_name , 
		data = enade_df
	)

summary( glm_result )
```

---

## Replication Example {-}

This example matches the tecnologia em gestão da tecnologia da informação test scores on PDF page 48 of [the 2021 final results document](https://download.inep.gov.br/enade/resultados/2021/apresentacao_resultados_finais.pdf#page=48):

```{r eval = FALSE , results = "hide" }

it_students <- subset( enade_df , co_grupo %in% 6409 )

results <- sapply( it_students[ c( 'nt_fg' , 'nt_ce' , 'nt_ger' ) ] , mean , na.rm = TRUE )

stopifnot( round( results[ 'nt_fg' ] , 1 ) == 30.4 )
stopifnot( round( results[ 'nt_ce' ] , 1 ) == 38.2 )
stopifnot( round( results[ 'nt_ger' ] , 1 ) == 36.3 )

```

---

## Analysis Examples with `dplyr` \ {-}

The R `dplyr` library offers an alternative grammar of data manipulation to base R and SQL syntax. [dplyr](https://github.com/tidyverse/dplyr/) offers many verbs, such as `summarize`, `group_by`, and `mutate`, the convenience of pipe-able functions, and the `tidyverse` style of non-standard evaluation. [This vignette](https://cran.r-project.org/web/packages/dplyr/vignettes/dplyr.html) details the available features. As a starting point for ENADE users, this code replicates previously-presented examples:

```{r eval = FALSE , results = "hide" }
library(dplyr)
enade_tbl <- as_tibble( enade_df )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
enade_tbl %>%
	summarize( mean = mean( nt_obj_fg , na.rm = TRUE ) )

enade_tbl %>%
	group_by( administrative_category ) %>%
	summarize( mean = mean( nt_obj_fg , na.rm = TRUE ) )
```

---

## Analysis Examples with `data.table` \ {-}

The R `data.table` library provides a high-performance version of base R's data.frame with syntax and feature enhancements for ease of use, convenience and programming speed. [data.table](https://r-datatable.com) offers concise syntax: fast to type, fast to read, fast speed, memory efficiency, a careful API lifecycle management, an active community, and a rich set of features. [This vignette](https://cran.r-project.org/web/packages/data.table/vignettes/datatable-intro.html) details the available features. As a starting point for ENADE users, this code replicates previously-presented examples:

```{r eval = FALSE , results = 'hide' }
library(data.table)
enade_dt <- data.table( enade_df )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = 'hide' }
enade_dt[ , mean( nt_obj_fg , na.rm = TRUE ) ]

enade_dt[ , mean( nt_obj_fg , na.rm = TRUE ) , by = administrative_category ]
```

---

## Analysis Examples with `duckdb` \ {-}

The R `duckdb` library provides an embedded analytical data management system with support for the Structured Query Language (SQL). [duckdb](https://duckdb.org) offers a simple, feature-rich, fast, and free SQL OLAP management system. [This vignette](https://duckdb.org/docs/api/r) details the available features. As a starting point for ENADE users, this code replicates previously-presented examples:

```{r eval = FALSE , results = 'hide' }
library(duckdb)
con <- dbConnect( duckdb::duckdb() , dbdir = 'my-db.duckdb' )
dbWriteTable( con , 'enade' , enade_df )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = 'hide' }
dbGetQuery( con , 'SELECT AVG( nt_obj_fg ) FROM enade' )

dbGetQuery(
	con ,
	'SELECT
		administrative_category ,
		AVG( nt_obj_fg )
	FROM
		enade
	GROUP BY
		administrative_category'
)
```
