# Exame Nacional do Ensino Medio (ENEM) {-}

[![License: GPL v3](https://img.shields.io/badge/License-GPLv3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0) <a href="https://github.com/asdfree/enem/actions"><img src="https://github.com/asdfree/enem/actions/workflows/r.yml/badge.svg" alt="Github Actions Badge"></a> <a href="https://www.eventbrite.com/e/1771826874879"><img src="https://img.shields.io/badge/Free%20R%20Training%20Washington%20DC-November%2019th,%202025-orange" alt="Washington DC: November 19th, 2025"></a>

The national student aptitude test, used to assess high school completion and university admission.

* One table with one row per test-taking student, a second of study habit questionnaire respondents.

* Updated annually since 1998.

* Maintained by Brazil's [Instituto Nacional de Estudos e Pesquisas Educacionais Anisio Teixeira](http://www.inep.gov.br/)

---

## Recommended Reading {-}

Two Methodology Documents:

> `Leia_Me_Enem` included in each [annual zipped file](https://download.inep.gov.br/microdados/microdados_enem_2022.zip)

> [Wikipedia Entry](https://pt.wikipedia.org/wiki/Exame_Nacional_do_Ensino_M%C3%A9dio)

<br>

One Haiku:

```{r}
# graduation stage
# shake hands, toss cap, unroll scroll,
# mais um exame?
```

---

## Download, Import, Preparation {-}

Download and unzip the 2022 file:

```{r eval = FALSE , results = "hide" }
library(httr)
library(archive)

tf <- tempfile()

this_url <-	"https://download.inep.gov.br/microdados/microdados_enem_2022.zip"

GET( this_url , write_disk( tf ) , progress() )

archive_extract( tf , dir = tempdir() )
```

Import the 2022 file:

```{r eval = FALSE , results = "hide" }
library(readr)

enem_fns <- list.files( tempdir() , recursive = TRUE , full.names = TRUE )

enem_fn <- grep( "MICRODADOS_ENEM_([0-9][0-9][0-9][0-9])\\.csv$" , enem_fns , value = TRUE )

enem_tbl <- read_csv2( enem_fn , locale = locale( encoding = 'latin1' ) )

enem_df <- data.frame( enem_tbl )

names( enem_df ) <- tolower( names( enem_df ) )
```

### Save Locally \ {-}

Save the object at any point:

```{r eval = FALSE , results = "hide" }
# enem_fn <- file.path( path.expand( "~" ) , "ENEM" , "this_file.rds" )
# saveRDS( enem_df , file = enem_fn , compress = FALSE )
```

Load the same object:

```{r eval = FALSE , results = "hide" }
# enem_df <- readRDS( enem_fn )
```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE , results = "hide" }
enem_df <- 
	transform( 
		enem_df , 
		
		domestic_worker = as.numeric( q007 %in% c( 'B' , 'C' , 'D' ) ) ,
		
		administrative_category =
			factor(
				tp_dependencia_adm_esc ,
				levels = 1:4 ,
				labels = c( 'Federal' , 'Estadual' , 'Municipal' , 'Privada' )
			) ,

		state_name = 
			factor( 
				co_uf_esc , 
				levels = c( 11:17 , 21:29 , 31:33 , 35 , 41:43 , 50:53 ) ,
				labels = c( "Rondonia" , "Acre" , "Amazonas" , 
				"Roraima" , "Para" , "Amapa" , "Tocantins" , 
				"Maranhao" , "Piaui" , "Ceara" , "Rio Grande do Norte" , 
				"Paraiba" , "Pernambuco" , "Alagoas" , "Sergipe" , 
				"Bahia" , "Minas Gerais" , "Espirito Santo" , 
				"Rio de Janeiro" , "Sao Paulo" , "Parana" , 
				"Santa Catarina" , "Rio Grande do Sul" , 
				"Mato Grosso do Sul" , "Mato Grosso" , "Goias" , 
				"Distrito Federal" )
			)

	)
	
```

---

## Analysis Examples with base R \ {-}

### Unweighted Counts {-}

Count the unweighted number of records in the table, overall and by groups:
```{r eval = FALSE , results = "hide" }
nrow( enem_df )

table( enem_df[ , "administrative_category" ] , useNA = "always" )
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
mean( enem_df[ , "nu_nota_mt" ] , na.rm = TRUE )

tapply(
	enem_df[ , "nu_nota_mt" ] ,
	enem_df[ , "administrative_category" ] ,
	mean ,
	na.rm = TRUE 
)
```

Calculate the distribution of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
prop.table( table( enem_df[ , "state_name" ] ) )

prop.table(
	table( enem_df[ , c( "state_name" , "administrative_category" ) ] ) ,
	margin = 2
)
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
sum( enem_df[ , "nu_nota_mt" ] , na.rm = TRUE )

tapply(
	enem_df[ , "nu_nota_mt" ] ,
	enem_df[ , "administrative_category" ] ,
	sum ,
	na.rm = TRUE 
)
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
quantile( enem_df[ , "nu_nota_mt" ] , 0.5 , na.rm = TRUE )

tapply(
	enem_df[ , "nu_nota_mt" ] ,
	enem_df[ , "administrative_category" ] ,
	quantile ,
	0.5 ,
	na.rm = TRUE 
)
```

### Subsetting {-}

Limit your `data.frame` to mother graduated from high school:
```{r eval = FALSE , results = "hide" }
sub_enem_df <- subset( enem_df , q002 %in% c( 'E' , 'F' , 'G' ) )
```
Calculate the mean (average) of this subset:
```{r eval = FALSE , results = "hide" }
mean( sub_enem_df[ , "nu_nota_mt" ] , na.rm = TRUE )
```

### Measures of Uncertainty {-}

Calculate the variance, overall and by groups:
```{r eval = FALSE , results = "hide" }
var( enem_df[ , "nu_nota_mt" ] , na.rm = TRUE )

tapply(
	enem_df[ , "nu_nota_mt" ] ,
	enem_df[ , "administrative_category" ] ,
	var ,
	na.rm = TRUE 
)
```

### Regression Models and Tests of Association {-}

Perform a t-test:
```{r eval = FALSE , results = "hide" }
t.test( nu_nota_mt ~ domestic_worker , enem_df )
```

Perform a chi-squared test of association:
```{r eval = FALSE , results = "hide" }
this_table <- table( enem_df[ , c( "domestic_worker" , "state_name" ) ] )

chisq.test( this_table )
```

Perform a generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	glm( 
		nu_nota_mt ~ domestic_worker + state_name , 
		data = enem_df
	)

summary( glm_result )
```

---

## Replication Example {-}
This example matches the registration counts in the [Sinopse ENEM 2022 Excel table](https://download.inep.gov.br/informacoes_estatisticas/sinopses_estatisticas/sinopses_enem/2022/sinopse_enem_2022.zip):

```{r eval = FALSE , results = "hide" }
stopifnot( nrow( enem_df ) == 3476105 )
```

---

## Analysis Examples with `dplyr` \ {-}

The R `dplyr` library offers an alternative grammar of data manipulation to base R and SQL syntax. [dplyr](https://github.com/tidyverse/dplyr/) offers many verbs, such as `summarize`, `group_by`, and `mutate`, the convenience of pipe-able functions, and the `tidyverse` style of non-standard evaluation. [This vignette](https://cran.r-project.org/web/packages/dplyr/vignettes/dplyr.html) details the available features. As a starting point for ENEM users, this code replicates previously-presented examples:

```{r eval = FALSE , results = "hide" }
library(dplyr)
enem_tbl <- as_tibble( enem_df )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
enem_tbl %>%
	summarize( mean = mean( nu_nota_mt , na.rm = TRUE ) )

enem_tbl %>%
	group_by( administrative_category ) %>%
	summarize( mean = mean( nu_nota_mt , na.rm = TRUE ) )
```

---

## Analysis Examples with `data.table` \ {-}

The R `data.table` library provides a high-performance version of base R's data.frame with syntax and feature enhancements for ease of use, convenience and programming speed. [data.table](https://r-datatable.com) offers concise syntax: fast to type, fast to read, fast speed, memory efficiency, a careful API lifecycle management, an active community, and a rich set of features. [This vignette](https://cran.r-project.org/web/packages/data.table/vignettes/datatable-intro.html) details the available features. As a starting point for ENEM users, this code replicates previously-presented examples:

```{r eval = FALSE , results = 'hide' }
library(data.table)
enem_dt <- data.table( enem_df )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = 'hide' }
enem_dt[ , mean( nu_nota_mt , na.rm = TRUE ) ]

enem_dt[ , mean( nu_nota_mt , na.rm = TRUE ) , by = administrative_category ]
```

---

## Analysis Examples with `duckdb` \ {-}

The R `duckdb` library provides an embedded analytical data management system with support for the Structured Query Language (SQL). [duckdb](https://duckdb.org) offers a simple, feature-rich, fast, and free SQL OLAP management system. [This vignette](https://duckdb.org/docs/api/r) details the available features. As a starting point for ENEM users, this code replicates previously-presented examples:

```{r eval = FALSE , results = 'hide' }
library(duckdb)
con <- dbConnect( duckdb::duckdb() , dbdir = 'my-db.duckdb' )
dbWriteTable( con , 'enem' , enem_df )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = 'hide' }
dbGetQuery( con , 'SELECT AVG( nu_nota_mt ) FROM enem' )

dbGetQuery(
	con ,
	'SELECT
		administrative_category ,
		AVG( nu_nota_mt )
	FROM
		enem
	GROUP BY
		administrative_category'
)
```
