# European Social Survey (ESS) {-}

```{r , echo = FALSE }

most_recent_build_date <- gsub( "-" , " " , as.Date( file.info( "ess.Rmd" )$mtime ) )

ess_badge <- paste0( "<img src='https://img.shields.io/badge/tested%20on%20my%20windows%20laptop:-" , most_recent_build_date , "-brightgreen' alt='Local Testing Badge'>" )

```

`r ess_badge`

The European Social Survey measures political opinion and behavior across the continent.

* One table per country with one row per sampled respondent.

* A complex sample survey designed to generalize to residents aged 15 and older in participating nations.

* Released biennially since 2002.

* Headquartered at [City, University of London](http://www.city.ac.uk/) and governed by [a scientific team across Europe](http://www.europeansocialsurvey.org/about/structure_and_governance.html).

---

Please skim before you begin:

1. 

2. 

3. This poem

```{r}
# 
# 
# 
```

---

## Download, Import, Preparation {-}

```{r , cache = TRUE }
library(foreign)

ess_be_df <- 
	read.spss( 
		file.path( 
			path.expand( "~" ) , 
			"ESS-Data-Wizard-subset-2023-01-17.sav/ESS-Data-Wizard-subset-2023-01-17.sav" 
		) ,
		to.data.frame = TRUE ,
		use.value.labels = FALSE
	)

ess_sddf_df <-
	read.spss(
		file.path(
			path.expand( "~" ) ,
			"ESS7SDDFe1_2.sav/ESS7SDDFe1_2.sav"
		) ,
		to.data.frame = TRUE ,
		use.value.labels = FALSE
	)	
```

---

## Analysis Examples with the `survey` library \ {-}

Construct a complex sample survey design:

```{r cache = TRUE }

```

```{r cache = TRUE }
library(survey)

ess_df <-
	merge( 
		ess_be_df , 
		ess_sddf_df , 
		by = c( 'cntry' , 'idno' ) 
	)

stopifnot( nrow( ess_df ) == nrow( ess_be_df ) )

ess_design <- 
	svydesign(
		ids = ~psu ,
		strata = ~stratum ,
		probs = ~prob ,
		data = ess_df
	)
```

### Variable Recoding {-}

Add new columns to the data set:
```{r cache = TRUE }
ess_design <- 
	update( 
		ess_design , 
		
		one = 1 ,
		
		non_european_immigrants =
			factor( impcntr ,
				labels = c( 'Allow many to come and live here' , 
				'Allow some' , 'Allow a few' , 'Allow none' )
			) ,
		
		sex = factor( icgndra , labels = c( 'male' , 'female' ) ) ,
			
		more_than_one_hour_tv_daily = as.numeric( tvtot >= 3 )
	)
```

### Unweighted Counts {-}

Count the unweighted number of records in the survey sample, overall and by groups:
```{r cache = TRUE , results = "hide" }
sum( weights( ess_design , "sampling" ) != 0 )

svyby( ~ one , ~ non_european_immigrants , ess_design , unwtd.count )
```

### Weighted Counts {-}
Count the weighted size of the generalizable population, overall and by groups:
```{r cache = TRUE , results = "hide" }
svytotal( ~ one , ess_design )

svyby( ~ one , ~ non_european_immigrants , ess_design , svytotal )
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r cache = TRUE , results = "hide" }
svymean( ~ ppltrst , ess_design )

svyby( ~ ppltrst , ~ non_european_immigrants , ess_design , svymean )
```

Calculate the distribution of a categorical variable, overall and by groups:
```{r cache = TRUE , results = "hide" }
svymean( ~ sex , ess_design , na.rm = TRUE )

svyby( ~ sex , ~ non_european_immigrants , ess_design , svymean , na.rm = TRUE )
```

Calculate the sum of a linear variable, overall and by groups:
```{r cache = TRUE , results = "hide" }
svytotal( ~ ppltrst , ess_design )

svyby( ~ ppltrst , ~ non_european_immigrants , ess_design , svytotal )
```

Calculate the weighted sum of a categorical variable, overall and by groups:
```{r cache = TRUE , results = "hide" }
svytotal( ~ sex , ess_design , na.rm = TRUE )

svyby( ~ sex , ~ non_european_immigrants , ess_design , svytotal , na.rm = TRUE )
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r cache = TRUE , results = "hide" }
svyquantile( ~ ppltrst , ess_design , 0.5 )

svyby( 
	~ ppltrst , 
	~ non_european_immigrants , 
	ess_design , 
	svyquantile , 
	0.5 ,
	ci = TRUE 
)
```

Estimate a ratio:
```{r cache = TRUE , results = "hide" }
svyratio( 
	numerator = ~ ppltrst , 
	denominator = ~ pplfair , 
	ess_design 
)
```

### Subsetting {-}

Restrict the survey design to voters:
```{r cache = TRUE , results = "hide" }
sub_ess_design <- subset( ess_design , vote == 1 )
```
Calculate the mean (average) of this subset:
```{r cache = TRUE , results = "hide" }
svymean( ~ ppltrst , sub_ess_design )
```

### Measures of Uncertainty {-}

Extract the coefficient, standard error, confidence interval, and coefficient of variation from any descriptive statistics function result, overall and by groups:
```{r cache = TRUE , results = "hide" }
this_result <- svymean( ~ ppltrst , ess_design )

coef( this_result )
SE( this_result )
confint( this_result )
cv( this_result )

grouped_result <-
	svyby( 
		~ ppltrst , 
		~ non_european_immigrants , 
		ess_design , 
		svymean 
	)
	
coef( grouped_result )
SE( grouped_result )
confint( grouped_result )
cv( grouped_result )
```

Calculate the degrees of freedom of any survey design object:
```{r cache = TRUE , results = "hide" }
degf( ess_design )
```

Calculate the complex sample survey-adjusted variance of any statistic:
```{r cache = TRUE , results = "hide" }
svyvar( ~ ppltrst , ess_design )
```

Include the complex sample design effect in the result for a specific statistic:
```{r cache = TRUE , results = "hide" }
# SRS without replacement
svymean( ~ ppltrst , ess_design , deff = TRUE )

# SRS with replacement
svymean( ~ ppltrst , ess_design , deff = "replace" )
```

Compute confidence intervals for proportions using methods that may be more accurate near 0 and 1. See `?svyciprop` for alternatives:
```{r cache = TRUE , results = "hide" }
svyciprop( ~ more_than_one_hour_tv_daily , ess_design ,
	method = "likelihood" , na.rm = TRUE )
```

### Regression Models and Tests of Association {-}

Perform a design-based t-test:
```{r cache = TRUE , results = "hide" }
svyttest( ppltrst ~ more_than_one_hour_tv_daily , ess_design )
```

Perform a chi-squared test of association for survey data:
```{r cache = TRUE , results = "hide" }
svychisq( 
	~ more_than_one_hour_tv_daily + sex , 
	ess_design 
)
```

Perform a survey-weighted generalized linear model:
```{r cache = TRUE , results = "hide" }
glm_result <- 
	svyglm( 
		ppltrst ~ more_than_one_hour_tv_daily + sex , 
		ess_design 
	)

summary( glm_result )
```

---

## Intermish {-}

---

## Analysis Examples with `srvyr` \ {-}

The R `srvyr` library calculates summary statistics from survey data, such as the mean, total or quantile using [dplyr](https://github.com/tidyverse/dplyr/)-like syntax. [srvyr](https://github.com/gergness/srvyr) allows for the use of many verbs, such as `summarize`, `group_by`, and `mutate`, the convenience of pipe-able functions, the `tidyverse` style of non-standard evaluation and more consistent return types than the `survey` package. [This vignette](https://cran.r-project.org/web/packages/srvyr/vignettes/srvyr-vs-survey.html) details the available features. As a starting point for ESS users, this code replicates previously-presented examples:

```{r cache = TRUE , results = "hide" }
library(srvyr)
ess_srvyr_design <- as_survey( ess_design )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r cache = TRUE , results = "hide" }
ess_srvyr_design %>%
	summarize( mean = survey_mean( ppltrst ) )

ess_srvyr_design %>%
	group_by( non_european_immigrants ) %>%
	summarize( mean = survey_mean( ppltrst ) )
```

---

## Replication Example {-}

```{r cache = TRUE , results = "hide" }

```

