# European Social Survey (ESS) {-}

[![License: GPL v3](https://img.shields.io/badge/License-GPLv3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0) <img src='https://img.shields.io/badge/Tested%20Locally-Windows%20Laptop-brightgreen' alt='Local Testing Badge'> <a href="https://www.eventbrite.com/e/1972097244947"><img src="https://img.shields.io/badge/Free%20Workshop%20Washington%20DC-December%203rd,%202025-orange" alt="Washington DC: December 3rd, 2025"></a>

The barometer of political opinion and behavior across the continent.

* One table per country with one row per sampled respondent.

* A complex sample designed to generalize to residents aged 15 and older in participating nations.

* Released biennially since 2002.

* Headquartered at [City, University of London](http://www.city.ac.uk/) and governed by [a scientific team across Europe](http://www.europeansocialsurvey.org/about/structure_and_governance.html).

---

## Recommended Reading {-}

Four Example Strengths & Limitations:

✔️ [Rotating modules allow external researchers to propose new questions](https://europeansocialsurvey.org/sites/default/files/2024-01/ESS_R13_module_call_specification.pdf)

✔️ [Sub-national geographies available](https://stessrelpubprodwe.blob.core.windows.net/data/round11/survey/ESS11_country_documentation_report_e02.pdf)

❌ [Country-specific differences in methodology](https://www.europeansocialsurvey.org/sites/default/files/2023-06/Country_specific_quality_control_checks_for_ess_weighting_procedures.pdf#page=10)

❌ [Questionnaires only translated into languages spoken by at least 5% of each country's population](https://www.europeansocialsurvey.org/sites/default/files/2023-06/ESS5_sampling_guidelines.pdf#page=11)

<br>

Three Example Findings:

1. [Childless adults aged 65 to 74 in 2002 were not more socially excluded than those in 2018](https://doi.org/10.1007/s10433-024-00831-z).

2. [Between 2002-2003 and 2016-2017, there was little change overall in the extent to which Europeans felt that their countries were made a better or worse place to live as a result of migration](https://doi.org/10.1787/0adf9e55-en).

3. [The 2022 Russian invasion of Ukraine reduced authoritarian attitudes across eight nations](https://doi.org/10.1093/pnasnexus/pgae292).

<br>

Two Methodology Documents:

> [Findings from the European Social Survey](https://www.europeansocialsurvey.org/findings/)

> [Wikipedia Entry](https://en.wikipedia.org/wiki/European_Social_Survey)

<br>

One Haiku:

```{r}
# pent up belief gauge
# open border monarchists
# survey for your thoughts
```

---

## Download, Import, Preparation {-}

1. Register at the ESS Data Portal at https://ess-search.nsd.no/.

2. Choose `ESS round 8 - 2016. Welfare attitudes, Attitudes to climate change`.

3. Download the integrated file and also the sample design (SDDF) files as `SAV` (SPSS) files:

```{r eval = FALSE , results = "hide" }
library(foreign)

ess_int_df <- 
	read.spss( 
		file.path( 
			path.expand( "~" ) , 
			"ESS8e02_2.sav" 
		) ,
		to.data.frame = TRUE ,
		use.value.labels = FALSE
	)

ess_sddf_df <-
	read.spss(
		file.path(
			path.expand( "~" ) ,
			"ESS8SDDFe01_1.sav"
		) ,
		to.data.frame = TRUE ,
		use.value.labels = FALSE
	)
	

ess_df <-
	merge( 
		ess_int_df , 
		ess_sddf_df , 
		by = c( 'cntry' , 'idno' ) 
	)

stopifnot( nrow( ess_df ) == nrow( ess_int_df ) )
```

### Save Locally \ {-}

Save the object at any point:

```{r eval = FALSE , results = "hide" }
# ess_fn <- file.path( path.expand( "~" ) , "ESS" , "this_file.rds" )
# saveRDS( ess_df , file = ess_fn , compress = FALSE )
```

Load the same object:

```{r eval = FALSE , results = "hide" }
# ess_df <- readRDS( ess_fn )
```

### Survey Design Definition {-}
Construct a complex sample survey design:

```{r eval = FALSE , results = "hide" }
library(survey)

options( survey.lonely.psu = "adjust" )

ess_df[ , 'anweight' ] <-
	ess_df[ , 'pspwght' ] *
	ess_df[ , 'pweight' ] *
	10000

ess_design <- 
	svydesign(
		ids = ~psu ,
		strata = ~stratum ,
		weights = ~anweight ,
		data = ess_df ,
		nest = TRUE
	)
```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE , results = "hide" }
ess_design <- 
	update( 
		ess_design , 
		
		one = 1 ,
		
		gndr = factor( gndr , labels = c( 'male' , 'female' ) ) ,
		
		netusoft =
			factor(
				netusoft ,
				levels = 1:5 ,
				labels = c( 'Never' , 'Only occasionally' ,
					'A few times a week' , 'Most days' , 'Every day' )
			) ,
			
		belonging_to_particular_religion = as.numeric( rlgblg == 1 )
	)
```

---

## Analysis Examples with the `survey` library \ {-}

### Unweighted Counts {-}

Count the unweighted number of records in the survey sample, overall and by groups:
```{r eval = FALSE , results = "hide" }
sum( weights( ess_design , "sampling" ) != 0 )

svyby( ~ one , ~ cntry , ess_design , unwtd.count )
```

### Weighted Counts {-}
Count the weighted size of the generalizable population, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ one , ess_design )

svyby( ~ one , ~ cntry , ess_design , svytotal )
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ ppltrst , ess_design , na.rm = TRUE )

svyby( ~ ppltrst , ~ cntry , ess_design , svymean , na.rm = TRUE )
```

Calculate the distribution of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ gndr , ess_design , na.rm = TRUE )

svyby( ~ gndr , ~ cntry , ess_design , svymean , na.rm = TRUE )
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ ppltrst , ess_design , na.rm = TRUE )

svyby( ~ ppltrst , ~ cntry , ess_design , svytotal , na.rm = TRUE )
```

Calculate the weighted sum of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ gndr , ess_design , na.rm = TRUE )

svyby( ~ gndr , ~ cntry , ess_design , svytotal , na.rm = TRUE )
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svyquantile( ~ ppltrst , ess_design , 0.5 , na.rm = TRUE )

svyby( 
	~ ppltrst , 
	~ cntry , 
	ess_design , 
	svyquantile , 
	0.5 ,
	ci = TRUE , na.rm = TRUE
)
```

Estimate a ratio:
```{r eval = FALSE , results = "hide" }
svyratio( 
	numerator = ~ ppltrst , 
	denominator = ~ pplfair , 
	ess_design ,
	na.rm = TRUE
)
```

### Subsetting {-}

Restrict the survey design to voters:
```{r eval = FALSE , results = "hide" }
sub_ess_design <- subset( ess_design , vote == 1 )
```
Calculate the mean (average) of this subset:
```{r eval = FALSE , results = "hide" }
svymean( ~ ppltrst , sub_ess_design , na.rm = TRUE )
```

### Measures of Uncertainty {-}

Extract the coefficient, standard error, confidence interval, and coefficient of variation from any descriptive statistics function result, overall and by groups:
```{r eval = FALSE , results = "hide" }
this_result <- svymean( ~ ppltrst , ess_design , na.rm = TRUE )

coef( this_result )
SE( this_result )
confint( this_result )
cv( this_result )

grouped_result <-
	svyby( 
		~ ppltrst , 
		~ cntry , 
		ess_design , 
		svymean ,
		na.rm = TRUE 
	)
	
coef( grouped_result )
SE( grouped_result )
confint( grouped_result )
cv( grouped_result )
```

Calculate the degrees of freedom of any survey design object:
```{r eval = FALSE , results = "hide" }
degf( ess_design )
```

Calculate the complex sample survey-adjusted variance of any statistic:
```{r eval = FALSE , results = "hide" }
svyvar( ~ ppltrst , ess_design , na.rm = TRUE )
```

Include the complex sample design effect in the result for a specific statistic:
```{r eval = FALSE , results = "hide" }
# SRS without replacement
svymean( ~ ppltrst , ess_design , na.rm = TRUE , deff = TRUE )

# SRS with replacement
svymean( ~ ppltrst , ess_design , na.rm = TRUE , deff = "replace" )
```

Compute confidence intervals for proportions using methods that may be more accurate near 0 and 1. See `?svyciprop` for alternatives:
```{r eval = FALSE , results = "hide" }
svyciprop( ~ belonging_to_particular_religion , ess_design ,
	method = "likelihood" , na.rm = TRUE )
```

### Regression Models and Tests of Association {-}

Perform a design-based t-test:
```{r eval = FALSE , results = "hide" }
svyttest( ppltrst ~ belonging_to_particular_religion , ess_design )
```

Perform a chi-squared test of association for survey data:
```{r eval = FALSE , results = "hide" }
svychisq( 
	~ belonging_to_particular_religion + gndr , 
	ess_design 
)
```

Perform a survey-weighted generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	svyglm( 
		ppltrst ~ belonging_to_particular_religion + gndr , 
		ess_design 
	)

summary( glm_result )
```

---

## Replication Example {-}

This example matches statistics and confidence intervals within 0.1% from the [Guide to Using Weights and Sample Design Indicators with ESS Data](https://stessrelpubprodwe.blob.core.windows.net/data/round10/survey/ESS_weighting_data_1_1.pdf#page=13):

```{r eval = FALSE , results = "hide" }
published_proportions <- c( 0.166 , 0.055 , 0.085 , 0.115 , 0.578 )

published_lb <- c( 0.146 , 0.045 , 0.072 , 0.099 , 0.550 )

published_ub <- c( 0.188 , 0.068 , 0.100 , 0.134 , 0.605 )

austrians <- subset( ess_design , cntry == 'AT' )

( results <- svymean( ~ netusoft , austrians , na.rm = TRUE ) )

stopifnot( all( round( coef( results ) , 3 ) == published_proportions ) )

( ci_results <- confint( results ) )

stopifnot( all( abs( ci_results[ , 1 ] - published_lb ) < 0.0015 ) )

stopifnot( all( abs( ci_results[ , 2 ] - published_ub ) < 0.0015 ) )
```

---

## Analysis Examples with `srvyr` \ {-}

The R `srvyr` library calculates summary statistics from survey data, such as the mean, total or quantile using [dplyr](https://github.com/tidyverse/dplyr/)-like syntax. [srvyr](https://github.com/gergness/srvyr) allows for the use of many verbs, such as `summarize`, `group_by`, and `mutate`, the convenience of pipe-able functions, the `tidyverse` style of non-standard evaluation and more consistent return types than the `survey` package. [This vignette](https://cran.r-project.org/web/packages/srvyr/vignettes/srvyr-vs-survey.html) details the available features. As a starting point for ESS users, this code replicates previously-presented examples:

```{r eval = FALSE , results = "hide" }
library(srvyr)
ess_srvyr_design <- as_survey( ess_design )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
ess_srvyr_design %>%
	summarize( mean = survey_mean( ppltrst , na.rm = TRUE ) )

ess_srvyr_design %>%
	group_by( cntry ) %>%
	summarize( mean = survey_mean( ppltrst , na.rm = TRUE ) )
```


