# FDA Adverse Event Reporting System (FAERS) {-}

[![License: GPL v3](https://img.shields.io/badge/License-GPLv3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0) <a href="https://github.com/asdfree/faers/actions"><img src="https://github.com/asdfree/faers/actions/workflows/r.yml/badge.svg" alt="Github Actions Badge"></a> <a href="https://www.eventbrite.com/e/1771826874879"><img src="https://img.shields.io/badge/Free%20R%20Training%20Washington%20DC-November%2019th,%202025-orange" alt="Washington DC: November 19th, 2025"></a>

The post-marketing safety surveillance program for drug and therapeutic biological products.

* Multiple tables linked by `primaryid` including demographics, outcomes, drug start and end dates.

* Voluntary reports from practitioners and patients, not representative, no verification of causality.

* Published quarterly since 2004, file structure revisions at 2012Q4 and 2014Q3.

* Maintained by the United States [Food and Drug Administration (FDA)](http://www.fda.gov/).

---

## Recommended Reading {-}

Two Methodology Documents:

> `ASC_NTS.DOC` included in each [quarterly zipped file](https://fis.fda.gov/content/Exports/faers_ascii_2023q1.zip), especially the Entity Relationship Diagram

> [Questions and Answers on FDA's Adverse Event Reporting System (FAERS)](https://www.fda.gov/drugs/surveillance/questions-and-answers-fdas-adverse-event-reporting-system-faers)

<br>

One Haiku:

```{r}
# side effect guestbook
# violet you're turning violet
# vi'lent dose response
```

---

## Function Definitions {-}

Define a function to import each text file:

```{r eval = FALSE , results = "hide" }
read_faers <-
	function( this_fn ){
		read.table( this_fn , sep = "$" , header = TRUE , comment.char = "" , quote = "" )
	}
```	

---

## Download, Import, Preparation {-}

Download the quarterly file:

```{r eval = FALSE , results = "hide" }
library(httr)

tf <- tempfile()

this_url <- "https://fis.fda.gov/content/Exports/faers_ascii_2023q1.zip"

GET( this_url , write_disk( tf ) , progress() )

unzipped_files <- unzip( tf , exdir = tempdir() )
```

Import multiple tables from the downloaded quarter of microdata:
```{r eval = FALSE , results = "hide" }
# one record per report
faers_demo_df <- read_faers( grep( 'DEMO23Q1\\.txt$' , unzipped_files , value = TRUE ) )

# one or more record per report
faers_drug_df <- read_faers( grep( 'DRUG23Q1\\.txt$' , unzipped_files , value = TRUE ) )

# zero or more records per report
faers_outcome_df <- read_faers( grep( 'OUTC23Q1\\.txt$' , unzipped_files , value = TRUE ) )
```

Construct an analysis file limited to reported deaths:
```{r eval = FALSE , results = "hide" }
# limit the outcome file to deaths
faers_deaths_df <- subset( faers_outcome_df , outc_cod == 'DE' )

# merge demographics with each reported death
faers_df <-	merge( faers_demo_df , faers_deaths_df )

# confirm that the analysis file matches the number of death outcomes
stopifnot( nrow( faers_deaths_df ) == nrow( faers_df ) )

# confirm zero reports include multiple deaths from the same reported adverse event
stopifnot( nrow( faers_df ) == length( unique( faers_df[ , 'primaryid' ] ) ) )
```

### Save Locally \ {-}

Save the object at any point:

```{r eval = FALSE , results = "hide" }
# faers_fn <- file.path( path.expand( "~" ) , "FAERS" , "this_file.rds" )
# saveRDS( faers_df , file = faers_fn , compress = FALSE )
```

Load the same object:

```{r eval = FALSE , results = "hide" }
# faers_df <- readRDS( faers_fn )
```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE , results = "hide" }
faers_df <- 
	transform( 
		faers_df , 
		
		physician_reported = as.numeric( occp_cod == "MD" ) ,
		
		reporter_country_categories = 
			ifelse( reporter_country == 'US' , 'USA' ,
			ifelse( reporter_country == 'COUNTRY NOT SPECIFIED' , 'missing' ,
			ifelse( reporter_country == 'JP' , 'Japan' ,
			ifelse( reporter_country == 'UK' , 'UK' ,
			ifelse( reporter_country == 'CA' , 'Canada' ,
			ifelse( reporter_country == 'FR' , 'France' ,
				'Other' ) ) ) ) ) ) ,
		
		init_fda_year = as.numeric( substr( init_fda_dt , 1 , 4 ) )
		
	)
	
```

---

## Analysis Examples with base R \ {-}

### Unweighted Counts {-}

Count the unweighted number of records in the table, overall and by groups:
```{r eval = FALSE , results = "hide" }
nrow( faers_df )

table( faers_df[ , "reporter_country_categories" ] , useNA = "always" )
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
mean( faers_df[ , "init_fda_year" ] , na.rm = TRUE )

tapply(
	faers_df[ , "init_fda_year" ] ,
	faers_df[ , "reporter_country_categories" ] ,
	mean ,
	na.rm = TRUE 
)
```

Calculate the distribution of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
prop.table( table( faers_df[ , "sex" ] ) )

prop.table(
	table( faers_df[ , c( "sex" , "reporter_country_categories" ) ] ) ,
	margin = 2
)
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
sum( faers_df[ , "init_fda_year" ] , na.rm = TRUE )

tapply(
	faers_df[ , "init_fda_year" ] ,
	faers_df[ , "reporter_country_categories" ] ,
	sum ,
	na.rm = TRUE 
)
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
quantile( faers_df[ , "init_fda_year" ] , 0.5 , na.rm = TRUE )

tapply(
	faers_df[ , "init_fda_year" ] ,
	faers_df[ , "reporter_country_categories" ] ,
	quantile ,
	0.5 ,
	na.rm = TRUE 
)
```

### Subsetting {-}

Limit your `data.frame` to elderly persons:
```{r eval = FALSE , results = "hide" }
sub_faers_df <- subset( faers_df , age_grp == "E" )
```
Calculate the mean (average) of this subset:
```{r eval = FALSE , results = "hide" }
mean( sub_faers_df[ , "init_fda_year" ] , na.rm = TRUE )
```

### Measures of Uncertainty {-}

Calculate the variance, overall and by groups:
```{r eval = FALSE , results = "hide" }
var( faers_df[ , "init_fda_year" ] , na.rm = TRUE )

tapply(
	faers_df[ , "init_fda_year" ] ,
	faers_df[ , "reporter_country_categories" ] ,
	var ,
	na.rm = TRUE 
)
```

### Regression Models and Tests of Association {-}

Perform a t-test:
```{r eval = FALSE , results = "hide" }
t.test( init_fda_year ~ physician_reported , faers_df )
```

Perform a chi-squared test of association:
```{r eval = FALSE , results = "hide" }
this_table <- table( faers_df[ , c( "physician_reported" , "sex" ) ] )

chisq.test( this_table )
```

Perform a generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	glm( 
		init_fda_year ~ physician_reported + sex , 
		data = faers_df
	)

summary( glm_result )
```

---

## Replication Example {-}

This example matches the death frequency counts in the `OUTC23Q1.pdf` file in [the downloaded quarter](https://fis.fda.gov/content/Exports/faers_ascii_2023q1.zip):

```{r eval = FALSE , results = "hide" }
stopifnot( nrow( faers_df ) == 37704 )
```

---

## Analysis Examples with `dplyr` \ {-}

The R `dplyr` library offers an alternative grammar of data manipulation to base R and SQL syntax. [dplyr](https://github.com/tidyverse/dplyr/) offers many verbs, such as `summarize`, `group_by`, and `mutate`, the convenience of pipe-able functions, and the `tidyverse` style of non-standard evaluation. [This vignette](https://cran.r-project.org/web/packages/dplyr/vignettes/dplyr.html) details the available features. As a starting point for FAERS users, this code replicates previously-presented examples:

```{r eval = FALSE , results = "hide" }
library(dplyr)
faers_tbl <- as_tibble( faers_df )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
faers_tbl %>%
	summarize( mean = mean( init_fda_year , na.rm = TRUE ) )

faers_tbl %>%
	group_by( reporter_country_categories ) %>%
	summarize( mean = mean( init_fda_year , na.rm = TRUE ) )
```

---

## Analysis Examples with `data.table` \ {-}

The R `data.table` library provides a high-performance version of base R's data.frame with syntax and feature enhancements for ease of use, convenience and programming speed. [data.table](https://r-datatable.com) offers concise syntax: fast to type, fast to read, fast speed, memory efficiency, a careful API lifecycle management, an active community, and a rich set of features. [This vignette](https://cran.r-project.org/web/packages/data.table/vignettes/datatable-intro.html) details the available features. As a starting point for FAERS users, this code replicates previously-presented examples:

```{r eval = FALSE , results = 'hide' }
library(data.table)
faers_dt <- data.table( faers_df )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = 'hide' }
faers_dt[ , mean( init_fda_year , na.rm = TRUE ) ]

faers_dt[ , mean( init_fda_year , na.rm = TRUE ) , by = reporter_country_categories ]
```

---

## Analysis Examples with `duckdb` \ {-}

The R `duckdb` library provides an embedded analytical data management system with support for the Structured Query Language (SQL). [duckdb](https://duckdb.org) offers a simple, feature-rich, fast, and free SQL OLAP management system. [This vignette](https://duckdb.org/docs/api/r) details the available features. As a starting point for FAERS users, this code replicates previously-presented examples:

```{r eval = FALSE , results = 'hide' }
library(duckdb)
con <- dbConnect( duckdb::duckdb() , dbdir = 'my-db.duckdb' )
dbWriteTable( con , 'faers' , faers_df )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = 'hide' }
dbGetQuery( con , 'SELECT AVG( init_fda_year ) FROM faers' )

dbGetQuery(
	con ,
	'SELECT
		reporter_country_categories ,
		AVG( init_fda_year )
	FROM
		faers
	GROUP BY
		reporter_country_categories'
)
```
