--- 
title: "Analyze Survey Data for Free"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: ajdamico/asdfree
description: ""
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  cache.lazy=FALSE
)
```

```{r echo = FALSE , eval = TRUE , results = "hide" }
file.copy( "CNAME" , "docs/CNAME" , overwrite = TRUE )
file.copy( "404.html" , "docs/404.html" , overwrite = TRUE )
```

# `r stringi::stri_trans_general(english::english(length(list.files("metadata/"))), id = "Title")` Public Microdatasets {-}

## One Easy To Type Website {.unlisted .unnumbered}

Please ask questions about this book on [cross validated](https://stats.stackexchange.com/questions/ask?tags=survey) for survey statistics or [stackoverflow](https://stackoverflow.com/questions/ask?tags=r+survey) for `R`.

This textbook replaces my [archived blog](http://usgsd.blogspot.com/), prior [code](https://github.com/ajdamico/asdfree/tree/archive), and the no longer maintained [lodown package](http://github.com/ajdamico/lodown).

_A work of R is never finished, merely abandoned._

&emsp;- Anthony Damico
