# Medical Expenditure Panel Survey (MEPS) {-}

<a href="https://github.com/asdfree/meps/actions"><img src="https://github.com/asdfree/meps/actions/workflows/r.yml/badge.svg" alt="Github Actions Badge"></a>

The Household Component captures person-level spending across service categories, coverage types.

* The consolidated file contains one row per individual within each sampled household, other tables contain one record per event (like prescription fills, hospitalizations), per job, per insurance policy.

* A complex sample survey designed to generalize to the U.S. civilian non-institutionalized population.

* Released annually since 1996.

* Administered by the [Agency for Healthcare Research and Quality](http://www.ahrq.gov/).

---

Please skim before you begin:

1. [MEPS HC-224 2020 Full Year Consolidated Data File](https://meps.ahrq.gov/data_stats/download_data/pufs/h224/h224doc.shtml)

2. [Wikipedia Entry](https://en.wikipedia.org/wiki/Medical_Expenditure_Panel_Survey)

3. This human-composed haiku or a bouquet of [artificial intelligence-generated limericks](https://www.gnod.com/search/ai#q=write%20a%20limerick-style%20poem%20about%20the%20Medical Expenditure Panel Survey)

```{r}
# king dumpty's horsemen
# ahrq stitches payors, bills, claims
# fractured health system
```

---

## Download, Import, Preparation {-}

Define a function to download, unzip, and import each sas file:

```{r messages = FALSE , eval = FALSE }
library(haven)

meps_sas_import <-
	function( this_url ){
		
		this_tf <- tempfile()
		
		download.file( this_url , this_tf , mode = 'wb' )
		
		this_tbl <- read_sas( this_tf )

		this_df <- data.frame( this_tbl )
		
		names( this_df ) <- tolower( names( this_df ) )
		
		this_df
	}

```	

Download and import the consolidated file and the replicate weights file:

```{r messages = FALSE , eval = FALSE }
meps_cons_df <-
	meps_sas_import( "https://meps.ahrq.gov/data_files/pufs/h224/h224v9.zip" )

meps_brr_df <-
	meps_sas_import( "https://meps.ahrq.gov/mepsweb/data_files/pufs/h036brr/h36brr20v9.zip" )
```

Merge the consolidated file with the replicate weights:

```{r messages = FALSE , eval = FALSE }
meps_df <- merge( meps_cons_df , meps_brr_df )

stopifnot( nrow( meps_df ) == nrow( meps_cons_df ) )

meps_df[ , 'one' ] <- 1
```

### Save locally \ {-}

Save the object at any point:

```{r eval = FALSE , results = "hide" }
# meps_fn <- file.path( path.expand( "~" ) , "MEPS" , "this_file.rds" )
# saveRDS( meps_df , file = meps_fn , compress = FALSE )
```

Load the same object:

```{r eval = FALSE , results = "hide" }
# meps_df <- readRDS( meps_fn )
```

### Survey Design Definition {-}
Construct a complex sample survey design:

```
library(survey)

meps_design <-
	svrepdesign(
		data = meps_df ,
		weights = ~ perwt20f ,
		type = "BRR" ,
		combined.weights = FALSE ,
		repweights = "brr[1-9]+" ,
		mse = TRUE
	)
```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE , results = "hide" }
meps_design <- 
	update( 
		meps_design , 
		
		one = 1 ,
		
		insured_december_31st = ifelse( ins20x %in% 1:2 , as.numeric( ins20x == 1 ) , NA )
		
	)
```

---

## Analysis Examples with the `survey` library \ {-}

### Unweighted Counts {-}

Count the unweighted number of records in the survey sample, overall and by groups:
```{r eval = FALSE , results = "hide" }
sum( weights( meps_design , "sampling" ) != 0 )

svyby( ~ one , ~ region20 , meps_design , unwtd.count )
```

### Weighted Counts {-}
Count the weighted size of the generalizable population, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ one , meps_design )

svyby( ~ one , ~ region20 , meps_design , svytotal )
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ totexp20 , meps_design )

svyby( ~ totexp20 , ~ region20 , meps_design , svymean )
```

Calculate the distribution of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ sex , meps_design )

svyby( ~ sex , ~ region20 , meps_design , svymean )
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ totexp20 , meps_design )

svyby( ~ totexp20 , ~ region20 , meps_design , svytotal )
```

Calculate the weighted sum of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ sex , meps_design )

svyby( ~ sex , ~ region20 , meps_design , svytotal )
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svyquantile( ~ totexp20 , meps_design , 0.5 )

svyby( 
	~ totexp20 , 
	~ region20 , 
	meps_design , 
	svyquantile , 
	0.5 ,
	ci = TRUE 
)
```

Estimate a ratio:
```{r eval = FALSE , results = "hide" }
svyratio( 
	numerator = ~ totmcd20 , 
	denominator = ~ totexp20 , 
	meps_design 
)
```

### Subsetting {-}

Restrict the survey design to seniors:
```{r eval = FALSE , results = "hide" }
sub_meps_design <- subset( meps_design , agelast >= 65 )
```
Calculate the mean (average) of this subset:
```{r eval = FALSE , results = "hide" }
svymean( ~ totexp20 , sub_meps_design )
```

### Measures of Uncertainty {-}

Extract the coefficient, standard error, confidence interval, and coefficient of variation from any descriptive statistics function result, overall and by groups:
```{r eval = FALSE , results = "hide" }
this_result <- svymean( ~ totexp20 , meps_design )

coef( this_result )
SE( this_result )
confint( this_result )
cv( this_result )

grouped_result <-
	svyby( 
		~ totexp20 , 
		~ region20 , 
		meps_design , 
		svymean 
	)
	
coef( grouped_result )
SE( grouped_result )
confint( grouped_result )
cv( grouped_result )
```

Calculate the degrees of freedom of any survey design object:
```{r eval = FALSE , results = "hide" }
degf( meps_design )
```

Calculate the complex sample survey-adjusted variance of any statistic:
```{r eval = FALSE , results = "hide" }
svyvar( ~ totexp20 , meps_design )
```

Include the complex sample design effect in the result for a specific statistic:
```{r eval = FALSE , results = "hide" }
# SRS without replacement
svymean( ~ totexp20 , meps_design , deff = TRUE )

# SRS with replacement
svymean( ~ totexp20 , meps_design , deff = "replace" )
```

Compute confidence intervals for proportions using methods that may be more accurate near 0 and 1. See `?svyciprop` for alternatives:
```{r eval = FALSE , results = "hide" }
svyciprop( ~ insured_december_31st , meps_design ,
	method = "likelihood" , na.rm = TRUE )
```

### Regression Models and Tests of Association {-}

Perform a design-based t-test:
```{r eval = FALSE , results = "hide" }
svyttest( totexp20 ~ insured_december_31st , meps_design )
```

Perform a chi-squared test of association for survey data:
```{r eval = FALSE , results = "hide" }
svychisq( 
	~ insured_december_31st + sex , 
	meps_design 
)
```

Perform a survey-weighted generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	svyglm( 
		totexp20 ~ insured_december_31st + sex , 
		meps_design 
	)

summary( glm_result )
```

---

## Intermish {-}

---

## Replication Example {-}

https://meps.ahrq.gov/mepsweb/survey_comp/standard_errors.jsp
```{r eval = FALSE , results = "hide" }
library(foreign)

xport_2002_tf <- tempfile()

xport_2002_url <- "https://meps.ahrq.gov/data_files/pufs/h70ssp.zip"

download.file( xport_2002_url , xport_2002_tf , mode = 'wb' )

unzipped_2002_xport <- unzip( xport_2002_tf , exdir = tempdir() )

meps_2002_df <- read.xport( unzipped_2002_xport )

names( meps_2002_df ) <- tolower( names( meps_2002_df ) )

meps_2002_design <-
	svydesign(
		~ varpsu ,
		strata = ~ varstr ,
		weights = ~ perwt02f ,
		data = meps_2002_df ,
		nest = TRUE
	)
			
result <- svymean( ~ totexp02 , meps_2002_design )
stopifnot( round( coef( result ) , 2 ) == 2813.24 )
stopifnot( round( SE( result ) , 2 ) == 58.99 )

```

---

## Analysis Examples with `srvyr` \ {-}

The R `srvyr` library calculates summary statistics from survey data, such as the mean, total or quantile using [dplyr](https://github.com/tidyverse/dplyr/)-like syntax. [srvyr](https://github.com/gergness/srvyr) allows for the use of many verbs, such as `summarize`, `group_by`, and `mutate`, the convenience of pipe-able functions, the `tidyverse` style of non-standard evaluation and more consistent return types than the `survey` package. [This vignette](https://cran.r-project.org/web/packages/srvyr/vignettes/srvyr-vs-survey.html) details the available features. As a starting point for MEPS users, this code replicates previously-presented examples:

```{r eval = FALSE , results = "hide" }
library(srvyr)
meps_srvyr_design <- as_survey( meps_design )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
meps_srvyr_design %>%
	summarize( mean = survey_mean( totexp20 ) )

meps_srvyr_design %>%
	group_by( region20 ) %>%
	summarize( mean = survey_mean( totexp20 ) )
```


