# Medical Large Claims Experience Study (MLCES) {-}

<a href="https://github.com/asdfree/mlces/actions"><img src="https://github.com/asdfree/mlces/actions/workflows/r.yml/badge.svg" alt="Github Actions Badge"></a>

The best private health insurance claims data available to the public. This data should be used to calibrate other data sets, probably nothing more.

* One table with one row per individual with nonzero total paid charges.

* A convenience sample of group (employer-sponsored) health insurance claims from seven private health insurers in the United States.

* 1997 thru 1999 with no expected updates in the future.

* Provided by the [Society of Actuaries (SOA)](http://www.soa.org/).

---

Please skim before you begin:

1. [Group Medical Insurance Claims Database Collection and Analysis Report](https://www.soa.org/4937d6/globalassets/assets/files/research/exp-study/large_claims_report.pdf)

2. [Claim Severities, Claim Relativities, and Age: Evidence from SOA Group Health Data](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=1412243)

3. This poem

```{r}
# skewed by black swan tails
# means, medians sing adieu
# claims distribution
```

---

## Download, Import, Preparation {-}

Download and import the 1999 medical claims file:

```{r eval = FALSE }
tf <- tempfile()

this_url <-	"https://www.soa.org/Files/Research/1999.zip"

download.file( this_url , tf , mode = 'wb' )

unzipped_files <- unzip( tf , exdir = tempdir() )

mlces_df <- read.csv( unzipped_files )

names( mlces_df ) <- tolower( names( mlces_df ) )
```

---

## Analysis Examples with base R \ {-}

Load a data frame:

```{r messages = FALSE , eval = FALSE }

```

```{r messages = FALSE , eval = FALSE }

```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE }
mlces_df <- 
	transform( 
		mlces_df , 
		
		one = 1 ,
		
		claimant_relationship_to_policyholder =
			ifelse( relation == "E" , "covered employee" ,
			ifelse( relation == "S" , "spouse of covered employee" ,
			ifelse( relation == "D" , "dependent of covered employee" , NA ) ) ) ,
			
		ppo_plan = as.numeric( ppo == 'Y' )
	)
	
```

### Unweighted Counts {-}

Count the unweighted number of records in the table, overall and by groups:
```{r eval = FALSE , results = "hide" }
nrow( mlces_df )

table( mlces_df[ , "claimant_relationship_to_policyholder" ] , useNA = "always" )
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
mean( mlces_df[ , "totpdchg" ] )

tapply(
	mlces_df[ , "totpdchg" ] ,
	mlces_df[ , "claimant_relationship_to_policyholder" ] ,
	mean 
)
```

Calculate the distribution of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
prop.table( table( mlces_df[ , "patsex" ] ) )

prop.table(
	table( mlces_df[ , c( "patsex" , "claimant_relationship_to_policyholder" ) ] ) ,
	margin = 2
)
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
sum( mlces_df[ , "totpdchg" ] )

tapply(
	mlces_df[ , "totpdchg" ] ,
	mlces_df[ , "claimant_relationship_to_policyholder" ] ,
	sum 
)
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
quantile( mlces_df[ , "totpdchg" ] , 0.5 )

tapply(
	mlces_df[ , "totpdchg" ] ,
	mlces_df[ , "claimant_relationship_to_policyholder" ] ,
	quantile ,
	0.5 
)
```

### Subsetting {-}

Limit your `data.frame` to persons under 18:
```{r eval = FALSE , results = "hide" }
sub_mlces_df <- subset( mlces_df , ( ( claimyr - patbrtyr ) < 18 ) )
```
Calculate the mean (average) of this subset:
```{r eval = FALSE , results = "hide" }
mean( sub_mlces_df[ , "totpdchg" ] )
```

### Measures of Uncertainty {-}

Calculate the variance, overall and by groups:
```{r eval = FALSE , results = "hide" }
var( mlces_df[ , "totpdchg" ] )

tapply(
	mlces_df[ , "totpdchg" ] ,
	mlces_df[ , "claimant_relationship_to_policyholder" ] ,
	var 
)
```

### Regression Models and Tests of Association {-}

Perform a t-test:
```{r eval = FALSE , results = "hide" }
t.test( totpdchg ~ ppo_plan , mlces_df )
```

Perform a chi-squared test of association:
```{r eval = FALSE , results = "hide" }
this_table <- table( mlces_df[ , c( "ppo_plan" , "patsex" ) ] )

chisq.test( this_table )
```

Perform a generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	glm( 
		totpdchg ~ ppo_plan + patsex , 
		data = mlces_df
	)

summary( glm_result )
```

---

## Intermish {-}

<center>https://en.wikipedia.org/wiki/Lorem_ipsum</center>

---

## Analysis Examples with `dplyr` \ {-}

The R `dplyr` library offers an alternative grammar of data manipulation to base R and SQL syntax. [dplyr](https://github.com/tidyverse/dplyr/) offers many verbs, such as `summarize`, `group_by`, and `mutate`, the convenience of pipe-able functions, and the `tidyverse` style of non-standard evaluation. [This vignette](https://cran.r-project.org/web/packages/dplyr/vignettes/dplyr.html) details the available features. As a starting point for MLCES users, this code replicates previously-presented examples:

```{r eval = FALSE , results = "hide" }
library(dplyr)
mlces_tbl <- tbl_df( mlces_df )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
mlces_tbl %>%
	summarize( mean = mean( totpdchg ) )

mlces_tbl %>%
	group_by( claimant_relationship_to_policyholder ) %>%
	summarize( mean = mean( totpdchg ) )
```

---

## Replication Example {-}

The example below matches statistics in Table II-A's 1999 row numbers 52 and 53 from the [Tables II – XI 1997–1999 Database](https://www.soa.org/4937cc/globalassets/assets/files/research/tables.zip):

Match Claimants Exceeding Deductible:

```{r eval = FALSE , results = "hide" }
# $0 deductible
stopifnot( nrow( mlces_df ) == 1591738 )

# $1,000 deductible
charges_above_1000 <- subset( mlces_df , totpdchg > 1000 )
stopifnot( nrow( charges_above_1000 ) == 402550 )
```

Match the Excess Charges Above Deductible:

```{r eval = FALSE , results = "hide" }
# $0 deductible
stopifnot( round( sum( mlces_df[ , 'totpdchg' ] ) , 0 ) == 2599356658 )

# $1,000 deductible
stopifnot( round( sum( charges_above_1000[ , 'totpdchg' ] - 1000 ) , 0 ) == 1883768786 )
```
