# National Health Interview Survey (NHIS) {-}

[![License: GPL v3](https://img.shields.io/badge/License-GPLv3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0) <a href="https://github.com/asdfree/nhis/actions"><img src="https://github.com/asdfree/nhis/actions/workflows/r.yml/badge.svg" alt="Github Actions Badge"></a> <a href="https://www.eventbrite.com/e/1972097244947"><img src="https://img.shields.io/badge/Free%20Workshop%20Washington%20DC-December%203rd,%202025-orange" alt="Washington DC: December 3rd, 2025"></a>

America's most detailed household survey of health status and medical experience.

* One table with one row per sampled adult (18+) within each sampled household, one table with one row per sample child (when available, same family not required), multiply-imputed income tables.

* A complex sample survey designed to generalize to the U.S. civilian non-institutionalized population.

* Released annually since 1963, the most recent major re-design in 2019.

* Conducted by the [National Center for Health Statistics](https://www.cdc.gov/nchs/) at the [Centers for Disease Control](http://www.cdc.gov/).

---

## Recommended Reading {-}

Four Example Strengths & Limitations:

✔️ [PRICSSA](https://ftp.cdc.gov/pub/Health_Statistics/NCHS/Dataset_Documentation/NHIS/2024/2024-NHIS-PRICSSA-508.pdf)

✔️ [Annual and rotating core questionnaires allow trend analysis](https://www.cdc.gov/nchs/data/nhis/Sample-Questionnaire-Brochure.pdf#page=3)

❌ [High rate of missingness for family income questions](https://ftp.cdc.gov/pub/Health_Statistics/NCHS/Dataset_Documentation/NHIS/2024/NHIS2024-imputation-techdoc-508.pdf)

❌ [2019 redesign interviews only one adult and one child (if present) per household](https://ftp.cdc.gov/pub/Health_Statistics/NCHS/Survey_Questionnaires/NHIS/2024/advance-letter-2024-508.pdf)

<br>

Three Example Findings:

1. [In 2022, 14% of US seniors met both aerobic and muscle-strengthening physical activity guidelines](https://www.cdc.gov/nchs/data/nhsr/nhsr215.pdf).

2. [Adults living alone in 2021 reported feelings of depression more often than those living with others](https://www.cdc.gov/nchs/data/nhsr/nhsr199.pdf).

3. [Among US adults aged 18+ in 2022, 3% were categorized as having severe anxiety symptoms](https://www.ncbi.nlm.nih.gov/books/NBK609621/).

<br>

Two Methodology Documents:

> [2024 Survey Description](https://ftp.cdc.gov/pub/Health_Statistics/NCHS/Dataset_Documentation/NHIS/2024/srvydesc-508.pdf)

> [Wikipedia Entry](https://en.wikipedia.org/wiki/National_Health_Interview_Survey)

<br>

One Haiku:

```{r}
# excellent health poor
# wealth. "sup, doc?" bugs, daft bills, free
# laughs best medicine
```

---

## Function Definitions {-}

Define a function to download, unzip, and import each comma-separated value file:

```{r eval = FALSE , results = "hide" }
nhis_csv_import <-
	function( this_url ){
		
		this_tf <- tempfile()
		
		download.file( this_url , this_tf , mode = 'wb' )
		
		unzipped_files <- unzip( this_tf , exdir = tempdir() )
		
		this_csv <- grep( '\\.csv$' , unzipped_files , value = TRUE )
		
		this_df <- read.csv( this_csv )
		
		file.remove( c( this_tf , unzipped_files ) )
		
		names( this_df ) <- tolower( names( this_df ) )
		
		this_df
	}

```

---

## Download, Import, Preparation {-}

Download and import the sample adult interview and imputed income files:

```{r eval = FALSE , results = "hide" }
nhis_df <-
	nhis_csv_import( 
		"https://ftp.cdc.gov/pub/Health_Statistics/NCHS/Datasets/NHIS/2024/adult24csv.zip" 
	)

imputed_income_df <- 
	nhis_csv_import( 
		"https://ftp.cdc.gov/pub/Health_Statistics/NCHS/Datasets/NHIS/2024/adultinc24csv.zip" 
	)
```

### Save Locally \ {-}

Save the object at any point:

```{r eval = FALSE , results = "hide" }
# nhis_fn <- file.path( path.expand( "~" ) , "NHIS" , "this_file.rds" )
# saveRDS( nhis_df , file = nhis_fn , compress = FALSE )
```

Load the same object:

```{r eval = FALSE , results = "hide" }
# nhis_df <- readRDS( nhis_fn )
```

### Survey Design Definition {-}
Construct a multiply-imputed, complex sample survey design:

Reshape the imputed income data.frame into a list based on the implicate number:
```{r eval = FALSE , results = "hide" }
imputed_income_list <- split( imputed_income_df , imputed_income_df[ , 'impnum_a' ] )
```

Remove overlapping columns except the merge variable:
```{r eval = FALSE , results = "hide" }
variables_to_remove <-
	setdiff( intersect( names( nhis_df ) , names( imputed_income_df ) ) , 'hhx' )

nhis_df <- nhis_df[ , !( names( nhis_df ) %in% variables_to_remove ) ]
```

Merge each implicate onto the sample adult table:
```{r eval = FALSE , results = "hide" }
nhis_list <-
	lapply( imputed_income_list ,
		function( w ){
			this_df <- merge( nhis_df , w )
			stopifnot( nrow( this_df ) == nrow( nhis_df ) )
			this_df
		} )
```

Define the design:
```{r eval = FALSE , results = "hide" }
library(survey)
library(mitools)

nhis_design <- 
	svydesign( 
		id = ~ ppsu , 
		strata = ~ pstrat ,
		nest = TRUE ,
		weights = ~ wtfa_a ,
		data = imputationList( nhis_list )
	)
```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE , results = "hide" }
nhis_design <- 
	update( 
		nhis_design , 
		
		one = 1 ,
		
		poverty_category =
			factor( 
				findInterval( povrattc_a , c( 1 , 2 , 4 ) ) ,
				labels = 
					c( "below poverty" , "100-199%" , "200-399%" , "400%+" )
			) ,
			
		fair_or_poor_reported_health = 
			ifelse( phstat_a %in% 1:5 , as.numeric( phstat_a >= 4 ) , NA ) ,
			
		sex_a = factor( sex_a , levels = 1:2 , labels = c( "male" , "female" ) ) ,
		
		annual_premium_first_plan = ifelse( hicostr1_a > 40000 , NA , hicostr1_a )

	)
```

---

## Analysis Examples with the `survey` library \ {-}

### Unweighted Counts {-}

Count the unweighted number of records in the survey sample, overall and by groups:
```{r eval = FALSE , results = "hide" }
MIcombine( with( nhis_design , svyby( ~ one , ~ one , unwtd.count ) ) )

MIcombine( with( nhis_design , svyby( ~ one , ~ poverty_category , unwtd.count ) ) )
```

### Weighted Counts {-}
Count the weighted size of the generalizable population, overall and by groups:
```{r eval = FALSE , results = "hide" }
MIcombine( with( nhis_design , svytotal( ~ one ) ) )

MIcombine( with( nhis_design ,
	svyby( ~ one , ~ poverty_category , svytotal )
) )
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
MIcombine( with( nhis_design , svymean( ~ agep_a ) ) )

MIcombine( with( nhis_design ,
	svyby( ~ agep_a , ~ poverty_category , svymean )
) )
```

Calculate the distribution of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
MIcombine( with( nhis_design , svymean( ~ sex_a ) ) )

MIcombine( with( nhis_design ,
	svyby( ~ sex_a , ~ poverty_category , svymean )
) )
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
MIcombine( with( nhis_design , svytotal( ~ agep_a ) ) )

MIcombine( with( nhis_design ,
	svyby( ~ agep_a , ~ poverty_category , svytotal )
) )
```

Calculate the weighted sum of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
MIcombine( with( nhis_design , svytotal( ~ sex_a ) ) )

MIcombine( with( nhis_design ,
	svyby( ~ sex_a , ~ poverty_category , svytotal )
) )
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
MIcombine( with( nhis_design ,
	svyquantile(
		~ agep_a ,
		0.5 , se = TRUE 
) ) )

MIcombine( with( nhis_design ,
	svyby(
		~ agep_a , ~ poverty_category , svyquantile ,
		0.5 , se = TRUE ,
		ci = TRUE 
) ) )
```

Estimate a ratio:
```{r eval = FALSE , results = "hide" }
MIcombine( with( nhis_design ,
	svyratio( numerator = ~ annual_premium_first_plan , denominator = ~ agep_a , na.rm = TRUE )
) )
```

### Subsetting {-}

Restrict the survey design to uninsured:
```{r eval = FALSE , results = "hide" }
sub_nhis_design <- subset( nhis_design , notcov_a == 1 )
```
Calculate the mean (average) of this subset:
```{r eval = FALSE , results = "hide" }
MIcombine( with( sub_nhis_design , svymean( ~ agep_a ) ) )
```

### Measures of Uncertainty {-}

Extract the coefficient, standard error, confidence interval, and coefficient of variation from any descriptive statistics function result, overall and by groups:
```{r eval = FALSE , results = "hide" }
this_result <-
	MIcombine( with( nhis_design ,
		svymean( ~ agep_a )
	) )

coef( this_result )
SE( this_result )
confint( this_result )
cv( this_result )

grouped_result <-
	MIcombine( with( nhis_design ,
		svyby( ~ agep_a , ~ poverty_category , svymean )
	) )

coef( grouped_result )
SE( grouped_result )
confint( grouped_result )
cv( grouped_result )
```

Calculate the degrees of freedom of any survey design object:
```{r eval = FALSE , results = "hide" }
degf( nhis_design$designs[[1]] )
```

Calculate the complex sample survey-adjusted variance of any statistic:
```{r eval = FALSE , results = "hide" }
MIcombine( with( nhis_design , svyvar( ~ agep_a ) ) )
```

Include the complex sample design effect in the result for a specific statistic:
```{r eval = FALSE , results = "hide" }
# SRS without replacement
MIcombine( with( nhis_design ,
	svymean( ~ agep_a , deff = TRUE )
) )

# SRS with replacement
MIcombine( with( nhis_design ,
	svymean( ~ agep_a , deff = "replace" )
) )
```

Compute confidence intervals for proportions using methods that may be more accurate near 0 and 1. See `?svyciprop` for alternatives:
```{r eval = FALSE , results = "hide" }
# MIsvyciprop( ~ fair_or_poor_reported_health , nhis_design ,
# 	method = "likelihood" , na.rm = TRUE )
```

### Regression Models and Tests of Association {-}

Perform a design-based t-test:
```{r eval = FALSE , results = "hide" }
# MIsvyttest( agep_a ~ fair_or_poor_reported_health , nhis_design )
```

Perform a chi-squared test of association for survey data:
```{r eval = FALSE , results = "hide" }
# MIsvychisq( ~ fair_or_poor_reported_health + sex_a , nhis_design )
```

Perform a survey-weighted generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	MIcombine( with( nhis_design ,
		svyglm( agep_a ~ fair_or_poor_reported_health + sex_a )
	) )
	
summary( glm_result )
```

---

## Replication Example {-}

This example matches statistics and standard errors from PDF page 34 of [Multiple Imputation of Family Income in 2024 National Health
Interview Survey](https://ftp.cdc.gov/pub/Health_Statistics/NCHS/Dataset_Documentation/NHIS/2024/NHIS2024-imputation-techdoc-508.pdf#page=34):

```{r eval = FALSE , results = "hide" }
results <-
	MIcombine( 
		with(
			subset( nhis_design , notcov_a %in% 1:2 ) , 
			svyby(
				~ as.numeric( notcov_a == 1 ) , 
				~ poverty_category , 
				svymean , 
				na.rm = TRUE 
			) 
		)
	)

stopifnot(
	all(
		as.numeric( round( coef( results ) , 4 ) ) == c( 0.1776 , 0.1693 , 0.0994 , 0.0333 )
	) 
)

stopifnot(
	all( 
		as.numeric( round( SE( results ) , 4 ) ) == c( 0.0096 , 0.0075 , 0.0044 , 0.0022 )
	) 
)

```


