# National Household Travel Survey (NHTS) {-}

<a href="https://github.com/asdfree/nhts/actions"><img src="https://github.com/asdfree/nhts/actions/workflows/r.yml/badge.svg" alt="Github Actions Badge"></a>

The authoritative source on travel behavior, the Federal Highway Administration records all modes of transport and the characteristics of both people and vehicles.

* 

* 

* 

* 

---

Please skim before you begin:

1. [2017 NHTS Data User Guide](https://nhts.ornl.gov/assets/NHTS2017_UsersGuide_04232019_1.pdf)

2. [2017 NHTS Weighting Report](https://nhts.ornl.gov/assets/2017%20NHTS%20Weighting%20Report.pdf)

3. This human-composed haiku or a bouquet of [artificial intelligence-generated limericks](https://www.gnod.com/search/ai#q=write%20a%20limerick-style%20poem%20about%20the%20National Household Travel Survey)

```{r}
# commuter patterns,
# truckin'. what a long strange trip
# who went when where why
```

---

## Download, Import, Preparation {-}

```{r messages = FALSE , eval = FALSE }
library(haven)

nhts_download_unzip <-
	function( this_url ){
	
		tf <- tempfile()

		download.file( this_url , tf , mode = 'wb' )

		unzip( tf , exdir = tempdir() )
	}
	
unzipped_survey_data <- nhts_download_unzip( "https://nhts.ornl.gov/assets/2016/download/sas.zip" )

unzipped_replicate_weights <- nhts_download_unzip( "https://nhts.ornl.gov/assets/2016/download/Replicates.zip" )

unzipped_trip_chains <- nhts_download_unzip( "https://nhts.ornl.gov/assets/2016/download/TripChain/TripChain17.zip" )
```

```{r messages = FALSE , eval = FALSE }
hhpub_tbl <- read_sas( grep( "hhpub\\.sas7bdat$" , unzipped_survey_data , value = TRUE ) )
perpub_tbl <- read_sas( grep( "perpub\\.sas7bdat$" , unzipped_survey_data , value = TRUE ) )
trippub_tbl <- read_sas( grep( "trippub\\.sas7bdat$" , unzipped_survey_data , value = TRUE ) )
vehpub_tbl <- read_sas( grep( "vehpub\\.sas7bdat$" , unzipped_survey_data , value = TRUE ) )

hhwgt_tbl <- read_sas( grep( "hhwgt\\.sas7bdat$" , unzipped_replicate_weights , value = TRUE ) )
perwgt_tbl <- read_sas( grep( "perwgt\\.sas7bdat$" , unzipped_replicate_weights , value = TRUE ) )

hhpub_df <- data.frame( hhpub_tbl )
perpub_df <- data.frame( perpub_tbl )
trippub_df <- data.frame( trippub_tbl )
vehpub_df <- data.frame( vehpub_tbl )

hhwgt_df <- data.frame( hhwgt_tbl )
perwgt_df <- data.frame( perwgt_tbl )

names( hhpub_df ) <- tolower( names( hhpub_df ) )
names( perpub_df ) <- tolower( names( perpub_df ) )
names( trippub_df ) <- tolower( names( trippub_df ) )
names( vehpub_df ) <- tolower( names( vehpub_df ) )

names( hhwgt_df ) <- tolower( names( hhwgt_df ) )
names( perwgt_df ) <- tolower( names( perwgt_df ) )
```	

```{r messages = FALSE , eval = FALSE }
hhpub_df[ , 'one' ] <- 1

perpub_df[ , 'one' ] <- 1

trippub_df[ , 'one' ] <- 1
```

```{r messages = FALSE , eval = FALSE }

trippub_df[ !( trippub_df[ , 'trpmiles' ] %in% -9 ) , 'tripmiles_no_nines' ] <-
	trippub_df[ !( trippub_df[ , 'trpmiles' ] %in% -9 ) , 'trpmiles' ]

# total trips
trips_per_person <- with( trippub_df , aggregate( cbind( one , tripmiles_no_nines ) , list( houseid , personid ) , sum , na.rm = TRUE ) )
names( trips_per_person ) <- c( 'houseid' , 'personid' , 'trips_per_person' , 'miles_per_person' )

nhts_df <- merge( perpub_df , trips_per_person , all.x = TRUE )
nhts_df[ is.na( nhts_df[ , 'trips_per_person' ] ) , 'trips_per_person' ] <- 0
nhts_df[ is.na( nhts_df[ , 'miles_per_person' ] ) , 'miles_per_person' ] <- 0

# walking trips

walks_per_person <- with( subset( trippub_df , trptrans == '01' ) , aggregate( cbind( one , tripmiles_no_nines ) , list( houseid , personid ) , sum , na.rm = TRUE ) )
names( walks_per_person ) <- c( 'houseid' , 'personid' , 'walks_per_person' , 'walking_miles_per_person' )

nhts_df <- merge( nhts_df , walks_per_person , all.x = TRUE )
nhts_df[ is.na( nhts_df[ , 'walks_per_person' ] ) , 'walks_per_person' ] <- 0
nhts_df[ is.na( nhts_df[ , 'walking_miles_per_person' ] ) , 'walking_miles_per_person' ] <- 0

stopifnot( nrow( nhts_df ) == nrow( perpub_df ) )

```

### Save locally \ {-}

Save the object at any point:

```{r eval = FALSE , results = "hide" }
# nhts_fn <- file.path( path.expand( "~" ) , "NHTS" , "this_file.rds" )
# saveRDS( nhts_df , file = nhts_fn , compress = FALSE )
```

Load the same object:

```{r eval = FALSE , results = "hide" }
# nhts_df <- readRDS( nhts_fn )
```

### Survey Design Definition {-}
Construct a complex sample survey design:

```{r messages = FALSE , eval = FALSE }
library(survey)

hhpub_df <- hhpub_df[ order( hhpub_df[ , 'houseid' ] ) , ]
hhwgt_df <- hhwgt_df[ order( hhwgt_df[ , 'houseid' ] ) , ]

hh_design <-
	svrepdesign(
		weight = ~ wthhfin ,
		repweights = hhwgt_df[ grep( 'wthhfin[0-9]' , names( hhwgt_df ) , value = TRUE ) ] ,
		scale = 6 / 7 ,
		rscales = 1 ,
		# degf = 98 ,
		type = 'JK1' ,
		mse = TRUE ,
		data = hhpub_df
	)

# matches overall
svytotal( ~ one , hh_design )
# https://nhts.ornl.gov/tables09/ae/work/Job160511.html
```

```{r messages = FALSE , eval = FALSE }
nhts_df <- nhts_df[ do.call( order , nhts_df[ , c( 'houseid' , 'personid' ) ] ) , ]
perwgt_df <- perwgt_df[ do.call( order , perwgt_df[ , c( 'houseid' , 'personid' ) ] ) , ]

nhts_design <-
	svrepdesign(
		weight = ~ wtperfin ,
		repweights = perwgt_df[ grep( 'wtperfin[0-9]' , names( perwgt_df ) , value = TRUE ) ] ,
		scale = 6 / 7 ,
		rscales = rep( 1 , 98 ) ,
		# degf = 98 ,
		type = 'JK1' ,
		# mse = TRUE ,
		data = nhts_df
	)
```

```{r messages = FALSE , eval = FALSE }

trippub_df <- trippub_df[ do.call( order , trippub_df[ , c( 'houseid' , 'personid' ) ] ) , ]
perwgt_df <- perwgt_df[ do.call( order , perwgt_df[ , c( 'houseid' , 'personid' ) ] ) , ]

trip_design <-
	svrepdesign(
		weight = ~ wttrdfin ,
		repweights = perwgt_df[ grep( 'wttrdfin[0-9]' , names( perwgt_df ) , value = TRUE ) ] ,
		scale = 6 / 7 ,
		rscales = 1 ,
		# degf = 99 ,
		type = 'JK1' ,
		mse = TRUE ,
		data = trippub_df
	)
```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE , results = "hide" }
hh_design <-
	update(
		hh_design ,
		hhsize_categories =
			factor(
				findInterval( hhsize , 1:4 ) ,
				levels = 1:4 ,
				labels = c( 1:3 , '4 or more' )
			)
	)
	

nhts_design <- 
	update( 
		nhts_design , 
		
		urban_area = as.numeric( urbrur == '01' )
		
	)
```

---

## Analysis Examples with the `survey` library \ {-}

### Unweighted Counts {-}

Count the unweighted number of records in the survey sample, overall and by groups:
```{r eval = FALSE , results = "hide" }
sum( weights( nhts_design , "sampling" ) != 0 )

svyby( ~ one , ~ r_sex_imp , nhts_design , unwtd.count )
```

### Weighted Counts {-}
Count the weighted size of the generalizable population, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ one , nhts_design )

svyby( ~ one , ~ r_sex_imp , nhts_design , svytotal )
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ miles_per_person , nhts_design )

svyby( ~ miles_per_person , ~ r_sex_imp , nhts_design , svymean )
```

Calculate the distribution of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ hhstate , nhts_design )

svyby( ~ hhstate , ~ r_sex_imp , nhts_design , svymean )
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ miles_per_person , nhts_design )

svyby( ~ miles_per_person , ~ r_sex_imp , nhts_design , svytotal )
```

Calculate the weighted sum of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ hhstate , nhts_design )

svyby( ~ hhstate , ~ r_sex_imp , nhts_design , svytotal )
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svyquantile( ~ miles_per_person , nhts_design , 0.5 )

svyby( 
	~ miles_per_person , 
	~ r_sex_imp , 
	nhts_design , 
	svyquantile , 
	0.5 ,
	ci = TRUE 
)
```

Estimate a ratio:
```{r eval = FALSE , results = "hide" }
svyratio( 
	numerator = ~ walking_miles_per_person , 
	denominator = ~ miles_per_person , 
	nhts_design 
)
```

### Subsetting {-}

Restrict the survey design to ever cyclists:
```{r eval = FALSE , results = "hide" }
sub_nhts_design <- subset( nhts_design , nbiketrp > 0 )
```
Calculate the mean (average) of this subset:
```{r eval = FALSE , results = "hide" }
svymean( ~ miles_per_person , sub_nhts_design )
```

### Measures of Uncertainty {-}

Extract the coefficient, standard error, confidence interval, and coefficient of variation from any descriptive statistics function result, overall and by groups:
```{r eval = FALSE , results = "hide" }
this_result <- svymean( ~ miles_per_person , nhts_design )

coef( this_result )
SE( this_result )
confint( this_result )
cv( this_result )

grouped_result <-
	svyby( 
		~ miles_per_person , 
		~ r_sex_imp , 
		nhts_design , 
		svymean 
	)
	
coef( grouped_result )
SE( grouped_result )
confint( grouped_result )
cv( grouped_result )
```

Calculate the degrees of freedom of any survey design object:
```{r eval = FALSE , results = "hide" }
degf( nhts_design )
```

Calculate the complex sample survey-adjusted variance of any statistic:
```{r eval = FALSE , results = "hide" }
svyvar( ~ miles_per_person , nhts_design )
```

Include the complex sample design effect in the result for a specific statistic:
```{r eval = FALSE , results = "hide" }
# SRS without replacement
svymean( ~ miles_per_person , nhts_design , deff = TRUE )

# SRS with replacement
svymean( ~ miles_per_person , nhts_design , deff = "replace" )
```

Compute confidence intervals for proportions using methods that may be more accurate near 0 and 1. See `?svyciprop` for alternatives:
```{r eval = FALSE , results = "hide" }
svyciprop( ~ urban_area , nhts_design ,
	method = "likelihood" )
```

### Regression Models and Tests of Association {-}

Perform a design-based t-test:
```{r eval = FALSE , results = "hide" }
svyttest( miles_per_person ~ urban_area , nhts_design )
```

Perform a chi-squared test of association for survey data:
```{r eval = FALSE , results = "hide" }
svychisq( 
	~ urban_area + hhstate , 
	nhts_design 
)
```

Perform a survey-weighted generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	svyglm( 
		miles_per_person ~ urban_area + hhstate , 
		nhts_design 
	)

summary( glm_result )
```

---

## Intermish {-}

<center><i><b>the author as interstate highway system</b></i></center>

---

## Replication Example {-}

This example matches...

```{r messages = FALSE , eval = FALSE }

# https://nhts.ornl.gov/assets/2017_nhts_summary_travel_trends.pdf#page=12
hhsize_counts <- svytotal( ~ hhsize_categories , hh_design )
stopifnot( all( round( coef( hhsize_counts ) / 1000 , 0 ) == c( 32952 , 40056 , 18521 , 26679 ) ) )

hhsize_moe <- 
	confint( hhsize_counts , df = ncol( hh_design$repweights ) )[ , 2 ] - coef( hhsize_counts )

stopifnot( all( round( hhsize_moe / 1000 , 0 ) == c( 0 , 0 , 97 , 97 ) ) )	

# https://nhts.ornl.gov/assets/2017_nhts_summary_travel_trends.pdf#page=2
# westat author of this workshop
# https://rawgit.com/Westat-Transportation/summarizeNHTS/master/inst/tutorials/workshop/Workshop.html#(38)
# matches
unwtd_n <- with(nhts_df,tapply(trips_per_person,worker,sum))
stopifnot( all( unwtd_n == c( 79295 , 28 , 497944 , 346305 ) ) )
surveyed_n <- with(nhts_df,tapply(trips_per_person,worker,mean))
stopifnot( all( round( surveyed_n , 2 ) == c( 2.84 , 1.65 , 3.88 , 3.21 ) ) )

# coefficients match
a <- svyby(~trips_per_person,~worker,nhts_design,svymean)
stopifnot( round( coef( a ) , 2 ) == c( 2.78 , 1.28 , 3.77 , 3.01 ) )
# margin of errors match
stopifnot( all( round(confint(a,df=ncol( nhts_design$repweights ) )[,2]-coef(a),2) == c( 0.06 , 2.21 , 0.03 , 0.06 ) ) )

```


