# National Household Travel Survey (NHTS) {-}

[![License: GPL v3](https://img.shields.io/badge/License-GPLv3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0) <a href="https://github.com/asdfree/nhts/actions"><img src="https://github.com/asdfree/nhts/actions/workflows/r.yml/badge.svg" alt="Github Actions Badge"></a> <a href="https://www.eventbrite.com/e/1634269818019"><img src="https://img.shields.io/badge/Free%20R%20Training%20Washington%20DC-October%2015th,%202025-yellow" alt="Washington DC: October 15th, 2025"></a> <a href="https://www.eventbrite.com/e/1458850233319"><img src="https://img.shields.io/badge/Free%20Workshop%20New%20York%20NY-October%2022nd,%202025-orange" alt="New York NY: October 22nd, 2025"></a> <a href="https://www.eventbrite.com/e/1560552678069"><img src="https://img.shields.io/badge/Free%20Workshop%20Philadelphia%20PA-October%2029th,%202025-orange" alt="Philadelphia PA: October 29th, 2025"></a> <a href="https://www.eventbrite.com/e/1710882017169"><img src="https://img.shields.io/badge/Free%20Workshop%20Washington%20DC-November%2012th,%202025-orange" alt="Washington DC: November 12th, 2025"></a> <a href="https://www.eventbrite.com/e/1771826874879"><img src="https://img.shields.io/badge/Free%20R%20Training%20Washington%20DC-November%2019th,%202025-yellow" alt="Washington DC: November 19th, 2025"></a>

The authoritative source on travel behavior, recording characteristics of people and vehicles of all modes.

* Four core linkable tables, with one record per household, person, trip, and vehicle, respectively.

* A complex sample survey designed to generalize to the civilian non-institutional U.S. population.

* Released every five to eight years since 1969.

* Funded by the [Federal Highway Administration](https://highways.dot.gov/), with data collected by [Ipsos Public Affairs](https://www.ipsos.com/).

---

## Recommended Reading {-}

Four Example Strengths & Limitations:

✔️ [Origin-Destination passively collected data complement traditional household survey](https://mti.umd.edu/sites/mti.umd.edu/files/Collecting_Multimodal_National_and_Metropolitan_Behavior_Data%20Report1_updJuly2019.pdf#page=69)

✔️ [Sample supports analysis of metro areas within census divisions](https://nhts.ornl.gov/assets/2022/doc/2022%20NextGen%20NHTS%20Derived%20Variables-PubUse.pdf#page=2)

❌ [2022 redesign uses retrospective recorded travel day (1 day prior) rather than travel log](https://nhts.ornl.gov/assets/2022/doc/2022%20NextGen%20NHTS%20Annotated%20Survey.pdf#page=7)

❌ [Long-distance trip questions do not estimate respondent's annual behavior or volume](https://nhts.ornl.gov/assets/NextGen%20NHTS_State%20of%20Practice_032423.pdf#page=9)

<br>

Three Example Findings:

1. [Online-purchased home deliveries grew over 2017-2022, from 2.5 to 5.4 per person per month](https://nhts.ornl.gov/assets/NextGen%20NHTS_Newsletter_Issue5_071524.pdf#page=5).

2. [In 2022, 53% of K-12 students were dropped off at school in a private vehicle or drove themselves](https://www.washingtonpost.com/business/2024/02/02/school-bus-era-ends/).

3. [Nearly 9 in 10 US households had a vehicle available to drive in 2022](https://nhts.ornl.gov/assets/2022/pub/2022_NHTS_Summary_Travel_Trends.pdf#page=49).

<br>

Two Methodology Documents:

> [2022 NHTS Data User Guide](https://nhts.ornl.gov/assets/2022/doc/2022%20NextGen%20NHTS%20User's%20Guide%20V2_PubUse.pdf)

> [2022 NHTS Weighting Memo](https://nhts.ornl.gov/assets/2022/doc/2022%20NextGen%20NHTS%20Weighting%20Memo.pdf)

<br>

One Haiku:

```{r}
# commuter patterns,
# truckin'. what a long strange trip
# who went when where why
```

---

## Download, Import, Preparation {-}

Download and unzip each the 2022 files:
```{r eval = FALSE , results = "hide" }
library(haven)

tf <- tempfile()

download.file( "https://nhts.ornl.gov/assets/2022/download/sas.zip" , tf , mode = 'wb' )

unzipped_files <- unzip( tf , exdir = tempdir() )
```

Import the tables containing one record per household, person, trip, and vehicle:
```{r eval = FALSE , results = "hide" }
nhts_import <-
	function( this_prefix , this_unzip ){
		
		this_sas7bdat <-
			grep( 
				paste0( this_prefix , "\\.sas7bdat$" ) , 
				this_unzip , 
				value = TRUE 
			)
		
		this_tbl <- read_sas( this_sas7bdat )
		
		this_df <- data.frame( this_tbl )
		
		names( this_df ) <- tolower( names( this_df ) )
		
		this_df
	}
	
hhpub_df <- nhts_import( "hhv2pub" , unzipped_files )
perpub_df <- nhts_import( "perv2pub" , unzipped_files )
trippub_df <- nhts_import( "tripv2pub" , unzipped_files )
vehpub_df <- nhts_import( "vehv2pub" , unzipped_files )
```	

Add a column of ones to three of those tables, then a column of non-missing mileage to the trips table:
```{r eval = FALSE , results = "hide" }
hhpub_df[ , 'one' ] <- 1

perpub_df[ , 'one' ] <- 1

trippub_df[ , 'one' ] <- 1

trippub_df[ !( trippub_df[ , 'trpmiles' ] %in% -9 ) , 'wtd_tripmiles_no_nines' ] <-
	trippub_df[ !( trippub_df[ , 'trpmiles' ] %in% -9 ) , 'trpmiles' ] *
	trippub_df[ !( trippub_df[ , 'trpmiles' ] %in% -9 ) , 'wttrdfin' ]
```

Sum the total trip count and mileage to the person-level, both overall and restricted to walking only:
```{r eval = FALSE , results = "hide" }
trips_per_person <- 
	with( 
		trippub_df , 
		aggregate( 
			cbind( wttrdfin , wtd_tripmiles_no_nines ) , 
			list( houseid , personid ) , 
			sum , 
			na.rm = TRUE 
		) 
	)

names( trips_per_person ) <-
	c( 'houseid' , 'personid' , 'wtd_trips' , 'wtd_miles' )

walks_per_person <- 
	with( 
		subset( trippub_df , trptrans == '20' ) , 
		aggregate( 
			cbind( wttrdfin , wtd_tripmiles_no_nines ) , 
			list( houseid , personid ) , 
			sum , 
			na.rm = TRUE 
		) 
	)

names( walks_per_person ) <-
	c( 'houseid' , 'personid' , 'wtd_walks' , 'wtd_walk_miles' )
```

Merge these trip count and mileage values on to the person-level file, replacing non-matches with zero:
```{r eval = FALSE , results = "hide" }
nhts_df <- merge( perpub_df , trips_per_person , all.x = TRUE )

nhts_df <- merge( nhts_df , walks_per_person , all.x = TRUE )

for( this_variable in c( 'wtd_trips' , 'wtd_miles' , 'wtd_walks' , 'wtd_walk_miles' ) ){
	nhts_df[ is.na( nhts_df[ , this_variable ] ) , this_variable ] <- 0
}

stopifnot( nrow( nhts_df ) == nrow( perpub_df ) )
```

### Save Locally \ {-}

Save the object at any point:

```{r eval = FALSE , results = "hide" }
# nhts_fn <- file.path( path.expand( "~" ) , "NHTS" , "this_file.rds" )
# saveRDS( nhts_df , file = nhts_fn , compress = FALSE )
```

Load the same object:

```{r eval = FALSE , results = "hide" }
# nhts_df <- readRDS( nhts_fn )
```

### Survey Design Definition {-}
Construct a complex sample survey design:

Define household-level, person-level, and trip-level designs:
```{r eval = FALSE , results = "hide" }
library(survey)

hh_design <-
	svydesign( 
		id = ~ houseid , 
		strata = ~ stratumid , 
		data = hhpub_df , 
		weights = ~ wthhfin
	)

nhts_design <-
	svydesign(
		id = ~ houseid ,
		strata = ~ stratumid ,
		data = nhts_df ,
		weights = ~ wtperfin
	)

trip_design <-
	svydesign(
		id = ~ houseid ,
		strata = ~ stratumid ,
		data = trippub_df ,
		weights = ~ wttrdfin
	)
```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE , results = "hide" }
hh_design <-
	update(
		hh_design ,
		hhsize_categories =
			factor(
				findInterval( hhsize , 1:4 ) ,
				levels = 1:4 ,
				labels = c( 1:3 , '4 or more' )
			)
	)
	

nhts_design <- 
	update( 
		nhts_design , 
		
		urban_area = as.numeric( urbrur == '01' ) ,
		
		daily_person_trips = ( wtd_trips / ( 365 * wtperfin ) ) ,
		
		daily_person_miles_of_travel = ( wtd_miles / ( 365 * wtperfin ) ) ,
		
		daily_person_walks = ( wtd_walks / ( 365 * wtperfin ) ) ,
		
		daily_person_walk_miles_of_travel = ( wtd_walk_miles / ( 365 * wtperfin ) ) ,
		
		work_status = 
			factor( 
				as.numeric( worker ) , 
				levels = 2:1 , 
				labels = c( 'non-worker' , 'worker' ) 
			)

	)
```

---

## Analysis Examples with the `survey` library \ {-}

### Unweighted Counts {-}

Count the unweighted number of records in the survey sample, overall and by groups:
```{r eval = FALSE , results = "hide" }
sum( weights( nhts_design , "sampling" ) != 0 )

svyby( ~ one , ~ r_sex_imp , nhts_design , unwtd.count )
```

### Weighted Counts {-}
Count the weighted size of the generalizable population, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ one , nhts_design )

svyby( ~ one , ~ r_sex_imp , nhts_design , svytotal )
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ daily_person_walks , nhts_design )

svyby( ~ daily_person_walks , ~ r_sex_imp , nhts_design , svymean )
```

Calculate the distribution of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ work_status , nhts_design , na.rm = TRUE )

svyby( ~ work_status , ~ r_sex_imp , nhts_design , svymean , na.rm = TRUE )
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ daily_person_walks , nhts_design )

svyby( ~ daily_person_walks , ~ r_sex_imp , nhts_design , svytotal )
```

Calculate the weighted sum of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ work_status , nhts_design , na.rm = TRUE )

svyby( ~ work_status , ~ r_sex_imp , nhts_design , svytotal , na.rm = TRUE )
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svyquantile( ~ daily_person_walks , nhts_design , 0.5 )

svyby( 
	~ daily_person_walks , 
	~ r_sex_imp , 
	nhts_design , 
	svyquantile , 
	0.5 ,
	ci = TRUE 
)
```

Estimate a ratio:
```{r eval = FALSE , results = "hide" }
svyratio( 
	numerator = ~ daily_person_walk_miles_of_travel , 
	denominator = ~ daily_person_miles_of_travel , 
	nhts_design 
)
```

### Subsetting {-}

Restrict the survey design to individuals who have used a bicycle in last 30 days:
```{r eval = FALSE , results = "hide" }
sub_nhts_design <- subset( nhts_design , last30_bike == '01' )
```
Calculate the mean (average) of this subset:
```{r eval = FALSE , results = "hide" }
svymean( ~ daily_person_walks , sub_nhts_design )
```

### Measures of Uncertainty {-}

Extract the coefficient, standard error, confidence interval, and coefficient of variation from any descriptive statistics function result, overall and by groups:
```{r eval = FALSE , results = "hide" }
this_result <- svymean( ~ daily_person_walks , nhts_design )

coef( this_result )
SE( this_result )
confint( this_result )
cv( this_result )

grouped_result <-
	svyby( 
		~ daily_person_walks , 
		~ r_sex_imp , 
		nhts_design , 
		svymean 
	)
	
coef( grouped_result )
SE( grouped_result )
confint( grouped_result )
cv( grouped_result )
```

Calculate the degrees of freedom of any survey design object:
```{r eval = FALSE , results = "hide" }
degf( nhts_design )
```

Calculate the complex sample survey-adjusted variance of any statistic:
```{r eval = FALSE , results = "hide" }
svyvar( ~ daily_person_walks , nhts_design )
```

Include the complex sample design effect in the result for a specific statistic:
```{r eval = FALSE , results = "hide" }
# SRS without replacement
svymean( ~ daily_person_walks , nhts_design , deff = TRUE )

# SRS with replacement
svymean( ~ daily_person_walks , nhts_design , deff = "replace" )
```

Compute confidence intervals for proportions using methods that may be more accurate near 0 and 1. See `?svyciprop` for alternatives:
```{r eval = FALSE , results = "hide" }
svyciprop( ~ urban_area , nhts_design ,
	method = "likelihood" )
```

### Regression Models and Tests of Association {-}

Perform a design-based t-test:
```{r eval = FALSE , results = "hide" }
svyttest( daily_person_walks ~ urban_area , nhts_design )
```

Perform a chi-squared test of association for survey data:
```{r eval = FALSE , results = "hide" }
svychisq( 
	~ urban_area + work_status , 
	nhts_design 
)
```

Perform a survey-weighted generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	svyglm( 
		daily_person_walks ~ urban_area + work_status , 
		nhts_design 
	)

summary( glm_result )
```

---

## Replication Example {-}

This example matches the 2022 Household Size counts from [Table 2-1](https://nhts.ornl.gov/assets/2022/pub/2022_NHTS_Summary_Travel_Trends.pdf#page=18):

```{r eval = FALSE , results = "hide" }
hhsize_counts <- svytotal( ~ hhsize_categories , hh_design )

stopifnot(
	all( round( coef( hhsize_counts ) / 1000 , 0 ) == c( 36409 , 44751 , 19001 , 27384 ) )
)

hhsize_ci <- confint( hhsize_counts )

hhsize_moe <- hhsize_ci[ , 2 ] - coef( hhsize_counts )

stopifnot( all( round( hhsize_moe / 1000 , 0 ) == c( 1807 , 1760 , 1448 , 1742 ) ) )	
```

This example matches the 2022 Average Daily Person Trips per Person from [Table 2-9](https://nhts.ornl.gov/assets/2022/pub/2022_NHTS_Summary_Travel_Trends.pdf#page=23):

```{r eval = FALSE , results = "hide" }
this_mean <- svymean( ~ daily_person_trips , nhts_design )

stopifnot( round( coef( this_mean ) , 2 ) == 2.28 )

this_ci <- confint( this_mean )

this_moe <- this_ci[ , 2 ] - coef( this_mean )

stopifnot( round( this_moe , 2 ) == 0.06 )

```

This example matches the 2022 Average Daily PMT per Person from [Table 2-9](https://nhts.ornl.gov/assets/2022/pub/2022_NHTS_Summary_Travel_Trends.pdf#page=23):

```{r eval = FALSE , results = "hide" }
this_mean <- svymean( ~ daily_person_miles_of_travel , nhts_design )

stopifnot( round( coef( this_mean ) , 2 ) == 28.55 )

this_ci <- confint( this_mean )

this_moe <- this_ci[ , 2 ] - coef( this_mean )

stopifnot( round( this_moe , 2 ) == 2.39 )
```

This example matches the 2022 Average Person Trip Length (Miles) from [Table 2-9](https://nhts.ornl.gov/assets/2022/pub/2022_NHTS_Summary_Travel_Trends.pdf#page=23):

```{r eval = FALSE , results = "hide" }
this_mean <- svymean( ~ trpmiles , subset( trip_design , trpmiles >= 0 ) )

stopifnot( round( coef( this_mean ) , 2 ) == 12.56 )

this_ci <- confint( this_mean )

this_moe <- this_ci[ , 2 ] - coef( this_mean )

stopifnot( round( this_moe , 2 ) == 1.04 )

```

---

## Analysis Examples with `srvyr` \ {-}

The R `srvyr` library calculates summary statistics from survey data, such as the mean, total or quantile using [dplyr](https://github.com/tidyverse/dplyr/)-like syntax. [srvyr](https://github.com/gergness/srvyr) allows for the use of many verbs, such as `summarize`, `group_by`, and `mutate`, the convenience of pipe-able functions, the `tidyverse` style of non-standard evaluation and more consistent return types than the `survey` package. [This vignette](https://cran.r-project.org/web/packages/srvyr/vignettes/srvyr-vs-survey.html) details the available features. As a starting point for NHTS users, this code replicates previously-presented examples:

```{r eval = FALSE , results = "hide" }
library(srvyr)
nhts_srvyr_design <- as_survey( nhts_design )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
nhts_srvyr_design %>%
	summarize( mean = survey_mean( daily_person_walks ) )

nhts_srvyr_design %>%
	group_by( r_sex_imp ) %>%
	summarize( mean = survey_mean( daily_person_walks ) )
```


