# National Survey on Drug Use and Health (NSDUH) {-}

[![License: GPL v3](https://img.shields.io/badge/License-GPLv3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0) <a href="https://github.com/asdfree/nsduh/actions"><img src="https://github.com/asdfree/nsduh/actions/workflows/r.yml/badge.svg" alt="Github Actions Badge"></a> <a href="https://www.eventbrite.com/e/1479539405189"><img src="https://img.shields.io/badge/Free%20Workshop%20London%20UK-September%2030th,%202025-orange" alt="London UK: September 30th, 2025"></a> <a href="https://www.eventbrite.com/e/1601880590969"><img src="https://img.shields.io/badge/Free%20Workshop%20Washington%20DC-October%208th,%202025-orange" alt="Washington DC: October 8th, 2025"></a> <a href="https://www.eventbrite.com/e/1634269818019"><img src="https://img.shields.io/badge/Free%20R%20Training%20Washington%20DC-October%2015th,%202025-yellow" alt="Washington DC: October 15th, 2025"></a> <a href="https://www.eventbrite.com/e/1458850233319"><img src="https://img.shields.io/badge/Free%20Workshop%20New%20York%20NY-October%2022nd,%202025-orange" alt="New York NY: October 22nd, 2025"></a> <a href="https://www.eventbrite.com/e/1560552678069"><img src="https://img.shields.io/badge/Free%20Workshop%20Philadelphia%20PA-October%2029th,%202025-orange" alt="Philadelphia PA: October 29th, 2025"></a> <a href="https://www.eventbrite.com/e/1710882017169"><img src="https://img.shields.io/badge/Free%20Workshop%20Washington%20DC-November%2012th,%202025-orange" alt="Washington DC: November 12th, 2025"></a>

The primary survey to measure of prevalence of substance use and its correlates in the United States.

* One table with one row per sampled respondent.

* A complex survey designed to generalize to civilian, non-institutional americans aged 12 and older.

* Released periodically since 1979 and annually since 1990.

* Administered by the [Substance Abuse and Mental Health Services Administration](http://www.samhsa.gov/).

---

## Recommended Reading {-}

Four Example Strengths & Limitations:

✔️ [Detailed questions about substance use and mental health](https://www.samhsa.gov/data/sites/default/files/reports/rpt39458/2023NSDUHmrbWebCAISpecs013123.pdf#page=3)

✔️ [Restricted data can be pooled for state estimates](https://www.samhsa.gov/data/nsduh/state-reports-NSDUH-2022)

❌ [Traditional surveillance systems have lags associated with data collection, organization, and release](https://doi.org/10.1073/pnas.2207391120)

❌ [Data from 2021 should not be compared to estimates from 2019 or earlier](https://www.samhsa.gov/data/taxonomy/term/1244#term-1272)

<br>

Three Example Findings:

1. [Across 2002-2014, 36% of Americans aged 13-25 with ADHD used cannibis in the past month](http://doi.org/10.1056/NEJMoa1813751).

2. [In 2021, 23% of older adults with past-year non-medical opioid use received public assistance and only 42% reported very good or excellent health vs. 16% and 68% among those without such use](http://doi.org/10.3386/w31884).

3. [Among adolescents 12-17 in 2023, 7% used tobacco products or vaped nicotine in the past month](https://www.samhsa.gov/data/release/2023-national-survey-drug-use-and-health-nsduh-releases#highlights).

<br>

Two Methodology Documents:

> [2023 National Survey on Drug Use and Health (NSDUH): Public Use File Codebook](https://www.samhsa.gov/data/system/files/media-puf-file/NSDUH-2023-DS0001-info-codebook_v1.pdf)

> [2023 National Survey on Drug Use and Health (NSDUH): Methodological Summary and Definitions](https://www.samhsa.gov/data/report/2023-methodological-summary-and-definitions)

<br>

One Haiku:

```{r}
# drinking and thinking
# about your first time, were you
# smoking and joking?
```

---

## Download, Import, Preparation {-}

Download and import the national file:
```{r eval = FALSE , results = "hide" }
zip_tf <- tempfile()

zip_url <-
	"https://www.samhsa.gov/data/system/files/media-puf-file/NSDUH-2023-DS0001-bndl-data-r_v1.zip"
	
download.file( zip_url , zip_tf , mode = 'wb' )

nsduh_rdata <- unzip( zip_tf , exdir = tempdir() )

nsduh_rdata_contents <- load( nsduh_rdata )

nsduh_df_name <- grep( 'PUF' , nsduh_rdata_contents , value = TRUE , ignore.case = TRUE )

nsduh_df <- get( nsduh_df_name )

names( nsduh_df ) <- tolower( names( nsduh_df ) )

nsduh_df[ , 'one' ] <- 1
```

### Save Locally \ {-}

Save the object at any point:

```{r eval = FALSE , results = "hide" }
# nsduh_fn <- file.path( path.expand( "~" ) , "NSDUH" , "this_file.rds" )
# saveRDS( nsduh_df , file = nsduh_fn , compress = FALSE )
```

Load the same object:

```{r eval = FALSE , results = "hide" }
# nsduh_df <- readRDS( nsduh_fn )
```

### Survey Design Definition {-}
Construct a complex sample survey design:

```{r eval = FALSE , results = "hide" }
library(survey)

nsduh_design <- 
	svydesign( 
		id = ~ verep , 
		strata = ~ vestr_c , 
		data = nsduh_df , 
		weights = ~ analwt2_c , 
		nest = TRUE 
	)
```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE , results = "hide" }
nsduh_design <- 
	update( 
		nsduh_design , 
		
		one = 1 ,
		
		health = 
			factor( 
				health , 
				levels = 1:5 , 
				labels = c( "excellent" , "very good" , "good" ,
					"fair" , "poor" )
			) ,
			
		age_first_cigarette = ifelse( cigtry > 99 , NA , cigtry ) ,
		
		age_tried_cocaine = ifelse( cocage > 99 , NA , cocage ) ,

		ever_used_marijuana = as.numeric( ifelse( mjever < 4 , mjever == 1 , NA ) ) ,
		
		county_type =
			factor(
				coutyp4 ,
				levels = 1:3 ,
				labels = c( "large metro" , "small metro" , "nonmetro" )
			)
			
	)
```

---

## Analysis Examples with the `survey` library \ {-}

### Unweighted Counts {-}

Count the unweighted number of records in the survey sample, overall and by groups:
```{r eval = FALSE , results = "hide" }
sum( weights( nsduh_design , "sampling" ) != 0 )

svyby( ~ one , ~ county_type , nsduh_design , unwtd.count )
```

### Weighted Counts {-}
Count the weighted size of the generalizable population, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ one , nsduh_design )

svyby( ~ one , ~ county_type , nsduh_design , svytotal )
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ age_first_cigarette , nsduh_design , na.rm = TRUE )

svyby( ~ age_first_cigarette , ~ county_type , nsduh_design , svymean , na.rm = TRUE )
```

Calculate the distribution of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ health , nsduh_design , na.rm = TRUE )

svyby( ~ health , ~ county_type , nsduh_design , svymean , na.rm = TRUE )
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ age_first_cigarette , nsduh_design , na.rm = TRUE )

svyby( ~ age_first_cigarette , ~ county_type , nsduh_design , svytotal , na.rm = TRUE )
```

Calculate the weighted sum of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ health , nsduh_design , na.rm = TRUE )

svyby( ~ health , ~ county_type , nsduh_design , svytotal , na.rm = TRUE )
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svyquantile( ~ age_first_cigarette , nsduh_design , 0.5 , na.rm = TRUE )

svyby( 
	~ age_first_cigarette , 
	~ county_type , 
	nsduh_design , 
	svyquantile , 
	0.5 ,
	ci = TRUE , na.rm = TRUE
)
```

Estimate a ratio:
```{r eval = FALSE , results = "hide" }
svyratio( 
	numerator = ~ age_first_cigarette , 
	denominator = ~ age_tried_cocaine , 
	nsduh_design ,
	na.rm = TRUE
)
```

### Subsetting {-}

Restrict the survey design to individuals who are pregnant:
```{r eval = FALSE , results = "hide" }
sub_nsduh_design <- subset( nsduh_design , preg == 1 )
```
Calculate the mean (average) of this subset:
```{r eval = FALSE , results = "hide" }
svymean( ~ age_first_cigarette , sub_nsduh_design , na.rm = TRUE )
```

### Measures of Uncertainty {-}

Extract the coefficient, standard error, confidence interval, and coefficient of variation from any descriptive statistics function result, overall and by groups:
```{r eval = FALSE , results = "hide" }
this_result <- svymean( ~ age_first_cigarette , nsduh_design , na.rm = TRUE )

coef( this_result )
SE( this_result )
confint( this_result )
cv( this_result )

grouped_result <-
	svyby( 
		~ age_first_cigarette , 
		~ county_type , 
		nsduh_design , 
		svymean ,
		na.rm = TRUE 
	)
	
coef( grouped_result )
SE( grouped_result )
confint( grouped_result )
cv( grouped_result )
```

Calculate the degrees of freedom of any survey design object:
```{r eval = FALSE , results = "hide" }
degf( nsduh_design )
```

Calculate the complex sample survey-adjusted variance of any statistic:
```{r eval = FALSE , results = "hide" }
svyvar( ~ age_first_cigarette , nsduh_design , na.rm = TRUE )
```

Include the complex sample design effect in the result for a specific statistic:
```{r eval = FALSE , results = "hide" }
# SRS without replacement
svymean( ~ age_first_cigarette , nsduh_design , na.rm = TRUE , deff = TRUE )

# SRS with replacement
svymean( ~ age_first_cigarette , nsduh_design , na.rm = TRUE , deff = "replace" )
```

Compute confidence intervals for proportions using methods that may be more accurate near 0 and 1. See `?svyciprop` for alternatives:
```{r eval = FALSE , results = "hide" }
svyciprop( ~ ever_used_marijuana , nsduh_design ,
	method = "likelihood" , na.rm = TRUE )
```

### Regression Models and Tests of Association {-}

Perform a design-based t-test:
```{r eval = FALSE , results = "hide" }
svyttest( age_first_cigarette ~ ever_used_marijuana , nsduh_design )
```

Perform a chi-squared test of association for survey data:
```{r eval = FALSE , results = "hide" }
svychisq( 
	~ ever_used_marijuana + health , 
	nsduh_design 
)
```

Perform a survey-weighted generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	svyglm( 
		age_first_cigarette ~ ever_used_marijuana + health , 
		nsduh_design 
	)

summary( glm_result )
```

---

## Replication Example {-}

This matches the prevalence and SE of alcohol use in the past month from [Codebook Table F.2](https://www.samhsa.gov/data/system/files/media-puf-file/NSDUH-2023-DS0001-info-codebook_v1.pdf#page=701):

```{r eval = FALSE , results = "hide" }
result <- svymean( ~ alcmon , nsduh_design )

stopifnot( round( coef( result ) , 3 ) == 0.477 )
stopifnot( round( SE( result ) , 4 ) == 0.0049 )
```

---

## Analysis Examples with `srvyr` \ {-}

The R `srvyr` library calculates summary statistics from survey data, such as the mean, total or quantile using [dplyr](https://github.com/tidyverse/dplyr/)-like syntax. [srvyr](https://github.com/gergness/srvyr) allows for the use of many verbs, such as `summarize`, `group_by`, and `mutate`, the convenience of pipe-able functions, the `tidyverse` style of non-standard evaluation and more consistent return types than the `survey` package. [This vignette](https://cran.r-project.org/web/packages/srvyr/vignettes/srvyr-vs-survey.html) details the available features. As a starting point for NSDUH users, this code replicates previously-presented examples:

```{r eval = FALSE , results = "hide" }
library(srvyr)
nsduh_srvyr_design <- as_survey( nsduh_design )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
nsduh_srvyr_design %>%
	summarize( mean = survey_mean( age_first_cigarette , na.rm = TRUE ) )

nsduh_srvyr_design %>%
	group_by( county_type ) %>%
	summarize( mean = survey_mean( age_first_cigarette , na.rm = TRUE ) )
```


