# New York City Housing and Vacancy Survey (NYCHVS) {-}

<a href="https://github.com/asdfree/nychvs/actions"><img src="https://github.com/asdfree/nychvs/actions/workflows/r.yml/badge.svg" alt="Github Actions Badge"></a>

A city-wide assessment of the rental vacancy rate and other characteristics related to housing stock.

* One table with one record per occupied housing unit, a second table with one record per person inside each occupied housing unit, and a third table with one record per unoccupied housing unit.

* A complex sample designed to generalize to occupied & unoccupied housing in the five boroughs.

* Released more or less triennially since 1991.

* Funded by the [NYC Dept. of Housing Preservation & Development](www.nyc.gov/hpd), run by the [Census Bureau](http://www.census.gov/).

---

Please skim before you begin:

1. [Public Use File User Guide and Codebook](https://www2.census.gov/programs-surveys/nychvs/technical-documentation/record-layouts/2021/2021_NYCHVS_PUF_User_Guide_and_Codebook.pdf)

2. [Sample Design, Weighting, and Error Estimation](https://www2.census.gov/programs-surveys/nychvs/technical-documentation/source-and-accuracy/2021_NYCHVS_Sample_Design_Weighting_and_Error_Estimation.pdf)

3. This human-composed haiku or a bouquet of [artificial intelligence-generated limericks](https://www.gnod.com/search/ai#q=write%20a%20limerick-style%20poem%20about%20the%20New York City Housing and Vacancy Survey)

```{r}
# all i want is a
# room somewhere / with clawfoot tub
# and a frigidaire
```

---

## Download, Import, Preparation {-}

Define a function to download and import each comma-separated value file:
```{r eval = FALSE , results = "hide" }
nychvs_csv_import <-
	function( this_url ){
		
		tf <- tempfile()
		
		download.file( this_url , tf , mode = 'wb' )
		
		this_df <- read.csv( tf )
		
		names( this_df ) <- tolower( names( this_df ) )
		
		this_df
	}
```
	

Download and import the all units, occupied units, person, and vacant units tables:

```{r eval = FALSE , results = "hide" }
datasets_url <- "https://www2.census.gov/programs-surveys/nychvs/datasets/2021/microdata/"

all_units_df <-
	nychvs_csv_import( paste0( datasets_url , "allunits_puf_21.csv" ) )

occupied_units_df <-
	nychvs_csv_import( paste0( datasets_url , "occupied_puf_21.csv" ) )

person_df <-
	nychvs_csv_import( paste0( datasets_url , "person_puf_21.csv" ) )

vacant_units_df <-
	nychvs_csv_import( paste0( datasets_url , "vacant_puf_21.csv" ) )

stopifnot( nrow( all_units_df ) == nrow( occupied_units_df ) + nrow( vacant_units_df ) )
```

Merge the information stored in the all units table onto both the occupied and vacant unit tables:

```{r eval = FALSE , results = "hide" }
before_nrow <- nrow( occupied_units_df )

occupied_units_df <- merge( all_units_df , occupied_units_df )

stopifnot( nrow( occupied_units_df ) == before_nrow )

before_nrow <- nrow( vacant_units_df )

vacant_units_df <- merge( all_units_df , vacant_units_df )

stopifnot( nrow( vacant_units_df ) == before_nrow )

stopifnot( nrow( nychvs_df ) == nrow( nychvs_df_person ) )

nychvs_df[ , 'one' ] <- 1
```

### Save locally \ {-}

Save the object at any point:

```{r eval = FALSE , results = "hide" }
# nychvs_fn <- file.path( path.expand( "~" ) , "NYCHVS" , "this_file.rds" )
# saveRDS( nychvs_df , file = nychvs_fn , compress = FALSE )
```

Load the same object:

```{r eval = FALSE , results = "hide" }
# nychvs_df <- readRDS( nychvs_fn )
```

### Survey Design Definition {-}
Construct a complex sample survey design:

```{r message = FALSE , eval = FALSE }
library(survey)

nychvs_design <-
	svrepdesign(
		weight = ~fw ,
		repweights = 'fw[0-9]+' ,
		scale = 4 / 80 ,
		rscales = rep( 1 , 80 ) ,
		mse = TRUE ,
		type = 'JK1' ,
		data = nychvs_df
	)
```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE , results = "hide" }
nychvs_design <- 
	update( 
		nychvs_design , 
		
		one = 1 ,
		
		home_owners = as.numeric( sc115 == 1 ) ,

		yearly_household_income = ifelse( uf42 == 9999999 , 0 , as.numeric( uf42 ) ) ,
		
		gross_monthly_rent = ifelse( uf17 == 99999 , NA , as.numeric( uf17 ) ) ,
		
		borough =
			factor( boro , levels = 1:5 , labels =
				c( 'Bronx' , 'Brooklyn' , 'Manhattan' , 
				'Queens' , 'Staten Island' )
			) ,
			
		householder_sex = factor( hhr2 , labels = c( 'male' , 'female' ) )
			
	)
```

---

## Analysis Examples with the `survey` library \ {-}

### Unweighted Counts {-}

Count the unweighted number of records in the survey sample, overall and by groups:
```{r eval = FALSE , results = "hide" }
sum( weights( nychvs_design , "sampling" ) != 0 )

svyby( ~ one , ~ borough , nychvs_design , unwtd.count )
```

### Weighted Counts {-}
Count the weighted size of the generalizable population, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ one , nychvs_design )

svyby( ~ one , ~ borough , nychvs_design , svytotal )
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ yearly_household_income , nychvs_design , na.rm = TRUE )

svyby( ~ yearly_household_income , ~ borough , nychvs_design , svymean , na.rm = TRUE )
```

Calculate the distribution of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ householder_sex , nychvs_design )

svyby( ~ householder_sex , ~ borough , nychvs_design , svymean )
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ yearly_household_income , nychvs_design , na.rm = TRUE )

svyby( ~ yearly_household_income , ~ borough , nychvs_design , svytotal , na.rm = TRUE )
```

Calculate the weighted sum of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ householder_sex , nychvs_design )

svyby( ~ householder_sex , ~ borough , nychvs_design , svytotal )
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svyquantile( ~ yearly_household_income , nychvs_design , 0.5 , na.rm = TRUE )

svyby( 
	~ yearly_household_income , 
	~ borough , 
	nychvs_design , 
	svyquantile , 
	0.5 ,
	ci = TRUE , na.rm = TRUE
)
```

Estimate a ratio:
```{r eval = FALSE , results = "hide" }
svyratio( 
	numerator = ~ gross_monthly_rent , 
	denominator = ~ yearly_household_income , 
	nychvs_design ,
	na.rm = TRUE
)
```

### Subsetting {-}

Restrict the survey design to Manhattan:
```{r eval = FALSE , results = "hide" }
sub_nychvs_design <- subset( nychvs_design , boro == 3 )
```
Calculate the mean (average) of this subset:
```{r eval = FALSE , results = "hide" }
svymean( ~ yearly_household_income , sub_nychvs_design , na.rm = TRUE )
```

### Measures of Uncertainty {-}

Extract the coefficient, standard error, confidence interval, and coefficient of variation from any descriptive statistics function result, overall and by groups:
```{r eval = FALSE , results = "hide" }
this_result <- svymean( ~ yearly_household_income , nychvs_design , na.rm = TRUE )

coef( this_result )
SE( this_result )
confint( this_result )
cv( this_result )

grouped_result <-
	svyby( 
		~ yearly_household_income , 
		~ borough , 
		nychvs_design , 
		svymean ,
		na.rm = TRUE 
	)
	
coef( grouped_result )
SE( grouped_result )
confint( grouped_result )
cv( grouped_result )
```

Calculate the degrees of freedom of any survey design object:
```{r eval = FALSE , results = "hide" }
degf( nychvs_design )
```

Calculate the complex sample survey-adjusted variance of any statistic:
```{r eval = FALSE , results = "hide" }
svyvar( ~ yearly_household_income , nychvs_design , na.rm = TRUE )
```

Include the complex sample design effect in the result for a specific statistic:
```{r eval = FALSE , results = "hide" }
# SRS without replacement
svymean( ~ yearly_household_income , nychvs_design , na.rm = TRUE , deff = TRUE )

# SRS with replacement
svymean( ~ yearly_household_income , nychvs_design , na.rm = TRUE , deff = "replace" )
```

Compute confidence intervals for proportions using methods that may be more accurate near 0 and 1. See `?svyciprop` for alternatives:
```{r eval = FALSE , results = "hide" }
svyciprop( ~ home_owners , nychvs_design ,
	method = "likelihood" )
```

### Regression Models and Tests of Association {-}

Perform a design-based t-test:
```{r eval = FALSE , results = "hide" }
svyttest( yearly_household_income ~ home_owners , nychvs_design )
```

Perform a chi-squared test of association for survey data:
```{r eval = FALSE , results = "hide" }
svychisq( 
	~ home_owners + householder_sex , 
	nychvs_design 
)
```

Perform a survey-weighted generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	svyglm( 
		yearly_household_income ~ home_owners + householder_sex , 
		nychvs_design 
	)

summary( glm_result )
```

---

## Intermish {-}

---

## Replication Example {-}

Match the statistic and standard error shown at 6:05 of https://www.nyc.gov/assets/hpd/downloads/misc/gettingstartedwithnychvs_r_video.mp4
```{r eval = FALSE , results = "hide" }
svytotal( ~ one , nychvs_design )
```

---

## Analysis Examples with `srvyr` \ {-}

The R `srvyr` library calculates summary statistics from survey data, such as the mean, total or quantile using [dplyr](https://github.com/tidyverse/dplyr/)-like syntax. [srvyr](https://github.com/gergness/srvyr) allows for the use of many verbs, such as `summarize`, `group_by`, and `mutate`, the convenience of pipe-able functions, the `tidyverse` style of non-standard evaluation and more consistent return types than the `survey` package. [This vignette](https://cran.r-project.org/web/packages/srvyr/vignettes/srvyr-vs-survey.html) details the available features. As a starting point for NYCHVS users, this code replicates previously-presented examples:

```{r eval = FALSE , results = "hide" }
library(srvyr)
nychvs_srvyr_design <- as_survey( nychvs_design )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
nychvs_srvyr_design %>%
	summarize( mean = survey_mean( yearly_household_income , na.rm = TRUE ) )

nychvs_srvyr_design %>%
	group_by( borough ) %>%
	summarize( mean = survey_mean( yearly_household_income , na.rm = TRUE ) )
```


