# Programme for the International Assessment of Adult Competencies (PIAAC) {-}

<a href="https://github.com/asdfree/piaac/actions"><img src="https://github.com/asdfree/piaac/actions/workflows/r.yml/badge.svg" alt="Github Actions Badge"></a>

A cross-national study designed to understand the skills of workers in advanced-nation labor markets.

* One row per sampled adult.

* A multiply-imputed, complex sample survey designed to generalize to the population aged 16 to 65 across thirty three OECD nations.

* No expected release timeline.

* Administered by the [Organisation for Economic Co-operation and Development](http://www.oecd.org/).

---

Please skim before you begin:

1. [Technical Report of the Survey of Adult Skills](https://www.oecd.org/skills/piaac/publications/PIAAC_Technical_Report_2019.pdf)

2. [Wikipedia Entry](https://en.wikipedia.org/wiki/Programme_for_the_International_Assessment_of_Adult_Competencies)

3. This poem

```{r}
# what color collar
# workforce poets, potters, or
# pythagoreans
```

---

## Download, Import, Preparation {-}

```{r eval = FALSE }
library(haven)
library(httr)

tf <- tempfile()

this_url <- "https://webfs.oecd.org/piaac/puf-data/SAS/SAS7BDAT/prgusap1_2012.sas7bdat"

GET( this_url , write_disk( tf ) )

piaac_tbl <- read_sas( tf )

piaac_df <- data.frame( piaac_tbl )

names( piaac_df ) <- tolower( names( piaac_df ) )
```

### Save locally \ {-}

Save the object at any point:

```{r eval = FALSE , results = "hide" , messages = FALSE }
# piaac_fn <- file.path( path.expand( "~" ) , "PIAAC" , "this_file.rds" )
# saveRDS( piaac_df , file = piaac_fn , compress = FALSE )
```

Load the same object:

```{r eval = FALSE , results = "hide" , messages = FALSE }
# piaac_df <- readRDS( piaac_fn )
```

---

## Analysis Examples with the `survey` library \ {-}

Construct a multiply-imputed, complex sample survey design:

```{r messages = FALSE , eval = FALSE }
library(survey)
library(mitools)

pvals <- c( "pvlit" , "pvnum" , "pvpsl" )
pvars <- outer( pvals , 1:10 , paste0 )
non.pvals <- names(piaac_df)[ !( names(piaac_df) %in% pvars ) ]

for(k in 1:10){
	
	piaac_imp <- piaac_df[ , c( non.pvals , paste0( pvals , k ) ) ]
	
	for( j in pvals ){
		
		piaac_imp[ , j ] <- piaac_imp[ , paste0( j , k ) ]
		
		piaac_imp[ , paste0( j , k ) ] <- NULL

	}
	
	if( k == 1 ){
		piaac_mi <- list( piaac_imp )
	} else {
		piaac_mi <- c( piaac_mi , list( piaac_imp ) )
	}
}

jk.method <- unique( piaac_df[ , 'vemethod' ] )

stopifnot(length(jk.method) == 1)

stopifnot(jk.method %in% c("JK1", "JK2"))

if (jk.method == "JK2") jk.method <- "JKn"

piaac_design <-
	svrepdesign(
		weights = ~spfwt0 ,
		repweights = "spfwt[1-9]" ,
		rscales = rep( 1 , 80 ) ,
		scale = ifelse( jk.method == "JKn" , 1 , 79/80 ) ,
		type = jk.method ,
		data = imputationList( piaac_mi ) ,
		mse = TRUE
	)
```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE }
piaac_design <-
	update(
		piaac_design ,
		
		one = 1 ,
		
		sex = factor( gender_r , labels = c( "male" , "female" ) ) ,

		age_categories = 
			factor( 
				ageg10lfs , 
				levels = 1:5 , 
				labels = c( "24 or less" , "25-34" , "35-44" , "45-54" , "55 plus" ) 
			) ,
		
		working_at_paid_job_last_week = as.numeric( c_q01a == 1 )
		
	)
```

### Unweighted Counts {-}

Count the unweighted number of records in the survey sample, overall and by groups:
```{r eval = FALSE , results = "hide" }
MIcombine( with( piaac_design , svyby( ~ one , ~ one , unwtd.count ) ) )

MIcombine( with( piaac_design , svyby( ~ one , ~ age_categories , unwtd.count ) ) )
```

### Weighted Counts {-}
Count the weighted size of the generalizable population, overall and by groups:
```{r eval = FALSE , results = "hide" }
MIcombine( with( piaac_design , svytotal( ~ one ) ) )

MIcombine( with( piaac_design ,
	svyby( ~ one , ~ age_categories , svytotal )
) )
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
MIcombine( with( piaac_design , svymean( ~ pvnum , na.rm = TRUE ) ) )

MIcombine( with( piaac_design ,
	svyby( ~ pvnum , ~ age_categories , svymean , na.rm = TRUE )
) )
```

Calculate the distribution of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
MIcombine( with( piaac_design , svymean( ~ sex ) ) )

MIcombine( with( piaac_design ,
	svyby( ~ sex , ~ age_categories , svymean )
) )
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
MIcombine( with( piaac_design , svytotal( ~ pvnum , na.rm = TRUE ) ) )

MIcombine( with( piaac_design ,
	svyby( ~ pvnum , ~ age_categories , svytotal , na.rm = TRUE )
) )
```

Calculate the weighted sum of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
MIcombine( with( piaac_design , svytotal( ~ sex ) ) )

MIcombine( with( piaac_design ,
	svyby( ~ sex , ~ age_categories , svytotal )
) )
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
MIcombine( with( piaac_design ,
	svyquantile(
		~ pvnum ,
		0.5 , se = TRUE , na.rm = TRUE 
) ) )

MIcombine( with( piaac_design ,
	svyby(
		~ pvnum , ~ age_categories , svyquantile ,
		0.5 , se = TRUE ,
		ci = TRUE , na.rm = TRUE
) ) )
```

Estimate a ratio:
```{r eval = FALSE , results = "hide" }
MIcombine( with( piaac_design ,
	svyratio( numerator = ~ pvnum , denominator = ~ pvlit , na.rm = TRUE )
) )
```

### Subsetting {-}

Restrict the survey design to self-reported fair or poor health:
```{r eval = FALSE , results = "hide" }
sub_piaac_design <- subset( piaac_design , i_q08 %in% 4:5 )
```
Calculate the mean (average) of this subset:
```{r eval = FALSE , results = "hide" }
MIcombine( with( sub_piaac_design , svymean( ~ pvnum , na.rm = TRUE ) ) )
```

### Measures of Uncertainty {-}

Extract the coefficient, standard error, confidence interval, and coefficient of variation from any descriptive statistics function result, overall and by groups:
```{r eval = FALSE , results = "hide" }
this_result <-
	MIcombine( with( piaac_design ,
		svymean( ~ pvnum , na.rm = TRUE )
	) )

coef( this_result )
SE( this_result )
confint( this_result )
cv( this_result )

grouped_result <-
	MIcombine( with( piaac_design ,
		svyby( ~ pvnum , ~ age_categories , svymean , na.rm = TRUE )
	) )

coef( grouped_result )
SE( grouped_result )
confint( grouped_result )
cv( grouped_result )
```

Calculate the degrees of freedom of any survey design object:
```{r eval = FALSE , results = "hide" }
degf( piaac_design$designs[[1]] )
```

Calculate the complex sample survey-adjusted variance of any statistic:
```{r eval = FALSE , results = "hide" }
MIcombine( with( piaac_design , svyvar( ~ pvnum , na.rm = TRUE ) ) )
```

Include the complex sample design effect in the result for a specific statistic:
```{r eval = FALSE , results = "hide" }
# SRS without replacement
MIcombine( with( piaac_design ,
	svymean( ~ pvnum , na.rm = TRUE , deff = TRUE )
) )

# SRS with replacement
MIcombine( with( piaac_design ,
	svymean( ~ pvnum , na.rm = TRUE , deff = "replace" )
) )
```

Compute confidence intervals for proportions using methods that may be more accurate near 0 and 1. See `?svyciprop` for alternatives:
```{r eval = FALSE , results = "hide" }
# MIsvyciprop( ~ working_at_paid_job_last_week , piaac_design ,
# 	method = "likelihood" )
```

### Regression Models and Tests of Association {-}

Perform a design-based t-test:
```{r eval = FALSE , results = "hide" }
# MIsvyttest( pvnum ~ working_at_paid_job_last_week , piaac_design )
```

Perform a chi-squared test of association for survey data:
```{r eval = FALSE , results = "hide" }
# MIsvychisq( ~ working_at_paid_job_last_week + sex , piaac_design )
```

Perform a survey-weighted generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	MIcombine( with( piaac_design ,
		svyglm( pvnum ~ working_at_paid_job_last_week + sex )
	) )
	
summary( glm_result )
```

---

## Intermish {-}

<center>https://en.wikipedia.org/wiki/Lorem_ipsum</center>

---

## Replication Example {-}

This example below matches the statistics and standard errors from the [OECD's Technical Report Table 18.9 on PDF page 455](https://www.oecd.org/skills/piaac/_Technical Report_17OCT13.pdf#page=455).

```{r eval = FALSE , results = "hide" }
usa_pvlit <-
	MIcombine( with( piaac_design , svymean( ~ pvlit , na.rm = TRUE ) ) )
	
usa_pvnum <-
	MIcombine( with( piaac_design , svymean( ~ pvnum , na.rm = TRUE ) ) )

usa_pvpsl <-
	MIcombine( with( piaac_design , svymean( ~ pvpsl , na.rm = TRUE ) ) )

stopifnot( round( coef( usa_pvlit ) ) == 270 )
stopifnot( round( SE( usa_pvlit ) , 1 ) == 1.0 )
stopifnot( round( coef( usa_pvnum ) ) == 253 )
stopifnot( round( SE( usa_pvnum ) , 1 ) == 1.2 )
stopifnot( round( coef( usa_pvpsl ) ) == 277 )
stopifnot( round( SE( usa_pvpsl ) , 1 ) == 1.1 )

```
