# Pesquisa de Orcamentos Familiares (POF) {-}

<a href="https://github.com/asdfree/pof/actions"><img src="https://github.com/asdfree/pof/actions/workflows/r.yml/badge.svg" alt="Github Actions Badge"></a>

The Pesquisa de Orcamentos Familiares is Brazil's national survey of household budgets.

* One table of survey responses per sampled household. Additional tables, many containing one record per expenditure.

* A complex sample survey designed to generalize to the civilian population of Brazil.

* Released at irregular intervals, with only 2002-2003 and 2008-2009 microdata available.

* Administered by the [Instituto Brasileiro de Geografia e Estatistica](http://www.ibge.gov.br/).

---

Please skim before you begin:

1. [Pesquisa de Orçamentos Familiares 2017-2018 Perfil das despesas no Brasil](https://biblioteca.ibge.gov.br/visualizacao/livros/liv101761.pdf)

2. 

3. This human-composed haiku or a bouquet of [artificial intelligence-generated limericks](https://www.gnod.com/search/ai#q=write%20a%20limerick-style%20poem%20about%20the%20Pesquisa de Orcamentos Familiares)

```{r}
# shopping na praia
# roupa, comida, pede
# tres havaianas
```

---

## Download, Import, Preparation {-}

Download and import the dictionary file:

```{r eval = FALSE , results = "hide" }
library(archive)

dictionary_tf <- tempfile()

dictionary_url <-
	paste0(
		"https://ftp.ibge.gov.br/Orcamentos_Familiares/" ,
		"Pesquisa_de_Orcamentos_Familiares_2017_2018/Microdados/Documentacao_20230713.zip"
	)

download.file( dictionary_url , dictionary_tf , mode = 'wb' )

dictionary_files <- archive_extract( dictionary_tf , dir = tempdir() )
```

Import the variable dictionary:

```{r eval = FALSE , results = "hide" }
library(readxl)

dictionary_fn <- file.path( tempdir() , "Dicionários de váriaveis.xls" )

dictionary_tbl <- read_excel( dictionary_fn , sheet = "Morador" , skip = 3 )

dictionary_df <- data.frame( dictionary_tbl )

names( dictionary_df ) <- c( 'position' , 'length' , 'decimals' , 'column_name' , 'description' , 'variable_labels' )

dictionary_df[ c( 'position' , 'length' , 'decimals' ) ] <- sapply( dictionary_df[ c( 'position' , 'length' , 'decimals' ) ] , as.integer )

dictionary_df <- subset( dictionary_df , !is.na( position ) )

```

Import the post-stratification totals:
```{r eval = FALSE , results = "hide" }

post_stratification_fn <- file.path( tempdir() , "Pos_estratos_totais.xlsx" )

post_stratification_tbl <- read_excel( post_stratification_fn , skip = 5 )
	
post_stratification_df <- data.frame( post_stratification_tbl )

names( post_stratification_df ) <- c( 'estrato_pof' , 'pos_estrato' , 'total_pessoas' , 'uf' , 'cod_upa' )
```

Download the latest quarterly file:

```{r eval = FALSE , results = "hide" }
this_tf <- tempfile()

this_url <-
	paste0(
		"https://ftp.ibge.gov.br/Orcamentos_Familiares/" ,
		"Pesquisa_de_Orcamentos_Familiares_2017_2018/Microdados/Dados_20230713.zip"
	)

download.file( this_url , this_tf , mode = 'wb' )

unzipped_files <- unzip( this_tf , exdir = tempdir() )

morador_fn <- grep( 'MORADOR\.txt$' , unzipped_files , value = TRUE )
```

Import the latest quarterly file:

```{r eval = FALSE , results = "hide" }
library(readr)

morador_tbl <-
	read_fwf(
		morador_fn ,
		fwf_widths( 
			widths = dictionary_df[ , 'length' ] , 
			col_names = dictionary_df[ , 'column_name' ] 
		) ,
		col_types = 
			paste0( ifelse( is.na( dictionary_df[ , 'decimals' ] ) , "c" , "d" ) , collapse = '' )
	)

morador_df <- data.frame( morador_tbl )

names( morador_df ) <- tolower( names( morador_df ) )

```

Import the latest quarterly file:

```{r eval = FALSE , results = "hide" }
pof_df <- merge( morador_df , post_stratification_df )

stopifnot( nrow( pof_df ) == nrow( morador_df ) )
```

### Save locally \ {-}

Save the object at any point:

```{r eval = FALSE , results = "hide" }
# pof_fn <- file.path( path.expand( "~" ) , "POF" , "this_file.rds" )
# saveRDS( pof_df , file = pof_fn , compress = FALSE )
```

Load the same object:

```{r eval = FALSE , results = "hide" }
# pof_df <- readRDS( pof_fn )
```

### Survey Design Definition {-}
Construct a complex sample survey design:

```{r eval = FALSE , results = "hide" }
options( survey.lonely.psu = "adjust" )

library(survey)

pre_stratified_design <- 
	svydesign(
		id = ~ cod_upa , 
		strata = ~ estrato_pof ,
		weights = ~ peso_final ,
		data = pof_df ,
		nest = TRUE
	)

population_totals <- 
	data.frame(
		pos_estrato = unique( pof_df[ , 'pos_estrato' ] ) , 
		Freq = unique( pof_df[ , 'total_pessoas' ] ) 
	)

pof_design <-
	postStratify(
		pre_stratified_design , 
		~ pos_estrato , 
		population_totals
	)

```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE , results = "hide" }

pof_design <- 
	update(
		pof_design , 
		
		one = 1 ,
		
		# centimeters instead of meters
		altura_imputado = altura_imputado / 100 ,
		
		age_categories =
			factor( 
				1 + findInterval( idade_anos , 
					c( 20 , 25 , 30 , 35 , 45 , 55 , 65 , 75 ) ) ,
				levels = 1:9 , labels = c( "under 20" , "20-24" , "25-29" ,
				"30-34" , "35-44" , "45-54" , "55-64" , "65-74" , "75+" )
			) ,
		
		# create a body mass index (bmi) variable, excluding babies (who have altura_imputado==0)			
		body_mass_index = ifelse( altura_imputado == 0 , 0 , peso_imputado / ( altura_imputado ^ 2 ) ) ,
		
		sexo = ifelse( cod_sexo == '01' , "masculino" , ifelse( cod_sexo == '02' , "feminino" , NA ) )
		
		
	)

pof_design <-
	transform(
		pof_design ,
		
		# individuals with a low bmi - underweight
		underweight = ifelse( body_mass_index < 18.5 , 1 , 0 ) ,
		
		# individuals with a high bmi - overweight
		overweight = ifelse( body_mass_index >= 25 , 1 , 0 ) ,
		
		# individuals with a very high bmi - obese
		obese = ifelse( body_mass_index >= 30 , 1 , 0 )
	)

```

---

## Analysis Examples with the `survey` library \ {-}

### Unweighted Counts {-}

Count the unweighted number of records in the survey sample, overall and by groups:
```{r eval = FALSE , results = "hide" }
sum( weights( pof_design , "sampling" ) != 0 )

svyby( ~ one , ~ sexo , pof_design , unwtd.count )
```

### Weighted Counts {-}
Count the weighted size of the generalizable population, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ one , pof_design )

svyby( ~ one , ~ sexo , pof_design , svytotal )
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ body_mass_index , pof_design , na.rm = TRUE )

svyby( ~ body_mass_index , ~ sexo , pof_design , svymean , na.rm = TRUE )
```

Calculate the distribution of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ age_categories , pof_design )

svyby( ~ age_categories , ~ sexo , pof_design , svymean )
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ body_mass_index , pof_design , na.rm = TRUE )

svyby( ~ body_mass_index , ~ sexo , pof_design , svytotal , na.rm = TRUE )
```

Calculate the weighted sum of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ age_categories , pof_design )

svyby( ~ age_categories , ~ sexo , pof_design , svytotal )
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svyquantile( ~ body_mass_index , pof_design , 0.5 , na.rm = TRUE )

svyby( 
	~ body_mass_index , 
	~ sexo , 
	pof_design , 
	svyquantile , 
	0.5 ,
	ci = TRUE , na.rm = TRUE
)
```

Estimate a ratio:
```{r eval = FALSE , results = "hide" }
svyratio( 
	numerator = ~ peso_imputado , 
	denominator = ~ altura_imputado , 
	pof_design ,
	na.rm = TRUE
)
```

### Subsetting {-}

Restrict the survey design to :
```{r eval = FALSE , results = "hide" }
sub_pof_design <- subset( pof_design , underweight == 1 )
```
Calculate the mean (average) of this subset:
```{r eval = FALSE , results = "hide" }
svymean( ~ body_mass_index , sub_pof_design , na.rm = TRUE )
```

### Measures of Uncertainty {-}

Extract the coefficient, standard error, confidence interval, and coefficient of variation from any descriptive statistics function result, overall and by groups:
```{r eval = FALSE , results = "hide" }
this_result <- svymean( ~ body_mass_index , pof_design , na.rm = TRUE )

coef( this_result )
SE( this_result )
confint( this_result )
cv( this_result )

grouped_result <-
	svyby( 
		~ body_mass_index , 
		~ sexo , 
		pof_design , 
		svymean ,
		na.rm = TRUE 
	)
	
coef( grouped_result )
SE( grouped_result )
confint( grouped_result )
cv( grouped_result )
```

Calculate the degrees of freedom of any survey design object:
```{r eval = FALSE , results = "hide" }
degf( pof_design )
```

Calculate the complex sample survey-adjusted variance of any statistic:
```{r eval = FALSE , results = "hide" }
svyvar( ~ body_mass_index , pof_design , na.rm = TRUE )
```

Include the complex sample design effect in the result for a specific statistic:
```{r eval = FALSE , results = "hide" }
# SRS without replacement
svymean( ~ body_mass_index , pof_design , na.rm = TRUE , deff = TRUE )

# SRS with replacement
svymean( ~ body_mass_index , pof_design , na.rm = TRUE , deff = "replace" )
```

Compute confidence intervals for proportions using methods that may be more accurate near 0 and 1. See `?svyciprop` for alternatives:
```{r eval = FALSE , results = "hide" }
svyciprop( ~ obese , pof_design ,
	method = "likelihood" , na.rm = TRUE )
```

### Regression Models and Tests of Association {-}

Perform a design-based t-test:
```{r eval = FALSE , results = "hide" }
svyttest( body_mass_index ~ obese , pof_design )
```

Perform a chi-squared test of association for survey data:
```{r eval = FALSE , results = "hide" }
svychisq( 
	~ obese + age_categories , 
	pof_design 
)
```

Perform a survey-weighted generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	svyglm( 
		body_mass_index ~ obese + age_categories , 
		pof_design 
	)

summary( glm_result )
```

---

## Intermish {-}

<center>https://en.wikipedia.org/wiki/Lorem_ipsum</center>

---

## Replication Example {-}

This example matches statistics and coefficients of variation from [Tabela 4092 - Pessoas de 14 anos ou mais de idade, por condição em relação à força de trabalho e condição de ocupação](https://sidra.ibge.gov.br/tabela/4092):

```{r eval = FALSE , results = "hide" }
nationwide_adult_population <- svytotal( ~ pia , pof_design , na.rm = TRUE )
	
stopifnot( round( coef( nationwide_adult_population ) / 1000000 , 3 ) == 174.228 )
stopifnot( round( cv( nationwide_adult_population ) / 1000000 , 3 ) == 0 )
	
nationwide_labor_force <- svytotal( ~ pea_c , pof_design , na.rm = TRUE )

stopifnot( round( coef( nationwide_labor_force ) / 1000000 , 3 ) == 107.257 )
stopifnot( round( cv( nationwide_labor_force ) * 100 , 1 ) == 0.2 )
	
nationwide_employed <- svytotal( ~ ocup_c , pof_design , na.rm = TRUE )

stopifnot( round( coef( nationwide_employed ) / 1000000 , 3 ) == 97.825 )
stopifnot( round( cv( nationwide_employed ) * 100 , 1 ) == 0.2 )
	
nationwide_unemployed <- svytotal( ~ desocup30 , pof_design , na.rm = TRUE )

stopifnot( round( coef( nationwide_unemployed ) / 1000000 , 3 ) == 9.432 )
stopifnot( round( cv( nationwide_unemployed ) * 100 , 1 ) == 1.2 )
	
nationwide_not_in_labor_force <-
	svytotal( ~ as.numeric( pia & !pea_c ) , pof_design , na.rm = TRUE )

stopifnot( round( coef( nationwide_not_in_labor_force ) / 1000000 , 3 ) == 66.972 )
stopifnot( round( cv( nationwide_not_in_labor_force ) * 100 , 1 ) == 0.3 )
	

```

---

## Poverty and Inequality Estimation with `convey` \ {-}

The R `convey` library estimates measures of income concentration, poverty, inequality, and wellbeing. [This textbook](https://guilhermejacob.github.io/context/) details the available features. As a starting point for POF users, this code calculates the gini coefficient on complex sample survey data:

```{r eval = FALSE , results = "hide" }
library(convey)
pof_design <- convey_prep( pof_design )

svygini( ~ vd4020n , pof_design , na.rm = TRUE )
```

---

## Analysis Examples with `srvyr` \ {-}

The R `srvyr` library calculates summary statistics from survey data, such as the mean, total or quantile using [dplyr](https://github.com/tidyverse/dplyr/)-like syntax. [srvyr](https://github.com/gergness/srvyr) allows for the use of many verbs, such as `summarize`, `group_by`, and `mutate`, the convenience of pipe-able functions, the `tidyverse` style of non-standard evaluation and more consistent return types than the `survey` package. [This vignette](https://cran.r-project.org/web/packages/srvyr/vignettes/srvyr-vs-survey.html) details the available features. As a starting point for POF users, this code replicates previously-presented examples:

```{r eval = FALSE , results = "hide" }
library(srvyr)
pof_srvyr_design <- as_survey( pof_design )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
pof_srvyr_design %>%
	summarize( mean = survey_mean( body_mass_index , na.rm = TRUE ) )

pof_srvyr_design %>%
	group_by( sexo ) %>%
	summarize( mean = survey_mean( body_mass_index , na.rm = TRUE ) )
```


