# Pesquisa de Orcamentos Familiares (POF) {-}

[![License: GPL v3](https://img.shields.io/badge/License-GPLv3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0) <a href="https://github.com/asdfree/pof/actions"><img src="https://github.com/asdfree/pof/actions/workflows/r.yml/badge.svg" alt="Github Actions Badge"></a> <a href="https://www.eventbrite.com/e/1972097244947"><img src="https://img.shields.io/badge/Free%20Workshop%20Washington%20DC-December%203rd,%202025-orange" alt="Washington DC: December 3rd, 2025"></a>

Brazil's household budget survey designed to guide major economic indicators like the Contas nacionais.

* Various tables with one record per sampled household, resident, job, expenditure.

* A complex sample survey designed to generalize to the civilian population of Brazil.

* Released at irregular intervals, 2002-2003, 2008-2009, and 2017-2018 microdata available.

* Administered by the [Instituto Brasileiro de Geografia e Estatistica](http://www.ibge.gov.br/).

---

Please skim before you begin:

1. [Pesquisa de Orçamentos Familiares 2017-2018 Perfil das despesas no Brasil](https://biblioteca.ibge.gov.br/visualizacao/livros/liv101761.pdf)

2. [Conceitos e métodos](https://www.ibge.gov.br/estatisticas/sociais/saude/24786-pesquisa-de-orcamentos-familiares-2.html?=&t=conceitos-e-metodos)

3. A haiku regarding this microdata:

```{r}
# shopping na praia
# roupa, comida, pede
# tres havaianas
```

---

## Download, Import, Preparation {-}

Download the dictionary files:

```{r eval = FALSE , results = "hide" }
library(archive)

dictionary_tf <- tempfile()

dictionary_url <-
	paste0(
		"https://ftp.ibge.gov.br/Orcamentos_Familiares/" ,
		"Pesquisa_de_Orcamentos_Familiares_2017_2018/Microdados/Documentacao_20230713.zip"
	)

download.file( dictionary_url , dictionary_tf , mode = 'wb' )

dictionary_files <- archive_extract( dictionary_tf , dir = tempdir() )
```

Import the household variable dictionary:

```{r eval = FALSE , results = "hide" }
library(readxl)

dictionary_fn <- file.path( tempdir() , "Dicionários de váriaveis.xls" )

domicilio_dictionary_tbl <- read_excel( dictionary_fn , sheet = "Domicílio" , skip = 3 )

domicilio_dictionary_df <- data.frame( domicilio_dictionary_tbl )

names( domicilio_dictionary_df ) <-
	c( 'position' , 'length' , 'decimals' , 'column_name' , 'description' , 'variable_labels' )

domicilio_dictionary_df[ c( 'position' , 'length' , 'decimals' ) ] <-
	sapply( domicilio_dictionary_df[ c( 'position' , 'length' , 'decimals' ) ] , as.integer )

domicilio_dictionary_df <- subset( domicilio_dictionary_df , !is.na( position ) )
```

Import the resident variable dictionary:

```{r eval = FALSE , results = "hide" }
morador_dictionary_tbl <- read_excel( dictionary_fn , sheet = "Morador" , skip = 3 )

morador_dictionary_df <- data.frame( morador_dictionary_tbl )

names( morador_dictionary_df ) <-
	c( 'position' , 'length' , 'decimals' , 'column_name' , 'description' , 'variable_labels' )

morador_dictionary_df[ c( 'position' , 'length' , 'decimals' ) ] <-
	sapply( morador_dictionary_df[ c( 'position' , 'length' , 'decimals' ) ] , as.integer )

morador_dictionary_df <- subset( morador_dictionary_df , !is.na( position ) )
```

Import the post-stratification totals:
```{r eval = FALSE , results = "hide" }

post_stratification_fn <- file.path( tempdir() , "Pos_estratos_totais.xlsx" )

post_stratification_tbl <- read_excel( post_stratification_fn , skip = 5 )
	
post_stratification_df <- data.frame( post_stratification_tbl )

names( post_stratification_df ) <-
	c( 'estrato_pof' , 'pos_estrato' , 'total_pessoas' , 'uf' , 'cod_upa' )
```

Download the full dataset:

```{r eval = FALSE , results = "hide" }
this_tf <- tempfile()

this_url <-
	paste0(
		"https://ftp.ibge.gov.br/Orcamentos_Familiares/" ,
		"Pesquisa_de_Orcamentos_Familiares_2017_2018/Microdados/Dados_20230713.zip"
	)

download.file( this_url , this_tf , mode = 'wb' )

unzipped_files <- unzip( this_tf , exdir = tempdir() )
```

Import the household table:

```{r eval = FALSE , results = "hide" }
library(readr)

domicilio_fn <- grep( 'DOMICILIO\\.txt$' , unzipped_files , value = TRUE )

domicilio_tbl <-
	read_fwf(
		domicilio_fn ,
		fwf_widths( 
			widths = domicilio_dictionary_df[ , 'length' ] , 
			col_names = domicilio_dictionary_df[ , 'column_name' ] 
		)
	)

domicilio_df <- data.frame( domicilio_tbl )

names( domicilio_df ) <- tolower( names( domicilio_df ) )

```

Import the resident table:

```{r eval = FALSE , results = "hide" }
morador_fn <- grep( 'MORADOR\\.txt$' , unzipped_files , value = TRUE )

morador_tbl <-
	read_fwf(
		morador_fn ,
		fwf_widths( 
			widths = morador_dictionary_df[ , 'length' ] , 
			col_names = morador_dictionary_df[ , 'column_name' ] 
		)
	)

morador_df <- data.frame( morador_tbl )

names( morador_df ) <- tolower( names( morador_df ) )

```

Merge one household-level variable and also the post-stratification info onto the person-level table:

```{r eval = FALSE , results = "hide" }
dom_mor_df <- merge( domicilio_df[ c( 'cod_upa' , 'num_dom' , 'v6199' ) ] , morador_df )

pof_df <- merge( dom_mor_df , post_stratification_df )

stopifnot( nrow( pof_df ) == nrow( morador_df ) )
```

### Save Locally \ {-}

Save the object at any point:

```{r eval = FALSE , results = "hide" }
# pof_fn <- file.path( path.expand( "~" ) , "POF" , "this_file.rds" )
# saveRDS( pof_df , file = pof_fn , compress = FALSE )
```

Load the same object:

```{r eval = FALSE , results = "hide" }
# pof_df <- readRDS( pof_fn )
```

### Survey Design Definition {-}
Construct a complex sample survey design:

```{r eval = FALSE , results = "hide" }
library(survey)

options( survey.lonely.psu = "adjust" )

pre_stratified_design <- 
	svydesign(
		id = ~ cod_upa , 
		strata = ~ estrato_pof ,
		weights = ~ peso ,
		data = pof_df ,
		nest = TRUE
	)

population_totals <- 
	aggregate( peso_final ~ pos_estrato , data = pof_df , sum )
	
names( population_totals ) <- c( 'pos_estrato' , 'Freq' )

pof_design <-
	postStratify(
		pre_stratified_design , 
		~ pos_estrato , 
		population_totals
	)

```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE , results = "hide" }

pof_design <-
	update(
		pof_design ,
		
		one = 1 ,
		
		food_security =
			factor( 
				v6199 , 
				levels = 1:4 , 
				labels = c( 'food secure' , 'mild' , 'moderate' , 'severe' ) 
			) ,
	
		age_categories =
			factor( 
				1 + findInterval( v0403 , 
					c( 20 , 25 , 30 , 35 , 45 , 55 , 65 , 75 ) ) ,
				levels = 1:9 , 
				labels =
					c( "under 20" , "20-24" , "25-29" , "30-34" , "35-44" , 
					"45-54" , "55-64" , "65-74" , "75+" )
			) ,
		
		sex = factor( v0404 , levels = 1:2 , labels = c( 'male' , 'female' ) ) ,
		
		urban = as.numeric( tipo_situacao_reg == 1 )

	)

```

---

## Analysis Examples with the `survey` library \ {-}

### Unweighted Counts {-}

Count the unweighted number of records in the survey sample, overall and by groups:
```{r eval = FALSE , results = "hide" }
sum( weights( pof_design , "sampling" ) != 0 )

svyby( ~ one , ~ sex , pof_design , unwtd.count )
```

### Weighted Counts {-}
Count the weighted size of the generalizable population, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ one , pof_design )

svyby( ~ one , ~ sex , pof_design , svytotal )
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ renda_total , pof_design )

svyby( ~ renda_total , ~ sex , pof_design , svymean )
```

Calculate the distribution of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ age_categories , pof_design )

svyby( ~ age_categories , ~ sex , pof_design , svymean )
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ renda_total , pof_design )

svyby( ~ renda_total , ~ sex , pof_design , svytotal )
```

Calculate the weighted sum of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ age_categories , pof_design )

svyby( ~ age_categories , ~ sex , pof_design , svytotal )
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svyquantile( ~ renda_total , pof_design , 0.5 )

svyby( 
	~ renda_total , 
	~ sex , 
	pof_design , 
	svyquantile , 
	0.5 ,
	ci = TRUE 
)
```

Estimate a ratio:
```{r eval = FALSE , results = "hide" }
svyratio( 
	numerator = ~ renda_total , 
	denominator = ~ anos_estudo , 
	pof_design ,
	na.rm = TRUE
)
```

### Subsetting {-}

Restrict the survey design to credit card holders:
```{r eval = FALSE , results = "hide" }
sub_pof_design <- subset( pof_design , v0409 > 0 )
```
Calculate the mean (average) of this subset:
```{r eval = FALSE , results = "hide" }
svymean( ~ renda_total , sub_pof_design )
```

### Measures of Uncertainty {-}

Extract the coefficient, standard error, confidence interval, and coefficient of variation from any descriptive statistics function result, overall and by groups:
```{r eval = FALSE , results = "hide" }
this_result <- svymean( ~ renda_total , pof_design )

coef( this_result )
SE( this_result )
confint( this_result )
cv( this_result )

grouped_result <-
	svyby( 
		~ renda_total , 
		~ sex , 
		pof_design , 
		svymean 
	)
	
coef( grouped_result )
SE( grouped_result )
confint( grouped_result )
cv( grouped_result )
```

Calculate the degrees of freedom of any survey design object:
```{r eval = FALSE , results = "hide" }
degf( pof_design )
```

Calculate the complex sample survey-adjusted variance of any statistic:
```{r eval = FALSE , results = "hide" }
svyvar( ~ renda_total , pof_design )
```

Include the complex sample design effect in the result for a specific statistic:
```{r eval = FALSE , results = "hide" }
# SRS without replacement
svymean( ~ renda_total , pof_design , deff = TRUE )

# SRS with replacement
svymean( ~ renda_total , pof_design , deff = "replace" )
```

Compute confidence intervals for proportions using methods that may be more accurate near 0 and 1. See `?svyciprop` for alternatives:
```{r eval = FALSE , results = "hide" }
svyciprop( ~ urban , pof_design ,
	method = "likelihood" )
```

### Regression Models and Tests of Association {-}

Perform a design-based t-test:
```{r eval = FALSE , results = "hide" }
svyttest( renda_total ~ urban , pof_design )
```

Perform a chi-squared test of association for survey data:
```{r eval = FALSE , results = "hide" }
svychisq( 
	~ urban + age_categories , 
	pof_design 
)
```

Perform a survey-weighted generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	svyglm( 
		renda_total ~ urban + age_categories , 
		pof_design 
	)

summary( glm_result )
```

---

## Replication Example {-}
This example matches the 2017-2018 person-level food security estimates from [Tabela 3](
https://biblioteca.ibge.gov.br/visualizacao/livros/liv101749.pdf#page=32):

```{r eval = FALSE , results = "hide" }
person_level_food_security <- svymean( ~ food_security , pof_design , na.rm = TRUE )
	
stopifnot(
	all.equal(
		round( coef( person_level_food_security ) , 2 ) , 
		c( 0.59 , 0.27 , 0.09 , 0.05 ) , 
		check.attributes = FALSE 
	)
)
```

---

## Poverty and Inequality Estimation with `convey` \ {-}

The R `convey` library estimates measures of income concentration, poverty, inequality, and wellbeing. [This textbook](https://guilhermejacob.github.io/context/) details the available features. As a starting point for POF users, this code calculates the gini coefficient on complex sample survey data:

```{r eval = FALSE , results = "hide" }
library(convey)
pof_design <- convey_prep( pof_design )

svygini( ~ renda_total , pof_design , na.rm = TRUE )
```

---

## Analysis Examples with `srvyr` \ {-}

The R `srvyr` library calculates summary statistics from survey data, such as the mean, total or quantile using [dplyr](https://github.com/tidyverse/dplyr/)-like syntax. [srvyr](https://github.com/gergness/srvyr) allows for the use of many verbs, such as `summarize`, `group_by`, and `mutate`, the convenience of pipe-able functions, the `tidyverse` style of non-standard evaluation and more consistent return types than the `survey` package. [This vignette](https://cran.r-project.org/web/packages/srvyr/vignettes/srvyr-vs-survey.html) details the available features. As a starting point for POF users, this code replicates previously-presented examples:

```{r eval = FALSE , results = "hide" }
library(srvyr)
pof_srvyr_design <- as_survey( pof_design )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
pof_srvyr_design %>%
	summarize( mean = survey_mean( renda_total ) )

pof_srvyr_design %>%
	group_by( sex ) %>%
	summarize( mean = survey_mean( renda_total ) )
```


