# Residential Energy Consumption Survey (RECS) {-}

<a href="https://github.com/asdfree/recs/actions"><img src="https://github.com/asdfree/recs/actions/workflows/r.yml/badge.svg" alt="Github Actions Badge"></a>

dataset introduction

* table structure

* generalizable population

* publication period

* administrative organization

---

Please skim before you begin:

1. 

2. [2020 Residential Energy Consumption Survey: Household Characteristics Technical Documentation Summary](https://www.eia.gov/consumption/residential/data/2020/pdf/2020%20RECS_Methodology%20Report.pdf)

3. This human-composed haiku or a bouquet of [artificial intelligence-generated limericks](https://www.gnod.com/search/ai#q=write%20a%20limerick-style%20poem%20about%20the%20Residential Energy Consumption Survey)

```{r}
# 
# 
# 
```

---

## Download, Import, Preparation {-}

Download the sas file:
```{r eval = FALSE , results = "hide" }
sas_tf <- tempfile()

sas_url <- "https://www.eia.gov/consumption/residential/data/2020/sas/recs2020_public_v2.zip"

download.file( sas_url , sas_tf , mode = 'wb' )
```

Edit then execute the import script provided by the CDC:
```{r eval = FALSE , results = "hide" }
library(haven)

recs_tbl <- read_sas( sas_tf )

recs_df <- data.frame( recs_tbl )

names( recs_df ) <- tolower( names( recs_df ) )

recs_df[ , 'one' ] <- 1
```

### Save locally \ {-}

Save the object at any point:

```{r eval = FALSE , results = "hide" }
# recs_fn <- file.path( path.expand( "~" ) , "RECS" , "this_file.rds" )
# saveRDS( recs_df , file = recs_fn , compress = FALSE )
```

Load the same object:

```{r eval = FALSE , results = "hide" }
# recs_df <- readRDS( recs_fn )
```

### Survey Design Definition {-}
Construct a complex sample survey design:

```{r messages = FALSE , eval = FALSE }
library(survey)

recs_design <-
	svrepdesign(
		data = recs_df ,
		weight = ~ nweight ,
		repweights = 'nweight[1-9]+' ,
		type = 'JK1' ,
		combined.weights = TRUE ,
		scale = 59 / 60 ,
		mse = TRUE
	)
```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE , results = "hide" }
recs_design <- 
	
	update( 
		
		recs_design , 
		
		main_heating_fuel = 
			factor(
				fuelheat ,
				levels = c( -2 , 5 , 1 , 2 , 3 , 7 , 99 ) ,
				labels = 
					c(
						'Not applicable' , 
						'Electricity' , 
						'Natural gas from underground pipes' , 
						'Propane (bottled gas)' , 
						'Fuel oil' , 
						'Wood or pellets' , 
						'Other' 
					)
			) ,
			
		swimpool_binary =
			ifelse( swimpool %in% 0:1 , swimpool , NA )
			
	)
```

---

## Analysis Examples with the `survey` library \ {-}

### Unweighted Counts {-}

Count the unweighted number of records in the survey sample, overall and by groups:
```{r eval = FALSE , results = "hide" }
sum( weights( recs_design , "sampling" ) != 0 )

svyby( ~ one , ~ main_heating_fuel , recs_design , unwtd.count )
```

### Weighted Counts {-}
Count the weighted size of the generalizable population, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ one , recs_design )

svyby( ~ one , ~ main_heating_fuel , recs_design , svytotal )
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ totsqft_en , recs_design )

svyby( ~ totsqft_en , ~ main_heating_fuel , recs_design , svymean )
```

Calculate the distribution of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ state_name , recs_design )

svyby( ~ state_name , ~ main_heating_fuel , recs_design , svymean )
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ totsqft_en , recs_design )

svyby( ~ totsqft_en , ~ main_heating_fuel , recs_design , svytotal )
```

Calculate the weighted sum of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ state_name , recs_design )

svyby( ~ state_name , ~ main_heating_fuel , recs_design , svytotal )
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svyquantile( ~ totsqft_en , recs_design , 0.5 )

svyby( 
	~ totsqft_en , 
	~ main_heating_fuel , 
	recs_design , 
	svyquantile , 
	0.5 ,
	ci = TRUE 
)
```

Estimate a ratio:
```{r eval = FALSE , results = "hide" }
svyratio( 
	numerator = ~ totcsqft , 
	denominator = ~ totsqft_en , 
	recs_design 
)
```

### Subsetting {-}

Restrict the survey design to households that cook three or more hot meals per day:
```{r eval = FALSE , results = "hide" }
sub_recs_design <- subset( recs_design , nummeal == 1 )
```
Calculate the mean (average) of this subset:
```{r eval = FALSE , results = "hide" }
svymean( ~ totsqft_en , sub_recs_design )
```

### Measures of Uncertainty {-}

Extract the coefficient, standard error, confidence interval, and coefficient of variation from any descriptive statistics function result, overall and by groups:
```{r eval = FALSE , results = "hide" }
this_result <- svymean( ~ totsqft_en , recs_design )

coef( this_result )
SE( this_result )
confint( this_result )
cv( this_result )

grouped_result <-
	svyby( 
		~ totsqft_en , 
		~ main_heating_fuel , 
		recs_design , 
		svymean 
	)
	
coef( grouped_result )
SE( grouped_result )
confint( grouped_result )
cv( grouped_result )
```

Calculate the degrees of freedom of any survey design object:
```{r eval = FALSE , results = "hide" }
degf( recs_design )
```

Calculate the complex sample survey-adjusted variance of any statistic:
```{r eval = FALSE , results = "hide" }
svyvar( ~ totsqft_en , recs_design )
```

Include the complex sample design effect in the result for a specific statistic:
```{r eval = FALSE , results = "hide" }
# SRS without replacement
svymean( ~ totsqft_en , recs_design , deff = TRUE )

# SRS with replacement
svymean( ~ totsqft_en , recs_design , deff = "replace" )
```

Compute confidence intervals for proportions using methods that may be more accurate near 0 and 1. See `?svyciprop` for alternatives:
```{r eval = FALSE , results = "hide" }
svyciprop( ~ swimpool_binary , recs_design ,
	method = "likelihood" , na.rm = TRUE )
```

### Regression Models and Tests of Association {-}

Perform a design-based t-test:
```{r eval = FALSE , results = "hide" }
svyttest( totsqft_en ~ swimpool_binary , recs_design )
```

Perform a chi-squared test of association for survey data:
```{r eval = FALSE , results = "hide" }
svychisq( 
	~ swimpool_binary + state_name , 
	recs_design 
)
```

Perform a survey-weighted generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	svyglm( 
		totsqft_en ~ swimpool_binary + state_name , 
		recs_design 
	)

summary( glm_result )
```

---

## Intermish {-}

<center>https://en.wikipedia.org/wiki/Lorem_ipsum</center>

---

## Replication Example {-}

This example matches the statistic, standard error, and relative standard error on PDF page 8 of [2020 Residential Energy Consumption Survey: Using the microdata file to compute estimates and relative standard errors (RSEs)](https://www.eia.gov/consumption/residential/data/2020/pdf/microdata-guide.pdf)

```{r eval = FALSE , results = "hide" }
sas_v1_tf <- tempfile()

sas_v1_url <- "https://www.eia.gov/consumption/residential/data/2020/sas/recs2020_public_v1.zip"

download.file( sas_v1_url , sas_v1_tf , mode = 'wb' )

recs_v1_tbl <- read_sas( sas_v1_tf )

recs_v1_df <- data.frame( recs_v1_tbl )

names( recs_v1_df ) <- tolower( names( recs_v1_df ) )

recs_v1_design <-
 svrepdesign(
 data = recs_v1_df ,
 weight = ~ nweight ,
 repweights = 'nweight[1-9]+' ,
 type = 'JK1' ,
 combined.weights = TRUE ,
 scale = 59 / 60 ,
 mse = TRUE
 )
	
recs_v1_design <- 
	
	update( 
		
		recs_v1_design , 
		
		natural_gas_mainspace_heat = as.numeric( fuelheat == 1 )
			
	)
	
result <-
	svytotal( 
		~ natural_gas_mainspace_heat , 
		recs_v1_design 
	)

stopifnot( round( coef( result ) , 0 ) == 56245389 )
stopifnot( round( SE( result ) , 0 ) == 545591 )
stopifnot( round( 100 * SE( result ) / coef( result ) , 2 ) == 0.97 )
```

---

## Analysis Examples with `srvyr` \ {-}

The R `srvyr` library calculates summary statistics from survey data, such as the mean, total or quantile using [dplyr](https://github.com/tidyverse/dplyr/)-like syntax. [srvyr](https://github.com/gergness/srvyr) allows for the use of many verbs, such as `summarize`, `group_by`, and `mutate`, the convenience of pipe-able functions, the `tidyverse` style of non-standard evaluation and more consistent return types than the `survey` package. [This vignette](https://cran.r-project.org/web/packages/srvyr/vignettes/srvyr-vs-survey.html) details the available features. As a starting point for RECS users, this code replicates previously-presented examples:

```{r eval = FALSE , results = "hide" }
library(srvyr)
recs_srvyr_design <- as_survey( recs_design )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
recs_srvyr_design %>%
	summarize( mean = survey_mean( totsqft_en ) )

recs_srvyr_design %>%
	group_by( main_heating_fuel ) %>%
	summarize( mean = survey_mean( totsqft_en ) )
```


