# Survey of Income and Program Participation (SIPP) {-}

<a href="https://github.com/asdfree/sipp/actions"><img src="https://github.com/asdfree/sipp/actions/workflows/r.yml/badge.svg" alt="Github Actions Badge"></a>

The US Census Bureau's longitudinal assessment of the dynamics of income, labor force participation, social program participation and eligibility, with general demographic characteristics to measure the effectiveness of existing federal, state, and local programs.

* 

*

*

*

---

Please skim before you begin:

1. [2022 Survey of Income and Program Participation Users' Guide](https://www2.census.gov/programs-surveys/sipp/tech-documentation/methodology/2022_SIPP_Users_Guide_JUN23.pdf#page=15)

2. [2022 Data User Notes](https://www.census.gov/programs-surveys/sipp/tech-documentation/user-notes/2022-usernotes.html)

3. This human-composed haiku or a bouquet of [artificial intelligence-generated limericks](https://www.gnod.com/search/ai#q=write%20a%20limerick-style%20poem%20about%20the%20Survey of Income and Program Participation)

```{r}
# federal programs
# poverty oversample
# monthly dynamics
```

---

## Download, Import, Preparation {-}

Determine which variables from the main table to import:
```{r eval = FALSE , results = "hide" }
variables_to_keep <-
	c( 'ssuid' , 'pnum' , 'monthcode' , 'spanel' , 'swave' , 'erelrpe' , 
	'tlivqtr' , 'wpfinwgt' , 'rmesr' , 'thcyincpov' , 'tfcyincpov' ,
	'tehc_st' , 'rhicovann' , 'rfpov' , 'thnetworth' , 'tftotinc' )
```

Download and import the latest main file:
```{r eval = FALSE , results = "hide" }
library(httr)
library(data.table)

main_tf <- tempfile()

main_url <-
	paste0(
		"https://www2.census.gov/programs-surveys/sipp/" ,
		"data/datasets/2022/pu2022_csv.zip"
	)

GET( main_url , write_disk( main_tf ) , progress() )

main_csv <- unzip( main_tf , exdir = tempdir() )

sipp_main_dt <- fread( main_csv , sep = "|" , select = toupper( variables_to_keep ) )

sipp_main_df <- data.frame( sipp_main_dt )

names( sipp_main_df ) <- tolower( names( sipp_main_df ) )

```

Download and import the appropriate replicate weights file:
```{r eval = FALSE , results = "hide" }
rw_tf <- tempfile()

rw_url <-
	paste0(
		"https://www2.census.gov/programs-surveys/sipp/" ,
		"data/datasets/2022/rw2022_csv.zip"
	)

GET( rw_url , write_disk( rw_tf ) , progress() )

rw_csv <- unzip( rw_tf , exdir = tempdir() )

sipp_rw_dt <- fread( rw_csv , sep = "|" )

sipp_rw_df <- data.frame( sipp_rw_dt )

names( sipp_rw_df ) <- tolower( names( sipp_rw_df ) )
```

Merge the december records from these two files:
```{r eval = FALSE , results = "hide" }

sipp_df <-
	merge(
		sipp_main_df[ sipp_main_df[ , 'monthcode' ] %in% 12 , ] ,
		sipp_rw_df[ sipp_rw_df[ , 'monthcode' ] %in% 12 , ] ,
		by = c( 'ssuid' , 'pnum' , 'monthcode' , 'spanel' , 'swave' )
	)
	
stopifnot( nrow( sipp_df ) == sum( sipp_rw_df[ , 'monthcode' ] %in% 12 ) )
```

### Save locally \ {-}

Save the object at any point:

```{r eval = FALSE , results = "hide" }
# sipp_fn <- file.path( path.expand( "~" ) , "SIPP" , "this_file.rds" )
# saveRDS( sipp_df , file = sipp_fn , compress = FALSE )
```

Load the same object:

```{r eval = FALSE , results = "hide" }
# sipp_df <- readRDS( sipp_fn )
```

### Survey Design Definition {-}
Construct a complex sample survey design:

```{r message = FALSE , eval = FALSE }
library(survey)

sipp_design <- 
	svrepdesign(
			data = sipp_df ,
			weights = ~ wpfinwgt ,
			repweights = "repwgt([1-9]+)" ,
			type = "Fay" ,
			rho = 0.5
		)
	
```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE , results = "hide" }
rmesr_values <-
	c( 
		"With a job entire month, worked all weeks",
		"With a job all month, absent from work without pay 1+ weeks, absence not due to layoff",
		"With a job all month, absent from work without pay 1+ weeks, absence due to layoff",
		"With a job at least 1 but not all weeks, no time on layoff and no time looking for work",
		"With a job at least 1 but not all weeks, some weeks on layoff or looking for work",
		"No job all month, on layoff or looking for work all weeks",
		"No job all month, at least one but not all weeks on layoff or looking for work",
		"No job all month, no time on layoff and no time looking for work"
	)

sipp_design <- 
	
	update( 
		
		sipp_design , 
		
		one = 1 ,
		
		employment_status = factor( rmesr , levels = 1:8 , labels = rmesr_values ) ,
			
		household_below_poverty = as.numeric( thcyincpov < 1 ) ,
		
		family_below_poverty = as.numeric( tfcyincpov < 1 ) ,
		
		state_name =
			
			factor(
				
				as.numeric( tehc_st ) ,
				
				levels = 
					c(1L, 2L, 4L, 5L, 6L, 8L, 9L, 10L, 
					11L, 12L, 13L, 15L, 16L, 17L, 18L, 
					19L, 20L, 21L, 22L, 23L, 24L, 25L, 
					26L, 27L, 28L, 29L, 30L, 31L, 32L, 
					33L, 34L, 35L, 36L, 37L, 38L, 39L, 
					40L, 41L, 42L, 44L, 45L, 46L, 47L, 
					48L, 49L, 50L, 51L, 53L, 54L, 55L, 
					56L, 60L, 61L) ,
		
				labels =
					c("Alabama", "Alaska", "Arizona", "Arkansas", "California", 
					"Colorado", "Connecticut", "Delaware", "District of Columbia", 
					"Florida", "Georgia", "Hawaii", "Idaho", "Illinois", "Indiana", 
					"Iowa", "Kansas", "Kentucky", "Louisiana", "Maine", "Maryland", 
					"Massachusetts", "Michigan", "Minnesota", "Mississippi", "Missouri", 
					"Montana", "Nebraska", "Nevada", "New Hampshire", "New Jersey", 
					"New Mexico", "New York", "North Carolina", "North Dakota", "Ohio", 
					"Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", "South Carolina", 
					"South Dakota", "Tennessee", "Texas", "Utah", "Vermont", "Virginia", 
					"Washington", "West Virginia", "Wisconsin", "Wyoming", "Puerto Rico",
					"Foreign Country")
			)
			
	)
```

---

## Analysis Examples with the `survey` library \ {-}

### Unweighted Counts {-}

Count the unweighted number of records in the survey sample, overall and by groups:
```{r eval = FALSE , results = "hide" }
sum( weights( sipp_design , "sampling" ) != 0 )

svyby( ~ one , ~ state_name , sipp_design , unwtd.count )
```

### Weighted Counts {-}
Count the weighted size of the generalizable population, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ one , sipp_design )

svyby( ~ one , ~ state_name , sipp_design , svytotal )
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ tftotinc , sipp_design , na.rm = TRUE )

svyby( ~ tftotinc , ~ state_name , sipp_design , svymean , na.rm = TRUE )
```

Calculate the distribution of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svymean( ~ employment_status , sipp_design , na.rm = TRUE )

svyby( ~ employment_status , ~ state_name , sipp_design , svymean , na.rm = TRUE )
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ tftotinc , sipp_design , na.rm = TRUE )

svyby( ~ tftotinc , ~ state_name , sipp_design , svytotal , na.rm = TRUE )
```

Calculate the weighted sum of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svytotal( ~ employment_status , sipp_design , na.rm = TRUE )

svyby( ~ employment_status , ~ state_name , sipp_design , svytotal , na.rm = TRUE )
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
svyquantile( ~ tftotinc , sipp_design , 0.5 , na.rm = TRUE )

svyby( 
	~ tftotinc , 
	~ state_name , 
	sipp_design , 
	svyquantile , 
	0.5 ,
	ci = TRUE , na.rm = TRUE
)
```

Estimate a ratio:
```{r eval = FALSE , results = "hide" }
svyratio( 
	numerator = ~ tftotinc , 
	denominator = ~ rfpov , 
	sipp_design ,
	na.rm = TRUE
)
```

### Subsetting {-}

Restrict the survey design to individuals ever covered by health insurance during the year:
```{r eval = FALSE , results = "hide" }
sub_sipp_design <- subset( sipp_design , rhicovann == 1 )
```
Calculate the mean (average) of this subset:
```{r eval = FALSE , results = "hide" }
svymean( ~ tftotinc , sub_sipp_design , na.rm = TRUE )
```

### Measures of Uncertainty {-}

Extract the coefficient, standard error, confidence interval, and coefficient of variation from any descriptive statistics function result, overall and by groups:
```{r eval = FALSE , results = "hide" }
this_result <- svymean( ~ tftotinc , sipp_design , na.rm = TRUE )

coef( this_result )
SE( this_result )
confint( this_result )
cv( this_result )

grouped_result <-
	svyby( 
		~ tftotinc , 
		~ state_name , 
		sipp_design , 
		svymean ,
		na.rm = TRUE 
	)
	
coef( grouped_result )
SE( grouped_result )
confint( grouped_result )
cv( grouped_result )
```

Calculate the degrees of freedom of any survey design object:
```{r eval = FALSE , results = "hide" }
degf( sipp_design )
```

Calculate the complex sample survey-adjusted variance of any statistic:
```{r eval = FALSE , results = "hide" }
svyvar( ~ tftotinc , sipp_design , na.rm = TRUE )
```

Include the complex sample design effect in the result for a specific statistic:
```{r eval = FALSE , results = "hide" }
# SRS without replacement
svymean( ~ tftotinc , sipp_design , na.rm = TRUE , deff = TRUE )

# SRS with replacement
svymean( ~ tftotinc , sipp_design , na.rm = TRUE , deff = "replace" )
```

Compute confidence intervals for proportions using methods that may be more accurate near 0 and 1. See `?svyciprop` for alternatives:
```{r eval = FALSE , results = "hide" }
svyciprop( ~ family_below_poverty , sipp_design ,
	method = "likelihood" , na.rm = TRUE )
```

### Regression Models and Tests of Association {-}

Perform a design-based t-test:
```{r eval = FALSE , results = "hide" }
svyttest( tftotinc ~ family_below_poverty , sipp_design )
```

Perform a chi-squared test of association for survey data:
```{r eval = FALSE , results = "hide" }
svychisq( 
	~ family_below_poverty + employment_status , 
	sipp_design 
)
```

Perform a survey-weighted generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	svyglm( 
		tftotinc ~ family_below_poverty + employment_status , 
		sipp_design 
	)

summary( glm_result )
```

---

## Intermish {-}

<center>https://en.wikipedia.org/wiki/Lorem_ipsum</center>

---

## Replication Example {-}

This example matches statistics and standard errors from the [Wealth and Asset Ownership for Households, by Type of Asset and Selected Characteristics: 2021](https://www2.census.gov/programs-surveys/demo/tables/wealth/2021/wealth-asset-ownership/Wealth_tables_dy2021.xlsx):

Match the number of households shown in Table 4:

```{r eval = FALSE , results = "hide" }
sipp_household_design <- subset( sipp_design , erelrpe %in% 1:2 & tlivqtr %in% 1:2 )

stopifnot( round( coef( svytotal( ~ one , sipp_household_design ) ) / 1000 , -2 ) == 132700 )
```
	
Compute Household Net Worth distribution and standard errors across the Total row of Tables 4 and 4A:

```{r eval = FALSE , results = "hide" }
sipp_household_design <-
	update(
		sipp_household_design ,
		thnetworth_category =
			factor(
				findInterval( 
					thnetworth , 
					c( 1 , 5000 , 10000 , 25000 , 50000 , 100000 , 250000 , 500000 ) 
				) ,
				levels = 0:8 ,
				labels = c( "Zero or Negative" , "$1 to $4,999" , "$5,000 to $9,999" , 
				"$10,000 to $24,999" , "$25,000 to $49,999" , "$50,000 to $99,999" , 
				"$100,000 to $249,999" , "$250,000 to $499,999" , "$500,000 or over" )
			)
	)

results <- svymean( ~ thnetworth_category , sipp_household_design )

stopifnot( 
	all.equal( as.numeric( round( coef( results ) * 100 , 1 ) ) , 
	c( 11.3 , 6.9 , 3.5 , 5.9 , 6.4 , 8.2 , 15.3 , 13.9 , 28.7 ) ) 
)

stopifnot(
	all.equal( as.numeric( round( SE( results ) * 100 , 1 ) ) ,
	c( 0.3 , 0.2 , 0.2 , 0.2 , 0.2 , 0.2 , 0.3 , 0.3 , 0.3 ) )
)

```

---

## Poverty and Inequality Estimation with `convey` \ {-}

The R `convey` library estimates measures of income concentration, poverty, inequality, and wellbeing. [This textbook](https://guilhermejacob.github.io/context/) details the available features. As a starting point for SIPP users, this code calculates the gini coefficient on complex sample survey data:

```{r eval = FALSE , results = "hide" }
library(convey)
sipp_design <- convey_prep( sipp_design )

svygini( ~ tftotinc , sipp_design , na.rm = TRUE )
```

---

## Analysis Examples with `srvyr` \ {-}

The R `srvyr` library calculates summary statistics from survey data, such as the mean, total or quantile using [dplyr](https://github.com/tidyverse/dplyr/)-like syntax. [srvyr](https://github.com/gergness/srvyr) allows for the use of many verbs, such as `summarize`, `group_by`, and `mutate`, the convenience of pipe-able functions, the `tidyverse` style of non-standard evaluation and more consistent return types than the `survey` package. [This vignette](https://cran.r-project.org/web/packages/srvyr/vignettes/srvyr-vs-survey.html) details the available features. As a starting point for SIPP users, this code replicates previously-presented examples:

```{r eval = FALSE , results = "hide" }
library(srvyr)
sipp_srvyr_design <- as_survey( sipp_design )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
sipp_srvyr_design %>%
	summarize( mean = survey_mean( tftotinc , na.rm = TRUE ) )

sipp_srvyr_design %>%
	group_by( state_name ) %>%
	summarize( mean = survey_mean( tftotinc , na.rm = TRUE ) )
```


