# Social Security Administration Public Use Microdata (SSA) {-}

<a href="https://github.com/asdfree/ssa/actions"><img src="https://github.com/asdfree/ssa/actions/workflows/r.yml/badge.svg" alt="Github Actions Badge"></a>

Research extracts provided by the Social Security Administration.

* Tables contain either one record per person or one record per person per year.

* The entire population of either social security number holders (most of the country) or social security recipients (just beneficiaries). One-percent samples should be multiplied by 100 to get accurate nationwide count statistics, five-percent samples by 20.

* No expected release timeline.

* Released by the United States [Social Security Administration (SSA)](http://www.ssa.gov/).

---

Please skim before you begin:

1.

2.

3. This poem

```{r}
# 
# 
# 
```

---

## Download, Import, Preparation {-}

Download and import the national file:
```{r eval = FALSE }
library(haven)
library(httr)

tf <- tempfile()

ssa_url <- "https://www.ssa.gov/policy/docs/microdata/epuf/epuf2006_sas_files.zip"

GET( ssa_url , write_disk( tf ) )

ssa_files <- unzip( tf , exdir = tempdir() )

demographic_fn <- grep( 'demographic' , ssa_files , value = TRUE )

annual_fn <- grep( 'annual' , ssa_files , value = TRUE )

demographic_tbl <- read_sas( demographic_fn )

annual_tbl <- read_sas( annual_fn )

demographic_df <- data.frame( demographic_tbl )

annual_df <- data.frame( annual_tbl )

names( demographic_df ) <- tolower( names( demographic_df ) )

names( annual_df ) <- tolower( names( annual_df ) )
```

### Save locally \ {-}

Save the object at any point:

```{r eval = FALSE , results = "hide" , messages = FALSE }
# ssa_fn <- file.path( path.expand( "~" ) , "SSA" , "this_file.rds" )
# saveRDS( ssa_df , file = ssa_fn , compress = FALSE )
```

Load the same object:

```{r eval = FALSE , results = "hide" , messages = FALSE }
# ssa_df <- readRDS( ssa_fn )
```

---

## Analysis Examples with base R \ {-}

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE }
ssa_df <- 
	transform( 
		ssa_df , 
		
		mental_disorder = as.numeric( diag %in% 1:2 ) ,
		
		program_eligibility =
			factor( 
				prel , 
				
				levels = 0:5 , 
				
				labels =
					c( "Unspecified" ,
					"Aged individual" ,
					"Aged spouse" ,
					"Disabled or blind individual" ,
					"Disabled or blind spouse" ,
					"Disabled or blind child" )
			)
			
	)
	
```

### Unweighted Counts {-}

Count the unweighted number of records in the table, overall and by groups:
```{r eval = FALSE , results = "hide" }
nrow( ssa_df )

table( ssa_df[ , "stat" ] , useNA = "always" )
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
mean( ssa_df[ , "fpmt" ] )

tapply(
	ssa_df[ , "fpmt" ] ,
	ssa_df[ , "stat" ] ,
	mean 
)
```

Calculate the distribution of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
prop.table( table( ssa_df[ , "program_eligibility" ] ) )

prop.table(
	table( ssa_df[ , c( "program_eligibility" , "stat" ) ] ) ,
	margin = 2
)
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
sum( ssa_df[ , "fpmt" ] )

tapply(
	ssa_df[ , "fpmt" ] ,
	ssa_df[ , "stat" ] ,
	sum 
)
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
quantile( ssa_df[ , "fpmt" ] , 0.5 )

tapply(
	ssa_df[ , "fpmt" ] ,
	ssa_df[ , "stat" ] ,
	quantile ,
	0.5 
)
```

### Subsetting {-}

Limit your `data.frame` to females:
```{r eval = FALSE , results = "hide" }
sub_ssa_df <- subset( ssa_df , sex == "F" )
```
Calculate the mean (average) of this subset:
```{r eval = FALSE , results = "hide" }
mean( sub_ssa_df[ , "fpmt" ] )
```

### Measures of Uncertainty {-}

Calculate the variance, overall and by groups:
```{r eval = FALSE , results = "hide" }
var( ssa_df[ , "fpmt" ] )

tapply(
	ssa_df[ , "fpmt" ] ,
	ssa_df[ , "stat" ] ,
	var 
)
```

### Regression Models and Tests of Association {-}

Perform a t-test:
```{r eval = FALSE , results = "hide" }
t.test( fpmt ~ mental_disorder , ssa_df )
```

Perform a chi-squared test of association:
```{r eval = FALSE , results = "hide" }
this_table <- table( ssa_df[ , c( "mental_disorder" , "program_eligibility" ) ] )

chisq.test( this_table )
```

Perform a generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	glm( 
		fpmt ~ mental_disorder + program_eligibility , 
		data = ssa_df
	)

summary( glm_result )
```

---

## Intermish {-}

<center>https://en.wikipedia.org/wiki/Lorem_ipsum</center>

---

## Analysis Examples with `dplyr` \ {-}

The R `dplyr` library offers an alternative grammar of data manipulation to base R and SQL syntax. [dplyr](https://github.com/tidyverse/dplyr/) offers many verbs, such as `summarize`, `group_by`, and `mutate`, the convenience of pipe-able functions, and the `tidyverse` style of non-standard evaluation. [This vignette](https://cran.r-project.org/web/packages/dplyr/vignettes/dplyr.html) details the available features. As a starting point for SSA users, this code replicates previously-presented examples:

```{r eval = FALSE , results = "hide" }
library(dplyr)
ssa_tbl <- as_tibble( ssa_df )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
ssa_tbl %>%
	summarize( mean = mean( fpmt ) )

ssa_tbl %>%
	group_by( stat ) %>%
	summarize( mean = mean( fpmt ) )
```


