# Social Security Administration Public Use Microdata (SSA) {-}

<a href="https://github.com/asdfree/ssa/actions"><img src="https://github.com/asdfree/ssa/actions/workflows/r.yml/badge.svg" alt="Github Actions Badge"></a>

Research extracts provided by the Social Security Administration.

* Tables contain either one record per person or one record per person per year.

* The entire population of either social security number holders (most of the country) or social security recipients (just beneficiaries). One-percent samples should be multiplied by 100 to get accurate nationwide count statistics, five-percent samples by 20.

* No expected release timeline.

* Released by the United States [Social Security Administration (SSA)](http://www.ssa.gov/).

---

Please skim before you begin:

1.

2.

3. This poem

```{r}
# 
# 
# 
```

---

## Download, Import, Preparation {-}

Download and import the 1951-2006 topcoded earnings one percent file:
```{r eval = FALSE }
library(haven)
library(httr)

tf <- tempfile()

ssa_url <- "https://www.ssa.gov/policy/docs/microdata/epuf/epuf2006_sas_files.zip"

GET( ssa_url , write_disk( tf ) )

ssa_files <- unzip( tf , exdir = tempdir() )

ssa_fn <- grep( 'demographic' , ssa_files , value = TRUE )

annual_fn <- grep( 'annual' , ssa_files , value = TRUE )

ssa_tbl <- read_sas( ssa_fn )

annual_tbl <- read_sas( annual_fn )

ssa_df <- data.frame( ssa_tbl )

annual_df <- data.frame( annual_tbl )

names( ssa_df ) <- tolower( names( ssa_df ) )

names( annual_df ) <- tolower( names( annual_df ) )
```

Sum up 1951-1951 earnings, 1953-2006 earnings, and 1953-2006 credits:
```{r eval = FALSE }
summed_earnings_5152 <-
	with( 
		subset( annual_df , year_earn %in% 1951:1952 ) , 
		aggregate( annual_earnings , list( id ) , sum )
	)
	
names( summed_earnings_5152 ) <- c( 'id' , 'tot_cov_earn5152' )

summed_earnings_5306 <-
	with( 
		subset( annual_df , year_earn > 1952 ) , 
		aggregate( annual_earnings , list( id ) , sum )
	)
	
names( summed_earnings_5306 ) <- c( 'id' , 'tot_cov_earn5306' )

summed_quarters_5306 <-
	with( 
		subset( annual_df , year_earn > 1952 ) , 
		aggregate( annual_qtrs , list( id ) , sum )
	)

names( summed_quarters_5306 ) <- c( 'id' , 'qc5306' )
```

Isolate a single year of earnings:
```{r eval = FALSE }
earnings_2006 <- annual_df[ annual_df[ , 'year_earn' ] == 2006 , c( 'id' , 'annual_earnings' ) ]

names( earnings_2006 ) <- c( 'id' , 'annual_earnings_2006' )
```

Merge each of these columns on to the demographic table:
```{r eval = FALSE }
stopifnot( all( !is.na( ssa_df ) ) )

before_nrow <- nrow( ssa_df )

ssa_df <- merge( ssa_df , summed_earnings_5152 , all.x = TRUE )

ssa_df <- merge( ssa_df , summed_earnings_5306 , all.x = TRUE )

ssa_df <- merge( ssa_df , summed_quarters_5306 , all.x = TRUE )

ssa_df <- merge( ssa_df , earnings_2006 , all.x = TRUE )

ssa_df[ is.na( ssa_df ) ] <- 0

stopifnot( nrow( ssa_df ) == before_nrow )

```

### Save locally \ {-}

Save the object at any point:

```{r eval = FALSE , results = "hide" , messages = FALSE }
# ssa_fn <- file.path( path.expand( "~" ) , "SSA" , "this_file.rds" )
# saveRDS( ssa_df , file = ssa_fn , compress = FALSE )
```

Load the same object:

```{r eval = FALSE , results = "hide" , messages = FALSE }
# ssa_df <- readRDS( ssa_fn )
```

---

## Analysis Examples with base R \ {-}

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE }
ssa_df <- 
	transform( 
		ssa_df ,

		decade_of_birth = floor( yob / 10 ) * 10 ,
		
		sex = factor( sex , levels = 1:2 , labels = c( 'male' , 'female' ) )
		
		tot_cov_earn3706 = ( tot_cov_earn3750 + tot_cov_earn5152 + tot_cov_earn5306 ) ,
		
		qc3706 = ( qc3750 + qc5152 + qc5306 ) ,
		
		any_earnings_2006 = ( annual_earnings_2006 > 0 )
	)
```

### Unweighted Counts {-}

Count the unweighted number of records in the table, overall and by groups:
```{r eval = FALSE , results = "hide" }
nrow( ssa_df )

table( ssa_df[ , "sex" ] , useNA = "always" )
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
mean( ssa_df[ , "tot_cov_earn3706" ] )

tapply(
	ssa_df[ , "tot_cov_earn3706" ] ,
	ssa_df[ , "sex" ] ,
	mean 
)
```

Calculate the distribution of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
prop.table( table( ssa_df[ , "decade_of_birth" ] ) )

prop.table(
	table( ssa_df[ , c( "decade_of_birth" , "sex" ) ] ) ,
	margin = 2
)
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
sum( ssa_df[ , "tot_cov_earn3706" ] )

tapply(
	ssa_df[ , "tot_cov_earn3706" ] ,
	ssa_df[ , "sex" ] ,
	sum 
)
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
quantile( ssa_df[ , "tot_cov_earn3706" ] , 0.5 )

tapply(
	ssa_df[ , "tot_cov_earn3706" ] ,
	ssa_df[ , "sex" ] ,
	quantile ,
	0.5 
)
```

### Subsetting {-}

Limit your `data.frame` to at least forty lifetime credits:
```{r eval = FALSE , results = "hide" }
sub_ssa_df <- subset( ssa_df , qc3706 >= 40 )
```
Calculate the mean (average) of this subset:
```{r eval = FALSE , results = "hide" }
mean( sub_ssa_df[ , "tot_cov_earn3706" ] )
```

### Measures of Uncertainty {-}

Calculate the variance, overall and by groups:
```{r eval = FALSE , results = "hide" }
var( ssa_df[ , "tot_cov_earn3706" ] )

tapply(
	ssa_df[ , "tot_cov_earn3706" ] ,
	ssa_df[ , "sex" ] ,
	var 
)
```

### Regression Models and Tests of Association {-}

Perform a t-test:
```{r eval = FALSE , results = "hide" }
t.test( tot_cov_earn3706 ~ any_earnings_2006 , ssa_df )
```

Perform a chi-squared test of association:
```{r eval = FALSE , results = "hide" }
this_table <- table( ssa_df[ , c( "any_earnings_2006" , "decade_of_birth" ) ] )

chisq.test( this_table )
```

Perform a generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	glm( 
		tot_cov_earn3706 ~ any_earnings_2006 + decade_of_birth , 
		data = ssa_df
	)

summary( glm_result )
```

---

## Intermish {-}

<center>https://en.wikipedia.org/wiki/Lorem_ipsum</center>

---

## Analysis Examples with `dplyr` \ {-}

The R `dplyr` library offers an alternative grammar of data manipulation to base R and SQL syntax. [dplyr](https://github.com/tidyverse/dplyr/) offers many verbs, such as `summarize`, `group_by`, and `mutate`, the convenience of pipe-able functions, and the `tidyverse` style of non-standard evaluation. [This vignette](https://cran.r-project.org/web/packages/dplyr/vignettes/dplyr.html) details the available features. As a starting point for SSA users, this code replicates previously-presented examples:

```{r eval = FALSE , results = "hide" }
library(dplyr)
ssa_tbl <- as_tibble( ssa_df )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
ssa_tbl %>%
	summarize( mean = mean( tot_cov_earn3706 ) )

ssa_tbl %>%
	group_by( sex ) %>%
	summarize( mean = mean( tot_cov_earn3706 ) )
```

---

## Replication Example {-}

```{r eval = FALSE , results = "hide" }

```
