# Social Security Public-Use Data Files (SSA) {-}

[![License: GPL v3](https://img.shields.io/badge/License-GPLv3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0) <a href="https://github.com/asdfree/ssa/actions"><img src="https://github.com/asdfree/ssa/actions/workflows/r.yml/badge.svg" alt="Github Actions Badge"></a> <a href="https://www.eventbrite.com/e/1972097244947"><img src="https://img.shields.io/badge/Free%20Workshop%20Washington%20DC-December%203rd,%202025-orange" alt="Washington DC: December 3rd, 2025"></a>

Microdata from administrative sources like the Master Beneficiary Record, Supplemental Security Record.

* Tables contain either one record per person or one record per person per year.

* A systematic sample of either social security number holders (most americans) or program recipients (current beneficiaries). Multiply 1% samples by 100 to get weighted statistics, 5% samples by 20.

* [No expected release timeline](https://www.ssa.gov/policy/pub_schedule.html).

* Released by the [Office of Research, Evaluation, and Statistics](https://www.ssa.gov/policy/index.html), US [Social Security Administration](http://www.ssa.gov/).

---

## Recommended Reading {-}

Two Methodology Documents:

> [The 2006 Earnings Public-Use Microdata File: An Introduction](https://www.ssa.gov/policy/docs/ssb/v71n4/v71n4p33.html)

> [Comparing Earnings Estimates from the 2006 Public-Use File and the Annual Statistical Supplement](https://www.ssa.gov/policy/docs/rsnotes/rsn2012-01.html)

<br>

One Haiku:

```{r}
# annual earnings.
# for pensioner payouts, see
# the '04 extract
```

---

## Download, Import, Preparation {-}

Download and import the 1951-2006 one percent files with one record per person and per person-year:
```{r eval = FALSE , results = "hide" }
library(haven)
library(httr)

tf <- tempfile()

ssa_url <- "https://www.ssa.gov/policy/docs/microdata/epuf/epuf2006_sas_files.zip"

GET( ssa_url , write_disk( tf ) , progress() )

ssa_files <- unzip( tf , exdir = tempdir() )

ssa_fn <- grep( 'demographic' , ssa_files , value = TRUE )

annual_fn <- grep( 'annual' , ssa_files , value = TRUE )

ssa_tbl <- read_sas( ssa_fn )

annual_tbl <- read_sas( annual_fn )

ssa_df <- data.frame( ssa_tbl )

annual_df <- data.frame( annual_tbl )

names( ssa_df ) <- tolower( names( ssa_df ) )

names( annual_df ) <- tolower( names( annual_df ) )
```

Sum up 1951-1952 and 1953-2006 earnings, and also 1953-2006 credits, copying the naming convention:
```{r eval = FALSE , results = "hide" }
summed_earnings_5152 <-
	with( 
		subset( annual_df , year_earn %in% 1951:1952 ) , 
		aggregate( annual_earnings , list( id ) , sum )
	)
	
names( summed_earnings_5152 ) <- c( 'id' , 'tot_cov_earn5152' )

summed_earnings_5306 <-
	with( 
		subset( annual_df , year_earn > 1952 ) , 
		aggregate( annual_earnings , list( id ) , sum )
	)
	
names( summed_earnings_5306 ) <- c( 'id' , 'tot_cov_earn5306' )

summed_quarters_5306 <-
	with( 
		subset( annual_df , year_earn > 1952 ) , 
		aggregate( annual_qtrs , list( id ) , sum )
	)

names( summed_quarters_5306 ) <- c( 'id' , 'qc5306' )
```

Isolate a single year of earnings:
```{r eval = FALSE , results = "hide" }
earnings_2006 <- annual_df[ annual_df[ , 'year_earn' ] == 2006 , c( 'id' , 'annual_earnings' ) ]

names( earnings_2006 ) <- c( 'id' , 'tot_cov_earn06' )
```

Merge each new column on to the person-level table, then add zeroes to person-years without earnings:
```{r eval = FALSE , results = "hide" }
stopifnot( all( !is.na( ssa_df ) ) )

before_nrow <- nrow( ssa_df )

ssa_df <- merge( ssa_df , summed_earnings_5152 , all.x = TRUE )

ssa_df <- merge( ssa_df , summed_earnings_5306 , all.x = TRUE )

ssa_df <- merge( ssa_df , summed_quarters_5306 , all.x = TRUE )

ssa_df <- merge( ssa_df , earnings_2006 , all.x = TRUE )

ssa_df[ is.na( ssa_df ) ] <- 0

stopifnot( nrow( ssa_df ) == before_nrow )

```

### Save Locally \ {-}

Save the object at any point:

```{r eval = FALSE , results = "hide" }
# ssa_fn <- file.path( path.expand( "~" ) , "SSA" , "this_file.rds" )
# saveRDS( ssa_df , file = ssa_fn , compress = FALSE )
```

Load the same object:

```{r eval = FALSE , results = "hide" }
# ssa_df <- readRDS( ssa_fn )
```

### Variable Recoding {-}

Add new columns to the data set:
```{r eval = FALSE , results = "hide" }
ssa_df <- 
	transform( 
		ssa_df ,

		decade_of_birth = floor( yob / 10 ) * 10 ,
		
		sex = factor( sex , levels = 1:2 , labels = c( 'male' , 'female' ) ) ,
		
		tot_cov_earn3706 = ( tot_cov_earn3750 + tot_cov_earn5152 + tot_cov_earn5306 ) ,
		
		qc3706 = ( qc3750 + qc5152 + qc5306 ) ,
		
		any_earnings_2006 = ( tot_cov_earn06 > 0 ) ,
		
		earnings_periods =
			factor(
				ifelse( ( tot_cov_earn5152 + tot_cov_earn5306 > 0 ) & tot_cov_earn3750 > 0 , 1 ,
				ifelse( tot_cov_earn5152 > 0 | tot_cov_earn5306 > 0 , 2 ,
				ifelse( tot_cov_earn3750 > 0 , 3 , 4 ) ) ) ,
				levels = 1:4 ,
				labels =
					c( 'Earnings in both periods' , 'Earnings during 1951-2006 only' ,
						'Earnings during 1937-1950 only' , 'No earnings' ) )
	)
```

---

## Analysis Examples with base R \ {-}

### Unweighted Counts {-}

Count the unweighted number of records in the table, overall and by groups:
```{r eval = FALSE , results = "hide" }
nrow( ssa_df )

table( ssa_df[ , "sex" ] , useNA = "always" )
```

### Descriptive Statistics {-}

Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
mean( ssa_df[ , "tot_cov_earn3706" ] )

tapply(
	ssa_df[ , "tot_cov_earn3706" ] ,
	ssa_df[ , "sex" ] ,
	mean 
)
```

Calculate the distribution of a categorical variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
prop.table( table( ssa_df[ , "decade_of_birth" ] ) )

prop.table(
	table( ssa_df[ , c( "decade_of_birth" , "sex" ) ] ) ,
	margin = 2
)
```

Calculate the sum of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
sum( ssa_df[ , "tot_cov_earn3706" ] )

tapply(
	ssa_df[ , "tot_cov_earn3706" ] ,
	ssa_df[ , "sex" ] ,
	sum 
)
```

Calculate the median (50th percentile) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
quantile( ssa_df[ , "tot_cov_earn3706" ] , 0.5 )

tapply(
	ssa_df[ , "tot_cov_earn3706" ] ,
	ssa_df[ , "sex" ] ,
	quantile ,
	0.5 
)
```

### Subsetting {-}

Limit your `data.frame` to individuals with at least forty lifetime credits:
```{r eval = FALSE , results = "hide" }
sub_ssa_df <- subset( ssa_df , qc3706 >= 40 )
```
Calculate the mean (average) of this subset:
```{r eval = FALSE , results = "hide" }
mean( sub_ssa_df[ , "tot_cov_earn3706" ] )
```

### Measures of Uncertainty {-}

Calculate the variance, overall and by groups:
```{r eval = FALSE , results = "hide" }
var( ssa_df[ , "tot_cov_earn3706" ] )

tapply(
	ssa_df[ , "tot_cov_earn3706" ] ,
	ssa_df[ , "sex" ] ,
	var 
)
```

### Regression Models and Tests of Association {-}

Perform a t-test:
```{r eval = FALSE , results = "hide" }
t.test( tot_cov_earn3706 ~ any_earnings_2006 , ssa_df )
```

Perform a chi-squared test of association:
```{r eval = FALSE , results = "hide" }
this_table <- table( ssa_df[ , c( "any_earnings_2006" , "decade_of_birth" ) ] )

chisq.test( this_table )
```

Perform a generalized linear model:
```{r eval = FALSE , results = "hide" }
glm_result <- 
	glm( 
		tot_cov_earn3706 ~ any_earnings_2006 + decade_of_birth , 
		data = ssa_df
	)

summary( glm_result )
```

---

## Replication Example {-}

This example matches statistics in [The 2006 Earnings Public-Use Microdata File: An Introduction](https://www.ssa.gov/policy/docs/ssb/v71n4/v71n4p33.html):

Chart 5. Percentage distribution of individuals in EPUF, by capped Social Security taxable earnings status: 
```{r eval = FALSE , results = "hide" }
chart_five_results <- prop.table( table( ssa_df[ , 'earnings_periods' ] ) )
chart_five_results <- round( 100 * chart_five_results )

stopifnot( chart_five_results[ 'Earnings in both periods' ] == 16 )
stopifnot( chart_five_results[ 'Earnings during 1951-2006 only' ] == 55 )
stopifnot( chart_five_results[ 'Earnings during 1937-1950 only' ] == 4 )
stopifnot( chart_five_results[ 'No earnings' ] == 25 )
```

Table 4. Average and median Social Security taxable earnings in EPUF, by sex, 1951–2006 (in dollars):
```{r eval = FALSE , results = "hide" }
nonzero_2006_earners <- ssa_df[ ssa_df[ , 'tot_cov_earn06' ] > 0 , 'tot_cov_earn06' ]
stopifnot( round( mean( nonzero_2006_earners ) , 0 ) == 30953 )
stopifnot( round( quantile( nonzero_2006_earners )[ 3 ] , 0 ) == 24000 )
```

Table A1. Number and percentage distribution of individuals with Social Security taxable earnings records in EPUF, by sex, 1951–2006:
```{r eval = FALSE , results = "hide" }
nonzero_2006_earners <- ssa_df[ ssa_df[ , 'tot_cov_earn06' ] > 0 , ]
stopifnot( round( mean( nonzero_2006_earners[ , 'tot_cov_earn06' ] ) , 0 ) == 30953 )
stopifnot( round( quantile( nonzero_2006_earners[ , 'tot_cov_earn06' ] )[ 3 ] , 0 ) == 24000 )
```

This example matches statistics in [Comparing Earnings Estimates from the 2006 Earnings Public-Use File and the Annual Statistical Supplement](https://www.ssa.gov/policy/docs/rsnotes/rsn2012-01.html):

Table 4. Comparing Supplement and EPUF estimates: Number of all, male, and female workers with any earnings during the year, 1951–2006:

```{r eval = FALSE , results = "hide" }
stopifnot( round( nrow( nonzero_2006_earners ) * 100 , -3 ) == 156280000 )
earners_in_2006_by_sex <- table( nonzero_2006_earners[ , 'sex' ] ) * 100
stopifnot( round( earners_in_2006_by_sex[ 'male' ] , -3 ) == 81576000 )
stopifnot( round( earners_in_2006_by_sex[ 'female' ] , -3 ) == 74681000 )
```

---

## Analysis Examples with `dplyr` \ {-}

The R `dplyr` library offers an alternative grammar of data manipulation to base R and SQL syntax. [dplyr](https://github.com/tidyverse/dplyr/) offers many verbs, such as `summarize`, `group_by`, and `mutate`, the convenience of pipe-able functions, and the `tidyverse` style of non-standard evaluation. [This vignette](https://cran.r-project.org/web/packages/dplyr/vignettes/dplyr.html) details the available features. As a starting point for SSA users, this code replicates previously-presented examples:

```{r eval = FALSE , results = "hide" }
library(dplyr)
ssa_tbl <- as_tibble( ssa_df )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = "hide" }
ssa_tbl %>%
	summarize( mean = mean( tot_cov_earn3706 ) )

ssa_tbl %>%
	group_by( sex ) %>%
	summarize( mean = mean( tot_cov_earn3706 ) )
```

---

## Analysis Examples with `data.table` \ {-}

The R `data.table` library provides a high-performance version of base R's data.frame with syntax and feature enhancements for ease of use, convenience and programming speed. [data.table](https://r-datatable.com) offers concise syntax: fast to type, fast to read, fast speed, memory efficiency, a careful API lifecycle management, an active community, and a rich set of features. [This vignette](https://cran.r-project.org/web/packages/data.table/vignettes/datatable-intro.html) details the available features. As a starting point for SSA users, this code replicates previously-presented examples:

```{r eval = FALSE , results = 'hide' }
library(data.table)
ssa_dt <- data.table( ssa_df )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = 'hide' }
ssa_dt[ , mean( tot_cov_earn3706 ) ]

ssa_dt[ , mean( tot_cov_earn3706 ) , by = sex ]
```

---

## Analysis Examples with `duckdb` \ {-}

The R `duckdb` library provides an embedded analytical data management system with support for the Structured Query Language (SQL). [duckdb](https://duckdb.org) offers a simple, feature-rich, fast, and free SQL OLAP management system. [This vignette](https://duckdb.org/docs/api/r) details the available features. As a starting point for SSA users, this code replicates previously-presented examples:

```{r eval = FALSE , results = 'hide' }
library(duckdb)
con <- dbConnect( duckdb::duckdb() , dbdir = 'my-db.duckdb' )
dbWriteTable( con , 'ssa' , ssa_df )
```
Calculate the mean (average) of a linear variable, overall and by groups:
```{r eval = FALSE , results = 'hide' }
dbGetQuery( con , 'SELECT AVG( tot_cov_earn3706 ) FROM ssa' )

dbGetQuery(
	con ,
	'SELECT
		sex ,
		AVG( tot_cov_earn3706 )
	FROM
		ssa
	GROUP BY
		sex'
)
```
